### 2017-04-20 template/Template.Rmd
I did intial set up

---
# IMPORTANT: Change settings here, but DO NOT change the spacing. 
# Remove comments and add values where applicable. 
# The descriptions below should be self-explanatory

title: "An Analysis for Why Fiona Would Still Uses Edge and Bing"
subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# Comment: ----- Follow this pattern for up to 5 authors
Author1: "Nico Katzke"  # First Author
Ref1: "Bureau for Economic Research, Stellenbosch University, South Africa" # First Author's Affiliation
Email1: "nfkatzke\\@gmail.com" # First Author's Email address

Author2: "John Smith"
Ref2: "Some other Institution, Cape Town, South Africa"
Email2: "John Smith\\@gmail.com"
CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

Author3: "John Doe"
Email3: "JohnSmith\\@gmail.com"

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
JELCodes: "L250 \\sep L100"

# Comment: ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage\\" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
RemovePreprintSubmittedTo: FALSE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if above parameter is set to TRUE.
toc: no                         # Add a table of contents
numbersections: yes             # Should sections (and thus figures and tables) be numbered?
fontsize: 12pt                  # Set fontsize
linestretch: 1.5                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
    include:
      in_header: Tex/packages.txt # Reference file with extra packages
abstract: |
  Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf. These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

```

<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

References are to be made as follows: @fama1997[p. 33] and @grinold2000 Such authors could also be referenced in brackets [@grinold2000] and together [@fama1997 \& @grinold2000]. Source the reference code from scholar.google.com by clicking on ``cite'' below article name. Then select BibTeX at the bottom of the Cite window, and proceed to copy and paste this code into your ref.bib file, located in the directory's Tex folder. Open this file in Rstudio for ease of management, else open it in your preferred Tex environment. Add and manage your article details here for simplicity - once saved, it will self-adjust in your paper.

> I suggest renaming the top line after \@article, as done in the template ref.bib file, to something more intuitive for you to remember. Do not change the rest of the code. Also, be mindful of the fact that bib references from google scholar may at times be incorrect. Reference Latex forums for correct bibtex notation.

To reference a section, you have to set a label using ``\\label'' in R, and then reference it in-text as e.g.: section \ref{Data}.

Writing in Rmarkdown is surprizingly easy - see [this website](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf) cheatsheet for a summary on writing Rmd writing tips.

# Data \label{Data}

Discussion of data should be thorough with a table of statistics and ideally a figure.

In your tempalte folder, you will find a Data and a Code folder. In order to keep your data files neat, store all of them in your Data folder. Also, I strongly suggest keeping this Rmd file for writing and executing commands, not writing out long pieces of data-wrangling. In the example below, I simply create a ggplot template for scatter plot consistency. I suggest keeping all your data in a data folder.

<!-- The following is a code chunk. It must have its own unique name (after the r), or no name. After the comma follows commands for R which are self-explanatory. By default, the code and messages will not be printed in your pdf, just the output: -->

```{r Figure1, warning =  FALSE, fig.align = 'center', fig.cap = "Caption Here \\label{Figure1}", fig.ext = 'png', fig.height = 3, fig.width = 6}

library("tidyverse")
library("ggplot2")
library("ggthemes")

if (!file.exists(file.path(getwd(), "Code/Example_Plot_Scatter.R"))) {
  # This is just a random plot to show you a plot. This is done if the getwd() does not point to your Template's directory.
  # Set WD below to your project's working directory.

  g <-
  ggplot(data = mtcars %>% tbl_df()) + # All data must be tbl_df()::: class(mtcars %>% tbl_df())
  geom_point(aes(x = disp, y = hp, color = cyl)) +
  theme_bw() +
  # theme(legend.position = "none") +
  theme(plot.title = element_text(size = 10)) +
  theme(axis.title = element_text(size = 10), 
        axis.title.x = element_text()) +
  ylab("y-axis") +
  xlab("x-axis")

print(g)

} else {

# ----------  Using your Code and Data folders to manage your Template:
#------------------- To make this work, change WD to your project's directory:

  WD <- getwd()

# Load Example Data using read_csv from Hadley Wickham (or other csv reading methods of your choice):
ExampleData <- read_csv( file.path(WD, "Data/ExampleData.csv") )

source( file.path(WD, "Code/Example_Plot_Scatter.R") ) # Load a function from your 'Code' folder. This e.g. plots scatters to your own preferences.
source( file.path(WD, "Code/Data_Create.R") ) #

ExampleData <- Data_Create(N = 100) # Same as the ExampleData loaded above - this is just a function to create the stored data.

g <- Example_Plot_Scatter(DataInput = ExampleData, 
                          X = "Height_Score",
                          Y = "Weight_Score",
                          Z = "Agility_Score",
                          Theme = theme_bw(), # run ?ggthemes for other themes. 
                          Title = "", # I prefer to use the caption set above. 
                          Ylab = "Height", 
                          Xlab = "Lenght",
                          LegendTitle = "Agility",
                          TitleSize = 10, 
                          LabSize = 10)

print(g)

}

# Although the functions above are really simple, the principle is simple: containing calculations and data wrangling in their own functions will make this template much cleaner and manageable.

# When you start working, delete these meaningless functions and replace with your own...

```

To reference the plot above, add a ``\\label'' after the caption in the chunk heading, as done above. Then reference the plot as such: As can be seen, figure \ref{Figure1} is excellent. The nice thing now is that it correctly numbers all your figures (and sections or tables) and will update if it moves. The links are also dynamic.

I very strongly suggest using ggplot2 (ideally in combination with dplyr) using the ggtheme package to change the themes of your figures.

Also note the information that I have placed above the chunks in the code chunks for the figures. You can edit any of these easily - visit the Rmarkdown webpage for more information.

Here follows another figure from built-in ggplot2 data:

```{r figure2, warning =  FALSE, fig.align = 'center', fig.cap = "Diamond Cut Plot \\label{lit}", fig.ext = 'png', fig.height = 3, fig.width = 6}

g <-
ggplot(diamonds, aes(depth, fill = cut, colour = cut)) +
  geom_density(alpha = 0.3) +
  theme_bw()

print(g)

```

#  Methodology

## Subsection
Ideally do not overuse subsections. It equates to bad writing.^[This is an example of a footnote by the way. Something that should also not be overused.]

## Math section

Equations should be written as such:

\begin{align} 
\beta = \sum_{i = 1}^{\infty}\frac{\alpha^2}{\sigma_{t-1}^2} \label{eq1} \\ 
\int_{x = 1}^{\infty}x_{i} = 1 \notag
\end{align}

If you would like to see the equations as you type in Rmarkdown, use $ symbols instead (see this for yourself by adjusted the equation):

$$
\beta = \sum_{i = 1}^{\infty}\frac{\alpha^2}{\sigma_{t-1}^2} \\ 
\int_{x = 1}^{\infty}x_{i} = 1
$$

Note again the reference to equation \ref{eq1}. Writing nice math requires practice. Note I used a forward slashes to make a space in the equations. I can also align equations using ``\&'', and set to numbering only the first line. Now I will have to type ``begin equation'' which is a native \LaTeX command. Here follows a more complicated equation:

\begin{align} 
	y_t &= c + B(L) y_{t-1} + e_t   \label{eq2}    \\ \notag 
	e_t &= H_t^{1/2}  z_t ; \quad z_t \sim  N(0,I_N) \quad \& \quad H_t = D_tR_tD_t \\ \notag
		D_t^2 &= {\sigma_{1,t}, \dots, \sigma_{N,t}}   \\ \notag
		\sigma_{i,t}^2 &= \gamma_i+\kappa_{i,t}  v_{i, t-1}^2 +\eta_i  \sigma_{i, t-1}^2, \quad \forall i \\ \notag
		R_{t, i, j} &= {diag(Q_{t, i, j}}^{-1}) . Q_{t, i, j} . diag(Q_{t, i, j}^{-1})  \\ \notag
		Q_{t, i, j} &= (1-\alpha-\beta)  \bar{Q} + \alpha  z_t  z_t'  + \beta  Q_{t, i, j} \notag
\end{align}

Noten that in \ref{eq2} I have aligned the equations by the equal signs. I also want only one tag, and I create spaces using ``quads''.

See if you can figure out how to do complex math using the two examples provided in \ref{eq1} and \ref{eq2}.

<!-- $$ -->
<!-- This is a commented out section in the writing part. -->
<!-- Comments are created by highlighting text, amnd pressing CTL+C -->
<!-- \\begin{align} -->
<!-- \\beta = \\alpha^2 -->
<!-- \end{align} -->
<!-- $$ -->

# Results

Tables can be included as follows. Use the _xtable_ (or kable) package for tables. Table placement = H implies Latex tries to place the table Here, and not on a new page (there are, however, very many ways to skin this cat. Luckily there are many forums online!).

```{r ShortTable, results = 'asis'}

library(xtable)
data = mtcars[1:5,] %>% tbl_df()

table <- xtable(data, caption = "Short Table Example \\label{tab1}")
  print.xtable(table, 
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H', 
             # scalebox = 0.3, 
             comment = FALSE,
             caption.placement = 'bottom'
             )

```

To reference calculations __in text__, _do this:_ From table \ref{tab1} we see the average value of mpg is `r mean(mtcars[1:5,]$mpg)`.

Including tables that span across pages, use e.g.:

```{r LongTable, results = 'asis'}

library(xtable)
data = mtcars %>% tbl_df()
table <- xtable(data, caption = "Long Table Example")
  print.xtable(table, 
             tabular.environment = "longtable",
             floating = FALSE,
             table.placement = 'H', 
             comment = FALSE, caption.placement = 'bottom'
             )
# See https://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf for table inspiration
```

\hfill

<!-- hfill can be used to create a space, like here between text and table. -->

# Lists

To add lists, simply using the following notation

* This is really simple

    + Just note the spaces here - writing in R you have to sometimes be pedantic about spaces...

* Note that Rmarkdown notation removes the pain of defining \LaTeX environments!

# Conclusion

I hope you find this template useful. Remember, stackoverflow is your friend - use it to find answers to questions. Feel free to write me a mail if you have any questions regarding the use of this package feel free to email me. To cite this package, simply type citation("Texevier") in Rstudio to get the citation for @Texevier (Note that uncited references in your bibtex file will not be included in References).

<!-- Make title of bibliography here: -->
<!-- \newpage -->
# References

### 2017-04-20 template/Template.tex
I did intial set up

\documentclass[12pt,preprint, authoryear]{elsarticle}

\usepackage{lmodern}
%%%% My spacing
\usepackage{setspace}
\setstretch{1.5}
\DeclareMathSizes{12}{14}{10}{10}

% Wrap around which gives all figures included the [H] command, or places it "here". This can be tedious to code in Rmarkdown.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\let\origtable\table
\let\endorigtable\endtable
\renewenvironment{table}[1][2] {
    \expandafter\origtable\expandafter[H]
} {
    \endorigtable
}

\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi

\usepackage{amssymb, amsmath, amsthm, amsfonts}

\usepackage[round]{natbib}
\bibliographystyle{natbib}
\def\bibsection{\section*{References}} %%% Make "References" appear before bibliography
\usepackage{longtable}
\usepackage[margin=2cm,bottom=4cm,top=2.5cm, includefoot]{geometry}
\usepackage{fancyhdr}
\usepackage[bottom, hang, flushmargin]{footmisc}
\usepackage{graphicx}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}
\setlength{\parindent}{0cm}
\setlength{\parskip}{1.3ex plus 0.5ex minus 0.3ex}
\usepackage{textcomp}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.3pt}

\usepackage{array}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

%%%%  Remove the "preprint submitted to" part. Don't worry about this either, it just looks better without it:
\journal{Journal of Finance}

 \def\tightlist{} % This allows for subbullets!

\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

%%% Include extra packages specified by user
% Insert custom packages here as follows
% \usepackage{tikz}

\begin{document}

\begin{frontmatter}  %

\title{An Analysis for Why Fiona Would Still Uses Edge and Bing}

\author[Add1]{Nico Katzke}
\ead{nfkatzke@gmail.com}

\author[Add1,Add2]{John Smith}
\ead{John Smith@gmail.com}

\author[Add1,Add2]{John Doe}
\ead{JohnSmith@gmail.com}

\address[Add1]{Bureau for Economic Research, Stellenbosch University, South Africa}
\address[Add2]{Some other Institution, Cape Town, South Africa}

\cortext[cor]{Corresponding author: Nico Katzke}

\begin{abstract}
\small{
Abstract to be written here. The abstract should not be too long and
should provide the reader with a good understanding what you are writing
about. Academic papers are not like novels where you keep the reader in
suspense. To be effective in getting others to read your paper, be as
open and concise about your findings here as possible. Ideally, upon
reading your abstract, the reader should feel he / she must read your
paper in entirety.
}
\end{abstract}

\vspace{1cm}

\begin{keyword}
\footnotesize{
Multivariate GARCH \sep Kalman Filter \sep Copula \\ \vspace{0.3cm}
\textit{JEL classification} L250 \sep L100
}
\end{keyword}
\vspace{0.5cm}
\end{frontmatter}

%________________________
% Header and Footers
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{fancy}
\chead{}
\rhead{This will appear as Right Header}
\lfoot{}
\rfoot{\footnotesize Page \thepage\\}
\lhead{}
%\rfoot{\footnotesize Page \thepage\ } % "e.g. Page 2"
\cfoot{}

%\setlength\headheight{30pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%________________________

\headsep 35pt % So that header does not go over title

\section{\texorpdfstring{Introduction
\label{Introduction}}{Introduction }}\label{introduction}

References are to be made as follows: Fama and French
(\protect\hyperlink{ref-fama1997}{1997}, 33) and Grinold and Kahn
(\protect\hyperlink{ref-grinold2000}{2000}) Such authors could also be
referenced in brackets (Grinold and Kahn
\protect\hyperlink{ref-grinold2000}{2000}) and together (Fama and French
\protect\hyperlink{ref-fama1997}{1997} \& Grinold and Kahn
(\protect\hyperlink{ref-grinold2000}{2000})). Source the reference code
from scholar.google.com by clicking on ``cite'' below article name. Then
select BibTeX at the bottom of the Cite window, and proceed to copy and
paste this code into your ref.bib file, located in the directory's Tex
folder. Open this file in Rstudio for ease of management, else open it
in your preferred Tex environment. Add and manage your article details
here for simplicity - once saved, it will self-adjust in your paper.

\begin{quote}
I suggest renaming the top line after @article, as done in the template
ref.bib file, to something more intuitive for you to remember. Do not
change the rest of the code. Also, be mindful of the fact that bib
references from google scholar may at times be incorrect. Reference
Latex forums for correct bibtex notation.
\end{quote}

To reference a section, you have to set a label using
``\textbackslash{}label'' in R, and then reference it in-text as e.g.:
section \ref{Data}.

Writing in Rmarkdown is surprizingly easy - see
\href{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}{this
website} cheatsheet for a summary on writing Rmd writing tips.

\section{\texorpdfstring{Data \label{Data}}{Data }}\label{data}

Discussion of data should be thorough with a table of statistics and
ideally a figure.

In your tempalte folder, you will find a Data and a Code folder. In
order to keep your data files neat, store all of them in your Data
folder. Also, I strongly suggest keeping this Rmd file for writing and
executing commands, not writing out long pieces of data-wrangling. In
the example below, I simply create a ggplot template for scatter plot
consistency. I suggest keeping all your data in a data folder.

\begin{figure}[H]

{\centering \includegraphics{Template_files/figure-latex/Figure1-1}

}

\caption{Caption Here \label{Figure1}}\label{fig:Figure1}
\end{figure}

To reference the plot above, add a ``\textbackslash{}label'' after the
caption in the chunk heading, as done above. Then reference the plot as
such: As can be seen, figure \ref{Figure1} is excellent. The nice thing
now is that it correctly numbers all your figures (and sections or
tables) and will update if it moves. The links are also dynamic.

I very strongly suggest using ggplot2 (ideally in combination with
dplyr) using the ggtheme package to change the themes of your figures.

Also note the information that I have placed above the chunks in the
code chunks for the figures. You can edit any of these easily - visit
the Rmarkdown webpage for more information.

Here follows another figure from built-in ggplot2 data:

\begin{figure}[H]

{\centering \includegraphics{Template_files/figure-latex/figure2-1}

}

\caption{Diamond Cut Plot \label{lit}}\label{fig:figure2}
\end{figure}

\section{Methodology}\label{methodology}

\subsection{Subsection}\label{subsection}

Ideally do not overuse subsections. It equates to bad writing.\footnote{This
  is an example of a footnote by the way. Something that should also not
  be overused.}

\subsection{Math section}\label{math-section}

Equations should be written as such:

\begin{align} 
\beta = \sum_{i = 1}^{\infty}\frac{\alpha^2}{\sigma_{t-1}^2} \label{eq1} \\ 
\int_{x = 1}^{\infty}x_{i} = 1 \notag
\end{align}

If you would like to see the equations as you type in Rmarkdown, use \$
symbols instead (see this for yourself by adjusted the equation):

\[
\beta = \sum_{i = 1}^{\infty}\frac{\alpha^2}{\sigma_{t-1}^2} \\ 
\int_{x = 1}^{\infty}x_{i} = 1
\]

Note again the reference to equation \ref{eq1}. Writing nice math
requires practice. Note I used a forward slashes to make a space in the
equations. I can also align equations using
\texttt{\textbackslash{}\&\textquotesingle{}\textquotesingle{},\ and\ set\ to\ numbering\ only\ the\ first\ line.\ Now\ I\ will\ have\ to\ type}begin
equation'' which is a native \LaTeX command. Here follows a more
complicated equation:

\begin{align} 
    y_t &= c + B(L) y_{t-1} + e_t   \label{eq2}    \\ \notag 
    e_t &= H_t^{1/2}  z_t ; \quad z_t \sim  N(0,I_N) \quad \& \quad H_t = D_tR_tD_t \\ \notag
        D_t^2 &= {\sigma_{1,t}, \dots, \sigma_{N,t}}   \\ \notag
        \sigma_{i,t}^2 &= \gamma_i+\kappa_{i,t}  v_{i, t-1}^2 +\eta_i  \sigma_{i, t-1}^2, \quad \forall i \\ \notag
        R_{t, i, j} &= {diag(Q_{t, i, j}}^{-1}) . Q_{t, i, j} . diag(Q_{t, i, j}^{-1})  \\ \notag
        Q_{t, i, j} &= (1-\alpha-\beta)  \bar{Q} + \alpha  z_t  z_t'  + \beta  Q_{t, i, j} \notag
\end{align}

Noten that in \ref{eq2} I have aligned the equations by the equal signs.
I also want only one tag, and I create spaces using ``quads''.

See if you can figure out how to do complex math using the two examples
provided in \ref{eq1} and \ref{eq2}.

\section{Results}\label{results}

Tables can be included as follows. Use the \emph{xtable} (or kable)
package for tables. Table placement = H implies Latex tries to place the
table Here, and not on a new page (there are, however, very many ways to
skin this cat. Luckily there are many forums online!).

\begin{table}[H]
\centering
\begin{tabular}{rrrrrrrrrrrr}
  \hline
 & mpg & cyl & disp & hp & drat & wt & qsec & vs & am & gear & carb \\ 
  \hline
Mazda RX4 & 21.00 & 6.00 & 160.00 & 110.00 & 3.90 & 2.62 & 16.46 & 0.00 & 1.00 & 4.00 & 4.00 \\ 
  Mazda RX4 Wag & 21.00 & 6.00 & 160.00 & 110.00 & 3.90 & 2.88 & 17.02 & 0.00 & 1.00 & 4.00 & 4.00 \\ 
  Datsun 710 & 22.80 & 4.00 & 108.00 & 93.00 & 3.85 & 2.32 & 18.61 & 1.00 & 1.00 & 4.00 & 1.00 \\ 
  Hornet 4 Drive & 21.40 & 6.00 & 258.00 & 110.00 & 3.08 & 3.21 & 19.44 & 1.00 & 0.00 & 3.00 & 1.00 \\ 
  Hornet Sportabout & 18.70 & 8.00 & 360.00 & 175.00 & 3.15 & 3.44 & 17.02 & 0.00 & 0.00 & 3.00 & 2.00 \\ 
   \hline
\end{tabular}
\caption{Short Table Example \label{tab1}} 
\end{table}

To reference calculations \textbf{in text}, \emph{do this:} From table
\ref{tab1} we see the average value of mpg is 20.98.

Including tables that span across pages, use e.g.:

\begin{longtable}{rrrrrrrrrrrr}
  \hline
 & mpg & cyl & disp & hp & drat & wt & qsec & vs & am & gear & carb \\ 
  \hline
Mazda RX4 & 21.00 & 6.00 & 160.00 & 110.00 & 3.90 & 2.62 & 16.46 & 0.00 & 1.00 & 4.00 & 4.00 \\ 
  Mazda RX4 Wag & 21.00 & 6.00 & 160.00 & 110.00 & 3.90 & 2.88 & 17.02 & 0.00 & 1.00 & 4.00 & 4.00 \\ 
  Datsun 710 & 22.80 & 4.00 & 108.00 & 93.00 & 3.85 & 2.32 & 18.61 & 1.00 & 1.00 & 4.00 & 1.00 \\ 
  Hornet 4 Drive & 21.40 & 6.00 & 258.00 & 110.00 & 3.08 & 3.21 & 19.44 & 1.00 & 0.00 & 3.00 & 1.00 \\ 
  Hornet Sportabout & 18.70 & 8.00 & 360.00 & 175.00 & 3.15 & 3.44 & 17.02 & 0.00 & 0.00 & 3.00 & 2.00 \\ 
  Valiant & 18.10 & 6.00 & 225.00 & 105.00 & 2.76 & 3.46 & 20.22 & 1.00 & 0.00 & 3.00 & 1.00 \\ 
  Duster 360 & 14.30 & 8.00 & 360.00 & 245.00 & 3.21 & 3.57 & 15.84 & 0.00 & 0.00 & 3.00 & 4.00 \\ 
  Merc 240D & 24.40 & 4.00 & 146.70 & 62.00 & 3.69 & 3.19 & 20.00 & 1.00 & 0.00 & 4.00 & 2.00 \\ 
  Merc 230 & 22.80 & 4.00 & 140.80 & 95.00 & 3.92 & 3.15 & 22.90 & 1.00 & 0.00 & 4.00 & 2.00 \\ 
  Merc 280 & 19.20 & 6.00 & 167.60 & 123.00 & 3.92 & 3.44 & 18.30 & 1.00 & 0.00 & 4.00 & 4.00 \\ 
  Merc 280C & 17.80 & 6.00 & 167.60 & 123.00 & 3.92 & 3.44 & 18.90 & 1.00 & 0.00 & 4.00 & 4.00 \\ 
  Merc 450SE & 16.40 & 8.00 & 275.80 & 180.00 & 3.07 & 4.07 & 17.40 & 0.00 & 0.00 & 3.00 & 3.00 \\ 
  Merc 450SL & 17.30 & 8.00 & 275.80 & 180.00 & 3.07 & 3.73 & 17.60 & 0.00 & 0.00 & 3.00 & 3.00 \\ 
  Merc 450SLC & 15.20 & 8.00 & 275.80 & 180.00 & 3.07 & 3.78 & 18.00 & 0.00 & 0.00 & 3.00 & 3.00 \\ 
  Cadillac Fleetwood & 10.40 & 8.00 & 472.00 & 205.00 & 2.93 & 5.25 & 17.98 & 0.00 & 0.00 & 3.00 & 4.00 \\ 
  Lincoln Continental & 10.40 & 8.00 & 460.00 & 215.00 & 3.00 & 5.42 & 17.82 & 0.00 & 0.00 & 3.00 & 4.00 \\ 
  Chrysler Imperial & 14.70 & 8.00 & 440.00 & 230.00 & 3.23 & 5.34 & 17.42 & 0.00 & 0.00 & 3.00 & 4.00 \\ 
  Fiat 128 & 32.40 & 4.00 & 78.70 & 66.00 & 4.08 & 2.20 & 19.47 & 1.00 & 1.00 & 4.00 & 1.00 \\ 
  Honda Civic & 30.40 & 4.00 & 75.70 & 52.00 & 4.93 & 1.61 & 18.52 & 1.00 & 1.00 & 4.00 & 2.00 \\ 
  Toyota Corolla & 33.90 & 4.00 & 71.10 & 65.00 & 4.22 & 1.83 & 19.90 & 1.00 & 1.00 & 4.00 & 1.00 \\ 
  Toyota Corona & 21.50 & 4.00 & 120.10 & 97.00 & 3.70 & 2.46 & 20.01 & 1.00 & 0.00 & 3.00 & 1.00 \\ 
  Dodge Challenger & 15.50 & 8.00 & 318.00 & 150.00 & 2.76 & 3.52 & 16.87 & 0.00 & 0.00 & 3.00 & 2.00 \\ 
  AMC Javelin & 15.20 & 8.00 & 304.00 & 150.00 & 3.15 & 3.44 & 17.30 & 0.00 & 0.00 & 3.00 & 2.00 \\ 
  Camaro Z28 & 13.30 & 8.00 & 350.00 & 245.00 & 3.73 & 3.84 & 15.41 & 0.00 & 0.00 & 3.00 & 4.00 \\ 
  Pontiac Firebird & 19.20 & 8.00 & 400.00 & 175.00 & 3.08 & 3.85 & 17.05 & 0.00 & 0.00 & 3.00 & 2.00 \\ 
  Fiat X1-9 & 27.30 & 4.00 & 79.00 & 66.00 & 4.08 & 1.94 & 18.90 & 1.00 & 1.00 & 4.00 & 1.00 \\ 
  Porsche 914-2 & 26.00 & 4.00 & 120.30 & 91.00 & 4.43 & 2.14 & 16.70 & 0.00 & 1.00 & 5.00 & 2.00 \\ 
  Lotus Europa & 30.40 & 4.00 & 95.10 & 113.00 & 3.77 & 1.51 & 16.90 & 1.00 & 1.00 & 5.00 & 2.00 \\ 
  Ford Pantera L & 15.80 & 8.00 & 351.00 & 264.00 & 4.22 & 3.17 & 14.50 & 0.00 & 1.00 & 5.00 & 4.00 \\ 
  Ferrari Dino & 19.70 & 6.00 & 145.00 & 175.00 & 3.62 & 2.77 & 15.50 & 0.00 & 1.00 & 5.00 & 6.00 \\ 
  Maserati Bora & 15.00 & 8.00 & 301.00 & 335.00 & 3.54 & 3.57 & 14.60 & 0.00 & 1.00 & 5.00 & 8.00 \\ 
  Volvo 142E & 21.40 & 4.00 & 121.00 & 109.00 & 4.11 & 2.78 & 18.60 & 1.00 & 1.00 & 4.00 & 2.00 \\ 
   \hline
\hline
\caption{Long Table Example} 
\end{longtable}

\hfill

\section{Lists}\label{lists}

To add lists, simply using the following notation

\begin{itemize}
\item
  This is really simple

  \begin{itemize}
  \tightlist
  \item
    Just note the spaces here - writing in R you have to sometimes be
    pedantic about spaces\ldots{}
  \end{itemize}
\item
  Note that Rmarkdown notation removes the pain of defining
  \LaTeX environments!
\end{itemize}

\section{Conclusion}\label{conclusion}

I hope you find this template useful. Remember, stackoverflow is your
friend - use it to find answers to questions. Feel free to write me a
mail if you have any questions regarding the use of this package feel
free to email me. To cite this package, simply type
citation(``Texevier'') in Rstudio to get the citation for Katzke
(\protect\hyperlink{ref-Texevier}{2017}) (Note that uncited references
in your bibtex file will not be included in References).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-fama1997}{}
Fama, Eugene F, and Kenneth R French. 1997. ``Industry Costs of
Equity.'' \emph{Journal of Financial Economics} 43 (2). Elsevier:
153--93.

\hypertarget{ref-grinold2000}{}
Grinold, Richard C, and Ronald N Kahn. 2000. ``Active Portfolio
Management.'' McGraw Hill New York, NY.

\hypertarget{ref-Texevier}{}
Katzke, N.F. 2017. \emph{Texevier: Package to Create Elsevier Templates
for Rmarkdown}. Stellenbosch, South Africa: Bureau for Economic
Research.

% Force include bibliography in my chosen format:
\newpage
\nocite{*}
\bibliography{}

\end{document}

### 2017-04-20 my_name.R
My name function

@@ -1,0 +1,4 @@
+my_name <- function(name){
+  cat(paste("Hello", name))
+}
+  

### 2017-04-20 my_name.R
I deleted my name function

@@ -1,4 +1,0 @@
-my_name <- function(name){
-  cat(paste("Hello", name))
-}
-  

### 2017-05-18 Literature Review/A Brief Literature Review.Rmd
Attempt at moving lit review to rmd

---
title: "Literature Review"

documentclass: "elsarticle"

Author1: "Fiona Ganie"  
Ref1: "University of Cape Town" 
Email1: "GNXFIO001\\@myuct.ac.za"

BottomRFooter: "\\footnotesize Page \\thepage\\" 
addtoprule: TRUE
addfootrule: TRUE 
bibliography: Tex/ref.bib  
RemovePreprintSubmittedTo: TRUE

toc: no                        
numbersections: yes

fontsize: 12pt                  
linestretch: 1.5                
link-citations: TRUE

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 
    fig_height: 3.5
    include:
      in_header: Tex/packages.txt 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

```

# Introduction \label{Introduction}

Forecasting is a core element of the processes of a business. Producing forecasts of high quality is essential as they play an important role in inventory control, purchasing decisions of new equipment and investment decisions, and are needed for effective planning [@granvik2010forecasting]. Automatic forecasting techniques have commonly been used in business to produce a wide variety of forecasts and are a useful tool for producing a large number of forecasts. Automatic techniques can be used by non-experts without training in the use of time series [@hyndman2007automatic]. Automatic forecasting algorithms must have the ability to select the appropriate model for forecasting and estimate model parameters, without the need for human intervention [@hyndman2007automatic]. The most common automatic forecasting techniques used are the ARIMA and Exponential Smoothing (ETS) models. ETS models have been used extensively in forecasting demand for inventories [@billah2006exponential]. They have performed well in forecasting competitions against more complex models and are particularly good at forecasting seasonal data over short horizons [@hyndman2002state].

Automatic forecasting techniques can be too inflexible and do not allow for analysts with specialized knowledge to incorporate their expertise into the model effectively. This leads to difficulties in producing forecasts of high quality, thus the demand for high quality forecasts far exceeds the rate at which they are produced [@taylor2017forecasting]. In December 2016, Facebook released their forecasting model, Prophet. According to @taylor2017forecasting, Prophet is a flexible model that can be adjusted by a large number of non-experts who have little knowledge about time series, however have expert knowledge on the data-generating process. Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared. This is Facebook’s definition of forecasting at scale, and provides a solution to the problems that automatic forecasting pose [@taylor2017forecasting].

# Exponential State Space Models \label{Exponential State Space Models}

## Exponential Smoothing Methods 
ETS methods originated in the 1950’s and is the key work of Brown , Holt, and Winters [@de200625]. It has been widely used for many applications in industry and has been the foundation of many successful forecasting methods [@de200625]. Exponential smoothing techniques produce forecasts by calculating them as a weighted average of past observations. Observations that are most recent are more heavily weighted, with these weights decaying exponentially as the observations get older.

Exponential smoothing methods were originally classified by Pegel according to the components present in the time series (trend and seasonality), and whether they were additive or multiplicative in nature [@de200625]. Gardner and Taylor further extended this classification to include a damped additive and multiplicative trend respectively [@de200625]. A taxonomy of exponential smoothing methods was later created by Hyndman et al. and extended by Taylor [@de200625]. It considered all possible combinations of the trend and seasonality components thus producing 15 methods of exponential smoothing as seen in table \ref{tab1}.

```{r ShortTable, results = 'asis'}

library(xtable)

WD <- getwd()
data <- read.csv( file.path(WD, "Data/ETSMethods.csv"), header=TRUE, sep=" ", check.names=FALSE)

table <- xtable(data, caption = "Taxonomy of Exponential Smoothing Methods \\label{tab1}")
  print.xtable(table, 
             table.placement = 'H', 
             comment = FALSE,
             caption.placement = 'bottom'
             )

```

\hfill

The most common of these methods are Simple Exponential Smoothing (N,N), Holt’s linear method (A,N) and Holt-Winters seasonal method (A,A).

Although exponential smoothing methods performed well and have been widely used in business, they did not have any statistical framework underpinning them [@de200625]. They produced point estimates rather than prediction intervals and selection of a forecasting method was specific to the time series in question. Efforts were made by Box and Jenkins to create a statistical framework that underlies exponential smoothing methods and they found that forecasts produced by linear methods are special cases of ARIMA models [@de200625]. This however does not hold for the multiplicative methods.

## State Space Models 
The motivation behind state space models was the need for a statistical framework that underlies all methods of exponential smoothing [@de200625]. @hyndman2002state provided this framework as an extension of Ord, Koehler, and Snyder’s class of exponential state space models which underpinned some of the smoothing methods [@de200625]. He proposed two state space models, one corresponding to an additive error and the other to a multiplicative error. By specifying a distribution for these error terms, he turned the previously deterministic methods into stochastic models. These two models were then applied to all 15 smoothing methods, producing 30 models in total.

The two state space models produce identical point estimates to each other as well as to the exponential smoothing methods, however they will generate differing prediction intervals [@hyndman2013forecasting]. Each model consists of a measurement equation and a transition equation. The measurement equation shows the relationship between the observations and the unobserved components (level, trend seasonality), while the transition equation shows how the unobserved components change over time.

The introduction of the exponential state space model has allowed for model selection using information criteria rather than using an ad hoc method. It has also made the calculation of the likelihood easier and simulation from the underlying model possible [@hyndman2002state].

## Application and Performance 
In a paper by @smith2015comparison, Holt-Winters exponential smoothing (HWES) model was applied to time series data relating to data on patents. The patent data was classified into three different groups; monthly forecasts were made and the results were compared to forecasts produced by ARIMA models. To avoid over-forecasting, the trend of the HWES was damped to produce forecasts with a flat trend for later observations. The results showed that all models forecasted the data adequately. The HWES model fitted the data well and performed better than the ARIMA models in the case where the time series for the patent group was fairly stationary. Based on the forecasting results, it could not be concluded that a specific time series model is better for forecasting patent data.

@hassani2015forecasting forecasted tourism demand in specific European countries over a short, medium and long term horizon. The ETS model was used to produce forecasts and was compared to six parametric and non-parametric techniques. The results showed that no single model is best for producing forecasts across all countries and time horizons. The ETS model together with the Neural Networks (NN) and Fractionalized ARIMA (ARFIMA) produced forecasts with the least amount of accuracy and are thus not relevant models for forecasting tourism demand in Europe. 24 steps-ahead error forecasts produced by the ETS model were found to be larger than the errors produced by the benchmark model - Recurrent Single Spectrum Analysis (SSA-R) across all European countries.

## Robustness 
@gardner2006exponential explained the robustness of ETS models by comparing them to equivalent models. Simple exponential smoothing which is equivalent to an ARIMA(0,1,1) model is the most robust method and has performed well in forecasting many types of time series. By comparing simple exponential smoothing to lower order ARIMA models, it is seen that errors that arise due to model specification is not as much of a problem as ARIMA models. Furthermore, ARIMA models tend to produce larger MSE due to model selection errors, and this is worsened if the errors are not normally distributed [@gardner2006exponential].

Satchell and Timmermann found that for time series with a finite history, the weights assigned to each observation in the simple exponential smoothing method are robust, provided that the variance of the random walk relative to the variance of the error component is not too small [@gardner2006exponential].

Forecasts produced by simple exponential smoothing have performed well in modelling annual sales of different products, as well as aggregated economic series with low sampling frequencies [@gardner2006exponential].

# Prophet

## Bayesian Generalized Additive Model
Prophet is the key work of @taylor2017forecasting and consists of a decomposable model with a component for growth, seasonality, and holidays. These components consist of linear and non-linear functions of time. This is similar to a Generalized Additive Model (GAM) which is a regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017forecasting]. Prophet, like GAM, frames the forecasting problem as a curve fitting exercise using backfitting to find the regression functions.

The growth component is modelled in a similar way to modelling population growths which uses a logistic growth model [@taylor2017forecasting]. Populations typically grow non-linearly (although the growth component could also be linear) up to an upper bound known as the carrying capacity, and remains constant thereafter. The rate at which the population grows changes over time, and this is accounted for by including changepoints in the model where the growth rate may be changed [@taylor2017forecasting]. These changepoints can be automatically selected and have a Laplace prior distribution placed on them [@taylor2017forecasting]. The parameter of the prior distribution can be used to adjust the growth rate of the changepoints. This allows non-experts with knowledge about events that may affect growth to use the parameter as a knob and adjust it to increase or decrease the number of changepoints [@taylor2017forecasting]. It also allows for the analyst to add changepoints which the automatic selection procedure may have missed [@taylor2017forecasting]. Furthermore, analysts may also specify the carrying capacity and adjust it based on their knowledge of the total market size [@taylor2017forecasting].

The decomposable form of the model allows for components to be easily added to it. This allows for multiple seasonality components with different periods to be added to the model. The seasonality component is modelled by a Fourier series, with the parameters of the Fourier series having a Normal prior distribution [@taylor2017forecasting]. The variance of the parameters can be adjusted by analysts to smooth the model and change how much of historical seasonality is projected to the future [@taylor2017forecasting].

The name, date, and country of past and future holidays and events may be inputted by the analyst into a list [@taylor2017forecasting]. By specifying the country in which the events take place or the holidays occur, separate lists can be populated for global events/holidays and country-specific events/holidays. The union of the two lists are then used for forecasting. Like seasonality, a Normal prior distribution is placed on the parameters of the holidays, and the variance of the parameters can be adjusted by analysts to smooth the model [@taylor2017forecasting].

## Application and Performance 
@taylor2017forecasting forecasted the number of events on Facebook using Prophet. The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend. The forecasts produced by Prophet were compared to forecasts produced by common automatic forecasting techniques such as ETS, ARIMA, and the seasonal naïve model, as well as to simple models such as the naïve model. While the ETS and seasonal naïve model were quite robust, the ARIMA forecasts were fragile. No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations. Hence, Prophet had lower forecasting errors compared to the automatic forecasting methods.

## Robustness 
Prophet is a flexible model and has intuitive parameters that can be easily interpreted by human beings. The GAM is fitted quickly, allowing the analyst to interactively change the model parameters [@taylor2017forecasting]. Unlike ARIMA models, Prophet can produce forecasts over different scales and allows for missing values in the time series without the need for interpolation [@taylor2017forecasting].

When a large number of forecasts are produced, manually identifying problematic forecasts becomes a time consuming and difficult task. Prophet provides a semi-automated forecast evaluation system that selects the best model that fits the data. When there are large forecast errors, the forecasts are flagged so that the analyst can explore the cause of the errors, identify and remove potential outliers and either adjust the model or choose a more appropriate model [@taylor2017forecasting].

# Evaluation of Forecasts

## Common Methods 
Model selection has historically been very subjective, with no techniques that offered mathematical rigour for choosing models [@smith2015comparison]. Progress has been made over time and many mathematical techniques have been introduced for model selection. Use of information criteria is one such method, with Akaike’s information criterion (AIC) and Bayes information criterion (BIC) being the most common [@smith2015comparison]. These information criteria minimize forecasting errors while penalizing it for overfitting. Information criteria should be used to compare models of similar structure only. To compare models of different structure, the most commonly used measures of forecast errors are the Root Mean Squared error (RMSE), Mean Absolute Error (MAE), Mean Absolute Percentage Error (MAPE), and Mean Absolute Scaled Error (MASE). A lower value of these statistics indicates that the forecasts are better, with a MASE less than one indicating that the forecasts are better than those produced by a naïve model [@smith2015comparison].

@smith2015comparison have used these methods in their paper when evaluating patent forecasts and making comparisons between the HWES and ARIMA models. In a paper by @hassani2015forecasting they chose their ETS model based on the selection that R automatically made for the time series. They compared models of different structure using the RMSE and a modified Diebold-Mariano test. @hyndman2002state suggests that ETS models should be selected using the AIC rather than the MSE or MAPE. The AIC provides a way of choosing between models with an additive error and models with a multiplicative error since it is based on the likelihood function. The MSE and MAPE are not able to select between the two error types because both models produce the same point estimates. According to @de200625, the best approach of selecting models with different structure is the Diebold-Mariano test. The MSE is scale dependent and should not be used to make comparisons while the MAPE encounters difficulties when the time series values are close to or equal to zero.

## Mincer-Zarnowitz Approach 
@mincer1969evaluation proposed two methods of measuring forecast accuracy - an absolute and a relative measure. Absolute measures of accuracy measure the distance between actual and predicted values. If the mean distance is equal to zero, the forecast is said to be unbiased [@mincer1969evaluation]. To analyse absolute forecast errors, the observed values are regressed against the predicted values. A joint hypothesis test is performed to check if the intercept is equal to zero and the slope is equal to one. If the intercept is equal to zero, the forecasts are unbiased and do not overestimate or underestimate the data. If the slope is equal to one, the residual errors are uncorrelated with the predictions and the forecast is said to be efficient [@mincer1969evaluation]. The MSE is a measure that can be decomposed into the residual variance and two components, one reflecting the bias and the other reflecting the inefficiency of the forecast. If the forecasts are unbiased and efficient, the MSE will reduce to the residual variance [@mincer1969evaluation]. Absolute forecast measures cannot be used to make comparisons between forecasts with different scales or economic variables. Thiel states that the consequences of forecasting errors and how they impact decisions is of greater importance than the size of the forecasting errors [@mincer1969evaluation].

Rather than using absolute measures of forecast accuracy, @mincer1969evaluation suggested the use of relative measures of forecast accuracy. Relative accuracy analysis allows for meaningful comparisons of different forecasting methods to be made. They proposed an index which considers the ratio of the MSE of the forecast to the MSE of a benchmark forecasting method. They used an extrapolation of the data history as a benchmark, as it is a cost effective and accessible method, however the benchmark may be any method which is relevant for comparison. This ratio is known as the Relative Mean Square Error (RM).

The numerator represents a return which is inversely proportional to the MSE error of the forecasts, while the denominator represents the cost of forecasting which is inversely proportional to the MSE of the benchmark [@mincer1969evaluation]. Hence the ratio is representative of a rate of return index and ranks the performance of forecasts as such. If the forecasts are better than the benchmark, the RM will be less than one.

\newpage

# References

### 2017-05-18 Literature Review/A_Brief_Literature_Review.tex
Attempt at moving lit review to rmd

\documentclass[12pt,preprint, authoryear]{elsarticle}

\usepackage{lmodern}
%%%% My spacing
\usepackage{setspace}
\setstretch{1.5}
\DeclareMathSizes{12}{14}{10}{10}

% Wrap around which gives all figures included the [H] command, or places it "here". This can be tedious to code in Rmarkdown.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\let\origtable\table
\let\endorigtable\endtable
\renewenvironment{table}[1][2] {
    \expandafter\origtable\expandafter[H]
} {
    \endorigtable
}

\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi

\usepackage{amssymb, amsmath, amsthm, amsfonts}

\usepackage[round]{natbib}
\bibliographystyle{natbib}
\def\bibsection{\section*{References}} %%% Make "References" appear before bibliography
\usepackage{longtable}
\usepackage[margin=2cm,bottom=4cm,top=2.5cm, includefoot]{geometry}
\usepackage{fancyhdr}
\usepackage[bottom, hang, flushmargin]{footmisc}
\usepackage{graphicx}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}
\setlength{\parindent}{0cm}
\setlength{\parskip}{1.3ex plus 0.5ex minus 0.3ex}
\usepackage{textcomp}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.3pt}

\usepackage{array}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

%%%%  Remove the "preprint submitted to" part. Don't worry about this either, it just looks better without it:
\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother

 \def\tightlist{} % This allows for subbullets!

\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

%%% Include extra packages specified by user
% Insert custom packages here as follows
% \usepackage{tikz}

\begin{document}

\begin{frontmatter}  %

\title{Literature Review}

\author[Add1]{Fiona Ganie}
\ead{GNXFIO001@myuct.ac.za}

\address[Add1]{University of Cape Town}

\vspace{1cm}

\begin{keyword}
\footnotesize{
 \\ \vspace{0.3cm}
\textit{JEL classification} 
}
\end{keyword}
\vspace{0.5cm}
\end{frontmatter}

%________________________
% Header and Footers
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{fancy}
\chead{}
\rhead{}
\lfoot{}
\rfoot{\footnotesize Page \thepage\\}
\lhead{}
%\rfoot{\footnotesize Page \thepage\ } % "e.g. Page 2"
\cfoot{}

%\setlength\headheight{30pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%________________________

\headsep 35pt % So that header does not go over title

\section{\texorpdfstring{Introduction
\label{Introduction}}{Introduction }}\label{introduction}

Forecasting is a core element of the processes of a business. Producing
forecasts of high quality is essential as they play an important role in
inventory control, purchasing decisions of new equipment and investment
decisions, and are needed for effective planning (Granvik
\protect\hyperlink{ref-granvik2010forecasting}{2010}). Automatic
forecasting techniques have commonly been used in business to produce a
wide variety of forecasts and are a useful tool for producing a large
number of forecasts. Automatic techniques can be used by non-experts
without training in the use of time series (R. J. Hyndman, Khandakar,
and others \protect\hyperlink{ref-hyndman2007automatic}{2007}).
Automatic forecasting algorithms must have the ability to select the
appropriate model for forecasting and estimate model parameters, without
the need for human intervention (R. J. Hyndman, Khandakar, and others
\protect\hyperlink{ref-hyndman2007automatic}{2007}). The most common
automatic forecasting techniques used are the ARIMA and Exponential
Smoothing (ETS) models. ETS models have been used extensively in
forecasting demand for inventories (Billah et al.
\protect\hyperlink{ref-billah2006exponential}{2006}). They have
performed well in forecasting competitions against more complex models
and are particularly good at forecasting seasonal data over short
horizons (R. J. Hyndman et al.
\protect\hyperlink{ref-hyndman2002state}{2002}).

Automatic forecasting techniques can be too inflexible and do not allow
for analysts with specialized knowledge to incorporate their expertise
into the model effectively. This leads to difficulties in producing
forecasts of high quality, thus the demand for high quality forecasts
far exceeds the rate at which they are produced (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). In December 2016,
Facebook released their forecasting model, Prophet. According to Taylor
and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017}),
Prophet is a flexible model that can be adjusted by a large number of
non-experts who have little knowledge about time series, however have
expert knowledge on the data-generating process. Prophet allows for a
large number of forecasts to be produced across a variety of problems
and consists of a robust evaluation system that allows for a large
number of forecasts be evaluated and compared. This is Facebook's
definition of forecasting at scale, and provides a solution to the
problems that automatic forecasting pose (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\section{\texorpdfstring{Exponential State Space Models
\label{Exponential State Space Models}}{Exponential State Space Models }}\label{exponential-state-space-models}

\subsection{Exponential Smoothing
Methods}\label{exponential-smoothing-methods}

ETS methods originated in the 1950's and is the key work of Brown ,
Holt, and Winters (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). It has been widely used for
many applications in industry and has been the foundation of many
successful forecasting methods (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). Exponential smoothing
techniques produce forecasts by calculating them as a weighted average
of past observations. Observations that are most recent are more heavily
weighted, with these weights decaying exponentially as the observations
get older.

Exponential smoothing methods were originally classified by Pegel
according to the components present in the time series (trend and
seasonality), and whether they were additive or multiplicative in nature
(De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). Gardner
and Taylor further extended this classification to include a damped
additive and multiplicative trend respectively (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). A taxonomy of exponential
smoothing methods was later created by Hyndman et al. and extended by
Taylor (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
It considered all possible combinations of the trend and seasonality
components thus producing 15 methods of exponential smoothing as seen in
table \ref{tab1}.

\begin{table}[H]
\centering
\begin{tabular}{rlll}
  \hline
 & Trend &  & Seasonality \\ 
  \hline
 & None(N) & Additive(A) & Multiplicative(M) \\ 
  None(N) & (N,N) & (N,A) & (N,M) \\ 
  Additive(A) & (A,N) & (A,A) & (A,M) \\ 
  Additive\_damped(Ad) & (Ad,N) & (Ad,A) & (Ad,M) \\ 
  Multiplicative(M) & (M,N) & (M,A) & (M,M) \\ 
  Multiplicative\_damped(Md) & (Md,N) & (Md,A) & (Md,M) \\ 
   \hline
\end{tabular}
\caption{Taxonomy of Exponential Smoothing Methods \label{tab1}} 
\end{table}

\hfill

The most common of these methods are Simple Exponential Smoothing (N,N),
Holt's linear method (A,N) and Holt-Winters seasonal method (A,A).

Although exponential smoothing methods performed well and have been
widely used in business, they did not have any statistical framework
underpinning them (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). They produced point estimates
rather than prediction intervals and selection of a forecasting method
was specific to the time series in question. Efforts were made by Box
and Jenkins to create a statistical framework that underlies exponential
smoothing methods and they found that forecasts produced by linear
methods are special cases of ARIMA models (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). This however does not hold for
the multiplicative methods.

\subsection{State Space Models}\label{state-space-models}

The motivation behind state space models was the need for a statistical
framework that underlies all methods of exponential smoothing (De
Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). R. J.
Hyndman et al. (\protect\hyperlink{ref-hyndman2002state}{2002}) provided
this framework as an extension of Ord, Koehler, and Snyder's class of
exponential state space models which underpinned some of the smoothing
methods (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
He proposed two state space models, one corresponding to an additive
error and the other to a multiplicative error. By specifying a
distribution for these error terms, he turned the previously
deterministic methods into stochastic models. These two models were then
applied to all 15 smoothing methods, producing 30 models in total.

The two state space models produce identical point estimates to each
other as well as to the exponential smoothing methods, however they will
generate differing prediction intervals (R. Hyndman and Athanasopoulos
\protect\hyperlink{ref-hyndman2013forecasting}{2013}). Each model
consists of a measurement equation and a transition equation. The
measurement equation shows the relationship between the observations and
the unobserved components (level, trend seasonality), while the
transition equation shows how the unobserved components change over
time.

The introduction of the exponential state space model has allowed for
model selection using information criteria rather than using an ad hoc
method. It has also made the calculation of the likelihood easier and
simulation from the underlying model possible (R. J. Hyndman et al.
\protect\hyperlink{ref-hyndman2002state}{2002}).

\subsection{Application and
Performance}\label{application-and-performance}

In a paper by Smith and Agrawal
(\protect\hyperlink{ref-smith2015comparison}{2015}), Holt-Winters
exponential smoothing (HWES) model was applied to time series data
relating to data on patents. The patent data was classified into three
different groups; monthly forecasts were made and the results were
compared to forecasts produced by ARIMA models. To avoid
over-forecasting, the trend of the HWES was damped to produce forecasts
with a flat trend for later observations. The results showed that all
models forecasted the data adequately. The HWES model fitted the data
well and performed better than the ARIMA models in the case where the
time series for the patent group was fairly stationary. Based on the
forecasting results, it could not be concluded that a specific time
series model is better for forecasting patent data.

Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
forecasted tourism demand in specific European countries over a short,
medium and long term horizon. The ETS model was used to produce
forecasts and was compared to six parametric and non-parametric
techniques. The results showed that no single model is best for
producing forecasts across all countries and time horizons. The ETS
model together with the Neural Networks (NN) and Fractionalized ARIMA
(ARFIMA) produced forecasts with the least amount of accuracy and are
thus not relevant models for forecasting tourism demand in Europe. 24
steps-ahead error forecasts produced by the ETS model were found to be
larger than the errors produced by the benchmark model - Recurrent
Single Spectrum Analysis (SSA-R) across all European countries.

\subsection{Robustness}\label{robustness}

Gardner (\protect\hyperlink{ref-gardner2006exponential}{2006}) explained
the robustness of ETS models by comparing them to equivalent models.
Simple exponential smoothing which is equivalent to an ARIMA(0,1,1)
model is the most robust method and has performed well in forecasting
many types of time series. By comparing simple exponential smoothing to
lower order ARIMA models, it is seen that errors that arise due to model
specification is not as much of a problem as ARIMA models. Furthermore,
ARIMA models tend to produce larger MSE due to model selection errors,
and this is worsened if the errors are not normally distributed (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

Satchell and Timmermann found that for time series with a finite
history, the weights assigned to each observation in the simple
exponential smoothing method are robust, provided that the variance of
the random walk relative to the variance of the error component is not
too small (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

Forecasts produced by simple exponential smoothing have performed well
in modelling annual sales of different products, as well as aggregated
economic series with low sampling frequencies (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

\section{Prophet}\label{prophet}

\subsection{Bayesian Generalized Additive
Model}\label{bayesian-generalized-additive-model}

Prophet is the key work of Taylor and Letham
(\protect\hyperlink{ref-taylor2017forecasting}{2017}) and consists of a
decomposable model with a component for growth, seasonality, and
holidays. These components consist of linear and non-linear functions of
time. This is similar to a Generalized Additive Model (GAM) which is a
regression model that consists of non-linear and linear regression
functions applied to predictor variables (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Prophet, like GAM,
frames the forecasting problem as a curve fitting exercise using
backfitting to find the regression functions.

The growth component is modelled in a similar way to modelling
population growths which uses a logistic growth model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Populations
typically grow non-linearly (although the growth component could also be
linear) up to an upper bound known as the carrying capacity, and remains
constant thereafter. The rate at which the population grows changes over
time, and this is accounted for by including changepoints in the model
where the growth rate may be changed (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). These changepoints
can be automatically selected and have a Laplace prior distribution
placed on them (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). The parameter of
the prior distribution can be used to adjust the growth rate of the
changepoints. This allows non-experts with knowledge about events that
may affect growth to use the parameter as a knob and adjust it to
increase or decrease the number of changepoints (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). It also allows for
the analyst to add changepoints which the automatic selection procedure
may have missed (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Furthermore,
analysts may also specify the carrying capacity and adjust it based on
their knowledge of the total market size (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

The decomposable form of the model allows for components to be easily
added to it. This allows for multiple seasonality components with
different periods to be added to the model. The seasonality component is
modelled by a Fourier series, with the parameters of the Fourier series
having a Normal prior distribution (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). The variance of
the parameters can be adjusted by analysts to smooth the model and
change how much of historical seasonality is projected to the future
(Taylor and Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}).

The name, date, and country of past and future holidays and events may
be inputted by the analyst into a list (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). By specifying the
country in which the events take place or the holidays occur, separate
lists can be populated for global events/holidays and country-specific
events/holidays. The union of the two lists are then used for
forecasting. Like seasonality, a Normal prior distribution is placed on
the parameters of the holidays, and the variance of the parameters can
be adjusted by analysts to smooth the model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\subsection{Application and
Performance}\label{application-and-performance-1}

Taylor and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017})
forecasted the number of events on Facebook using Prophet. The time
series was impacted by holidays, had strong multi-period seasonality,
and a piecewise trend. The forecasts produced by Prophet were compared
to forecasts produced by common automatic forecasting techniques such as
ETS, ARIMA, and the seasonal naïve model, as well as to simple models
such as the naïve model. While the ETS and seasonal naïve model were
quite robust, the ARIMA forecasts were fragile. No model besides Prophet
accounted for the dips around holidays and the upward trend of the time
series towards later observations. Hence, Prophet had lower forecasting
errors compared to the automatic forecasting methods.

\subsection{Robustness}\label{robustness-1}

Prophet is a flexible model and has intuitive parameters that can be
easily interpreted by human beings. The GAM is fitted quickly, allowing
the analyst to interactively change the model parameters (Taylor and
Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}). Unlike
ARIMA models, Prophet can produce forecasts over different scales and
allows for missing values in the time series without the need for
interpolation (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

When a large number of forecasts are produced, manually identifying
problematic forecasts becomes a time consuming and difficult task.
Prophet provides a semi-automated forecast evaluation system that
selects the best model that fits the data. When there are large forecast
errors, the forecasts are flagged so that the analyst can explore the
cause of the errors, identify and remove potential outliers and either
adjust the model or choose a more appropriate model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\section{Evaluation of Forecasts}\label{evaluation-of-forecasts}

\subsection{Common Methods}\label{common-methods}

Model selection has historically been very subjective, with no
techniques that offered mathematical rigour for choosing models (Smith
and Agrawal \protect\hyperlink{ref-smith2015comparison}{2015}). Progress
has been made over time and many mathematical techniques have been
introduced for model selection. Use of information criteria is one such
method, with Akaike's information criterion (AIC) and Bayes information
criterion (BIC) being the most common (Smith and Agrawal
\protect\hyperlink{ref-smith2015comparison}{2015}). These information
criteria minimize forecasting errors while penalizing it for
overfitting. Information criteria should be used to compare models of
similar structure only. To compare models of different structure, the
most commonly used measures of forecast errors are the Root Mean Squared
error (RMSE), Mean Absolute Error (MAE), Mean Absolute Percentage Error
(MAPE), and Mean Absolute Scaled Error (MASE). A lower value of these
statistics indicates that the forecasts are better, with a MASE less
than one indicating that the forecasts are better than those produced by
a naïve model (Smith and Agrawal
\protect\hyperlink{ref-smith2015comparison}{2015}).

Smith and Agrawal (\protect\hyperlink{ref-smith2015comparison}{2015})
have used these methods in their paper when evaluating patent forecasts
and making comparisons between the HWES and ARIMA models. In a paper by
Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
they chose their ETS model based on the selection that R automatically
made for the time series. They compared models of different structure
using the RMSE and a modified Diebold-Mariano test. R. J. Hyndman et al.
(\protect\hyperlink{ref-hyndman2002state}{2002}) suggests that ETS
models should be selected using the AIC rather than the MSE or MAPE. The
AIC provides a way of choosing between models with an additive error and
models with a multiplicative error since it is based on the likelihood
function. The MSE and MAPE are not able to select between the two error
types because both models produce the same point estimates. According to
De Gooijer and Hyndman (\protect\hyperlink{ref-de200625}{2006}), the
best approach of selecting models with different structure is the
Diebold-Mariano test. The MSE is scale dependent and should not be used
to make comparisons while the MAPE encounters difficulties when the time
series values are close to or equal to zero.

\subsection{Mincer-Zarnowitz Approach}\label{mincer-zarnowitz-approach}

Mincer and Zarnowitz
(\protect\hyperlink{ref-mincer1969evaluation}{1969}) proposed two
methods of measuring forecast accuracy - an absolute and a relative
measure. Absolute measures of accuracy measure the distance between
actual and predicted values. If the mean distance is equal to zero, the
forecast is said to be unbiased (Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). To analyse absolute
forecast errors, the observed values are regressed against the predicted
values. A joint hypothesis test is performed to check if the intercept
is equal to zero and the slope is equal to one. If the intercept is
equal to zero, the forecasts are unbiased and do not overestimate or
underestimate the data. If the slope is equal to one, the residual
errors are uncorrelated with the predictions and the forecast is said to
be efficient (Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). The MSE is a
measure that can be decomposed into the residual variance and two
components, one reflecting the bias and the other reflecting the
inefficiency of the forecast. If the forecasts are unbiased and
efficient, the MSE will reduce to the residual variance (Mincer and
Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}). Absolute
forecast measures cannot be used to make comparisons between forecasts
with different scales or economic variables. Thiel states that the
consequences of forecasting errors and how they impact decisions is of
greater importance than the size of the forecasting errors (Mincer and
Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}).

Rather than using absolute measures of forecast accuracy, Mincer and
Zarnowitz (\protect\hyperlink{ref-mincer1969evaluation}{1969}) suggested
the use of relative measures of forecast accuracy. Relative accuracy
analysis allows for meaningful comparisons of different forecasting
methods to be made. They proposed an index which considers the ratio of
the MSE of the forecast to the MSE of a benchmark forecasting method.
They used an extrapolation of the data history as a benchmark, as it is
a cost effective and accessible method, however the benchmark may be any
method which is relevant for comparison. This ratio is known as the
Relative Mean Square Error (RM).

The numerator represents a return which is inversely proportional to the
MSE error of the forecasts, while the denominator represents the cost of
forecasting which is inversely proportional to the MSE of the benchmark
(Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). Hence the ratio is
representative of a rate of return index and ranks the performance of
forecasts as such. If the forecasts are better than the benchmark, the
RM will be less than one.

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-billah2006exponential}{}
Billah, Baki, Maxwell L King, Ralph D Snyder, and Anne B Koehler. 2006.
``Exponential Smoothing Model Selection for Forecasting.''
\emph{International Journal of Forecasting} 22 (2). Elsevier: 239--47.

\hypertarget{ref-de200625}{}
De Gooijer, Jan G, and Rob J Hyndman. 2006. ``25 Years of Time Series
Forecasting.'' \emph{International Journal of Forecasting} 22 (3).
Elsevier: 443--73.

\hypertarget{ref-gardner2006exponential}{}
Gardner, Everette S. 2006. ``Exponential Smoothing: The State of the
Art---Part Ii.'' \emph{International Journal of Forecasting} 22 (4).
Elsevier: 637--66.

\hypertarget{ref-granvik2010forecasting}{}
Granvik, Ashley. 2010. ``Forecasting Exchange Rates.'' Arcada-Nylands
svenska yrkeshögskola.

\hypertarget{ref-hassani2015forecasting}{}
Hassani, Hossein, Emmanuel Sirimal Silva, Nikolaos Antonakakis, George
Filis, Rangan Gupta, and others. 2015. ``Forecasting Accuracy Evaluation
of Tourist Arrivals: Evidence from Parametric and Non-Parametric
Techniques.''

\hypertarget{ref-hyndman2013forecasting}{}
Hyndman, RJ, and G Athanasopoulos. 2013. ``Forecasting: Principles and
Practice, Otexts, Australia.'' \emph{Available via: Www. OTexts.
Com/Fpp}.

\hypertarget{ref-hyndman2007automatic}{}
Hyndman, Rob J, Yeasmin Khandakar, and others. 2007. ``Automatic Time
Series for Forecasting: The Forecast Package for R.'' Monash University,
Department of Econometrics; Business Statistics.

\hypertarget{ref-hyndman2002state}{}
Hyndman, Rob J, Anne B Koehler, Ralph D Snyder, and Simone Grose. 2002.
``A State Space Framework for Automatic Forecasting Using Exponential
Smoothing Methods.'' \emph{International Journal of Forecasting} 18 (3).
Elsevier: 439--54.

\hypertarget{ref-mincer1969evaluation}{}
Mincer, Jacob A, and Victor Zarnowitz. 1969. ``The Evaluation of
Economic Forecasts.'' In \emph{Economic Forecasts and Expectations:
Analysis of Forecasting Behavior and Performance}, 3--46. NBER.

\hypertarget{ref-smith2015comparison}{}
Smith, Mick, and Rajeev Agrawal. 2015. ``A Comparison of Time Series
Model Forecasting Methods on Patent Groups.'' In \emph{MAICS}, 167--73.

\hypertarget{ref-taylor2017forecasting}{}
Taylor, Sean J, and Benjamin Letham. 2017. ``Forecasting at Scale.''

% Force include bibliography in my chosen format:
\newpage
\nocite{*}
\bibliography{}

\end{document}

### 2017-06-13 Literature Review/A Brief Literature Review.Rmd
Literature Review updated for final submission

@@
 
-Forecasting is a core element of the processes of a business.
-Producing forecasts of high quality is essential as they play an important role in inventory control, purchasing decisions of new equipment and investment decisions, and are needed for effective planning [@granvik2010forecasting].
-Automatic forecasting techniques have commonly been used in business to produce a wide variety of forecasts and are a useful tool for producing a large number of forecasts.
-Automatic techniques can be used by non-experts without training in the use of time series [@hyndman2007automatic].
-Automatic forecasting algorithms must have the ability to select the appropriate model for forecasting and estimate model parameters, without the need for human intervention [@hyndman2007automatic].
-The most common automatic forecasting techniques used are the ARIMA and Exponential Smoothing (ETS) models.
-ETS models have been used extensively in forecasting demand for inventories [@billah2006exponential].
-They have performed well in forecasting competitions against more complex models and are particularly good at forecasting seasonal data over short horizons [@hyndman2002state].
+The ability to forecast exchange rates is of great importance to policy makers, businesses and speculators.
+Policy makers are interested in forecasting exchange rates to assess how its movement will impact inflation, and how monetary policy will need to be adjusted in the future to keep inflation within the inflation target [@fahimifard2009].
+Businesses need to assess how future movements in exchange rates will affect the cost of importing goods and the demand for goods exported, while speculators aim to identify opportunities in price movements which can be exploited to make a profit.
 
-Automatic forecasting techniques can be too inflexible and do not allow for analysts with specialized knowledge to incorporate their expertise into the model effectively.
-This leads to difficulties in producing forecasts of high quality, thus the demand for high quality forecasts far exceeds the rate at which they are produced [@taylor2017forecasting].
-In December 2016, Facebook released their forecasting model, Prophet.
-According to @taylor2017forecasting, Prophet is a flexible model that can be adjusted by a large number of non-experts who have little knowledge about time series, however have expert knowledge on the data-generating process.
+Producing accurate forecasts of exchange rates has proven to be a difficult task due to the noise of the time series [@abu1996].
+Some economists argue that the exchange rate market is very efficient and future movements in price are unpredictable [@abu1996].
+This belief in the efficient market hypothesis has led many researchers to model exchange rates using a Random Walk (RW) model [@abu1996].
+Although the RW model has produced superior forecasts, some researchers believe that future prices depend on present and past events, and other techniques such as the autoregressive integrated moving average (ARIMA) model has been applied to forecast exchange rates [@abu1996].
+
+The ARIMA model is one of the most commonly used methods in financial forecasting [@fahimifard2009].
+It is known for its attractive quality of being a rich class of processes, making it possible to find a model that adequately fits the data [@fahimifard2009].
+Some research has shown that the nature of exchange rate data is non-linear, and this non-linearity may not be captured adequately by the ARIMA model [@zhang1998].
+Linear parametric forecasting techniques such as the generalised autoregressive conditional heteroskedasticity (GARCH) model has been proposed as a solution to this problem [@zhang1998].
+Although the GARCH model takes care of the non-linearity in the data, the pre-specification of the model may prevent it from capturing all the different non-linearities which may be present in the data.
+
+With the evolution of computational power, other non-linear non-parametric techniques such as Artificial Neural Networks (ANN), Support Vector Regression (SVR), Fuzzy Logistic Systems and hybrid versions of the ARIMA model have been proposed.
+The ARIMA model has been used as the standard benchmark model against which these other methods have been compared.
+Although they have produced significantly better results than the ARIMA model, building these models is a complex task.
+The ARIMA model is more tractable, less computationally expensive, and has been used as the building blocks for more advanced models such as the autoregressive fractionally integrated (ARFIMA), Fuzzy ARIMA, and seasonal ARIMA (SARIMA) model which have been used for exchange rate forecasting.
+
+Facebook recently open sourced their forecasting model Prophet.
+Prophet is said to be preferable over the ARIMA model as it has the ability to forecast at scale, is easily interpretable and has parameters that can be intuitively adjusted [@taylor2017].
+This paper aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of Bitcoin/ZAR, using a Mincer-Zarnowitz approach of measuring forecast accuracy.
+Section 2 will consider the origins of the ARIMA model and how it has performed and evolved over time by exploring empirical studies where it has been used for forecasting exchange rates.
+Section 3 will also consider other techniques that have been used to forecast exchange rates and how they have been compared to the ARIMA model.
+Section 4 will look at hybrid models that were inspired by the ARIMA model, while section 5 looks at forecasting with Prophet.
+Finally, section 6 will look at the Mincer-Zarnowitz approach to evaluating forecasts.
+
+# Forecasting Exchange Rates with the ARIMA Model
+
+## Birth of ARIMA
+ARIMA models became highly popular since its introduction by Box and Jenkins, when it was shown that they could outperform complex econometric models in a variety of situations [@spyros1997].
+The ARIMA model expresses the process {$y_t$} as a function of the weighted average of past values of the process and lagged values of the residuals.
+The weighted average of the past p values of the process represents an autoregressive (AR) process of order p.
+It feeds back past values of the process into the current value, inducing correlation between all lags of the process.
+The weighted average of the q lagged residuals represents a moving average (MA) process of order q.
+The purpose of mixing the MA process with the AR process is to reduce the large number of past values required by AR processes and to control for the autocorrelation which it creates between lagged values of the process.
+The combination of the AR(p) and MA(q) process results in a more parsimonious model, and forms a stationary autoregressive moving average (ARMA(p,q)) process defined as:
+
+$$ 
+	y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
+$$
+where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.
+
+The ARIMA(p,d,q) model generalises the ARMA model in that it includes both stationary and non-stationary processes.
+The parameter d is the degree of differencing required to render the process stationary.
+If d is equal to zero the process is stationary and equivalent to an ARMA model, and if d is strictly positive the process requires differencing to make it stationary.
+The ARIMA model can be defined succinctly using the backward shift operator B, which shifts the process back by one unit of time, and is defined as $By_t = y_{t-1}$.
+The ARIMA model has the form:
+
+$$ 
+	(1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
+$$
+where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.
+
+## The Box-Jenkins Methodology and Automatic Selection
+Box and Jenkins proposed a set of guidelines that can be followed when selecting ARIMA models.
+This systematic procedure to designing ARIMA models has made them highly popular [@spyros1997].
+It consists of a four-stage iterative process in which: 1) the process is either transformed or differenced to detrend and stabilise the variance of the data, 2) the autocorrelation and partial autocorrelation plots are used to determine the order of p and q, 3) the parameters of the model are estimated, 4) a diagnostic check is performed to ensure that the residuals are a white noise process.
+If the residuals are not white noise steps 2-4 are repeated until a satisfactory model is identified.
+On the contrary, if the diagnostic check shows that the residuals are random then the developed model is the final model used for forecasting.
+
+Due to the large number of forecasts made, it is useful to have an automatic procedure which is able to select the appropriate ARIMA model to fit the data.
+The auto.arima function in R is able to automatically choose the order of the parameters p, q and d.
+The order of differencing is determined by using the KPSS test [@ruppert2015].
+The KPSS test checks the null hypothesis of stationarity and sets d to zero if the null hypothesis is accepted, otherwise it iteratively increases d by 1 and tests the null hypothesis until it is accepted [@ruppert2015].
+Once the order of differencing has been determined, the order of p and q are chosen based on Akaike’s Information Criterion (AIC) or the Bayesian Information Criterion (BIC) [@ruppert2015].
+
+## Empirical Applications
+ARIMA models have been commonly used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012].
+They have been used as a benchmark to compare new forecasting techniques which have emerged over time.
+@kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies.
+Its results were used to evaluate the performance of three Neural Network models.
+The Normalised Mean Square Error (NMSE) and Mean Absolute Error (MAE) were used to measure the forecast errors while the Directional Symmetry (DS), Correct Up trend (CU) and Correct Down trend (CD) were used to measure the accuracy of the direction of the forecasts.
+Similarly, @lin2012 employed an ARIMA model to forecast the Taiwan New Dollar exchange rate.
+The aim of applying the ARIMA model was to analyse the ability of a proposed hybrid model created from the Squares Support Vector Regression (LSSVR) and Empirical Mode Decomposition (EMD) model.
+@lin2012 used the same indicators as @kamruzzaman2003 to evaluate the accuracy of the direction of the forecasts.
+To measure the forecast errors produced they applied the mean absolute percentage error (MAPE), root‐mean‐square error (RMSE) and mean absolute difference (MAD).
+@khashei2012 also investigated the ability to forecast exchange rates by considering the British pound and US dollar, over a short and long forecast horizon.
+They compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN), to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark, by claiming that it is the most important linear model.
+The ARIMA model has also yielded satisfactory results when predicting exchange rates.
+@nwankwo2014 forecasted the Nairo/dollar rate and diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on the AIC.
+
+The ARIMA model is attractive as it is tractable and produces good forecasts in short time periods where more than 100 observations are used [@tseng2001]. 
+Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data.
+
+## Evolution of ARIMA
+Over time, the ARIMA model has evolved to cater for a wider variety of data that it can be applied to, and to compensate for some of its shortcomings.
+The most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).
+
+The SARIMA model was introduced by Box and Jenkins as a generalisation of the ARIMA model to capture the periodic behaviour of data.
+SARIMA models extend the ARIMA class by including a term for seasonal differencing.
+Exchange rate data that has been observed have shown seasonality, and have been attempted to be fitted by SARIMA models [@etuk2013].
+@etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit to the data.
+Their results showed that the SARIMA model adequately described the variation in the exchange rate series.
+This agrees with the results of a study by @etuk2012 which fitted a multiplicative SARIMA model to the Naira/Dollar exchange rates.
+
+The ARFIMA model is another generalisation of the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain.
+While the ARIMA model displays exponentially decaying autocorrelation between observations the further apart they are in time, the ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993].
+This makes the ARFIMA model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993].
+@cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.
+Although the ARFIMA model fitted the data well, it failed to beat the random walk in out-of sample forecasts.
+
+GARCH models were later developed due to the failure of ARIMA models to capture the volatility in financial markets [@anastasakis2009].
+They have been widely used to forecast the volatility of stock returns, interest rates and exchange rates and are useful for portfolio allocation decisions as well as option pricing [@cermak2017].
+@cermak2017 employed a GARCH(1,1) model to analyse the volatility of Bitcoin in countries in which Bitcoin is commonly traded, while Hsieh (1989) applied a GARCH model to investigate five exchange rates.
+Hsieh’s (1989) results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.
+In a similar study, @fahimifard2009 illustrated that the GARCH model outperformed the ARIMA, while Neural Networks outperformed both the GARCH and ARIMA model.
+
+# Other techniques used to forecast exchange rates
+Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques.
+These more complex techniques are non-linear and can fit complex time series more easily [@castillo2002].
+Unlike regression analysis and the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002].
+Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and the Fuzzy Logistic Systems.
+
+## Neural Networks
+Artificial Neural Networks are commonly used in financial forecasting and are more advantageous than other non-linear forecasting methods.
+They can approximate any continuous function, are data driven, and can adapt to non-stationary environments [@khashei2011].
+Neural Networks mimic the structure of the brain and consists of three layers: the input, hidden and output layer, each of which contain nodes which are representative of the neurons in our brain.
+The job of the nodes in the input layer is to send signals along connections to the nodes in the hidden layer, where the connections are representative of the synapses in our brain.
+The nodes in the hidden layer uses an activation function and assigns a weight to each connection according to the level of importance of information.
+The value in the hidden layer is then sent along connections to the output layer, where weights are again assigned to each connection.
+The error is then computed in the output layer and is backpropagated through the network.
+The weights are adjusted according to the error that they contribute, allowing the ANN to be trained and learn.
+This learning process is repeated until the network achieves the desired output.
+
+ANN’s have had many successful applications in forecasting exchange rates.
+In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates.
+Superior results were produced when compared to the ARIMA and GARCH model using the RMSE, MSE and MAD as measure of performance.
+Similar results were found by @zhang1998 however the ANN failed to outperform the random walk over long forecast horizons.
+Contrary to the findings above, @kuan1995 attempted to investigate how well feedforward and recurrent networks capture the non-linearity of exchange rate data and obtained conflicting results with neural networks having significantly better market timing ability or lower out-of-sample Mean Square Prediction Error (MSPE) compared to the ARMA models for some exchange rates only.
+@franses1998 suggested that the non-linear feature of exchange rates that are picked up by ANNs may actually be due to neglected GARCH effects.
+They investigated the application of ANNs to forecast exchange rates and found that the presence of GARCH in the data may mistakenly lead one to believe that the returns can be forecasted on the exchange rates themselves.
+Their results show that there is no gain in producing out-of-sample forecasts using ANN if the data is not truly non-linear.
+
+Although neural networks have been broadly applied in financial forecasting, the process of building neural networks is a complex task and there is no consistent method to design them compared to the traditional Box-Jenkins ARIMA model.
+Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes and the training sample size [@zhang1998].
+
+## Fuzzy Logistic Systems
+Fuzzy Logistic Systems were initially developed to solve problems involving linguistic terms, and have been used successful in financial forecasting [@khashei2009].
+Unlike computers, the human decision-making process includes a range of possibilities spanning ‘yes’ and ‘no’.
+Fuzzy logic tries to imitate this reasoning and decision-making process by taking a discrete decision space that consists of a large number of states and turning it into a continuous space.
+This allows for more finer decisions to be provided as opposed to discrete decisions.
+
+@santos2007 investigated how well fuzzy systems and neural networks perform compared to the traditional ARMA and GARCH model.
+He examined the forecasts of  Brazilian exchange rate returns looking at different frequencies of the series and compared their one-step-ahead forecasts.
+By analysing the RMSE, U-Theil inequality index, percentage of corrected predicted signals (CPS), and the Pesaran-Timmermann (PT) predictive failure statistic, he found that the Fuzzy Systems and Neural Networks achieved higher returns based on the forecasts they produced.
+Similar results were found by @khashei2009 when he analysed the predictive capabilities of Fuzzy Systems, Neural Networks, the traditional ARIMA model, and a Fuzzy ARIMA model.
+The MAE and MSE of the ARIMA model were higher compared to the other models.
+
+Although Fuzzy Logistic Systems has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate reasoning and are more suitable for problems which do not require high accuracy.
+Like ANN’s there is no systematic approach to designing Fuzzy Logistic Systems and they are only understandable when simple.
+
+# Hybrid ARIMA Models
+Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA, and of more complex non-linear models.
+By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations.
+Some of the hybrid models which have been commonly used are the Fuzzy ARIMA and ANN-ARIMA model.
+
+##Fuzzy ARIMA
+The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data [@tseng2001].
+In a world that is constantly changing and with the rapid advancement of technology, access to large amounts of historical data is difficult to obtain.
+On the contrary, a fuzzy regression model requires little historical data however produces wide prediction intervals if extreme values are present in the data.
+The Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to exploit the advantages of both models while simultaneously overcoming their limitations.
+@tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model’s appropriateness and power.
+The Fuzzy ARIMA not only produced forecasts that were superior to the ARIMA and fuzzy time series models but also provided an upper and lower bound which can be used by decision makers to determine the best and worst possible situations.
+
+##ANN-ARIMA
+Although ARIMA models are powerful, they require non-stationary data to be differenced and impose prior assumptions onto the distribution of the data [@ince2006].
+In contrast, machine learning techniques such as ANNs do not impose any assumptions onto the data generating process however, being data-driven, are sensitive to the number of input nodes used.
+The ANN-ARIMA model draws on the strengths of both models and overcomes these individual difficulties.
+@ince2006 created an ANN-ARIMA model by using the ARIMA to determine the number input nodes required by an ANN for three exchange rates.
+When the ANN-ARIMA model was compared to the pure ARIMA, the hybrid model outperformed the ARIMA based on the MSE.
+The findings of @khashei2011 study agree with these results.
+They implemented an ARIMA model and used its residuals together with past observations of the data as inputs for the ANN.
+Their hybrid model had superior in-sample and out-sample forecasts compared to the random walk, linear AR and ANN model.
+
+Although the traditional ARIMA models produce less superior forecasts than its hybrid forms and other complex non-linear techniques, its forecasts are still satisfactory.
+They are simple models that are easy to implement and have a consistent method of model design and selection.
+ARIMA models are also more robust and efficient than complex structural models in relation to short-run forecasting.
+The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison justifies it as a good starting point to compare it to Facebook’s forecasting method, Prophet, that was recently released.
+
+# Forecasting with Prophet 
+The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series.
+Prophet differs to traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process.
 Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared.
-This is Facebook’s definition of forecasting at scale, and provides a solution to the problems that automatic forecasting pose [@taylor2017forecasting].
+This is Facebook’s definition of forecasting at scale.
 
-# Exponential State Space Models \label{Exponential State Space Models}
+## Construction of Prophet
+Prophet consists of a decomposable model of the form:
 
-## Exponential Smoothing Methods 
-ETS methods originated in the 1950’s and is the key work of Brown , Holt, and Winters [@de200625].
-It has been widely used for many applications in industry and has been the foundation of many successful forecasting methods [@de200625].
-Exponential smoothing techniques produce forecasts by calculating them as a weighted average of past observations.
-Observations that are most recent are more heavily weighted, with these weights decaying exponentially as the observations get older.
+$$ 
+	y(t) = g(t) + s(t) + h(t) + e_t
+$$
+where the components of the model represent the growth, seasonality and holiday respectively, and $e_t$ is white noise.
 
-Exponential smoothing methods were originally classified by Pegel according to the components present in the time series (trend and seasonality), and whether they were additive or multiplicative in nature [@de200625].
-Gardner and Taylor further extended this classification to include a damped additive and multiplicative trend respectively [@de200625].
-A taxonomy of exponential smoothing methods was later created by Hyndman et al.
-and extended by Taylor [@de200625].
-It considered all possible combinations of the trend and seasonality components thus producing 15 methods of exponential smoothing as seen in table \ref{tab1}.
+These components consist of linear and non-linear functions of time.
+This differs to ARIMA models in which future values of the process are linear functions of previous observations and lagged residuals.
+Prophet is more like a Generalized Additive Model (GAM) which is a regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017].
+Prophet, like the GAM, frames the forecasting problem as a curve fitting exercise, using backfitting to find the regression functions.
+The GAM is fitted quickly, allowing the analyst to interactively change the model parameters [@taylor2017].
 
-```{r ShortTable, results = 'asis'}
-
-library(xtable)
-
-WD <- getwd()
-data <- read.csv( file.path(WD, "Data/ETSMethods.csv"), header=TRUE, sep=" ", check.names=FALSE)
-
-table <- xtable(data, caption = "Taxonomy of Exponential Smoothing Methods \\label{tab1}")
-  print.xtable(table, 
-             table.placement = 'H', 
-             comment = FALSE,
-             caption.placement = 'bottom'
-             )
-
-```
-
-\hfill
-
-The most common of these methods are Simple Exponential Smoothing (N,N), Holt’s linear method (A,N) and Holt-Winters seasonal method (A,A).
-
-Although exponential smoothing methods performed well and have been widely used in business, they did not have any statistical framework underpinning them [@de200625].
-They produced point estimates rather than prediction intervals and selection of a forecasting method was specific to the time series in question.
-Efforts were made by Box and Jenkins to create a statistical framework that underlies exponential smoothing methods and they found that forecasts produced by linear methods are special cases of ARIMA models [@de200625].
-This however does not hold for the multiplicative methods.
-
-## State Space Models 
-The motivation behind state space models was the need for a statistical framework that underlies all methods of exponential smoothing [@de200625].
-@hyndman2002state provided this framework as an extension of Ord, Koehler, and Snyder’s class of exponential state space models which underpinned some of the smoothing methods [@de200625].
-He proposed two state space models, one corresponding to an additive error and the other to a multiplicative error.
-By specifying a distribution for these error terms, he turned the previously deterministic methods into stochastic models.
-These two models were then applied to all 15 smoothing methods, producing 30 models in total.
-
-The two state space models produce identical point estimates to each other as well as to the exponential smoothing methods, however they will generate differing prediction intervals [@hyndman2013forecasting].
-Each model consists of a measurement equation and a transition equation.
-The measurement equation shows the relationship between the observations and the unobserved components (level, trend seasonality), while the transition equation shows how the unobserved components change over time.
-
-The introduction of the exponential state space model has allowed for model selection using information criteria rather than using an ad hoc method.
-It has also made the calculation of the likelihood easier and simulation from the underlying model possible [@hyndman2002state].
-
-## Application and Performance 
-In a paper by @smith2015comparison, Holt-Winters exponential smoothing (HWES) model was applied to time series data relating to data on patents.
-The patent data was classified into three different groups; monthly forecasts were made and the results were compared to forecasts produced by ARIMA models.
-To avoid over-forecasting, the trend of the HWES was damped to produce forecasts with a flat trend for later observations.
-The results showed that all models forecasted the data adequately.
-The HWES model fitted the data well and performed better than the ARIMA models in the case where the time series for the patent group was fairly stationary.
-Based on the forecasting results, it could not be concluded that a specific time series model is better for forecasting patent data.
-
-@hassani2015forecasting forecasted tourism demand in specific European countries over a short, medium and long term horizon.
-The ETS model was used to produce forecasts and was compared to six parametric and non-parametric techniques.
-The results showed that no single model is best for producing forecasts across all countries and time horizons.
-The ETS model together with the Neural Networks (NN) and Fractionalized ARIMA (ARFIMA) produced forecasts with the least amount of accuracy and are thus not relevant models for forecasting tourism demand in Europe.
-24 steps-ahead error forecasts produced by the ETS model were found to be larger than the errors produced by the benchmark model - Recurrent Single Spectrum Analysis (SSA-R) across all European countries.
-
-## Robustness 
-@gardner2006exponential explained the robustness of ETS models by comparing them to equivalent models.
-Simple exponential smoothing which is equivalent to an ARIMA(0,1,1) model is the most robust method and has performed well in forecasting many types of time series.
-By comparing simple exponential smoothing to lower order ARIMA models, it is seen that errors that arise due to model specification is not as much of a problem as ARIMA models.
-Furthermore, ARIMA models tend to produce larger MSE due to model selection errors, and this is worsened if the errors are not normally distributed [@gardner2006exponential].
-
-Satchell and Timmermann found that for time series with a finite history, the weights assigned to each observation in the simple exponential smoothing method are robust, provided that the variance of the random walk relative to the variance of the error component is not too small [@gardner2006exponential].
-
-Forecasts produced by simple exponential smoothing have performed well in modelling annual sales of different products, as well as aggregated economic series with low sampling frequencies [@gardner2006exponential].
-
-# Prophet
-
-## Bayesian Generalized Additive Model
-Prophet is the key work of @taylor2017forecasting and consists of a decomposable model with a component for growth, seasonality, and holidays.
-These components consist of linear and non-linear functions of time.
-This is similar to a Generalized Additive Model (GAM) which is a regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017forecasting].
-Prophet, like GAM, frames the forecasting problem as a curve fitting exercise using backfitting to find the regression functions.
-
-The growth component is modelled in a similar way to modelling population growths which uses a logistic growth model [@taylor2017forecasting].
+The growth component is modelled in a similar way to population growths which use a logistic growth model [@taylor2017].
 Populations typically grow non-linearly (although the growth component could also be linear) up to an upper bound known as the carrying capacity, and remains constant thereafter.
-The rate at which the population grows changes over time, and this is accounted for by including changepoints in the model where the growth rate may be changed [@taylor2017forecasting].
-These changepoints can be automatically selected and have a Laplace prior distribution placed on them [@taylor2017forecasting].
-The parameter of the prior distribution can be used to adjust the growth rate of the changepoints.
-This allows non-experts with knowledge about events that may affect growth to use the parameter as a knob and adjust it to increase or decrease the number of changepoints [@taylor2017forecasting].
-It also allows for the analyst to add changepoints which the automatic selection procedure may have missed [@taylor2017forecasting].
-Furthermore, analysts may also specify the carrying capacity and adjust it based on their knowledge of the total market size [@taylor2017forecasting].
+The rate at which the population grows changes over time, and this is accounted for by including changepoints in the model where the growth rate can be automatically selected and may be adjusted [@taylor2017].
+This allows non-experts with knowledge about events that may affect growth to use the parameter as a knob and adjust it to increase or decrease the number of changepoints [@taylor2017].
+It also allows for the analyst to add changepoints which the automatic selection procedure may have missed [@taylor2017].
+Furthermore, analysts may also specify the carrying capacity and adjust it based on their knowledge of the total market size [@taylor2017].
 
@@
 This allows for multiple seasonality components with different periods to be added to the model.
-The seasonality component is modelled by a Fourier series, with the parameters of the Fourier series having a Normal prior distribution [@taylor2017forecasting].
-The variance of the parameters can be adjusted by analysts to smooth the model and change how much of historical seasonality is projected to the future [@taylor2017forecasting].
+The variance of the parameters of the seasonality component’s distribution can be adjusted by analysts to smooth the model and change how much of historical seasonality is projected to the future [@taylor2017].
 
-The name, date, and country of past and future holidays and events may be inputted by the analyst into a list [@taylor2017forecasting].
+The name, date, and country of past and future holidays and events may also be inputted by the analyst into a list [@taylor2017].
 By specifying the country in which the events take place or the holidays occur, separate lists can be populated for global events/holidays and country-specific events/holidays.
 The union of the two lists are then used for forecasting.
-Like seasonality, a Normal prior distribution is placed on the parameters of the holidays, and the variance of the parameters can be adjusted by analysts to smooth the model [@taylor2017forecasting].
+Like seasonality, the variance of the parameters of the holiday component’s distribution, can be adjusted by analysts to smooth the model [@taylor2017].
 
-## Application and Performance 
-@taylor2017forecasting forecasted the number of events on Facebook using Prophet.
+## Empirical Applications
+Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge in the model building process and has produced significantly improved forecasts compared to the ARIMA model.
+@taylor2017 forecasted the number of events on Facebook using Prophet.
 The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend.
-The forecasts produced by Prophet were compared to forecasts produced by common automatic forecasting techniques such as ETS, ARIMA, and the seasonal naïve model, as well as to simple models such as the naïve model.
-While the ETS and seasonal naïve model were quite robust, the ARIMA forecasts were fragile.
+The forecasts produced by Prophet were compared to common automatic forecasting techniques such as exponential smoothing, ARIMA, and the seasonal naïve model, as well as to simple models such as the naïve model.
+While exponential smoothing and the seasonal naïve model were quite robust, the ARIMA forecasts were fragile.
 No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.
-Hence, Prophet had lower forecasting errors compared to the automatic forecasting methods.
 
-## Robustness 
-Prophet is a flexible model and has intuitive parameters that can be easily interpreted by human beings.
-The GAM is fitted quickly, allowing the analyst to interactively change the model parameters [@taylor2017forecasting].
-Unlike ARIMA models, Prophet can produce forecasts over different scales and allows for missing values in the time series without the need for interpolation [@taylor2017forecasting].
-
+## Semi-Automatic Selection
 When a large number of forecasts are produced, manually identifying problematic forecasts becomes a time consuming and difficult task.
 Prophet provides a semi-automated forecast evaluation system that selects the best model that fits the data.
-When there are large forecast errors, the forecasts are flagged so that the analyst can explore the cause of the errors, identify and remove potential outliers and either adjust the model or choose a more appropriate model [@taylor2017forecasting].
+When there are large forecast errors, the forecasts are flagged so that the analyst can explore the cause of the errors, identify and remove potential outliers and either adjust the model or choose a more appropriate model [@taylor2017].
+Unlike ARIMA’s fully-automated evaluation system, Prophet provides interactive feedback and keeps the analyst in the loop.
 
-# Evaluation of Forecasts
+Prophet’s ability to forecast at scale enables it to model a wide variety of data and may be able to capture the non-linearity present in exchange rates.
+In contrast, the ARIMA model fails to capture the non-linearity inherent in exchange rate data and only produces good forecasts over short time horizons.
+Prophet is also a flexible model that has intuitive parameters which can be easily interpreted and modified by human beings [@taylor2017].
+This may be useful to non-experts who have domain knowledge about factors that may affect the movement of exchange rate prices.
+Furthermore, unlike the ARIMA model, Prophet can produce forecasts over irregular time intervals and allows for missing values in the time series without the need for interpolation [@taylor2017].
+If Prophet produces forecasts that are as good as the ARIMA model when forecasting Bitcoin/ZAR, it can be compared to more complex forecasting methods that are less tractable such as the hybrid models and machine learning techniques seen earlier.
 
-## Common Methods 
-Model selection has historically been very subjective, with no techniques that offered mathematical rigour for choosing models [@smith2015comparison].
-Progress has been made over time and many mathematical techniques have been introduced for model selection.
-Use of information criteria is one such method, with Akaike’s information criterion (AIC) and Bayes information criterion (BIC) being the most common [@smith2015comparison].
-These information criteria minimize forecasting errors while penalizing it for overfitting.
-Information criteria should be used to compare models of similar structure only.
-To compare models of different structure, the most commonly used measures of forecast errors are the Root Mean Squared error (RMSE), Mean Absolute Error (MAE), Mean Absolute Percentage Error (MAPE), and Mean Absolute Scaled Error (MASE).
-A lower value of these statistics indicates that the forecasts are better, with a MASE less than one indicating that the forecasts are better than those produced by a naïve model [@smith2015comparison].
+# Mincer-Zarnowitz Approach to Forecast Evaluation 
+The Mincer-Zarnowitz approach to evaluating forecast accuracy is commonly used and can be useful when comparing the forecasts produced by Prophet and the ARIMA model.
+@mincer1969 proposed an absolute and relative measure to evaluate forecast accuracy.
+Absolute measures consider the distance between actual and predicted values.
+To analyse the absolute errors produced by the forecasts, the observed values are regressed against the predicted values.
+The intercept of the regression equation represents the mean distance between the observed and predicted values while the slope represents the correlation between the residual errors and the predicted values.
+A zero intercept implies that the forecasts are unbiased and do not overestimate or underestimate the data, while a unity slope implies that the forecasts are efficient and uncorrelated with the residual errors.
+A joint hypothesis test is performed to check this efficiency and bias of the forecasts, and models which produce the best results are selected.
 
-@smith2015comparison have used these methods in their paper when evaluating patent forecasts and making comparisons between the HWES and ARIMA models.
-In a paper by @hassani2015forecasting they chose their ETS model based on the selection that R automatically made for the time series.
-They compared models of different structure using the RMSE and a modified Diebold-Mariano test.
-@hyndman2002state suggests that ETS models should be selected using the AIC rather than the MSE or MAPE.
-The AIC provides a way of choosing between models with an additive error and models with a multiplicative error since it is based on the likelihood function.
-The MSE and MAPE are not able to select between the two error types because both models produce the same point estimates.
-According to @de200625, the best approach of selecting models with different structure is the Diebold-Mariano test.
-The MSE is scale dependent and should not be used to make comparisons while the MAPE encounters difficulties when the time series values are close to or equal to zero.
+Absolute forecast measures cannot be used to make comparisons between forecasts with different scales or economic variables.
+Furthermore, the size of forecasting errors is not as significant as the consequences of forecasting errors and how they impact the decision making process.
+Hence, rather than using absolute measures of forecast accuracy, @mincer1969 suggested the use of relative measures.
+Relative accuracy analysis allows for meaningful comparisons of different forecasting methods to be made.
+It uses an index which considers the ratio of the MSE of the forecast to the MSE of a benchmark forecasting method.
+A useful benchmark that may be used is an extrapolation of the data history, as it is a cost effective and accessible method, however any method which is relevant for comparison may be used.
+This ratio is known as the Relative Mean Square Error (RM) for forecast evaluation.
+The numerator can be viewed as a return which is inversely proportional to the MSE error of the forecasts, while the denominator can be viewed as the cost of forecasting which is inversely proportional to the MSE of the benchmark [@mincer1969].
+Hence the ratio is representative of a rate of return index and ranks the performance of forecasts as such.
+Models that result in forecasts with a RM that is less than one are said to produce superior forecasts to the benchmark model being considered.
 
-## Mincer-Zarnowitz Approach 
-@mincer1969evaluation proposed two methods of measuring forecast accuracy - an absolute and a relative measure.
-Absolute measures of accuracy measure the distance between actual and predicted values.
-If the mean distance is equal to zero, the forecast is said to be unbiased [@mincer1969evaluation].
-To analyse absolute forecast errors, the observed values are regressed against the predicted values.
-A joint hypothesis test is performed to check if the intercept is equal to zero and the slope is equal to one.
-If the intercept is equal to zero, the forecasts are unbiased and do not overestimate or underestimate the data.
-If the slope is equal to one, the residual errors are uncorrelated with the predictions and the forecast is said to be efficient [@mincer1969evaluation].
-The MSE is a measure that can be decomposed into the residual variance and two components, one reflecting the bias and the other reflecting the inefficiency of the forecast.
-If the forecasts are unbiased and efficient, the MSE will reduce to the residual variance [@mincer1969evaluation].
-Absolute forecast measures cannot be used to make comparisons between forecasts with different scales or economic variables.
-Thiel states that the consequences of forecasting errors and how they impact decisions is of greater importance than the size of the forecasting errors [@mincer1969evaluation].
-
-Rather than using absolute measures of forecast accuracy, @mincer1969evaluation suggested the use of relative measures of forecast accuracy.
-Relative accuracy analysis allows for meaningful comparisons of different forecasting methods to be made.
-They proposed an index which considers the ratio of the MSE of the forecast to the MSE of a benchmark forecasting method.
-They used an extrapolation of the data history as a benchmark, as it is a cost effective and accessible method, however the benchmark may be any method which is relevant for comparison.
-This ratio is known as the Relative Mean Square Error (RM).
-
-The numerator represents a return which is inversely proportional to the MSE error of the forecasts, while the denominator represents the cost of forecasting which is inversely proportional to the MSE of the benchmark [@mincer1969evaluation].
-Hence the ratio is representative of a rate of return index and ranks the performance of forecasts as such.
-If the forecasts are better than the benchmark, the RM will be less than one.
+# Conclusion
+ARIMA models have been highly popular since its introduction by Box and Jenkins and are widely used for financial forecasting.
+They are easy to implement, have a systematic procedure for model design, and produce good forecasts over short time horizons.
+Although robust, they have the limitation of requiring large amounts of historical observations and they fail to capture the non-linearity inherent in exchange rate data.
+With the evolution of computational power and the advancement of statistics, more complex techniques such as ANN’s, Fuzzy Logistic Systems, and Hybrid ARIMA models have been proposed and compared to the traditional ARIMA model.
+Although these models produce superior results to the ARIMA model, they are less tractable and require the analyst to understand the complex statistical intricacies involved.
+Prophet’s Bayesian approach to forecasting is flexible and allows for non-experts to incorporate their domain knowledge in the model building process.
+If Prophet produces forecasts that are as good as the ARIMA model when forecasting Bitcoin/ZAR, it can be compared to these more complex forecasting methods that are less tractable and scalable.
 

### 2017-06-13 Literature Review/A_Brief_Literature_Review.tex
Literature Review updated for final submission

@@
 
-Forecasting is a core element of the processes of a business. Producing
-forecasts of high quality is essential as they play an important role in
-inventory control, purchasing decisions of new equipment and investment
-decisions, and are needed for effective planning (Granvik
-\protect\hyperlink{ref-granvik2010forecasting}{2010}). Automatic
-forecasting techniques have commonly been used in business to produce a
-wide variety of forecasts and are a useful tool for producing a large
-number of forecasts. Automatic techniques can be used by non-experts
-without training in the use of time series (R. J. Hyndman, Khandakar,
-and others \protect\hyperlink{ref-hyndman2007automatic}{2007}).
-Automatic forecasting algorithms must have the ability to select the
-appropriate model for forecasting and estimate model parameters, without
-the need for human intervention (R. J. Hyndman, Khandakar, and others
-\protect\hyperlink{ref-hyndman2007automatic}{2007}). The most common
-automatic forecasting techniques used are the ARIMA and Exponential
-Smoothing (ETS) models. ETS models have been used extensively in
-forecasting demand for inventories (Billah et al.
-\protect\hyperlink{ref-billah2006exponential}{2006}). They have
-performed well in forecasting competitions against more complex models
-and are particularly good at forecasting seasonal data over short
-horizons (R. J. Hyndman et al.
-\protect\hyperlink{ref-hyndman2002state}{2002}).
+The ability to forecast exchange rates is of great importance to policy
+makers, businesses and speculators. Policy makers are interested in
+forecasting exchange rates to assess how its movement will impact
+inflation, and how monetary policy will need to be adjusted in the
+future to keep inflation within the inflation target (Fahimifard et al.
+\protect\hyperlink{ref-fahimifard2009}{2009}). Businesses need to assess
+how future movements in exchange rates will affect the cost of importing
+goods and the demand for goods exported, while speculators aim to
+identify opportunities in price movements which can be exploited to make
+a profit.
 
-Automatic forecasting techniques can be too inflexible and do not allow
-for analysts with specialized knowledge to incorporate their expertise
-into the model effectively. This leads to difficulties in producing
-forecasts of high quality, thus the demand for high quality forecasts
-far exceeds the rate at which they are produced (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). In December 2016,
-Facebook released their forecasting model, Prophet. According to Taylor
-and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017}),
-Prophet is a flexible model that can be adjusted by a large number of
-non-experts who have little knowledge about time series, however have
-expert knowledge on the data-generating process. Prophet allows for a
-large number of forecasts to be produced across a variety of problems
-and consists of a robust evaluation system that allows for a large
-number of forecasts be evaluated and compared. This is Facebook's
-definition of forecasting at scale, and provides a solution to the
-problems that automatic forecasting pose (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}).
+Producing accurate forecasts of exchange rates has proven to be a
+difficult task due to the noise of the time series (Abu-Mostafa and
+Atiya \protect\hyperlink{ref-abu1996}{1996}). Some economists argue that
+the exchange rate market is very efficient and future movements in price
+are unpredictable (Abu-Mostafa and Atiya
+\protect\hyperlink{ref-abu1996}{1996}). This belief in the efficient
+market hypothesis has led many researchers to model exchange rates using
+a Random Walk (RW) model (Abu-Mostafa and Atiya
+\protect\hyperlink{ref-abu1996}{1996}). Although the RW model has
+produced superior forecasts, some researchers believe that future prices
+depend on present and past events, and other techniques such as the
+autoregressive integrated moving average (ARIMA) model has been applied
+to forecast exchange rates (Abu-Mostafa and Atiya
+\protect\hyperlink{ref-abu1996}{1996}).
 
-\section{\texorpdfstring{Exponential State Space Models
-\label{Exponential State Space Models}}{Exponential State Space Models }}\label{exponential-state-space-models}
+The ARIMA model is one of the most commonly used methods in financial
+forecasting (Fahimifard et al.
+\protect\hyperlink{ref-fahimifard2009}{2009}). It is known for its
+attractive quality of being a rich class of processes, making it
+possible to find a model that adequately fits the data (Fahimifard et
+al. \protect\hyperlink{ref-fahimifard2009}{2009}). Some research has
+shown that the nature of exchange rate data is non-linear, and this
+non-linearity may not be captured adequately by the ARIMA model (Zhang
+and Hu \protect\hyperlink{ref-zhang1998}{1998}). Linear parametric
+forecasting techniques such as the generalised autoregressive
+conditional heteroskedasticity (GARCH) model has been proposed as a
+solution to this problem (Zhang and Hu
+\protect\hyperlink{ref-zhang1998}{1998}). Although the GARCH model takes
+care of the non-linearity in the data, the pre-specification of the
+model may prevent it from capturing all the different non-linearities
+which may be present in the data.
 
-\subsection{Exponential Smoothing
-Methods}\label{exponential-smoothing-methods}
+With the evolution of computational power, other non-linear
+non-parametric techniques such as Artificial Neural Networks (ANN),
+Support Vector Regression (SVR), Fuzzy Logistic Systems and hybrid
+versions of the ARIMA model have been proposed. The ARIMA model has been
+used as the standard benchmark model against which these other methods
+have been compared. Although they have produced significantly better
+results than the ARIMA model, building these models is a complex task.
+The ARIMA model is more tractable, less computationally expensive, and
+has been used as the building blocks for more advanced models such as
+the autoregressive fractionally integrated (ARFIMA), Fuzzy ARIMA, and
+seasonal ARIMA (SARIMA) model which have been used for exchange rate
+forecasting.
 
-ETS methods originated in the 1950's and is the key work of Brown ,
-Holt, and Winters (De Gooijer and Hyndman
-\protect\hyperlink{ref-de200625}{2006}). It has been widely used for
-many applications in industry and has been the foundation of many
-successful forecasting methods (De Gooijer and Hyndman
-\protect\hyperlink{ref-de200625}{2006}). Exponential smoothing
-techniques produce forecasts by calculating them as a weighted average
-of past observations. Observations that are most recent are more heavily
-weighted, with these weights decaying exponentially as the observations
-get older.
+Facebook recently open sourced their forecasting model Prophet. Prophet
+is said to be preferable over the ARIMA model as it has the ability to
+forecast at scale, is easily interpretable and has parameters that can
+be intuitively adjusted (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}). This paper aims to compare
+the forecasts produced by the traditional ARIMA model and Prophet
+through an evaluation of Bitcoin/ZAR, using a Mincer-Zarnowitz approach
+of measuring forecast accuracy. Section 2 will consider the origins of
+the ARIMA model and how it has performed and evolved over time by
+exploring empirical studies where it has been used for forecasting
+exchange rates. Section 3 will also consider other techniques that have
+been used to forecast exchange rates and how they have been compared to
+the ARIMA model. Section 4 will look at hybrid models that were inspired
+by the ARIMA model, while section 5 looks at forecasting with Prophet.
+Finally, section 6 will look at the Mincer-Zarnowitz approach to
+evaluating forecasts.
 
-Exponential smoothing methods were originally classified by Pegel
-according to the components present in the time series (trend and
-seasonality), and whether they were additive or multiplicative in nature
-(De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). Gardner
-and Taylor further extended this classification to include a damped
-additive and multiplicative trend respectively (De Gooijer and Hyndman
-\protect\hyperlink{ref-de200625}{2006}). A taxonomy of exponential
-smoothing methods was later created by Hyndman et al. and extended by
-Taylor (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
-It considered all possible combinations of the trend and seasonality
-components thus producing 15 methods of exponential smoothing as seen in
-table \ref{tab1}.
+\section{Forecasting Exchange Rates with the ARIMA
+Model}\label{forecasting-exchange-rates-with-the-arima-model}
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rlll}
-  \hline
- & Trend &  & Seasonality \\ 
-  \hline
- & None(N) & Additive(A) & Multiplicative(M) \\ 
-  None(N) & (N,N) & (N,A) & (N,M) \\ 
-  Additive(A) & (A,N) & (A,A) & (A,M) \\ 
-  Additive\_damped(Ad) & (Ad,N) & (Ad,A) & (Ad,M) \\ 
-  Multiplicative(M) & (M,N) & (M,A) & (M,M) \\ 
-  Multiplicative\_damped(Md) & (Md,N) & (Md,A) & (Md,M) \\ 
-   \hline
-\end{tabular}
-\caption{Taxonomy of Exponential Smoothing Methods \label{tab1}} 
-\end{table}
+\subsection{Birth of ARIMA}\label{birth-of-arima}
 
-\hfill
+ARIMA models became highly popular since its introduction by Box and
+Jenkins, when it was shown that they could outperform complex
+econometric models in a variety of situations (Hibon and Makridakis
+\protect\hyperlink{ref-spyros1997}{1997}). The ARIMA model expresses the
+process \{\(y_t\)\} as a function of the weighted average of past values
+of the process and lagged values of the residuals. The weighted average
+of the past p values of the process represents an autoregressive (AR)
+process of order p.~It feeds back past values of the process into the
+current value, inducing correlation between all lags of the process. The
+weighted average of the q lagged residuals represents a moving average
+(MA) process of order q. The purpose of mixing the MA process with the
+AR process is to reduce the large number of past values required by AR
+processes and to control for the autocorrelation which it creates
+between lagged values of the process. The combination of the AR(p) and
+MA(q) process results in a more parsimonious model, and forms a
+stationary autoregressive moving average (ARMA(p,q)) process defined as:
 
-The most common of these methods are Simple Exponential Smoothing (N,N),
-Holt's linear method (A,N) and Holt-Winters seasonal method (A,A).
+\[ 
+    y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
+\] where c is a constant and \{\(e_t\)\} is a white noise process with
+zero mean and variance \(\sigma^2\).
 
-Although exponential smoothing methods performed well and have been
-widely used in business, they did not have any statistical framework
-underpinning them (De Gooijer and Hyndman
-\protect\hyperlink{ref-de200625}{2006}). They produced point estimates
-rather than prediction intervals and selection of a forecasting method
-was specific to the time series in question. Efforts were made by Box
-and Jenkins to create a statistical framework that underlies exponential
-smoothing methods and they found that forecasts produced by linear
-methods are special cases of ARIMA models (De Gooijer and Hyndman
-\protect\hyperlink{ref-de200625}{2006}). This however does not hold for
-the multiplicative methods.
+The ARIMA(p,d,q) model generalises the ARMA model in that it includes
+both stationary and non-stationary processes. The parameter d is the
+degree of differencing required to render the process stationary. If d
+is equal to zero the process is stationary and equivalent to an ARMA
+model, and if d is strictly positive the process requires differencing
+to make it stationary. The ARIMA model can be defined succinctly using
+the backward shift operator B, which shifts the process back by one unit
+of time, and is defined as \(By_t = y_{t-1}\). The ARIMA model has the
+form:
 
-\subsection{State Space Models}\label{state-space-models}
+\[ 
+    (1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
+\] where c is a constant and \{\(e_t\)\} is a white noise process with
+zero mean and variance \(\sigma^2\).
 
-The motivation behind state space models was the need for a statistical
-framework that underlies all methods of exponential smoothing (De
-Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). R. J.
-Hyndman et al. (\protect\hyperlink{ref-hyndman2002state}{2002}) provided
-this framework as an extension of Ord, Koehler, and Snyder's class of
-exponential state space models which underpinned some of the smoothing
-methods (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
-He proposed two state space models, one corresponding to an additive
-error and the other to a multiplicative error. By specifying a
-distribution for these error terms, he turned the previously
-deterministic methods into stochastic models. These two models were then
-applied to all 15 smoothing methods, producing 30 models in total.
+\subsection{The Box-Jenkins Methodology and Automatic
+Selection}\label{the-box-jenkins-methodology-and-automatic-selection}
 
-The two state space models produce identical point estimates to each
-other as well as to the exponential smoothing methods, however they will
-generate differing prediction intervals (R. Hyndman and Athanasopoulos
-\protect\hyperlink{ref-hyndman2013forecasting}{2013}). Each model
-consists of a measurement equation and a transition equation. The
-measurement equation shows the relationship between the observations and
-the unobserved components (level, trend seasonality), while the
-transition equation shows how the unobserved components change over
-time.
+Box and Jenkins proposed a set of guidelines that can be followed when
+selecting ARIMA models. This systematic procedure to designing ARIMA
+models has made them highly popular (Hibon and Makridakis
+\protect\hyperlink{ref-spyros1997}{1997}). It consists of a four-stage
+iterative process in which: 1) the process is either transformed or
+differenced to detrend and stabilise the variance of the data, 2) the
+autocorrelation and partial autocorrelation plots are used to determine
+the order of p and q, 3) the parameters of the model are estimated, 4) a
+diagnostic check is performed to ensure that the residuals are a white
+noise process. If the residuals are not white noise steps 2-4 are
+repeated until a satisfactory model is identified. On the contrary, if
+the diagnostic check shows that the residuals are random then the
+developed model is the final model used for forecasting.
 
-The introduction of the exponential state space model has allowed for
-model selection using information criteria rather than using an ad hoc
-method. It has also made the calculation of the likelihood easier and
-simulation from the underlying model possible (R. J. Hyndman et al.
-\protect\hyperlink{ref-hyndman2002state}{2002}).
+Due to the large number of forecasts made, it is useful to have an
+automatic procedure which is able to select the appropriate ARIMA model
+to fit the data. The auto.arima function in R is able to automatically
+choose the order of the parameters p, q and d. The order of differencing
+is determined by using the KPSS test (Ruppert and Matteson
+\protect\hyperlink{ref-ruppert2015}{2015}). The KPSS test checks the
+null hypothesis of stationarity and sets d to zero if the null
+hypothesis is accepted, otherwise it iteratively increases d by 1 and
+tests the null hypothesis until it is accepted (Ruppert and Matteson
+\protect\hyperlink{ref-ruppert2015}{2015}). Once the order of
+differencing has been determined, the order of p and q are chosen based
+on Akaike's Information Criterion (AIC) or the Bayesian Information
+Criterion (BIC) (Ruppert and Matteson
+\protect\hyperlink{ref-ruppert2015}{2015}).
 
-\subsection{Application and
-Performance}\label{application-and-performance}
+\subsection{Empirical Applications}\label{empirical-applications}
 
-In a paper by Smith and Agrawal
-(\protect\hyperlink{ref-smith2015comparison}{2015}), Holt-Winters
-exponential smoothing (HWES) model was applied to time series data
-relating to data on patents. The patent data was classified into three
-different groups; monthly forecasts were made and the results were
-compared to forecasts produced by ARIMA models. To avoid
-over-forecasting, the trend of the HWES was damped to produce forecasts
-with a flat trend for later observations. The results showed that all
-models forecasted the data adequately. The HWES model fitted the data
-well and performed better than the ARIMA models in the case where the
-time series for the patent group was fairly stationary. Based on the
-forecasting results, it could not be concluded that a specific time
-series model is better for forecasting patent data.
+ARIMA models have been commonly used in financial forecasting and are
+popular for observing stock prices and exchange rates due to its power
+and statistical properties (C.-S. Lin, Chiu, and Lin
+\protect\hyperlink{ref-lin2012}{2012}). They have been used as a
+benchmark to compare new forecasting techniques which have emerged over
+time. Kamruzzaman and Sarker
+(\protect\hyperlink{ref-kamruzzaman2003}{2003}) applied an ARIMA model
+to forecast the exchange rate between the Australian dollar and six
+other currencies. Its results were used to evaluate the performance of
+three Neural Network models. The Normalised Mean Square Error (NMSE) and
+Mean Absolute Error (MAE) were used to measure the forecast errors while
+the Directional Symmetry (DS), Correct Up trend (CU) and Correct Down
+trend (CD) were used to measure the accuracy of the direction of the
+forecasts. Similarly, C.-S. Lin, Chiu, and Lin
+(\protect\hyperlink{ref-lin2012}{2012}) employed an ARIMA model to
+forecast the Taiwan New Dollar exchange rate. The aim of applying the
+ARIMA model was to analyse the ability of a proposed hybrid model
+created from the Squares Support Vector Regression (LSSVR) and Empirical
+Mode Decomposition (EMD) model. C.-S. Lin, Chiu, and Lin
+(\protect\hyperlink{ref-lin2012}{2012}) used the same indicators as
+Kamruzzaman and Sarker (\protect\hyperlink{ref-kamruzzaman2003}{2003})
+to evaluate the accuracy of the direction of the forecasts. To measure
+the forecast errors produced they applied the mean absolute percentage
+error (MAPE), root‐mean‐square error (RMSE) and mean absolute difference
+(MAD). Khashei, Bijari, and Ardali
+(\protect\hyperlink{ref-khashei2012}{2012}) also investigated the
+ability to forecast exchange rates by considering the British pound and
+US dollar, over a short and long forecast horizon. They compared the
+predictive capability of their proposed hybrid model consisting of an
+ARIMA and a Probabilistic Neural Network (PNN), to the traditional ARIMA
+model, and justified the use of the ARIMA as a benchmark, by claiming
+that it is the most important linear model. The ARIMA model has also
+yielded satisfactory results when predicting exchange rates. Nwankwo
+(\protect\hyperlink{ref-nwankwo2014}{2014}) forecasted the Nairo/dollar
+rate and diagnostic testing revealed that the ARIMA(1,0,0) model was the
+best fit for the data based on the AIC.
 
-Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
-forecasted tourism demand in specific European countries over a short,
-medium and long term horizon. The ETS model was used to produce
-forecasts and was compared to six parametric and non-parametric
-techniques. The results showed that no single model is best for
-producing forecasts across all countries and time horizons. The ETS
-model together with the Neural Networks (NN) and Fractionalized ARIMA
-(ARFIMA) produced forecasts with the least amount of accuracy and are
-thus not relevant models for forecasting tourism demand in Europe. 24
-steps-ahead error forecasts produced by the ETS model were found to be
-larger than the errors produced by the benchmark model - Recurrent
-Single Spectrum Analysis (SSA-R) across all European countries.
+The ARIMA model is attractive as it is tractable and produces good
+forecasts in short time periods where more than 100 observations are
+used (Tseng et al. \protect\hyperlink{ref-tseng2001}{2001}). Although
+the ARIMA model has the advantage of ease of implementation and
+flexibility, it fails to capture the non-linearity and volatility
+present in exchange rate data.
 
-\subsection{Robustness}\label{robustness}
+\subsection{Evolution of ARIMA}\label{evolution-of-arima}
 
-Gardner (\protect\hyperlink{ref-gardner2006exponential}{2006}) explained
-the robustness of ETS models by comparing them to equivalent models.
-Simple exponential smoothing which is equivalent to an ARIMA(0,1,1)
-model is the most robust method and has performed well in forecasting
-many types of time series. By comparing simple exponential smoothing to
-lower order ARIMA models, it is seen that errors that arise due to model
-specification is not as much of a problem as ARIMA models. Furthermore,
-ARIMA models tend to produce larger MSE due to model selection errors,
-and this is worsened if the errors are not normally distributed (Gardner
-\protect\hyperlink{ref-gardner2006exponential}{2006}).
+Over time, the ARIMA model has evolved to cater for a wider variety of
+data that it can be applied to, and to compensate for some of its
+shortcomings. The most popular versions of the ARIMA model that has been
+implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA)
+and Fractional ARIMA model (FARIMA).
 
-Satchell and Timmermann found that for time series with a finite
-history, the weights assigned to each observation in the simple
-exponential smoothing method are robust, provided that the variance of
-the random walk relative to the variance of the error component is not
-too small (Gardner
-\protect\hyperlink{ref-gardner2006exponential}{2006}).
+The SARIMA model was introduced by Box and Jenkins as a generalisation
+of the ARIMA model to capture the periodic behaviour of data. SARIMA
+models extend the ARIMA class by including a term for seasonal
+differencing. Exchange rate data that has been observed have shown
+seasonality, and have been attempted to be fitted by SARIMA models
+(Etuk, Wokoma, and Moffat \protect\hyperlink{ref-etuk2013}{2013}). Etuk,
+Wokoma, and Moffat (\protect\hyperlink{ref-etuk2013}{2013}) modelled the
+Naira/CFA Franc exchange rate which exhibited monthly seasonality using
+an additive SARIMA model, to demonstrate that it can be a useful fit to
+the data. Their results showed that the SARIMA model adequately
+described the variation in the exchange rate series. This agrees with
+the results of a study by Etuk and others
+(\protect\hyperlink{ref-etuk2012}{2012}) which fitted a multiplicative
+SARIMA model to the Naira/Dollar exchange rates.
 
-Forecasts produced by simple exponential smoothing have performed well
-in modelling annual sales of different products, as well as aggregated
-economic series with low sampling frequencies (Gardner
-\protect\hyperlink{ref-gardner2006exponential}{2006}).
+The ARFIMA model is another generalisation of the ARIMA model in that
+the degree required to make the data stationary can assume any real
+value, and is no longer restricted to the integer domain. While the
+ARIMA model displays exponentially decaying autocorrelation between
+observations the further apart they are in time, the ARFIMA model has
+the ability to capture the dependence between observations that are
+widely spread apart in time (Cheung
+\protect\hyperlink{ref-cheung1993}{1993}). This makes the ARFIMA model
+more parsimonious since it can capture long memory in data as well as
+short term dynamics (Cheung \protect\hyperlink{ref-cheung1993}{1993}).
+Cheung (\protect\hyperlink{ref-cheung1993}{1993}) fitted the ARFIMA
+model to examine five exchange rates, and found that there was strong
+evidence of long memory in the exchange rate time series. Although the
+ARFIMA model fitted the data well, it failed to beat the random walk in
+out-of sample forecasts.
 
-\section{Prophet}\label{prophet}
+GARCH models were later developed due to the failure of ARIMA models to
+capture the volatility in financial markets (Anastasakis and Mort
+\protect\hyperlink{ref-anastasakis2009}{2009}). They have been widely
+used to forecast the volatility of stock returns, interest rates and
+exchange rates and are useful for portfolio allocation decisions as well
+as option pricing (Cermak \protect\hyperlink{ref-cermak2017}{2017}).
+Cermak (\protect\hyperlink{ref-cermak2017}{2017}) employed a GARCH(1,1)
+model to analyse the volatility of Bitcoin in countries in which Bitcoin
+is commonly traded, while Hsieh (1989) applied a GARCH model to
+investigate five exchange rates. Hsieh's (1989) results showed that
+although the GARCH model outperformed the random walk, some non-linear
+information still remained in the residuals. In a similar study,
+Fahimifard et al. (\protect\hyperlink{ref-fahimifard2009}{2009})
+illustrated that the GARCH model outperformed the ARIMA, while Neural
+Networks outperformed both the GARCH and ARIMA model.
 
-\subsection{Bayesian Generalized Additive
-Model}\label{bayesian-generalized-additive-model}
+\section{Other techniques used to forecast exchange
+rates}\label{other-techniques-used-to-forecast-exchange-rates}
 
-Prophet is the key work of Taylor and Letham
-(\protect\hyperlink{ref-taylor2017forecasting}{2017}) and consists of a
-decomposable model with a component for growth, seasonality, and
-holidays. These components consist of linear and non-linear functions of
-time. This is similar to a Generalized Additive Model (GAM) which is a
-regression model that consists of non-linear and linear regression
-functions applied to predictor variables (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). Prophet, like GAM,
-frames the forecasting problem as a curve fitting exercise using
-backfitting to find the regression functions.
+Over time, financial forecasting methods have moved away from linear
+models like ARIMA and GARCH, to soft computing techniques. These more
+complex techniques are non-linear and can fit complex time series more
+easily (Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}).
+Unlike regression analysis and the ARIMA model, soft computing
+techniques do not impose structural assumptions on the model apriori
+(Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}). Some of
+the most commonly used artificial intelligence methods used to forecast
+exchange rate data are Neural Networks and the Fuzzy Logistic Systems.
 
-The growth component is modelled in a similar way to modelling
-population growths which uses a logistic growth model (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). Populations
-typically grow non-linearly (although the growth component could also be
-linear) up to an upper bound known as the carrying capacity, and remains
-constant thereafter. The rate at which the population grows changes over
-time, and this is accounted for by including changepoints in the model
-where the growth rate may be changed (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). These changepoints
-can be automatically selected and have a Laplace prior distribution
-placed on them (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). The parameter of
-the prior distribution can be used to adjust the growth rate of the
-changepoints. This allows non-experts with knowledge about events that
-may affect growth to use the parameter as a knob and adjust it to
-increase or decrease the number of changepoints (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). It also allows for
-the analyst to add changepoints which the automatic selection procedure
-may have missed (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). Furthermore,
-analysts may also specify the carrying capacity and adjust it based on
-their knowledge of the total market size (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}).
+\subsection{Neural Networks}\label{neural-networks}
+
+Artificial Neural Networks are commonly used in financial forecasting
+and are more advantageous than other non-linear forecasting methods.
+They can approximate any continuous function, are data driven, and can
+adapt to non-stationary environments (Khashei and Bijari
+\protect\hyperlink{ref-khashei2011}{2011}). Neural Networks mimic the
+structure of the brain and consists of three layers: the input, hidden
+and output layer, each of which contain nodes which are representative
+of the neurons in our brain. The job of the nodes in the input layer is
+to send signals along connections to the nodes in the hidden layer,
+where the connections are representative of the synapses in our brain.
+The nodes in the hidden layer uses an activation function and assigns a
+weight to each connection according to the level of importance of
+information. The value in the hidden layer is then sent along
+connections to the output layer, where weights are again assigned to
+each connection. The error is then computed in the output layer and is
+backpropagated through the network. The weights are adjusted according
+to the error that they contribute, allowing the ANN to be trained and
+learn. This learning process is repeated until the network achieves the
+desired output.
+
+ANN's have had many successful applications in forecasting exchange
+rates. In a study done by Fahimifard et al.
+(\protect\hyperlink{ref-fahimifard2009}{2009}), the ANN was found as an
+effective way to improve the forecasts of exchange rates. Superior
+results were produced when compared to the ARIMA and GARCH model using
+the RMSE, MSE and MAD as measure of performance. Similar results were
+found by Zhang and Hu (\protect\hyperlink{ref-zhang1998}{1998}) however
+the ANN failed to outperform the random walk over long forecast
+horizons. Contrary to the findings above, Kuan and Liu
+(\protect\hyperlink{ref-kuan1995}{1995}) attempted to investigate how
+well feedforward and recurrent networks capture the non-linearity of
+exchange rate data and obtained conflicting results with neural networks
+having significantly better market timing ability or lower out-of-sample
+Mean Square Prediction Error (MSPE) compared to the ARMA models for some
+exchange rates only. Franses and Van Homelen
+(\protect\hyperlink{ref-franses1998}{1998}) suggested that the
+non-linear feature of exchange rates that are picked up by ANNs may
+actually be due to neglected GARCH effects. They investigated the
+application of ANNs to forecast exchange rates and found that the
+presence of GARCH in the data may mistakenly lead one to believe that
+the returns can be forecasted on the exchange rates themselves. Their
+results show that there is no gain in producing out-of-sample forecasts
+using ANN if the data is not truly non-linear.
+
+Although neural networks have been broadly applied in financial
+forecasting, the process of building neural networks is a complex task
+and there is no consistent method to design them compared to the
+traditional Box-Jenkins ARIMA model. Unlike ARIMA models, the
+performance of ANNs is sensitive to many modelling factors such as the
+number of input nodes and the training sample size (Zhang and Hu
+\protect\hyperlink{ref-zhang1998}{1998}).
+
+\subsection{Fuzzy Logistic Systems}\label{fuzzy-logistic-systems}
+
+Fuzzy Logistic Systems were initially developed to solve problems
+involving linguistic terms, and have been used successful in financial
+forecasting (Khashei, Bijari, and Ardali
+\protect\hyperlink{ref-khashei2009}{2009}). Unlike computers, the human
+decision-making process includes a range of possibilities spanning `yes'
+and `no'. Fuzzy logic tries to imitate this reasoning and
+decision-making process by taking a discrete decision space that
+consists of a large number of states and turning it into a continuous
+space. This allows for more finer decisions to be provided as opposed to
+discrete decisions.
+
+Santos, Costa, and Santos Coelho
+(\protect\hyperlink{ref-santos2007}{2007}) investigated how well fuzzy
+systems and neural networks perform compared to the traditional ARMA and
+GARCH model. He examined the forecasts of Brazilian exchange rate
+returns looking at different frequencies of the series and compared
+their one-step-ahead forecasts. By analysing the RMSE, U-Theil
+inequality index, percentage of corrected predicted signals (CPS), and
+the Pesaran-Timmermann (PT) predictive failure statistic, he found that
+the Fuzzy Systems and Neural Networks achieved higher returns based on
+the forecasts they produced. Similar results were found by Khashei,
+Bijari, and Ardali (\protect\hyperlink{ref-khashei2009}{2009}) when he
+analysed the predictive capabilities of Fuzzy Systems, Neural Networks,
+the traditional ARIMA model, and a Fuzzy ARIMA model. The MAE and MSE of
+the ARIMA model were higher compared to the other models.
+
+Although Fuzzy Logistic Systems has the advantage over ARIMA models that
+they can be applied to data with few observations available, it gives
+acceptable rather than accurate reasoning and are more suitable for
+problems which do not require high accuracy. Like ANN's there is no
+systematic approach to designing Fuzzy Logistic Systems and they are
+only understandable when simple.
+
+\section{Hybrid ARIMA Models}\label{hybrid-arima-models}
+
+Over time, many researchers began to think of ways in which to harness
+the advantages of tractable linear models such as the ARIMA, and of more
+complex non-linear models. By combining the ARIMA model with other
+forecasting methods, the advantages of both forecasting methods are
+leveraged while simultaneously improving their limitations. Some of the
+hybrid models which have been commonly used are the Fuzzy ARIMA and
+ANN-ARIMA model.
+
+\subsection{Fuzzy ARIMA}\label{fuzzy-arima}
+
+The ARIMA model produces very accurate forecasts over short time
+horizons however it has the limitation of requiring more than 100
+observations of historical data (Tseng et al.
+\protect\hyperlink{ref-tseng2001}{2001}). In a world that is constantly
+changing and with the rapid advancement of technology, access to large
+amounts of historical data is difficult to obtain. On the contrary, a
+fuzzy regression model requires little historical data however produces
+wide prediction intervals if extreme values are present in the data. The
+Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to
+exploit the advantages of both models while simultaneously overcoming
+their limitations. Tseng et al.
+(\protect\hyperlink{ref-tseng2001}{2001}) proposed applying a Fuzzy
+ARIMA model to forecast the exchange rate of Taiwan dollars to US
+dollars to demonstrate the model's appropriateness and power. The Fuzzy
+ARIMA not only produced forecasts that were superior to the ARIMA and
+fuzzy time series models but also provided an upper and lower bound
+which can be used by decision makers to determine the best and worst
+possible situations.
+
+\subsection{ANN-ARIMA}\label{ann-arima}
+
+Although ARIMA models are powerful, they require non-stationary data to
+be differenced and impose prior assumptions onto the distribution of the
+data (Ince and Trafalis \protect\hyperlink{ref-ince2006}{2006}). In
+contrast, machine learning techniques such as ANNs do not impose any
+assumptions onto the data generating process however, being data-driven,
+are sensitive to the number of input nodes used. The ANN-ARIMA model
+draws on the strengths of both models and overcomes these individual
+difficulties. Ince and Trafalis (\protect\hyperlink{ref-ince2006}{2006})
+created an ANN-ARIMA model by using the ARIMA to determine the number
+input nodes required by an ANN for three exchange rates. When the
+ANN-ARIMA model was compared to the pure ARIMA, the hybrid model
+outperformed the ARIMA based on the MSE. The findings of Khashei and
+Bijari (\protect\hyperlink{ref-khashei2011}{2011}) study agree with
+these results. They implemented an ARIMA model and used its residuals
+together with past observations of the data as inputs for the ANN. Their
+hybrid model had superior in-sample and out-sample forecasts compared to
+the random walk, linear AR and ANN model.
+
+Although the traditional ARIMA models produce less superior forecasts
+than its hybrid forms and other complex non-linear techniques, its
+forecasts are still satisfactory. They are simple models that are easy
+to implement and have a consistent method of model design and selection.
+ARIMA models are also more robust and efficient than complex structural
+models in relation to short-run forecasting. The fact that they have
+been used as the foundation for more advanced models and have commonly
+been used as a benchmark for comparison justifies it as a good starting
+point to compare it to Facebook's forecasting method, Prophet, that was
+recently released.
+
+\section{Forecasting with Prophet}\label{forecasting-with-prophet}
+
+The techniques that have been considered for exchange rate forecasting
+thus far, require the analyst to have vocational knowledge about time
+series. Prophet differs to traditional time series models in that it is
+flexible and can be customised by a large number of non-experts who have
+little knowledge about time series, however have domain knowledge about
+the data generating process. Prophet allows for a large number of
+forecasts to be produced across a variety of problems and consists of a
+robust evaluation system that allows for a large number of forecasts be
+evaluated and compared. This is Facebook's definition of forecasting at
+scale.
+
+\subsection{Construction of Prophet}\label{construction-of-prophet}
+
+Prophet consists of a decomposable model of the form:
+
+\[ 
+    y(t) = g(t) + s(t) + h(t) + e_t
+\] where the components of the model represent the growth, seasonality
+and holiday respectively, and \(e_t\) is white noise.
+
+These components consist of linear and non-linear functions of time.
+This differs to ARIMA models in which future values of the process are
+linear functions of previous observations and lagged residuals. Prophet
+is more like a Generalized Additive Model (GAM) which is a regression
+model that consists of non-linear and linear regression functions
+applied to predictor variables (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}). Prophet, like the GAM, frames
+the forecasting problem as a curve fitting exercise, using backfitting
+to find the regression functions. The GAM is fitted quickly, allowing
+the analyst to interactively change the model parameters (Taylor and
+Letham \protect\hyperlink{ref-taylor2017}{2017}).
+
+The growth component is modelled in a similar way to population growths
+which use a logistic growth model (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}). Populations typically grow
+non-linearly (although the growth component could also be linear) up to
+an upper bound known as the carrying capacity, and remains constant
+thereafter. The rate at which the population grows changes over time,
+and this is accounted for by including changepoints in the model where
+the growth rate can be automatically selected and may be adjusted
+(Taylor and Letham \protect\hyperlink{ref-taylor2017}{2017}). This
+allows non-experts with knowledge about events that may affect growth to
+use the parameter as a knob and adjust it to increase or decrease the
+number of changepoints (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}). It also allows for the
+analyst to add changepoints which the automatic selection procedure may
+have missed (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}). Furthermore, analysts may
+also specify the carrying capacity and adjust it based on their
+knowledge of the total market size (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}).
 
@@
 added to it. This allows for multiple seasonality components with
-different periods to be added to the model. The seasonality component is
-modelled by a Fourier series, with the parameters of the Fourier series
-having a Normal prior distribution (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). The variance of
-the parameters can be adjusted by analysts to smooth the model and
-change how much of historical seasonality is projected to the future
-(Taylor and Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}).
+different periods to be added to the model. The variance of the
+parameters of the seasonality component's distribution can be adjusted
+by analysts to smooth the model and change how much of historical
+seasonality is projected to the future (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}).
 
 The name, date, and country of past and future holidays and events may
-be inputted by the analyst into a list (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}). By specifying the
-country in which the events take place or the holidays occur, separate
-lists can be populated for global events/holidays and country-specific
+also be inputted by the analyst into a list (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}). By specifying the country in
+which the events take place or the holidays occur, separate lists can be
+populated for global events/holidays and country-specific
 events/holidays. The union of the two lists are then used for
-forecasting. Like seasonality, a Normal prior distribution is placed on
-the parameters of the holidays, and the variance of the parameters can
-be adjusted by analysts to smooth the model (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}).
+forecasting. Like seasonality, the variance of the parameters of the
+holiday component's distribution, can be adjusted by analysts to smooth
+the model (Taylor and Letham \protect\hyperlink{ref-taylor2017}{2017}).
 
-\subsection{Application and
-Performance}\label{application-and-performance-1}
+\subsection{Empirical Applications}\label{empirical-applications-1}
 
-Taylor and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017})
-forecasted the number of events on Facebook using Prophet. The time
-series was impacted by holidays, had strong multi-period seasonality,
-and a piecewise trend. The forecasts produced by Prophet were compared
-to forecasts produced by common automatic forecasting techniques such as
-ETS, ARIMA, and the seasonal naïve model, as well as to simple models
-such as the naïve model. While the ETS and seasonal naïve model were
-quite robust, the ARIMA forecasts were fragile. No model besides Prophet
-accounted for the dips around holidays and the upward trend of the time
-series towards later observations. Hence, Prophet had lower forecasting
-errors compared to the automatic forecasting methods.
+Prophet's Bayesian approach to forecasting allows the analyst to
+incorporate their expert knowledge in the model building process and has
+produced significantly improved forecasts compared to the ARIMA model.
+Taylor and Letham (\protect\hyperlink{ref-taylor2017}{2017}) forecasted
+the number of events on Facebook using Prophet. The time series was
+impacted by holidays, had strong multi-period seasonality, and a
+piecewise trend. The forecasts produced by Prophet were compared to
+common automatic forecasting techniques such as exponential smoothing,
+ARIMA, and the seasonal naïve model, as well as to simple models such as
+the naïve model. While exponential smoothing and the seasonal naïve
+model were quite robust, the ARIMA forecasts were fragile. No model
+besides Prophet accounted for the dips around holidays and the upward
+trend of the time series towards later observations.
 
-\subsection{Robustness}\label{robustness-1}
-
-Prophet is a flexible model and has intuitive parameters that can be
-easily interpreted by human beings. The GAM is fitted quickly, allowing
-the analyst to interactively change the model parameters (Taylor and
-Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}). Unlike
-ARIMA models, Prophet can produce forecasts over different scales and
-allows for missing values in the time series without the need for
-interpolation (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}).
+\subsection{Semi-Automatic Selection}\label{semi-automatic-selection}
 
@@
 adjust the model or choose a more appropriate model (Taylor and Letham
-\protect\hyperlink{ref-taylor2017forecasting}{2017}).
+\protect\hyperlink{ref-taylor2017}{2017}). Unlike ARIMA's
+fully-automated evaluation system, Prophet provides interactive feedback
+and keeps the analyst in the loop.
 
-\section{Evaluation of Forecasts}\label{evaluation-of-forecasts}
+Prophet's ability to forecast at scale enables it to model a wide
+variety of data and may be able to capture the non-linearity present in
+exchange rates. In contrast, the ARIMA model fails to capture the
+non-linearity inherent in exchange rate data and only produces good
+forecasts over short time horizons. Prophet is also a flexible model
+that has intuitive parameters which can be easily interpreted and
+modified by human beings (Taylor and Letham
+\protect\hyperlink{ref-taylor2017}{2017}). This may be useful to
+non-experts who have domain knowledge about factors that may affect the
+movement of exchange rate prices. Furthermore, unlike the ARIMA model,
+Prophet can produce forecasts over irregular time intervals and allows
+for missing values in the time series without the need for interpolation
+(Taylor and Letham \protect\hyperlink{ref-taylor2017}{2017}). If Prophet
+produces forecasts that are as good as the ARIMA model when forecasting
+Bitcoin/ZAR, it can be compared to more complex forecasting methods that
+are less tractable such as the hybrid models and machine learning
+techniques seen earlier.
 
-\subsection{Common Methods}\label{common-methods}
+\section{Mincer-Zarnowitz Approach to Forecast
+Evaluation}\label{mincer-zarnowitz-approach-to-forecast-evaluation}
 
-Model selection has historically been very subjective, with no
-techniques that offered mathematical rigour for choosing models (Smith
-and Agrawal \protect\hyperlink{ref-smith2015comparison}{2015}). Progress
-has been made over time and many mathematical techniques have been
-introduced for model selection. Use of information criteria is one such
-method, with Akaike's information criterion (AIC) and Bayes information
-criterion (BIC) being the most common (Smith and Agrawal
-\protect\hyperlink{ref-smith2015comparison}{2015}). These information
-criteria minimize forecasting errors while penalizing it for
-overfitting. Information criteria should be used to compare models of
-similar structure only. To compare models of different structure, the
-most commonly used measures of forecast errors are the Root Mean Squared
-error (RMSE), Mean Absolute Error (MAE), Mean Absolute Percentage Error
-(MAPE), and Mean Absolute Scaled Error (MASE). A lower value of these
-statistics indicates that the forecasts are better, with a MASE less
-than one indicating that the forecasts are better than those produced by
-a naïve model (Smith and Agrawal
-\protect\hyperlink{ref-smith2015comparison}{2015}).
+The Mincer-Zarnowitz approach to evaluating forecast accuracy is
+commonly used and can be useful when comparing the forecasts produced by
+Prophet and the ARIMA model. Mincer and Zarnowitz
+(\protect\hyperlink{ref-mincer1969}{1969}) proposed an absolute and
+relative measure to evaluate forecast accuracy. Absolute measures
+consider the distance between actual and predicted values. To analyse
+the absolute errors produced by the forecasts, the observed values are
+regressed against the predicted values. The intercept of the regression
+equation represents the mean distance between the observed and predicted
+values while the slope represents the correlation between the residual
+errors and the predicted values. A zero intercept implies that the
+forecasts are unbiased and do not overestimate or underestimate the
+data, while a unity slope implies that the forecasts are efficient and
+uncorrelated with the residual errors. A joint hypothesis test is
+performed to check this efficiency and bias of the forecasts, and models
+which produce the best results are selected.
 
-Smith and Agrawal (\protect\hyperlink{ref-smith2015comparison}{2015})
-have used these methods in their paper when evaluating patent forecasts
-and making comparisons between the HWES and ARIMA models. In a paper by
-Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
-they chose their ETS model based on the selection that R automatically
-made for the time series. They compared models of different structure
-using the RMSE and a modified Diebold-Mariano test. R. J. Hyndman et al.
-(\protect\hyperlink{ref-hyndman2002state}{2002}) suggests that ETS
-models should be selected using the AIC rather than the MSE or MAPE. The
-AIC provides a way of choosing between models with an additive error and
-models with a multiplicative error since it is based on the likelihood
-function. The MSE and MAPE are not able to select between the two error
-types because both models produce the same point estimates. According to
-De Gooijer and Hyndman (\protect\hyperlink{ref-de200625}{2006}), the
-best approach of selecting models with different structure is the
-Diebold-Mariano test. The MSE is scale dependent and should not be used
-to make comparisons while the MAPE encounters difficulties when the time
-series values are close to or equal to zero.
+Absolute forecast measures cannot be used to make comparisons between
+forecasts with different scales or economic variables. Furthermore, the
+size of forecasting errors is not as significant as the consequences of
+forecasting errors and how they impact the decision making process.
+Hence, rather than using absolute measures of forecast accuracy, Mincer
+and Zarnowitz (\protect\hyperlink{ref-mincer1969}{1969}) suggested the
+use of relative measures. Relative accuracy analysis allows for
+meaningful comparisons of different forecasting methods to be made. It
+uses an index which considers the ratio of the MSE of the forecast to
+the MSE of a benchmark forecasting method. A useful benchmark that may
+be used is an extrapolation of the data history, as it is a cost
+effective and accessible method, however any method which is relevant
+for comparison may be used. This ratio is known as the Relative Mean
+Square Error (RM) for forecast evaluation. The numerator can be viewed
+as a return which is inversely proportional to the MSE error of the
+forecasts, while the denominator can be viewed as the cost of
+forecasting which is inversely proportional to the MSE of the benchmark
+(Mincer and Zarnowitz \protect\hyperlink{ref-mincer1969}{1969}). Hence
+the ratio is representative of a rate of return index and ranks the
+performance of forecasts as such. Models that result in forecasts with a
+RM that is less than one are said to produce superior forecasts to the
+benchmark model being considered.
 
-\subsection{Mincer-Zarnowitz Approach}\label{mincer-zarnowitz-approach}
+\section{Conclusion}\label{conclusion}
 
-Mincer and Zarnowitz
-(\protect\hyperlink{ref-mincer1969evaluation}{1969}) proposed two
-methods of measuring forecast accuracy - an absolute and a relative
-measure. Absolute measures of accuracy measure the distance between
-actual and predicted values. If the mean distance is equal to zero, the
-forecast is said to be unbiased (Mincer and Zarnowitz
-\protect\hyperlink{ref-mincer1969evaluation}{1969}). To analyse absolute
-forecast errors, the observed values are regressed against the predicted
-values. A joint hypothesis test is performed to check if the intercept
-is equal to zero and the slope is equal to one. If the intercept is
-equal to zero, the forecasts are unbiased and do not overestimate or
-underestimate the data. If the slope is equal to one, the residual
-errors are uncorrelated with the predictions and the forecast is said to
-be efficient (Mincer and Zarnowitz
-\protect\hyperlink{ref-mincer1969evaluation}{1969}). The MSE is a
-measure that can be decomposed into the residual variance and two
-components, one reflecting the bias and the other reflecting the
-inefficiency of the forecast. If the forecasts are unbiased and
-efficient, the MSE will reduce to the residual variance (Mincer and
-Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}). Absolute
-forecast measures cannot be used to make comparisons between forecasts
-with different scales or economic variables. Thiel states that the
-consequences of forecasting errors and how they impact decisions is of
-greater importance than the size of the forecasting errors (Mincer and
-Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}).
-
-Rather than using absolute measures of forecast accuracy, Mincer and
-Zarnowitz (\protect\hyperlink{ref-mincer1969evaluation}{1969}) suggested
-the use of relative measures of forecast accuracy. Relative accuracy
-analysis allows for meaningful comparisons of different forecasting
-methods to be made. They proposed an index which considers the ratio of
-the MSE of the forecast to the MSE of a benchmark forecasting method.
-They used an extrapolation of the data history as a benchmark, as it is
-a cost effective and accessible method, however the benchmark may be any
-method which is relevant for comparison. This ratio is known as the
-Relative Mean Square Error (RM).
-
-The numerator represents a return which is inversely proportional to the
-MSE error of the forecasts, while the denominator represents the cost of
-forecasting which is inversely proportional to the MSE of the benchmark
-(Mincer and Zarnowitz
-\protect\hyperlink{ref-mincer1969evaluation}{1969}). Hence the ratio is
-representative of a rate of return index and ranks the performance of
-forecasts as such. If the forecasts are better than the benchmark, the
-RM will be less than one.
+ARIMA models have been highly popular since its introduction by Box and
+Jenkins and are widely used for financial forecasting. They are easy to
+implement, have a systematic procedure for model design, and produce
+good forecasts over short time horizons. Although robust, they have the
+limitation of requiring large amounts of historical observations and
+they fail to capture the non-linearity inherent in exchange rate data.
+With the evolution of computational power and the advancement of
+statistics, more complex techniques such as ANN's, Fuzzy Logistic
+Systems, and Hybrid ARIMA models have been proposed and compared to the
+traditional ARIMA model. Although these models produce superior results
+to the ARIMA model, they are less tractable and require the analyst to
+understand the complex statistical intricacies involved. Prophet's
+Bayesian approach to forecasting is flexible and allows for non-experts
+to incorporate their domain knowledge in the model building process. If
+Prophet produces forecasts that are as good as the ARIMA model when
+forecasting Bitcoin/ZAR, it can be compared to these more complex
+forecasting methods that are less tractable and scalable.
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-billah2006exponential}{}
-Billah, Baki, Maxwell L King, Ralph D Snyder, and Anne B Koehler. 2006.
-``Exponential Smoothing Model Selection for Forecasting.''
-\emph{International Journal of Forecasting} 22 (2). Elsevier: 239--47.
+\hypertarget{ref-abu1996}{}
+Abu-Mostafa, Yaser S, and Amir F Atiya. 1996. ``Introduction to
+Financial Forecasting.'' \emph{Applied Intelligence} 6 (3). Springer:
+205--13.
 
-\hypertarget{ref-de200625}{}
-De Gooijer, Jan G, and Rob J Hyndman. 2006. ``25 Years of Time Series
-Forecasting.'' \emph{International Journal of Forecasting} 22 (3).
-Elsevier: 443--73.
+\hypertarget{ref-anastasakis2009}{}
+Anastasakis, Leonidas, and Neil Mort. 2009. ``Exchange Rate Forecasting
+Using a Combined Parametric and Nonparametric Self-Organising Modelling
+Approach.'' \emph{Expert Systems with Applications} 36 (10). Elsevier:
+12001--11.
 
-\hypertarget{ref-gardner2006exponential}{}
-Gardner, Everette S. 2006. ``Exponential Smoothing: The State of the
-Art---Part Ii.'' \emph{International Journal of Forecasting} 22 (4).
-Elsevier: 637--66.
+\hypertarget{ref-castillo2002}{}
+Castillo, Oscar, and Patricia Melin. 2002. ``Hybrid Intelligent Systems
+for Time Series Prediction Using Neural Networks, Fuzzy Logic, and
+Fractal Theory.'' \emph{IEEE Transactions on Neural Networks} 13 (6).
+IEEE: 1395--1408.
 
-\hypertarget{ref-granvik2010forecasting}{}
-Granvik, Ashley. 2010. ``Forecasting Exchange Rates.'' Arcada-Nylands
-svenska yrkeshögskola.
+\hypertarget{ref-cermak2017}{}
+Cermak, Vavrinec. 2017. ``Can Bitcoin Become a Viable Alternative to
+Fiat Currencies? An Empirical Analysis of Bitcoin's Volatility Based on
+a Garch Model.''
 
-\hypertarget{ref-hassani2015forecasting}{}
-Hassani, Hossein, Emmanuel Sirimal Silva, Nikolaos Antonakakis, George
-Filis, Rangan Gupta, and others. 2015. ``Forecasting Accuracy Evaluation
-of Tourist Arrivals: Evidence from Parametric and Non-Parametric
-Techniques.''
+\hypertarget{ref-cheung1993}{}
+Cheung, Yin-Wong. 1993. ``Long Memory in Foreign-Exchange Rates.''
+\emph{Journal of Business \& Economic Statistics} 11 (1). Taylor \&
+Francis: 93--101.
 
-\hypertarget{ref-hyndman2013forecasting}{}
-Hyndman, RJ, and G Athanasopoulos. 2013. ``Forecasting: Principles and
-Practice, Otexts, Australia.'' \emph{Available via: Www. OTexts.
-Com/Fpp}.
+\hypertarget{ref-etuk2012}{}
+Etuk, Ette Harrison, and others. 2012. ``A Seasonal Arima Model for
+Daily Nigerian Naira-Us Dollar Exchange Rates.'' \emph{Asian Journal of
+Empirical Research} 2 (6). Asian Economic; Social Society: 219--27.
 
-\hypertarget{ref-hyndman2007automatic}{}
-Hyndman, Rob J, Yeasmin Khandakar, and others. 2007. ``Automatic Time
-Series for Forecasting: The Forecast Package for R.'' Monash University,
-Department of Econometrics; Business Statistics.
+\hypertarget{ref-etuk2013}{}
+Etuk, Ette Harrison, Dagogo SA Wokoma, and Imoh Udo Moffat. 2013.
+``Additive Sarima Modelling of Monthly Nigerian Naira-Cfa Franc Exchange
+Rates.'' \emph{European Journal of Statistics and Probability} 1 (1):
+1--12.
 
-\hypertarget{ref-hyndman2002state}{}
-Hyndman, Rob J, Anne B Koehler, Ralph D Snyder, and Simone Grose. 2002.
-``A State Space Framework for Automatic Forecasting Using Exponential
-Smoothing Methods.'' \emph{International Journal of Forecasting} 18 (3).
-Elsevier: 439--54.
+\hypertarget{ref-fahimifard2009}{}
+Fahimifard, SM, Masuod Homayounifar, M Sabouhi, and AR Moghaddamnia.
+2009. ``Comparison of Anfis, Ann, Garch and Arima Techniques to Exchange
+Rate Forecasting.'' \emph{Journal of Applied Sciences} 9 (20): 3641--51.
 
-\hypertarget{ref-mincer1969evaluation}{}
+\hypertarget{ref-franses1998}{}
+Franses, Philip Hans, and Paul Van Homelen. 1998. ``On Forecasting
+Exchange Rates Using Neural Networks.'' \emph{Applied Financial
+Economics} 8 (6). Taylor \& Francis: 589--96.
+
+\hypertarget{ref-spyros1997}{}
+Hibon, Michael, and Spyros Makridakis. 1997. ``ARMA Models and the
+Box--Jenkins Methodology.'' John Wiley \& Sons, Ltd.
+
+\hypertarget{ref-ince2006}{}
+Ince, Huseyin, and Theodore B Trafalis. 2006. ``A Hybrid Model for
+Exchange Rate Prediction.'' \emph{Decision Support Systems} 42 (2).
+Elsevier: 1054--62.
+
+\hypertarget{ref-kamruzzaman2003}{}
+Kamruzzaman, Joarder, and Ruhul A Sarker. 2003. ``Forecasting of
+Currency Exchange Rates Using Ann: A Case Study.'' In \emph{Neural
+Networks and Signal Processing, 2003. Proceedings of the 2003
+International Conference on}, 1:793--97. IEEE.
+
+\hypertarget{ref-khashei2011}{}
+Khashei, Mehdi, and Hehdi Bijari. 2011. ``Exchange Rate Forecasting
+Better with Hybrid Artificial Neural Networks Models.'' \emph{Journal of
+Mathematical and Computational Science} 1 (1). Science \& Knowledge
+Publishing Corporation Limited (SCIK): 103.
+
+\hypertarget{ref-khashei2009}{}
+Khashei, Mehdi, Mehdi Bijari, and Gholam Ali Raissi Ardali. 2009.
+``Improvement of Auto-Regressive Integrated Moving Average Models Using
+Fuzzy Logic and Artificial Neural Networks (Anns).''
+\emph{Neurocomputing} 72 (4). Elsevier: 956--67.
+
+\hypertarget{ref-khashei2012}{}
+---------. 2012. ``Hybridization of Autoregressive Integrated Moving
+Average (Arima) with Probabilistic Neural Networks (Pnns).''
+\emph{Computers \& Industrial Engineering} 63 (1). Elsevier: 37--45.
+
+\hypertarget{ref-kuan1995}{}
+Kuan, Chung-Ming, and Tung Liu. 1995. ``Forecasting Exchange Rates Using
+Feedforward and Recurrent Neural Networks.'' \emph{Journal of Applied
+Econometrics} 10 (4). Wiley Online Library: 347--64.
+
+\hypertarget{ref-lin2012}{}
+Lin, Chiun-Sin, Sheng-Hsiung Chiu, and Tzu-Yu Lin. 2012. ``Empirical
+Mode Decomposition--based Least Squares Support Vector Regression for
+Foreign Exchange Rate Forecasting.'' \emph{Economic Modelling} 29 (6).
+Elsevier: 2583--90.
+
+\hypertarget{ref-mincer1969}{}
 Mincer, Jacob A, and Victor Zarnowitz. 1969. ``The Evaluation of
@@
 
-\hypertarget{ref-smith2015comparison}{}
-Smith, Mick, and Rajeev Agrawal. 2015. ``A Comparison of Time Series
-Model Forecasting Methods on Patent Groups.'' In \emph{MAICS}, 167--73.
+\hypertarget{ref-nwankwo2014}{}
+Nwankwo, Steve C. 2014. ``Autoregressive Integrated Moving Average
+(Arima) Model for Exchange Rate (Naira to Dollar).'' \emph{Academic
+Journal of Interdisciplinary Studies} 3 (4): 429.
 
-\hypertarget{ref-taylor2017forecasting}{}
+\hypertarget{ref-ruppert2015}{}
+Ruppert, David, and David S Matteson. 2015. \emph{Statistics and Data
+Analysis for Financial Engineering: With R Examples}. Springer.
+
+\hypertarget{ref-santos2007}{}
+Santos, André Alves Portela, Newton Carneiro Affonso da Costa, and
+Leandro dos Santos Coelho. 2007. ``Computational Intelligence Approaches
+and Linear Models in Case Studies of Forecasting Exchange Rates.''
+\emph{Expert Systems with Applications} 33 (4). Elsevier: 816--23.
+
+\hypertarget{ref-taylor2017}{}
 Taylor, Sean J, and Benjamin Letham. 2017. ``Forecasting at Scale.''
+
+\hypertarget{ref-tseng2001}{}
+Tseng, Fang-Mei, Gwo-Hshiung Tzeng, Hsiao-Cheng Yu, and Benjamin JC
+Yuan. 2001. ``Fuzzy Arima Model for Forecasting the Foreign Exchange
+Market.'' \emph{Fuzzy Sets and Systems} 118 (1). Elsevier: 9--19.
+
+\hypertarget{ref-zhang1998}{}
+Zhang, Gioqinang, and Michael Y Hu. 1998. ``Neural Network Forecasting
+of the British Pound/Us Dollar Exchange Rate.'' \emph{Omega} 26 (4).
+Elsevier: 495--506.
 

### 2017-08-27 Draft/Draft.Rmd
updated code

--- 2017-06-13 Literature Review/A Brief Literature Review.Rmd
@@
 ---
-title: "Literature Review"
+title: "Draft"
 
@@
 
-The ability to forecast exchange rates is of great importance to policy makers, businesses and speculators.
-Policy makers are interested in forecasting exchange rates to assess how its movement will impact inflation, and how monetary policy will need to be adjusted in the future to keep inflation within the inflation target [@fahimifard2009].
-Businesses need to assess how future movements in exchange rates will affect the cost of importing goods and the demand for goods exported, while speculators aim to identify opportunities in price movements which can be exploited to make a profit.
+The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has attracted attention in financial market forecasting (Kashei and Bijari 2007).
+The ARIMA model is known for its attractive quality of being a rich class of processes, making it possible to find a suitable model to adequately fit the data [@fahimifard2009].
+Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that these markets are not efficient and believe that future prices depend on current and past events [@abu1996].
+This has led to the application of the ARIMA model to exchange rate problems, where it has since gained popularity due to its ease of implementation and tractability.
+Although the ARIMA model yields good forecasts over short forecast horizons and is easy to use, the linearity of the ARIMA model fails to adequately capture the non-linearity inherent in exchange rate data [@zhang1998].
 
-Producing accurate forecasts of exchange rates has proven to be a difficult task due to the noise of the time series [@abu1996].
-Some economists argue that the exchange rate market is very efficient and future movements in price are unpredictable [@abu1996].
-This belief in the efficient market hypothesis has led many researchers to model exchange rates using a Random Walk (RW) model [@abu1996].
-Although the RW model has produced superior forecasts, some researchers believe that future prices depend on present and past events, and other techniques such as the autoregressive integrated moving average (ARIMA) model has been applied to forecast exchange rates [@abu1996].
+With the evolution of computational power, non-linear, soft computing techniques have been proposed as a solution.
+The ARIMA model has been used as the standard benchmark model against which these more complex methods have been compared.
+Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task.
+The ARIMA model is more tractable, less computationally expensive, has been used as the building blocks for more advanced models and has provided the inspiration for hybrid versions of the model which have been used in exchange rate forecasting.
 
-The ARIMA model is one of the most commonly used methods in financial forecasting [@fahimifard2009].
-It is known for its attractive quality of being a rich class of processes, making it possible to find a model that adequately fits the data [@fahimifard2009].
-Some research has shown that the nature of exchange rate data is non-linear, and this non-linearity may not be captured adequately by the ARIMA model [@zhang1998].
-Linear parametric forecasting techniques such as the generalised autoregressive conditional heteroskedasticity (GARCH) model has been proposed as a solution to this problem [@zhang1998].
-Although the GARCH model takes care of the non-linearity in the data, the pre-specification of the model may prevent it from capturing all the different non-linearities which may be present in the data.
+In December 2016, Facebook open-sourced their forecasting model Prophet.
+Prophet differs to traditional time series models like ARIMA, in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process.
+This can then be easily incorporated into the model through intuitively adjustable parameters [@taylor2017].
+If Prophet produces forecasts that are as good as the ARIMA model, it can be compared to more complex forecasting methods that are less tractable and flexible.
 
-With the evolution of computational power, other non-linear non-parametric techniques such as Artificial Neural Networks (ANN), Support Vector Regression (SVR), Fuzzy Logistic Systems and hybrid versions of the ARIMA model have been proposed.
-The ARIMA model has been used as the standard benchmark model against which these other methods have been compared.
-Although they have produced significantly better results than the ARIMA model, building these models is a complex task.
-The ARIMA model is more tractable, less computationally expensive, and has been used as the building blocks for more advanced models such as the autoregressive fractionally integrated (ARFIMA), Fuzzy ARIMA, and seasonal ARIMA (SARIMA) model which have been used for exchange rate forecasting.
+This paper broadly aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of Bitcoin/ZAR, using a Mincer-Zarnowitz approach of measuring forecast accuracy.
+More specifically, this paper aims to:
 
-Facebook recently open sourced their forecasting model Prophet.
-Prophet is said to be preferable over the ARIMA model as it has the ability to forecast at scale, is easily interpretable and has parameters that can be intuitively adjusted [@taylor2017].
-This paper aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of Bitcoin/ZAR, using a Mincer-Zarnowitz approach of measuring forecast accuracy.
-Section 2 will consider the origins of the ARIMA model and how it has performed and evolved over time by exploring empirical studies where it has been used for forecasting exchange rates.
-Section 3 will also consider other techniques that have been used to forecast exchange rates and how they have been compared to the ARIMA model.
-Section 4 will look at hybrid models that were inspired by the ARIMA model, while section 5 looks at forecasting with Prophet.
-Finally, section 6 will look at the Mincer-Zarnowitz approach to evaluating forecasts.
+1) Compare the out-of-sample forecasts over varying forecast horizons.
 
-# Forecasting Exchange Rates with the ARIMA Model
+2) Compare the forecasts produced by ARIMA's and Prophet's automatic selection procedures.
 
-## Birth of ARIMA
+3) Compare the out-of-sample forecast performance with missing values and outliers present in the data.
+
+4) Investigate the efficient market hypothesis (EMH) and the trading profit to be made with Prophet.
+
+Section \ref{Background} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting, as well as the theory behind the ARIMA and Prophet model.
+This will be followed by a brief explanation and motivation for the dataset used in this paper in Section \ref{Data}.
+Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring forecast accuracy.
+The results from applying the ARIMA and Prophet model to the dataset is presented in Section \ref{Results}.
+Finally, Section \ref{Discussion and Conclusions} discusses the results in a statistical and financially meaningful context.
+
+# Background \label{Background}
+
+## Forecasting Exchange Rates with the ARIMA model
 ARIMA models became highly popular since its introduction by Box and Jenkins, when it was shown that they could outperform complex econometric models in a variety of situations [@spyros1997].
@@
 $$
-where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.
+where c is a constant, $\theta$ and $\phi$ are parameters set such that ... and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.
 
@@
 $$
-where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.
+where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$ (otext.org/fpp/8/5)...need to get this reference hmmm
 
-## The Box-Jenkins Methodology and Automatic Selection
-Box and Jenkins proposed a set of guidelines that can be followed when selecting ARIMA models.
-This systematic procedure to designing ARIMA models has made them highly popular [@spyros1997].
-It consists of a four-stage iterative process in which: 1) the process is either transformed or differenced to detrend and stabilise the variance of the data, 2) the autocorrelation and partial autocorrelation plots are used to determine the order of p and q, 3) the parameters of the model are estimated, 4) a diagnostic check is performed to ensure that the residuals are a white noise process.
-If the residuals are not white noise steps 2-4 are repeated until a satisfactory model is identified.
-On the contrary, if the diagnostic check shows that the residuals are random then the developed model is the final model used for forecasting.
+ARIMA models have been commonly used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012].
+They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time.
+@kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies, and used the forecast errors as well as the accuracy of the direction of the forecasts to evaluate the performance of three Neural Network models.
+In a similar study, @khashei2012 compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN), to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark by claiming that it is the most important linear model.
+Their comparison was made by investigating the forecasts between the British pound and US dollar over varying forecast horizons.
+Not only has the ARIMA model been useful as a benchmark, but it has also yielded satisfactory results when predicting exchange rates.
+When @nwankwo2014 forecasted the rate between the Nairo and dollar, diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on Akaike's Information Criterion (AIC).
 
-Due to the large number of forecasts made, it is useful to have an automatic procedure which is able to select the appropriate ARIMA model to fit the data.
-The auto.arima function in R is able to automatically choose the order of the parameters p, q and d.
-The order of differencing is determined by using the KPSS test [@ruppert2015].
-The KPSS test checks the null hypothesis of stationarity and sets d to zero if the null hypothesis is accepted, otherwise it iteratively increases d by 1 and tests the null hypothesis until it is accepted [@ruppert2015].
-Once the order of differencing has been determined, the order of p and q are chosen based on Akaike’s Information Criterion (AIC) or the Bayesian Information Criterion (BIC) [@ruppert2015].
-
-## Empirical Applications
-ARIMA models have been commonly used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012].
-They have been used as a benchmark to compare new forecasting techniques which have emerged over time.
-@kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies.
-Its results were used to evaluate the performance of three Neural Network models.
-The Normalised Mean Square Error (NMSE) and Mean Absolute Error (MAE) were used to measure the forecast errors while the Directional Symmetry (DS), Correct Up trend (CU) and Correct Down trend (CD) were used to measure the accuracy of the direction of the forecasts.
-Similarly, @lin2012 employed an ARIMA model to forecast the Taiwan New Dollar exchange rate.
-The aim of applying the ARIMA model was to analyse the ability of a proposed hybrid model created from the Squares Support Vector Regression (LSSVR) and Empirical Mode Decomposition (EMD) model.
-@lin2012 used the same indicators as @kamruzzaman2003 to evaluate the accuracy of the direction of the forecasts.
-To measure the forecast errors produced they applied the mean absolute percentage error (MAPE), root‐mean‐square error (RMSE) and mean absolute difference (MAD).
-@khashei2012 also investigated the ability to forecast exchange rates by considering the British pound and US dollar, over a short and long forecast horizon.
-They compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN), to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark, by claiming that it is the most important linear model.
-The ARIMA model has also yielded satisfactory results when predicting exchange rates.
-@nwankwo2014 forecasted the Nairo/dollar rate and diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on the AIC.
-
-The ARIMA model is attractive as it is tractable and produces good forecasts in short time periods where more than 100 observations are used [@tseng2001]. 
+The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001].
 Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data.
-
-## Evolution of ARIMA
-Over time, the ARIMA model has evolved to cater for a wider variety of data that it can be applied to, and to compensate for some of its shortcomings.
+Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings.
 The most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).
 
-The SARIMA model was introduced by Box and Jenkins as a generalisation of the ARIMA model to capture the periodic behaviour of data.
-SARIMA models extend the ARIMA class by including a term for seasonal differencing.
-Exchange rate data that has been observed have shown seasonality, and have been attempted to be fitted by SARIMA models [@etuk2013].
-@etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit to the data.
+The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing.
+@etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality.
 Their results showed that the SARIMA model adequately described the variation in the exchange rate series.
-This agrees with the results of a study by @etuk2012 which fitted a multiplicative SARIMA model to the Naira/Dollar exchange rates.
+The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain.
+The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993].
+This makes the model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993].
+@cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.
 
-The ARFIMA model is another generalisation of the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain.
-While the ARIMA model displays exponentially decaying autocorrelation between observations the further apart they are in time, the ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993].
-This makes the ARFIMA model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993].
-@cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.
-Although the ARFIMA model fitted the data well, it failed to beat the random walk in out-of sample forecasts.
+Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed in an attempt to capture the volatility in financial markets [@anastasakis2009].
+Hsieh (1989) applied a GARCH model to investigate five exchange rates and his results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.
 
-GARCH models were later developed due to the failure of ARIMA models to capture the volatility in financial markets [@anastasakis2009].
-They have been widely used to forecast the volatility of stock returns, interest rates and exchange rates and are useful for portfolio allocation decisions as well as option pricing [@cermak2017].
-@cermak2017 employed a GARCH(1,1) model to analyse the volatility of Bitcoin in countries in which Bitcoin is commonly traded, while Hsieh (1989) applied a GARCH model to investigate five exchange rates.
-Hsieh’s (1989) results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.
-In a similar study, @fahimifard2009 illustrated that the GARCH model outperformed the ARIMA, while Neural Networks outperformed both the GARCH and ARIMA model.
-
-# Other techniques used to forecast exchange rates
+## Other techniques used to forecast exchange rates
 Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques.
 These more complex techniques are non-linear and can fit complex time series more easily [@castillo2002].
-Unlike regression analysis and the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002].
-Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and the Fuzzy Logistic Systems.
+Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002].
+Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.
 
-## Neural Networks
-Artificial Neural Networks are commonly used in financial forecasting and are more advantageous than other non-linear forecasting methods.
-They can approximate any continuous function, are data driven, and can adapt to non-stationary environments [@khashei2011].
-Neural Networks mimic the structure of the brain and consists of three layers: the input, hidden and output layer, each of which contain nodes which are representative of the neurons in our brain.
-The job of the nodes in the input layer is to send signals along connections to the nodes in the hidden layer, where the connections are representative of the synapses in our brain.
-The nodes in the hidden layer uses an activation function and assigns a weight to each connection according to the level of importance of information.
-The value in the hidden layer is then sent along connections to the output layer, where weights are again assigned to each connection.
-The error is then computed in the output layer and is backpropagated through the network.
-The weights are adjusted according to the error that they contribute, allowing the ANN to be trained and learn.
-This learning process is repeated until the network achieves the desired output.
-
-ANN’s have had many successful applications in forecasting exchange rates.
+Artificial Neural Networks (ANNs) have had many successful applications in forecasting exchange rates and are more advantageous than other non-linear forecasting methods.
+They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011].
 In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates.
-Superior results were produced when compared to the ARIMA and GARCH model using the RMSE, MSE and MAD as measure of performance.
-Similar results were found by @zhang1998 however the ANN failed to outperform the random walk over long forecast horizons.
-Contrary to the findings above, @kuan1995 attempted to investigate how well feedforward and recurrent networks capture the non-linearity of exchange rate data and obtained conflicting results with neural networks having significantly better market timing ability or lower out-of-sample Mean Square Prediction Error (MSPE) compared to the ARMA models for some exchange rates only.
+Superior results were produced when compared to the ARIMA and GARCH model using the root mean square error (RMSE), mean square error (MSE) and mean absolute difference as a measure of performance.
 @franses1998 suggested that the non-linear feature of exchange rates that are picked up by ANNs may actually be due to neglected GARCH effects.
 They investigated the application of ANNs to forecast exchange rates and found that the presence of GARCH in the data may mistakenly lead one to believe that the returns can be forecasted on the exchange rates themselves.
-Their results show that there is no gain in producing out-of-sample forecasts using ANN if the data is not truly non-linear.
+Their results show that there is no gain in producing out-of-sample forecasts using ANNs if the data is not truly non-linear.
+Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms, and have been successfully used in financial forecasting [@khashei2009].
+Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather discrete decisions to be provided.
+@santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model.
+He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one step-ahead forecasts.
+By analysing the RMSE, U-Theil inequality index, percentage of corrected predicted signals (CPS), and the Pesaran-Timmermann (PT) predictive failure statistic, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced.
+Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and a Fuzzy ARIMA model.
 
-Although neural networks have been broadly applied in financial forecasting, the process of building neural networks is a complex task and there is no consistent method to design them compared to the traditional Box-Jenkins ARIMA model.
-Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes and the training sample size [@zhang1998].
+Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model.
+Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998].
+Like ANNs there is no systematic approach to designing FLSs and they are only understandable when simple.
+Although FLSs has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.
 
-## Fuzzy Logistic Systems
-Fuzzy Logistic Systems were initially developed to solve problems involving linguistic terms, and have been used successful in financial forecasting [@khashei2009].
-Unlike computers, the human decision-making process includes a range of possibilities spanning ‘yes’ and ‘no’.
-Fuzzy logic tries to imitate this reasoning and decision-making process by taking a discrete decision space that consists of a large number of states and turning it into a continuous space.
-This allows for more finer decisions to be provided as opposed to discrete decisions.
+## Hybrid ARIMA Models
+Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA model, and of more complex non-linear models.
+By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations.
+Some of the hybrid models which have commonly been used in exchange rate forecasting are the Fuzzy ARIMA and ANN-ARIMA model.
 
-@santos2007 investigated how well fuzzy systems and neural networks perform compared to the traditional ARMA and GARCH model.
-He examined the forecasts of  Brazilian exchange rate returns looking at different frequencies of the series and compared their one-step-ahead forecasts.
-By analysing the RMSE, U-Theil inequality index, percentage of corrected predicted signals (CPS), and the Pesaran-Timmermann (PT) predictive failure statistic, he found that the Fuzzy Systems and Neural Networks achieved higher returns based on the forecasts they produced.
-Similar results were found by @khashei2009 when he analysed the predictive capabilities of Fuzzy Systems, Neural Networks, the traditional ARIMA model, and a Fuzzy ARIMA model.
-The MAE and MSE of the ARIMA model were higher compared to the other models.
-
-Although Fuzzy Logistic Systems has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate reasoning and are more suitable for problems which do not require high accuracy.
-Like ANN’s there is no systematic approach to designing Fuzzy Logistic Systems and they are only understandable when simple.
-
-# Hybrid ARIMA Models
-Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA, and of more complex non-linear models.
-By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations.
-Some of the hybrid models which have been commonly used are the Fuzzy ARIMA and ANN-ARIMA model.
-
-##Fuzzy ARIMA
-The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data [@tseng2001].
+The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data to yield accurate results [@tseng2001].
 In a world that is constantly changing and with the rapid advancement of technology, access to large amounts of historical data is difficult to obtain.
@@
 The Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to exploit the advantages of both models while simultaneously overcoming their limitations.
-@tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model’s appropriateness and power.
+@tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model's appropriateness and power.
 The Fuzzy ARIMA not only produced forecasts that were superior to the ARIMA and fuzzy time series models but also provided an upper and lower bound which can be used by decision makers to determine the best and worst possible situations.
 
-##ANN-ARIMA
 Although ARIMA models are powerful, they require non-stationary data to be differenced and impose prior assumptions onto the distribution of the data [@ince2006].
@@
 The ANN-ARIMA model draws on the strengths of both models and overcomes these individual difficulties.
-@ince2006 created an ANN-ARIMA model by using the ARIMA to determine the number input nodes required by an ANN for three exchange rates.
-When the ANN-ARIMA model was compared to the pure ARIMA, the hybrid model outperformed the ARIMA based on the MSE.
+@ince2006 created an ANN-ARIMA model by using the ARIMA model to determine the number input nodes required by an ANN for three exchange rates.
+When the ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model outperformed the ARIMA based on the MSE.
 The findings of @khashei2011 study agree with these results.
 They implemented an ARIMA model and used its residuals together with past observations of the data as inputs for the ANN.
-Their hybrid model had superior in-sample and out-sample forecasts compared to the random walk, linear AR and ANN model.
+Their hybrid model had superior in-sample and out-sample forecasts compared to the random walk, linear autoregressive and ANN model.
 
@@
 ARIMA models are also more robust and efficient than complex structural models in relation to short-run forecasting.
-The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison justifies it as a good starting point to compare it to Facebook’s forecasting method, Prophet, that was recently released.
+The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison justifies it as a good starting point to compare it to Facebook's forecasting method, Prophet, that was recently released.
 
-# Forecasting with Prophet 
+## Forecasting with Prophet 
 The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series.
@@
 
-## Construction of Prophet
 Prophet consists of a decomposable model of the form:
@@
 
-## Empirical Applications
-Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge in the model building process and has produced significantly improved forecasts compared to the ARIMA model.
-@taylor2017 forecasted the number of events on Facebook using Prophet.
-The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend.
-The forecasts produced by Prophet were compared to common automatic forecasting techniques such as exponential smoothing, ARIMA, and the seasonal naïve model, as well as to simple models such as the naïve model.
-While exponential smoothing and the seasonal naïve model were quite robust, the ARIMA forecasts were fragile.
-No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.
+# Data \label{Data}
+
+# Methodology \label{Methodology}
+
+## The Box-Jenkins Methodology and Automatic Selection
+Box and Jenkins proposed a set of guidelines that can be followed when selecting ARIMA models.
+This systematic procedure to designing ARIMA models has made them highly popular [@spyros1997].
+It consists of a four-stage iterative process in which: 1) the process is either transformed or differenced to detrend and stabilise the variance of the data, 2) the autocorrelation and partial autocorrelation plots are used to determine the order of p and q, 3) the parameters of the model are estimated, 4) a diagnostic check is performed to ensure that the residuals are a white noise process.
+If the residuals are not white noise steps 2-4 are repeated until a satisfactory model is identified.
+On the contrary, if the diagnostic check shows that the residuals are random then the developed model is the final model used for forecasting.
+
+Due to the large number of forecasts made, it is useful to have an automatic procedure which is able to select the appropriate ARIMA model to fit the data.
+The auto.arima function in R is able to automatically choose the order of the parameters p, q and d.
+The order of differencing is determined by using the KPSS test [@ruppert2015].
+The KPSS test checks the null hypothesis of stationarity and sets d to zero if the null hypothesis is accepted, otherwise it iteratively increases d by 1 and tests the null hypothesis until it is accepted [@ruppert2015].
+Once the order of differencing has been determined, the order of p and q are chosen based on Akaike’s Information Criterion (AIC) or the Bayesian Information Criterion (BIC) [@ruppert2015].
 
@@
 
-Prophet’s ability to forecast at scale enables it to model a wide variety of data and may be able to capture the non-linearity present in exchange rates.
-In contrast, the ARIMA model fails to capture the non-linearity inherent in exchange rate data and only produces good forecasts over short time horizons.
-Prophet is also a flexible model that has intuitive parameters which can be easily interpreted and modified by human beings [@taylor2017].
-This may be useful to non-experts who have domain knowledge about factors that may affect the movement of exchange rate prices.
-Furthermore, unlike the ARIMA model, Prophet can produce forecasts over irregular time intervals and allows for missing values in the time series without the need for interpolation [@taylor2017].
-If Prophet produces forecasts that are as good as the ARIMA model when forecasting Bitcoin/ZAR, it can be compared to more complex forecasting methods that are less tractable such as the hybrid models and machine learning techniques seen earlier.
-
-# Mincer-Zarnowitz Approach to Forecast Evaluation 
+## Mincer-Zarnowitz Approach to Forecast Evaluation 
 The Mincer-Zarnowitz approach to evaluating forecast accuracy is commonly used and can be useful when comparing the forecasts produced by Prophet and the ARIMA model.
@@
 
-# Conclusion
-ARIMA models have been highly popular since its introduction by Box and Jenkins and are widely used for financial forecasting.
-They are easy to implement, have a systematic procedure for model design, and produce good forecasts over short time horizons.
-Although robust, they have the limitation of requiring large amounts of historical observations and they fail to capture the non-linearity inherent in exchange rate data.
-With the evolution of computational power and the advancement of statistics, more complex techniques such as ANN’s, Fuzzy Logistic Systems, and Hybrid ARIMA models have been proposed and compared to the traditional ARIMA model.
-Although these models produce superior results to the ARIMA model, they are less tractable and require the analyst to understand the complex statistical intricacies involved.
-Prophet’s Bayesian approach to forecasting is flexible and allows for non-experts to incorporate their domain knowledge in the model building process.
-If Prophet produces forecasts that are as good as the ARIMA model when forecasting Bitcoin/ZAR, it can be compared to these more complex forecasting methods that are less tractable and scalable.
+# Results \label{Results}
+
+# Discussion and Conclusions \label{Discussion and Conclusions}
 

### 2017-08-27 Draft/Code/draftv2.R

@@ -1,0 +1,596 @@
+
+# -------------------------------READING IN DATASET------------------------------------------------
+
+# Loading required packages
+library(rmsfuns)
+packagestoload <- c("stats", "xts", "dplyr", "Dplyr2Xts", "PerformanceAnalytics", "ggplot2", "readr", "gridExtra", "lmtest")
+load_pkg(packagelist = packagestoload) 
+
+setwd("~/R/UCT_hons_proj/Draft/Data")
+
+# Set relative path 
+ExchangeRates <- read_csv2("prices.csv", col_names = TRUE)
+
+# Looking at first few rows of dataset
+ExchangeRates %>% 
+  data.frame %>% 
+  head
+
+# Looking at structure of dataset
+dim(ExchangeRates)
+str(ExchangeRates)
+
+#--------------------------------- CLEANING THE DATASET--------------------------------------------
+
+# Order dates in ascending order
+ExchangeRates <- ExchangeRates %>% 
+  arrange(TradeDate)
+
+# Make date column time-based
+ExchangeRates$TradeDate <- as.Date(ExchangeRates$TradeDate)
+
+# Checking if there are any missing values
+na_true <- is.na.data.frame(ExchangeRates)
+which(na_true)
+
+# --------------------------------EXPLORATORY DATA ANALYSIS---------------------------------------
+
+# Plot of closing prices
+plot_prices <- ggplot(ExchangeRates, aes(TradeDate, ClosePrice)) + 
+  geom_line() +
+  xlab("Trade Date") +
+  ylab("Closing Prices") + 
+  theme_minimal() + 
+  labs(title  = paste("BTC Price from", 
+                      ExchangeRates$TradeDate[1],
+                      "to",
+                      ExchangeRates$TradeDate[nrow(ExchangeRates)]))
+
+plot_prices 
+# no unusual observations
+# there's a certain degree of variability and an upward trend
+
+# Calculating daily log returns
+ExchangeRates <- 
+  ExchangeRates %>%
+  mutate(DailyLogReturns = c(NA, diff(log(ClosePrice), lag = 1)))
+
+# Plot of daily log returns
+plot_returns <- ggplot(ExchangeRates, aes(TradeDate, DailyLogReturns)) + 
+  geom_line() +
+  xlab("Trade Date") +
+  ylab("Daily Log Returns") +
[…]

### 2017-08-27 Draft/Draft.tex
updated figures

--- 2017-06-13 Literature Review/A_Brief_Literature_Review.tex
@@
 
-\title{Literature Review}
+\title{Draft}
 
@@
 
-The ability to forecast exchange rates is of great importance to policy
-makers, businesses and speculators. Policy makers are interested in
-forecasting exchange rates to assess how its movement will impact
-inflation, and how monetary policy will need to be adjusted in the
-future to keep inflation within the inflation target (Fahimifard et al.
-\protect\hyperlink{ref-fahimifard2009}{2009}). Businesses need to assess
-how future movements in exchange rates will affect the cost of importing
-goods and the demand for goods exported, while speculators aim to
-identify opportunities in price movements which can be exploited to make
-a profit.
+The Autoregressive Integrated Moving Average (ARIMA) model is one of the
+most widely used time series models that has attracted attention in
+financial market forecasting (Kashei and Bijari 2007). The ARIMA model
+is known for its attractive quality of being a rich class of processes,
+making it possible to find a suitable model to adequately fit the data
+(Fahimifard et al. \protect\hyperlink{ref-fahimifard2009}{2009}).
+Although the Random Walk model has typically been applied to foreign
+exchange markets and has produced superior results, some researchers
+have contended that these markets are not efficient and believe that
+future prices depend on current and past events (Abu-Mostafa and Atiya
+\protect\hyperlink{ref-abu1996}{1996}). This has led to the application
+of the ARIMA model to exchange rate problems, where it has since gained
+popularity due to its ease of implementation and tractability. Although
+the ARIMA model yields good forecasts over short forecast horizons and
+is easy to use, the linearity of the ARIMA model fails to adequately
+capture the non-linearity inherent in exchange rate data (Zhang and Hu
+\protect\hyperlink{ref-zhang1998}{1998}).
 
-Producing accurate forecasts of exchange rates has proven to be a
-difficult task due to the noise of the time series (Abu-Mostafa and
-Atiya \protect\hyperlink{ref-abu1996}{1996}). Some economists argue that
-the exchange rate market is very efficient and future movements in price
-are unpredictable (Abu-Mostafa and Atiya
-\protect\hyperlink{ref-abu1996}{1996}). This belief in the efficient
-market hypothesis has led many researchers to model exchange rates using
-a Random Walk (RW) model (Abu-Mostafa and Atiya
-\protect\hyperlink{ref-abu1996}{1996}). Although the RW model has
-produced superior forecasts, some researchers believe that future prices
-depend on present and past events, and other techniques such as the
-autoregressive integrated moving average (ARIMA) model has been applied
-to forecast exchange rates (Abu-Mostafa and Atiya
-\protect\hyperlink{ref-abu1996}{1996}).
-
-The ARIMA model is one of the most commonly used methods in financial
-forecasting (Fahimifard et al.
-\protect\hyperlink{ref-fahimifard2009}{2009}). It is known for its
-attractive quality of being a rich class of processes, making it
-possible to find a model that adequately fits the data (Fahimifard et
-al. \protect\hyperlink{ref-fahimifard2009}{2009}). Some research has
-shown that the nature of exchange rate data is non-linear, and this
-non-linearity may not be captured adequately by the ARIMA model (Zhang
-and Hu \protect\hyperlink{ref-zhang1998}{1998}). Linear parametric
-forecasting techniques such as the generalised autoregressive
-conditional heteroskedasticity (GARCH) model has been proposed as a
-solution to this problem (Zhang and Hu
-\protect\hyperlink{ref-zhang1998}{1998}). Although the GARCH model takes
-care of the non-linearity in the data, the pre-specification of the
-model may prevent it from capturing all the different non-linearities
-which may be present in the data.
-
-With the evolution of computational power, other non-linear
-non-parametric techniques such as Artificial Neural Networks (ANN),
-Support Vector Regression (SVR), Fuzzy Logistic Systems and hybrid
-versions of the ARIMA model have been proposed. The ARIMA model has been
-used as the standard benchmark model against which these other methods
-have been compared. Although they have produced significantly better
-results than the ARIMA model, building these models is a complex task.
-The ARIMA model is more tractable, less computationally expensive, and
-has been used as the building blocks for more advanced models such as
-the autoregressive fractionally integrated (ARFIMA), Fuzzy ARIMA, and
-seasonal ARIMA (SARIMA) model which have been used for exchange rate
+With the evolution of computational power, non-linear, soft computing
+techniques have been proposed as a solution. The ARIMA model has been
+used as the standard benchmark model against which these more complex
+methods have been compared. Although they have produced significantly
+better results than the ARIMA model, these models lack interpretability
+and building them is a challenging task. The ARIMA model is more
+tractable, less computationally expensive, has been used as the building
+blocks for more advanced models and has provided the inspiration for
+hybrid versions of the model which have been used in exchange rate
 forecasting.
 
-Facebook recently open sourced their forecasting model Prophet. Prophet
-is said to be preferable over the ARIMA model as it has the ability to
-forecast at scale, is easily interpretable and has parameters that can
-be intuitively adjusted (Taylor and Letham
-\protect\hyperlink{ref-taylor2017}{2017}). This paper aims to compare
-the forecasts produced by the traditional ARIMA model and Prophet
-through an evaluation of Bitcoin/ZAR, using a Mincer-Zarnowitz approach
-of measuring forecast accuracy. Section 2 will consider the origins of
-the ARIMA model and how it has performed and evolved over time by
-exploring empirical studies where it has been used for forecasting
-exchange rates. Section 3 will also consider other techniques that have
-been used to forecast exchange rates and how they have been compared to
-the ARIMA model. Section 4 will look at hybrid models that were inspired
-by the ARIMA model, while section 5 looks at forecasting with Prophet.
-Finally, section 6 will look at the Mincer-Zarnowitz approach to
-evaluating forecasts.
+In December 2016, Facebook open-sourced their forecasting model Prophet.
+Prophet differs to traditional time series models like ARIMA, in that it
+can be utilised by non-experts who have little knowledge about the
+statistical intricacies involved in the model, however have domain
+knowledge about the data generating process. This can then be easily
+incorporated into the model through intuitively adjustable parameters
+(Taylor and Letham \protect\hyperlink{ref-taylor2017}{2017}). If Prophet
+produces forecasts that are as good as the ARIMA model, it can be
+compared to more complex forecasting methods that are less tractable and
+flexible.
 
-\section{Forecasting Exchange Rates with the ARIMA
-Model}\label{forecasting-exchange-rates-with-the-arima-model}
+This paper broadly aims to compare the forecasts produced by the
+traditional ARIMA model and Prophet through an evaluation of
+Bitcoin/ZAR, using a Mincer-Zarnowitz approach of measuring forecast
+accuracy. More specifically, this paper aims to:
 
-\subsection{Birth of ARIMA}\label{birth-of-arima}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  Compare the out-of-sample forecasts over varying forecast horizons.
+\item
+  Compare the forecasts produced by ARIMA's and Prophet's automatic
+  selection procedures.
+\item
+  Compare the out-of-sample forecast performance with missing values and
+  outliers present in the data.
+\item
+  Investigate the efficient market hypothesis (EMH) and the trading
+  profit to be made with Prophet.
+\end{enumerate}
+
+Section \ref{Background} will present the key findings from the major
+works in which comparisons have been made to the ARIMA model in exchange
+rate forecasting, as well as the theory behind the ARIMA and Prophet
+model. This will be followed by a brief explanation and motivation for
+the dataset used in this paper in Section \ref{Data}. Section
+\ref{Methodology} will describe the methodology used in selecting the
+ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring
+forecast accuracy. The results from applying the ARIMA and Prophet model
+to the dataset is presented in Section \ref{Results}. Finally, Section
+\ref{Discussion and Conclusions} discusses the results in a statistical
+and financially meaningful context.
+
+\section{\texorpdfstring{Background
+\label{Background}}{Background }}\label{background}
+
+\subsection{Forecasting Exchange Rates with the ARIMA
+model}\label{forecasting-exchange-rates-with-the-arima-model}
 
@@
     y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
-\] where c is a constant and \{\(e_t\)\} is a white noise process with
-zero mean and variance \(\sigma^2\).
+\] where c is a constant, \(\theta\) and \(\phi\) are parameters set
+such that \ldots{} and \{\(e_t\)\} is a white noise process with zero
+mean and variance \(\sigma^2\).
 
@@
 \] where c is a constant and \{\(e_t\)\} is a white noise process with
-zero mean and variance \(\sigma^2\).
-
-\subsection{The Box-Jenkins Methodology and Automatic
-Selection}\label{the-box-jenkins-methodology-and-automatic-selection}
-
-Box and Jenkins proposed a set of guidelines that can be followed when
-selecting ARIMA models. This systematic procedure to designing ARIMA
-models has made them highly popular (Hibon and Makridakis
-\protect\hyperlink{ref-spyros1997}{1997}). It consists of a four-stage
-iterative process in which: 1) the process is either transformed or
-differenced to detrend and stabilise the variance of the data, 2) the
-autocorrelation and partial autocorrelation plots are used to determine
-the order of p and q, 3) the parameters of the model are estimated, 4) a
-diagnostic check is performed to ensure that the residuals are a white
-noise process. If the residuals are not white noise steps 2-4 are
-repeated until a satisfactory model is identified. On the contrary, if
-the diagnostic check shows that the residuals are random then the
-developed model is the final model used for forecasting.
-
-Due to the large number of forecasts made, it is useful to have an
-automatic procedure which is able to select the appropriate ARIMA model
-to fit the data. The auto.arima function in R is able to automatically
-choose the order of the parameters p, q and d. The order of differencing
-is determined by using the KPSS test (Ruppert and Matteson
-\protect\hyperlink{ref-ruppert2015}{2015}). The KPSS test checks the
-null hypothesis of stationarity and sets d to zero if the null
-hypothesis is accepted, otherwise it iteratively increases d by 1 and
-tests the null hypothesis until it is accepted (Ruppert and Matteson
-\protect\hyperlink{ref-ruppert2015}{2015}). Once the order of
-differencing has been determined, the order of p and q are chosen based
-on Akaike's Information Criterion (AIC) or the Bayesian Information
-Criterion (BIC) (Ruppert and Matteson
-\protect\hyperlink{ref-ruppert2015}{2015}).
-
-\subsection{Empirical Applications}\label{empirical-applications}
+zero mean and variance \(\sigma^2\) (otext.org/fpp/8/5)\ldots{}need to
+get this reference hmmm
 
@@
 and statistical properties (C.-S. Lin, Chiu, and Lin
-\protect\hyperlink{ref-lin2012}{2012}). They have been used as a
-benchmark to compare new forecasting techniques which have emerged over
-time. Kamruzzaman and Sarker
+\protect\hyperlink{ref-lin2012}{2012}). They have frequently been used
+as a benchmark to compare new forecasting techniques that have emerged
+over time. Kamruzzaman and Sarker
 (\protect\hyperlink{ref-kamruzzaman2003}{2003}) applied an ARIMA model
 to forecast the exchange rate between the Australian dollar and six
-other currencies. Its results were used to evaluate the performance of
-three Neural Network models. The Normalised Mean Square Error (NMSE) and
-Mean Absolute Error (MAE) were used to measure the forecast errors while
-the Directional Symmetry (DS), Correct Up trend (CU) and Correct Down
-trend (CD) were used to measure the accuracy of the direction of the
-forecasts. Similarly, C.-S. Lin, Chiu, and Lin
-(\protect\hyperlink{ref-lin2012}{2012}) employed an ARIMA model to
-forecast the Taiwan New Dollar exchange rate. The aim of applying the
-ARIMA model was to analyse the ability of a proposed hybrid model
-created from the Squares Support Vector Regression (LSSVR) and Empirical
-Mode Decomposition (EMD) model. C.-S. Lin, Chiu, and Lin
-(\protect\hyperlink{ref-lin2012}{2012}) used the same indicators as
-Kamruzzaman and Sarker (\protect\hyperlink{ref-kamruzzaman2003}{2003})
-to evaluate the accuracy of the direction of the forecasts. To measure
-the forecast errors produced they applied the mean absolute percentage
-error (MAPE), root‐mean‐square error (RMSE) and mean absolute difference
-(MAD). Khashei, Bijari, and Ardali
-(\protect\hyperlink{ref-khashei2012}{2012}) also investigated the
-ability to forecast exchange rates by considering the British pound and
-US dollar, over a short and long forecast horizon. They compared the
-predictive capability of their proposed hybrid model consisting of an
-ARIMA and a Probabilistic Neural Network (PNN), to the traditional ARIMA
-model, and justified the use of the ARIMA as a benchmark, by claiming
-that it is the most important linear model. The ARIMA model has also
-yielded satisfactory results when predicting exchange rates. Nwankwo
-(\protect\hyperlink{ref-nwankwo2014}{2014}) forecasted the Nairo/dollar
-rate and diagnostic testing revealed that the ARIMA(1,0,0) model was the
-best fit for the data based on the AIC.
+other currencies, and used the forecast errors as well as the accuracy
+of the direction of the forecasts to evaluate the performance of three
+Neural Network models. In a similar study, Khashei, Bijari, and Ardali
+(\protect\hyperlink{ref-khashei2012}{2012}) compared the predictive
+capability of their proposed hybrid model consisting of an ARIMA and a
+Probabilistic Neural Network (PNN), to the traditional ARIMA model, and
+justified the use of the ARIMA as a benchmark by claiming that it is the
+most important linear model. Their comparison was made by investigating
+the forecasts between the British pound and US dollar over varying
+forecast horizons. Not only has the ARIMA model been useful as a
+benchmark, but it has also yielded satisfactory results when predicting
+exchange rates. When Nwankwo (\protect\hyperlink{ref-nwankwo2014}{2014})
+forecasted the rate between the Nairo and dollar, diagnostic testing
+revealed that the ARIMA(1,0,0) model was the best fit for the data based
+on Akaike's Information Criterion (AIC).
 
 The ARIMA model is attractive as it is tractable and produces good
-forecasts in short time periods where more than 100 observations are
-used (Tseng et al. \protect\hyperlink{ref-tseng2001}{2001}). Although
-the ARIMA model has the advantage of ease of implementation and
-flexibility, it fails to capture the non-linearity and volatility
-present in exchange rate data.
+short-term forecasts when more than 100 observations are used (Tseng et
+al. \protect\hyperlink{ref-tseng2001}{2001}). Although the ARIMA model
+has the advantage of ease of implementation and flexibility, it fails to
+capture the non-linearity and volatility present in exchange rate data.
+Over time, the ARIMA model has evolved to cater for a wider variety of
+data and to compensate for some of its shortcomings. The most popular
+versions of the ARIMA model that has been implemented in exchange rate
+forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model
+(FARIMA).
 
-\subsection{Evolution of ARIMA}\label{evolution-of-arima}
+The SARIMA model was introduced to capture the periodic behaviour of
+data and extends the ARIMA class by including a term for seasonal
+differencing. Etuk, Wokoma, and Moffat
+(\protect\hyperlink{ref-etuk2013}{2013}) modelled the Naira/CFA Franc
+exchange rate which exhibited monthly seasonality using an additive
+SARIMA model, to demonstrate that it can be a useful fit for exchange
+rate data which displays seasonality. Their results showed that the
+SARIMA model adequately described the variation in the exchange rate
+series. The ARFIMA model generalises the ARIMA model in that the degree
+required to make the data stationary can assume any real value, and is
+no longer restricted to the integer domain. The ARFIMA model has the
+ability to capture the dependence between observations that are widely
+spread apart in time (Cheung \protect\hyperlink{ref-cheung1993}{1993}).
+This makes the model more parsimonious since it can capture long memory
+in data as well as short term dynamics (Cheung
+\protect\hyperlink{ref-cheung1993}{1993}). Cheung
+(\protect\hyperlink{ref-cheung1993}{1993}) fitted the ARFIMA model to
+examine five exchange rates, and found that there was strong evidence of
+long memory in the exchange rate time series.
 
-Over time, the ARIMA model has evolved to cater for a wider variety of
-data that it can be applied to, and to compensate for some of its
-shortcomings. The most popular versions of the ARIMA model that has been
-implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA)
-and Fractional ARIMA model (FARIMA).
+Generalised autoregressive conditional heteroskedasticity (GARCH) models
+were later developed in an attempt to capture the volatility in
+financial markets (Anastasakis and Mort
+\protect\hyperlink{ref-anastasakis2009}{2009}). Hsieh (1989) applied a
+GARCH model to investigate five exchange rates and his results showed
+that although the GARCH model outperformed the random walk, some
+non-linear information still remained in the residuals.
 
-The SARIMA model was introduced by Box and Jenkins as a generalisation
-of the ARIMA model to capture the periodic behaviour of data. SARIMA
-models extend the ARIMA class by including a term for seasonal
-differencing. Exchange rate data that has been observed have shown
-seasonality, and have been attempted to be fitted by SARIMA models
-(Etuk, Wokoma, and Moffat \protect\hyperlink{ref-etuk2013}{2013}). Etuk,
-Wokoma, and Moffat (\protect\hyperlink{ref-etuk2013}{2013}) modelled the
-Naira/CFA Franc exchange rate which exhibited monthly seasonality using
-an additive SARIMA model, to demonstrate that it can be a useful fit to
-the data. Their results showed that the SARIMA model adequately
-described the variation in the exchange rate series. This agrees with
-the results of a study by Etuk and others
-(\protect\hyperlink{ref-etuk2012}{2012}) which fitted a multiplicative
-SARIMA model to the Naira/Dollar exchange rates.
-
-The ARFIMA model is another generalisation of the ARIMA model in that
-the degree required to make the data stationary can assume any real
-value, and is no longer restricted to the integer domain. While the
-ARIMA model displays exponentially decaying autocorrelation between
-observations the further apart they are in time, the ARFIMA model has
-the ability to capture the dependence between observations that are
-widely spread apart in time (Cheung
-\protect\hyperlink{ref-cheung1993}{1993}). This makes the ARFIMA model
-more parsimonious since it can capture long memory in data as well as
-short term dynamics (Cheung \protect\hyperlink{ref-cheung1993}{1993}).
-Cheung (\protect\hyperlink{ref-cheung1993}{1993}) fitted the ARFIMA
-model to examine five exchange rates, and found that there was strong
-evidence of long memory in the exchange rate time series. Although the
-ARFIMA model fitted the data well, it failed to beat the random walk in
-out-of sample forecasts.
-
-GARCH models were later developed due to the failure of ARIMA models to
-capture the volatility in financial markets (Anastasakis and Mort
-\protect\hyperlink{ref-anastasakis2009}{2009}). They have been widely
-used to forecast the volatility of stock returns, interest rates and
-exchange rates and are useful for portfolio allocation decisions as well
-as option pricing (Cermak \protect\hyperlink{ref-cermak2017}{2017}).
-Cermak (\protect\hyperlink{ref-cermak2017}{2017}) employed a GARCH(1,1)
-model to analyse the volatility of Bitcoin in countries in which Bitcoin
-is commonly traded, while Hsieh (1989) applied a GARCH model to
-investigate five exchange rates. Hsieh's (1989) results showed that
-although the GARCH model outperformed the random walk, some non-linear
-information still remained in the residuals. In a similar study,
-Fahimifard et al. (\protect\hyperlink{ref-fahimifard2009}{2009})
-illustrated that the GARCH model outperformed the ARIMA, while Neural
-Networks outperformed both the GARCH and ARIMA model.
-
-\section{Other techniques used to forecast exchange
+\subsection{Other techniques used to forecast exchange
 rates}\label{other-techniques-used-to-forecast-exchange-rates}
@@
 easily (Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}).
-Unlike regression analysis and the ARIMA model, soft computing
-techniques do not impose structural assumptions on the model apriori
-(Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}). Some of
-the most commonly used artificial intelligence methods used to forecast
-exchange rate data are Neural Networks and the Fuzzy Logistic Systems.
+Unlike the ARIMA model, soft computing techniques do not impose
+structural assumptions on the model apriori (Castillo and Melin
+\protect\hyperlink{ref-castillo2002}{2002}). Some of the most commonly
+used artificial intelligence methods used to forecast exchange rate data
+are Neural Networks and Fuzzy Logistic Systems.
 
-\subsection{Neural Networks}\label{neural-networks}
-
-Artificial Neural Networks are commonly used in financial forecasting
-and are more advantageous than other non-linear forecasting methods.
-They can approximate any continuous function, are data driven, and can
-adapt to non-stationary environments (Khashei and Bijari
-\protect\hyperlink{ref-khashei2011}{2011}). Neural Networks mimic the
-structure of the brain and consists of three layers: the input, hidden
-and output layer, each of which contain nodes which are representative
-of the neurons in our brain. The job of the nodes in the input layer is
-to send signals along connections to the nodes in the hidden layer,
-where the connections are representative of the synapses in our brain.
-The nodes in the hidden layer uses an activation function and assigns a
-weight to each connection according to the level of importance of
-information. The value in the hidden layer is then sent along
-connections to the output layer, where weights are again assigned to
-each connection. The error is then computed in the output layer and is
-backpropagated through the network. The weights are adjusted according
-to the error that they contribute, allowing the ANN to be trained and
-learn. This learning process is repeated until the network achieves the
-desired output.
-
-ANN's have had many successful applications in forecasting exchange
-rates. In a study done by Fahimifard et al.
+Artificial Neural Networks (ANNs) have had many successful applications
+in forecasting exchange rates and are more advantageous than other
+non-linear forecasting methods. They are data driven, can adapt to
+non-stationary environments and can approximate any continuous function
+(Khashei and Bijari \protect\hyperlink{ref-khashei2011}{2011}). In a
+study done by Fahimifard et al.
 (\protect\hyperlink{ref-fahimifard2009}{2009}), the ANN was found as an
@@
 results were produced when compared to the ARIMA and GARCH model using
-the RMSE, MSE and MAD as measure of performance. Similar results were
-found by Zhang and Hu (\protect\hyperlink{ref-zhang1998}{1998}) however
-the ANN failed to outperform the random walk over long forecast
-horizons. Contrary to the findings above, Kuan and Liu
-(\protect\hyperlink{ref-kuan1995}{1995}) attempted to investigate how
-well feedforward and recurrent networks capture the non-linearity of
-exchange rate data and obtained conflicting results with neural networks
-having significantly better market timing ability or lower out-of-sample
-Mean Square Prediction Error (MSPE) compared to the ARMA models for some
-exchange rates only. Franses and Van Homelen
+the root mean square error (RMSE), mean square error (MSE) and mean
+absolute difference as a measure of performance. Franses and Van Homelen
 (\protect\hyperlink{ref-franses1998}{1998}) suggested that the
@@
 results show that there is no gain in producing out-of-sample forecasts
-using ANN if the data is not truly non-linear.
+using ANNs if the data is not truly non-linear. Fuzzy Logistic Systems
+(FLSs) were initially developed to solve problems involving linguistic
+terms, and have been successfully used in financial forecasting
+(Khashei, Bijari, and Ardali \protect\hyperlink{ref-khashei2009}{2009}).
+Fuzzy logic tries to imitate human reasoning and the decision-making
+process and allows for finer rather discrete decisions to be provided.
+Santos, Costa, and Santos Coelho
+(\protect\hyperlink{ref-santos2007}{2007}) investigated how well FLSs
+and ANNs perform compared to the traditional ARMA and GARCH model. He
+examined the forecasts of Brazilian exchange rate returns by considering
+different frequencies of the series and comparing their one step-ahead
+forecasts. By analysing the RMSE, U-Theil inequality index, percentage
+of corrected predicted signals (CPS), and the Pesaran-Timmermann (PT)
+predictive failure statistic, he found that FLSs and ANNs achieved
+higher returns based on the forecasts they produced. Similar results
+were found by Khashei, Bijari, and Ardali
+(\protect\hyperlink{ref-khashei2009}{2009}) when he analysed the
+predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and
+a Fuzzy ARIMA model.
 
-Although neural networks have been broadly applied in financial
-forecasting, the process of building neural networks is a complex task
-and there is no consistent method to design them compared to the
-traditional Box-Jenkins ARIMA model. Unlike ARIMA models, the
-performance of ANNs is sensitive to many modelling factors such as the
-number of input nodes and the training sample size (Zhang and Hu
-\protect\hyperlink{ref-zhang1998}{1998}).
+Although ANNs have been broadly applied in financial forecasting, the
+process of building them is a complex task and there is no consistent
+method of design compared to the traditional Box-Jenkins ARIMA model.
+Unlike ARIMA models, the performance of ANNs is sensitive to many
+modelling factors such as the number of input nodes included and the
+size of the training sample chosen (Zhang and Hu
+\protect\hyperlink{ref-zhang1998}{1998}). Like ANNs there is no
+systematic approach to designing FLSs and they are only understandable
+when simple. Although FLSs has the advantage over ARIMA models that they
+can be applied to data with few observations available, it gives
+acceptable rather than accurate results and are more suitable for
+problems which do not require high accuracy.
 
-\subsection{Fuzzy Logistic Systems}\label{fuzzy-logistic-systems}
-
-Fuzzy Logistic Systems were initially developed to solve problems
-involving linguistic terms, and have been used successful in financial
-forecasting (Khashei, Bijari, and Ardali
-\protect\hyperlink{ref-khashei2009}{2009}). Unlike computers, the human
-decision-making process includes a range of possibilities spanning `yes'
-and `no'. Fuzzy logic tries to imitate this reasoning and
-decision-making process by taking a discrete decision space that
-consists of a large number of states and turning it into a continuous
-space. This allows for more finer decisions to be provided as opposed to
-discrete decisions.
-
-Santos, Costa, and Santos Coelho
-(\protect\hyperlink{ref-santos2007}{2007}) investigated how well fuzzy
-systems and neural networks perform compared to the traditional ARMA and
-GARCH model. He examined the forecasts of Brazilian exchange rate
-returns looking at different frequencies of the series and compared
-their one-step-ahead forecasts. By analysing the RMSE, U-Theil
-inequality index, percentage of corrected predicted signals (CPS), and
-the Pesaran-Timmermann (PT) predictive failure statistic, he found that
-the Fuzzy Systems and Neural Networks achieved higher returns based on
-the forecasts they produced. Similar results were found by Khashei,
-Bijari, and Ardali (\protect\hyperlink{ref-khashei2009}{2009}) when he
-analysed the predictive capabilities of Fuzzy Systems, Neural Networks,
-the traditional ARIMA model, and a Fuzzy ARIMA model. The MAE and MSE of
-the ARIMA model were higher compared to the other models.
-
-Although Fuzzy Logistic Systems has the advantage over ARIMA models that
-they can be applied to data with few observations available, it gives
-acceptable rather than accurate reasoning and are more suitable for
-problems which do not require high accuracy. Like ANN's there is no
-systematic approach to designing Fuzzy Logistic Systems and they are
-only understandable when simple.
-
-\section{Hybrid ARIMA Models}\label{hybrid-arima-models}
+\subsection{Hybrid ARIMA Models}\label{hybrid-arima-models}
 
 Over time, many researchers began to think of ways in which to harness
-the advantages of tractable linear models such as the ARIMA, and of more
-complex non-linear models. By combining the ARIMA model with other
-forecasting methods, the advantages of both forecasting methods are
-leveraged while simultaneously improving their limitations. Some of the
-hybrid models which have been commonly used are the Fuzzy ARIMA and
-ANN-ARIMA model.
-
-\subsection{Fuzzy ARIMA}\label{fuzzy-arima}
+the advantages of tractable linear models such as the ARIMA model, and
+of more complex non-linear models. By combining the ARIMA model with
+other forecasting methods, the advantages of both forecasting methods
+are leveraged while simultaneously improving their limitations. Some of
+the hybrid models which have commonly been used in exchange rate
+forecasting are the Fuzzy ARIMA and ANN-ARIMA model.
 
@@
 horizons however it has the limitation of requiring more than 100
-observations of historical data (Tseng et al.
+observations of historical data to yield accurate results (Tseng et al.
 \protect\hyperlink{ref-tseng2001}{2001}). In a world that is constantly
@@
 
-\subsection{ANN-ARIMA}\label{ann-arima}
-
 Although ARIMA models are powerful, they require non-stationary data to
@@
 difficulties. Ince and Trafalis (\protect\hyperlink{ref-ince2006}{2006})
-created an ANN-ARIMA model by using the ARIMA to determine the number
-input nodes required by an ANN for three exchange rates. When the
-ANN-ARIMA model was compared to the pure ARIMA, the hybrid model
+created an ANN-ARIMA model by using the ARIMA model to determine the
+number input nodes required by an ANN for three exchange rates. When the
+ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model
 outperformed the ARIMA based on the MSE. The findings of Khashei and
@@
 hybrid model had superior in-sample and out-sample forecasts compared to
-the random walk, linear AR and ANN model.
+the random walk, linear autoregressive and ANN model.
 
@@
 
-\section{Forecasting with Prophet}\label{forecasting-with-prophet}
+\subsection{Forecasting with Prophet}\label{forecasting-with-prophet}
 
@@
 scale.
-
-\subsection{Construction of Prophet}\label{construction-of-prophet}
 
@@
 
-\subsection{Empirical Applications}\label{empirical-applications-1}
+\section{\texorpdfstring{Data \label{Data}}{Data }}\label{data}
 
-Prophet's Bayesian approach to forecasting allows the analyst to
-incorporate their expert knowledge in the model building process and has
-produced significantly improved forecasts compared to the ARIMA model.
-Taylor and Letham (\protect\hyperlink{ref-taylor2017}{2017}) forecasted
-the number of events on Facebook using Prophet. The time series was
-impacted by holidays, had strong multi-period seasonality, and a
-piecewise trend. The forecasts produced by Prophet were compared to
-common automatic forecasting techniques such as exponential smoothing,
-ARIMA, and the seasonal naïve model, as well as to simple models such as
-the naïve model. While exponential smoothing and the seasonal naïve
-model were quite robust, the ARIMA forecasts were fragile. No model
-besides Prophet accounted for the dips around holidays and the upward
-trend of the time series towards later observations.
+\section{\texorpdfstring{Methodology
+\label{Methodology}}{Methodology }}\label{methodology}
+
+\subsection{The Box-Jenkins Methodology and Automatic
+Selection}\label{the-box-jenkins-methodology-and-automatic-selection}
+
+Box and Jenkins proposed a set of guidelines that can be followed when
+selecting ARIMA models. This systematic procedure to designing ARIMA
+models has made them highly popular (Hibon and Makridakis
+\protect\hyperlink{ref-spyros1997}{1997}). It consists of a four-stage
+iterative process in which: 1) the process is either transformed or
+differenced to detrend and stabilise the variance of the data, 2) the
+autocorrelation and partial autocorrelation plots are used to determine
+the order of p and q, 3) the parameters of the model are estimated, 4) a
+diagnostic check is performed to ensure that the residuals are a white
+noise process. If the residuals are not white noise steps 2-4 are
+repeated until a satisfactory model is identified. On the contrary, if
+the diagnostic check shows that the residuals are random then the
+developed model is the final model used for forecasting.
+
+Due to the large number of forecasts made, it is useful to have an
+automatic procedure which is able to select the appropriate ARIMA model
+to fit the data. The auto.arima function in R is able to automatically
+choose the order of the parameters p, q and d. The order of differencing
+is determined by using the KPSS test (Ruppert and Matteson
+\protect\hyperlink{ref-ruppert2015}{2015}). The KPSS test checks the
+null hypothesis of stationarity and sets d to zero if the null
+hypothesis is accepted, otherwise it iteratively increases d by 1 and
+tests the null hypothesis until it is accepted (Ruppert and Matteson
+\protect\hyperlink{ref-ruppert2015}{2015}). Once the order of
+differencing has been determined, the order of p and q are chosen based
+on Akaike's Information Criterion (AIC) or the Bayesian Information
+Criterion (BIC) (Ruppert and Matteson
+\protect\hyperlink{ref-ruppert2015}{2015}).
 
@@
 
-Prophet's ability to forecast at scale enables it to model a wide
-variety of data and may be able to capture the non-linearity present in
-exchange rates. In contrast, the ARIMA model fails to capture the
-non-linearity inherent in exchange rate data and only produces good
-forecasts over short time horizons. Prophet is also a flexible model
-that has intuitive parameters which can be easily interpreted and
-modified by human beings (Taylor and Letham
-\protect\hyperlink{ref-taylor2017}{2017}). This may be useful to
-non-experts who have domain knowledge about factors that may affect the
-movement of exchange rate prices. Furthermore, unlike the ARIMA model,
-Prophet can produce forecasts over irregular time intervals and allows
-for missing values in the time series without the need for interpolation
-(Taylor and Letham \protect\hyperlink{ref-taylor2017}{2017}). If Prophet
-produces forecasts that are as good as the ARIMA model when forecasting
-Bitcoin/ZAR, it can be compared to more complex forecasting methods that
-are less tractable such as the hybrid models and machine learning
-techniques seen earlier.
-
-\section{Mincer-Zarnowitz Approach to Forecast
+\subsection{Mincer-Zarnowitz Approach to Forecast
 Evaluation}\label{mincer-zarnowitz-approach-to-forecast-evaluation}
@@
 
-\section{Conclusion}\label{conclusion}
+\section{\texorpdfstring{Results
+\label{Results}}{Results }}\label{results}
 
-ARIMA models have been highly popular since its introduction by Box and
-Jenkins and are widely used for financial forecasting. They are easy to
-implement, have a systematic procedure for model design, and produce
-good forecasts over short time horizons. Although robust, they have the
-limitation of requiring large amounts of historical observations and
-they fail to capture the non-linearity inherent in exchange rate data.
-With the evolution of computational power and the advancement of
-statistics, more complex techniques such as ANN's, Fuzzy Logistic
-Systems, and Hybrid ARIMA models have been proposed and compared to the
-traditional ARIMA model. Although these models produce superior results
-to the ARIMA model, they are less tractable and require the analyst to
-understand the complex statistical intricacies involved. Prophet's
-Bayesian approach to forecasting is flexible and allows for non-experts
-to incorporate their domain knowledge in the model building process. If
-Prophet produces forecasts that are as good as the ARIMA model when
-forecasting Bitcoin/ZAR, it can be compared to these more complex
-forecasting methods that are less tractable and scalable.
+\section{\texorpdfstring{Discussion and Conclusions
+\label{Discussion and Conclusions}}{Discussion and Conclusions }}\label{discussion-and-conclusions}
 
@@
 
-\hypertarget{ref-cermak2017}{}
-Cermak, Vavrinec. 2017. ``Can Bitcoin Become a Viable Alternative to
-Fiat Currencies? An Empirical Analysis of Bitcoin's Volatility Based on
-a Garch Model.''
-
 \hypertarget{ref-cheung1993}{}
@@
 Francis: 93--101.
-
-\hypertarget{ref-etuk2012}{}
-Etuk, Ette Harrison, and others. 2012. ``A Seasonal Arima Model for
-Daily Nigerian Naira-Us Dollar Exchange Rates.'' \emph{Asian Journal of
-Empirical Research} 2 (6). Asian Economic; Social Society: 219--27.
 
@@
 
-\hypertarget{ref-kuan1995}{}
-Kuan, Chung-Ming, and Tung Liu. 1995. ``Forecasting Exchange Rates Using
-Feedforward and Recurrent Neural Networks.'' \emph{Journal of Applied
-Econometrics} 10 (4). Wiley Online Library: 347--64.
-
 \hypertarget{ref-lin2012}{}

### 2017-08-27 Draft/draftv2.R
Highly confused about how to push what I want. I think I'm now adding the modifications to figures. Who knows. Sigh

@@ -1,685 +1,0 @@
-
-# -------------------------------READING IN DATASET------------------------------------------------
-
-# Loading required packages
-library(rmsfuns)
-packagestoload <- c("stats", "xts", "readr", "dplyr", "Dplyr2Xts", "PerformanceAnalytics", "ggplot2", "readr", "gridExtra", "lmtest")
-load_pkg(packagelist = packagestoload) 
-
-# Hanjo - add library readr
-
-setwd("~/R/UCT_hons_proj/Draft/Data")
-
-# Hanjo - Set relative path thanks to proj
-ExchangeRates <- read_csv2("Draft/Data/prices.csv")
-
-# Looking at first few rows of dataset
-ExchangeRates %>% 
-  data.frame %>% 
-  head
-
-# Looking at structure of dataset
-dim(ExchangeRates)
-str(ExchangeRates)
-
-#--------------------------------- CLEANING THE DATASET--------------------------------------------
-
-# Order dates in ascending order
-ExchangeRates <- ExchangeRates %>% 
-  arrange(TradeDate)
-
-# Make date column time-based
-ExchangeRates$TradeDate <- as.Date(ExchangeRates$TradeDate)
-
-# Checking if there are any missing values
-na_true <- is.na.data.frame(ExchangeRates)
-which(na_true)
-
-# --------------------------------EXPLORATORY DATA ANALYSIS---------------------------------------
-
-# Plot of closing prices
-plot_prices <- ggplot(ExchangeRates, aes(TradeDate, ClosePrice)) + 
-  geom_line() +
-  xlab("Trade Date") +
-  ylab("Closing Prices") + 
-  theme_minimal() + 
-  labs(title  = paste("BTC Price from", 
-                      ExchangeRates$TradeDate[1],
-                      "to",
-                      ExchangeRates$TradeDate[nrow(ExchangeRates)]))
-
-plot_prices 
-# no unusual observations
-# there's a certain degree of variability and an upward trend
-
-# Calculating daily log returns
-ExchangeRates <- 
-  ExchangeRates %>%
-  mutate(DailyLogReturns = c(NA, diff(log(ClosePrice), lag = 1)))
-
-# Plot of daily log returns
-plot_returns <- ggplot(ExchangeRates, aes(TradeDate, DailyLogReturns)) + 
-  geom_line() +
-  xlab("Trade Date") +
[…]

### 2017-09-14 Draft/Write-up/Draft.Rmd
Final Draft. Sigh

---
title: "How does the latest Prophet forecasting model compare to the ARIMA model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz approach"
Author: "Fiona Ganie"  
Student: "GNXFIO001"
bibliography: Tex/ref.bib       
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 5 
    fig_height: 4
abstract: |
  The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has gained popularity in the exchange rate market due to its ease of implementation and tractability. With the evolution of computational power, soft computing techniques have since been used in exchange rate markets. The ARIMA model has been used as the standard benchmark model against which these complex methods have been compared. Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task. Prophet is a sophisticated model that differs to traditional time series models in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. Prophet is also robust to outliers and can handle missing values in the time series without the need for interpolation. This paper compares the out-of-sample forecasts of Prophet and ARIMA over varying forecast horizons, using a Mincer-Zarnowitz approach for forecast evaluation. A complete dataset of the prices of Bitcoin/Rand as well as a dataset which has missing prices and outliers present is used in the analysis. It is found that the ARIMA model produces the most efficient and unbiased forecasts over all forecast horizons, and that the ARIMA model is more robust when faced with missing values and outliers present in the data. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

```

# Introduction \label{Introduction}
The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has attracted attention in financial market forecasting [@khashei2009]. Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that foreign exchange markets are not efficient and believe that future prices depend on current and past events [@abu1996]. This has led to the application of the ARIMA model to exchange rate problems, where it has since gained popularity due to its ease of implementation and tractability. The ARIMA model is easy to use and yields good forecasts over short forecast horizons, however the linearity of the ARIMA model fails to adequately capture the non-linearity inherent in exchange rate data [@zhang1998].

With the evolution of computational power, non-linear, soft computing techniques have been proposed as a solution. The ARIMA model has been used as the standard benchmark model against which these more complex methods have been compared. Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task. The ARIMA model is tractable and less computationally expensive. It has been used as the building blocks for more advanced models and has provided the inspiration for hybrid versions of the model which have been used in exchange rate forecasting.

In December 2016, Facebook open-sourced their forecasting model Prophet. Prophet differs to traditional time series model such as the ARIMA, in that it can produce high quality forecasts in a straightforward way. Prophet is a sophisticated model that provides informative results, however is configurable and easy to use.  It can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. This knowledge can be easily incorporated into the model through its intuitively adjustable parameters [@taylor2017]. Furthermore, Prophet is robust to outliers and can handle missing values in the time series without the need for interpolation [@taylor2017]. This allows analysts without knowledge on how to pre-process the data, to utilise the model without sacrificing predictive accuracy.  If Prophet produces forecasts that are as good as the ARIMA model, it can be compared to more complex forecasting methods that are less tractable and flexible.

This paper broadly aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of the exchange rate between Bitcoin and the Rand (BTC/ZAR), using a Mincer-Zarnowitz approach of measuring forecast accuracy. More specifically, this paper aims to:

1. Compare the out-of-sample forecasts of ARIMA and Prophet over varying forecast horizons.

2. Compare the out-of-sample forecast performance with missing values and outliers present in the data.

Section \ref{Background and Theory} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting, as well as the theory behind the ARIMA and Prophet model. Section \ref{Data} will provide a brief explanation and motivation for the dataset used in this paper. Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring forecast accuracy. The results from applying the ARIMA and Prophet model to the dataset is presented in Section \ref{Results}. Finally, Section \ref{Discussion and Conclusions discusses the results in a statistical context.

# Background and Theory \label{Background and Theory}

## Forecasting Exchange Rates with the ARIMA model
ARIMA models became highly popular since its introduction by Box and Jenkins, when it was shown that they could outperform complex econometric models in a variety of situations [@spyros1997]. The ARIMA model expresses the process {$y_t$} as a function of the weighted average of past values of the process and lagged values of the residuals. The weighted average of the past p values of the process represents an autoregressive (AR) process of order p. It feeds back past values of the process into the current value, inducing correlation between all lags of the process. The weighted average of the q lagged residuals represents a moving average (MA) process of order q. The purpose of mixing the MA process with the AR process is to reduce the large number of past values required by AR processes and to control for the autocorrelation which it creates between lagged values of the process. The combination of the AR(p) and MA(q) process results in a more parsimonious model, and forms a stationary autoregressive moving average (ARMA(p,q)) process defined as:

$$ 
	y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
$$
where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.

The ARIMA(p,d,q) model generalises the ARMA model in that it includes both stationary and non-stationary processes. The parameter d is the degree of differencing required to render the process stationary. If d is equal to zero the process is stationary and equivalent to an ARMA model, and if d is strictly positive the process requires differencing to make it stationary. The ARIMA model can be defined succinctly using the backward shift operator B, which shifts the process back by one unit of time, and is defined as $By_t = y_{t-1}$. The ARIMA model has the form:

$$ 
	(1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
$$
where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.

ARIMA models have commonly been used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012]. They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time. @kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies, and used the forecast errors as well as the accuracy of the direction of the forecasts to evaluate the performance of three Neural Network models. In a similar study, @khashei2012 compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN) to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark by claiming that it is the most important linear model. Their comparison was made by investigating the forecasts between the British pound and US dollar over varying forecast horizons. Not only has the ARIMA model been useful as a benchmark, but it has also yielded satisfactory results when predicting exchange rates. When @nwankwo2014 forecasted the rate between the Nairo and dollar, diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on Akaike’s Information Criterion (AIC).

The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001]. Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data. Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings. The most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).

The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing. @etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality. Their results showed that the SARIMA model adequately described the variation in the exchange rate series. The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain. The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993]. This makes the model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993]. @cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.

Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed in an attempt to capture the volatility in financial markets [@anastasakis2009]. Hsieh (1989) applied a GARCH model to investigate five exchange rates and his results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.

## Other techniques used to forecast exchange rates
Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques. These more complex techniques are non-linear and can fit complex time series more easily [@castillo2002]. Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002]. Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.

Artificial Neural Networks (ANNs) have had many successful applications in forecasting exchange rates and are more advantageous than other non-linear forecasting methods. They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011]. In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates. Superior results were produced when compared to the ARIMA and GARCH model using the root mean square error (RMSE), mean square error (MSE) and mean absolute difference as a measure of performance. Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms, and have been successfully used in financial forecasting [@khashei2009]. Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather than discrete decisions to be provided. @santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model. He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one step-ahead forecasts. By analysing accuracy statistics, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced. Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and a Fuzzy ARIMA model.

Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model. Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998]. Like ANNs there is no systematic approach to designing FLSs and they are only understandable when simple. Although FLSs has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.

## Hybrid ARIMA Models
Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA model, and of more complex non-linear models. By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations. Some of the hybrid models which have commonly been used in exchange rate forecasting are the Fuzzy ARIMA and ANN-ARIMA model.

The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data to yield accurate results [@tseng2001]. In a world that is constantly changing and with the rapid advancement of technology, access to large amounts of historical data is difficult to obtain. On the contrary, a fuzzy regression model requires little historical data however produces wide prediction intervals if extreme values are present in the data. The Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to exploit the advantages of both models while simultaneously overcoming their limitations. @tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model’s appropriateness and power. The Fuzzy ARIMA not only produced forecasts that were superior to the ARIMA and fuzzy time series models but also provided an upper and lower bound which can be used by decision makers to determine the best and worst possible situations.

Although ARIMA models are powerful, they require non-stationary data to be differenced and impose prior assumptions onto the distribution of the data [@ince2006]. In contrast, machine learning techniques such as ANNs do not impose any assumptions onto the data generating process however, being data-driven, are sensitive to the number of input nodes used. The ANN-ARIMA model draws on the strengths of both models and overcomes these individual difficulties. @ince2006 created an ANN-ARIMA model by using the ARIMA model to determine the number input nodes required by an ANN for three exchange rates. When the ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model outperformed the ARIMA based on the MSE.

Although the traditional ARIMA models produce less superior forecasts than its hybrid forms and other complex non-linear techniques, its forecasts are still satisfactory. They are simple models that are easy to implement and have a consistent method of model design and selection. ARIMA models are also more robust and efficient than complex structural models in relation to short-run forecasting. The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison, justifies it as a good starting point to compare it to Facebook’s forecasting method, Prophet, that was recently released.

## Forecasting with Prophet 
The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series. Prophet differs to traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process. Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared. This is Facebook’s definition of forecasting at scale.
Prophet is similar to a Generalized Additive Model (GAM) - an additive regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017]. The decomposable model is of the form:

$$ 
	y(t) = g(t) + s(t) + h(t) + e_t
$$
where the components of the model represent the growth, seasonality and holiday respectively, and $e_t$ is white noise.

Prophet, like the GAM, frames the forecasting problem as a curve fitting exercise and uses backfitting to find the regression functions. This allows for the model to be fitted quickly and missing values and large outliers to be handled elegantly. The regression model also provides model flexibility and allows the analyst to interactively change model parameters [@taylor2017].

The growth component of Prophet may be modelled as a linear or non-linear function of time. Linear growth is modelled by a piecewise constant function while non-linear growth is modelled similar to population growths which use a logistic growth model [@taylor2017]. A time-varying upper limit may be specified for logistic growth, at which point the forecasts will saturate. This carrying capacity allows the analyst to incorporate their prior knowledge about the maximum obtainable growth level such as the total market or population size into the model. Prophet accounts for changes in the trajectory of this trend by automatically detecting and selecting changepoints in the data at which the growth rate is allowed to change. These changepoints have a Laplace prior distribution placed on them and its scale parameter may be used to adjust the flexibility of the trend and to choose how aggressively the model should follow historical trend changes. Analysts may also adjust the number of potential changepoints included or manually specify their location. This allows non-experts with knowledge about events that may affect the growth rate to use the parameter as a knob to either increase or decrease the number of changepoints included [@taylor2017]. It also allows for the analyst to add changepoints which the automatic selection procedure may have missed or remove changepoints when the model is overfitting historical trends [@taylor2017].

The decomposable form of the model allows for multiple seasonality components with different periods to be added to the model. Seasonality components are modelled by a Fourier series and have a Normal prior distribution placed on its parameters [@taylor2017]. The spread parameter can be adjusted by analysts to smooth the seasonality and change how much of historical seasonality is projected into the future [@taylor2017].

The analyst may also provide a list of important events and holidays which have impacted the time series in the past or which they know might impact it in the future. The list could include the name, date, and country in which they have taken place or are expected to take place [@taylor2017]. By specifying the country of occurrence, separate lists can be populated for global events and holidays, and country-specific events and holidays. The union of the global and country-specific lists can then be used for forecasting. Like seasonality, a Normal prior distribution is placed on the parameters of the holiday component, and the scale parameter can be adjusted by analysts to smooth the holidays [@taylor2017].

Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge into the model building process and has produced significantly improved forecasts compared to the ARIMA model. @taylor2017 forecasted the number of events on Facebook using Prophet. The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend. The forecasts produced by Prophet were compared to common forecasting techniques such as exponential smoothing, ARIMA, the seasonal naïve, and the naïve model. While exponential smoothing and the seasonal naïve model were quite robust, the ARIMA forecasts were fragile. No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.

Prophet’s ability to forecast at scale enables it to model a wide variety of data and may be able to adequately fit exchange rate data. Its non-linear components could capture the non-linearity present in exchange rates. In contrast, the ARIMA model is a linear function of previous observations and lagged residuals and fails to capture the non-linearity inherent in exchange rate data. Prophet performs well on data with strong multiple “human scale” seasonalities and historical trend changes. This differs to the ARIMA model which requires the data to be de-trended and the variance stabilised before the model can be fitted. Hence Prophet could model the weekly seasonality of closing prices due to low trading activity which occurs around the weekend and high trading activity which occurs mid-week.

Choosing the correct combination of parameters for the ARIMA model is a challenging task due to the array of possible choices. Although the auto.arima function in R may be used to automatically select an ARIMA model that best fits the data, completely automatic forecasting methods are too brittle and do not allow for useful assumptions to be incorporated into the model. Prophet makes use of a semi-automatic forecasting technique that keeps the analyst-in-the-loop. Its default settings are said to generate forecasts that are as accurate as those produced by skilled forecasters. If the forecasts produced are unsatisfactory, they can be improved by the analyst by configuring the model through its easily interpretable parameters. Furthermore, non-experts who have domain knowledge about factors that affect Bitcoin or if the dates of events which could impact the price of Bitcoin are known, it may be incorporated into the model by the analyst. Prophet can produce forecasts over irregular time intervals and allows for missing values in the time series without the need for interpolation [@taylor2017]. The ARIMA model on the other hand requires large outliers to be removed and handles missing values by interpolation. If Prophet produces forecasts that are as good as the ARIMA model when forecasting BTC/ZAR, it can be compared to more complex forecasting methods that are less tractable such as the hybrid models and machine learning techniques seen earlier.

# Data \label{Data}

The data set used in this study consists of the daily closing prices of BTC/ZAR over the period 24 January 2016 to 17 July 2017. This comprises of a total of 541 trading days and was the chosen time period for analysis due to constraints in obtaining data over a longer period. The data was obtained from Bitcoincharts.

Bitcoin is of specific interest as it presents an interesting parallel to traditional exchange rate markets. The cryptocurrency is built on a decentralised system and as a result its value cannot be directly influenced by a central authority (Fantazzini and Nigmatullin 2016).  In addition, the Bitcoin market has attracted attention worldwide and is currently the leading cryptocurrency, with awareness and adoption of the currency growing over time (Fantazzini and Nigmatullin 2016). Its novelty makes it a highly volatile hence speculative market and provides an opportunity for forecasting.

In this study, the analysis of BTC/ZAR forecasts are based on the daily continuous log returns. The daily closing prices are transformed into returns by taking the log difference at each time t and is calculated as follows:

$$ r_t = ln(\frac{p_t}{p_{t-1}}) $$
where $p_t$ represents the closing price and $r_t$ represents the log return for time t = 1,2,.,T.

# Methodology \label{Methodology}
In this paper, the out-of-sample forecasts of ARIMA and Prophet are compared over varying forecast horizons. The data is first split into a training set and test set. The training set starts on the 24th January 2016 and ends on the 16th January 2017, while the test set starts on the 17th January 2017 and ends on the 17th July 2017. The Box-Jenkins methodology is then used to select the correct ARIMA model while Prophet’s semi-automatic procedure is used for model selection. The selected ARIMA and Prophet model are then fitted to the training set and rolling window forecasts are made 1 day, 30 days and 90 days ahead. This allows us to examine the forecast horizon effect. The Mincer-Zarnowitz test is then used to evaluate the forecasts produced by ARIMA and Prophet against the test set. Since there is no consensus on which accuracy statistic best measures the performance of forecasting techniques, the most popular criteria, the Mean Squared Error (MSE) is employed. The results obtained from the ARIMA model will be used as a benchmark for comparison. This process is then repeated based on data which has outliers and missing values present.

## The Box-Jenkins Methodology 
Box and Jenkins proposed a set of guidelines that can be followed when selecting ARIMA models. This systematic procedure to designing ARIMA models has made them highly popular [@spyros1997]. It consists of a four-stage iterative process in which:

1. The process is either transformed or differenced to de-trend and stabilise the variance of the data.

2. The autocorrelation and partial autocorrelation plots are used to determine the order of p and q. 
3. The parameters of the model are then estimated.

4. A diagnostic check is performed to ensure that the residuals are a white noise process.

If the residuals are not white noise, steps 2-4 are repeated until a satisfactory model is identified. On the contrary, if the diagnostic check shows that the residuals are random, the developed model will be the final model used for forecasting.

## Semi-Automatic Selection \label{Semi-Automatic Selection}
When a large number of forecasts are produced, manually identifying problematic forecasts becomes a time consuming and difficult task. Prophet provides a semi-automated forecast evaluation system that selects the best model which fits the data. When there are large forecast errors, the forecasts are flagged so that the analyst can explore the cause of the errors, identify and remove potential outliers and either adjust the model or choose a more appropriate model [@taylor2017]. This keeps the analyst-in-the-loop. Prophet has the following default settings:

* The trend is set to be linear
* The width of the uncertainty intervals is set to 80% 
* Weekly and yearly seasonality are automatically detected and included in the model if present 
* The smoothing parameter for holidays and seasonality is set at 10 
* The smoothing parameter for the trend is set to 0.05
* The number of potential changepoints is set to 25

This paper makes use of Prophet’s default settings to fit the data, however weekly seasonality is included in the model. This allows us to account for our prior knowledge about the closing prices of Bitcoin which tend to be lower around the weekend however higher mid-week, due to fluctuations in trading activity. Furthermore, a linear trend is appropriate since Bitcoin does not have an upper limit on its closing prices.

## Rolling Window Forecasts
Rolling window forecasts are useful in evaluating the robustness of a forecasting method. In a rolling window forecast, the forecasts are made h-steps at a time and the actual observation rather than the predicted value is used for the next prediction in the forecast horizon. In this way, a poor forecast will not have negative consequences on future forecasts to be made, since the observed values will be used to correct itself for the remaining forecasts [@zhang1998]. In this paper, we first fit the ARIMA and Prophet model to the training set and forecast 1 day ahead. We then refit the model to the training set, increasing its size by one observation, and the next 1 day ahead forecast is made. This process is repeated until all forecasts have been made into the test set. The 1 day ahead forecasts are then compared to the 1 day ahead observed values in the test set and the error statistics are calculated. This procedure is then repeated for the 30 day and 90 day ahead forecasts.

## Mincer-Zarnowitz Approach to Forecast Evaluation 
The Mincer-Zarnowitz approach to evaluating forecast accuracy is commonly used and can be useful when comparing the forecasts produced by Prophet and the ARIMA model. @mincer1969 proposed an absolute measure to evaluate forecast accuracy which considers the distance between the actual and predicted values. To analyse the absolute errors produced by the forecasts, the observed values $r_t$ are regressed against the predicted values $r_t$, i.e.

$$r_t = \alpha + \beta\cdot\hat{r_t} + e_t$$

where $\alpha$ represents the mean distance between the observed and predicted values while $\beta$ represents the correlation between the forecasted errors and the predicted values. When alpha = 0 it implies that the forecasts are unbiased and do not systematically overestimate or underestimate the data. When b=1 it implies that the forecasts are efficient and uncorrelated with the forecasted errors. A joint hypothesis test of $H_0: \alpha = 0 \cup \beta = 1$ is performed to check the efficiency and bias of the forecasts and the model which produces the best results generates superior forecasts.

## Evaluation of forecasts based on data with missing values and outliers
The original dataset is modified in order to examine how robust Prophet and ARIMA are when fitted to data with missing values and outliers. 20 values are randomly removed from the training set and 5 outliers are randomly inserted in place of existing values. The models are then fitted to this “dirty” dataset and evaluating of forecasts are carried out as before.

# Results \label{Results}

## Evaluation over varying forecast horizons
Although the Box-Jenkins methodology resulted in an ARIMA(0,0,1) model being the best fit to the training set, an ARIMA(1,0,0) is employed as the benchmark for comparison so that a rolling window forecast may be conducted. The AIC of the ARIMA(1,0,0) is marginally higher than that of the ARIMA(0,0,1) and analysis of the p-value indicates that the AR parameter is statistically significant in the explanation of the movement of BTC/ZAR. The Prophet model fitted to the training set included a linear trend and weekly seasonality, with all other parameters set to the default values as described in section \ref{Semi-Automatic Selection}.  The automatic-selection procedure for ARIMA and Prophet selected the same model as those chosen manually.

### 1-step ahead Forecasts
Figure 1 depicts the resulting forecasts from fitting both models. The forecasts produced by the ARIMA model and Prophet lie within a band centred around zero. Both models fail to capture the large movement of returns away from zero. The 80% uncertainty intervals produced by both models are approximately the same width and begin to widen as forecasts are made further into the test set.

![1-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\1-step ahead forecasts.png)

Table 1 displays the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the one-step ahead forecasts. The intercept estimate is small and positive for ARIMA and Prophet, with the intercept for ARIMA being closer to zero than Prophet. This indicates that both models systematically underestimate the 1-step ahead returns, with ARIMA yielding mean forecasts closer to the observed mean. The slope estimate is negative for both models indicating that the observed returns are actually the opposite of the suggested forecast. The F-test for the joint hypothesis of a unity slope and zero intercept is rejected at the 5% significance level for both models, with the p-value for ARIMA being much larger than Prophet. Hence the forecasts generated by both ARIMA and Prophet are biased and/or inefficient. The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.

![Mincer-Zarnowitz results for 1-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 1.png)

### 30-steps ahead Forecasts
Figure 2 illustrates the resulting 30-step ahead forecasts produced from fitting ARIMA and Prophet. As in the 1-step ahead case, both models fail to capture the large movement of returns away from zero with its forecasts lying within a band centred around zero. The 80% uncertainty intervals produced by both models are approximately the same width and marginally wider than the uncertainty intervals produced by the 1-step ahead forecasts.

![30-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\30-step ahead forecasts.png)

Table 2 displays the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 30-step ahead forecasts. The intercept estimate is small and positive for ARIMA and Prophet, with a smaller intercept estimated for Prophet. This indicates that both models systematically underestimate the 30-step ahead returns, with Prophet yielding mean forecasts closer to the observed mean. The slope estimate is large and negative for both models with the slope estimated to be as low as -12.74 for ARIMA. This suggests that the observed returns are actually the opposite of the proposed forecast. The F-test for the joint hypothesis of a unity slope and zero intercept is rejected at the 5% significance level for Prophet only. The p-value of 0.0614 from the joint hypothesis test is slightly significant for ARIMA. Hence the forecasts generated by Prophet are biased and/or inefficient while the forecasts generated by ARIMA are unbiased and/or efficient. The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.

![Mincer-Zarnowitz results for 30-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 2.png)

### 90-steps ahead Forecasts
Figure 3 illustrates the 90-step ahead forecasts produced by ARIMA and Prophet. Both models fail to capture large movement of returns away from zero as in the case of the 1-step ahead and 30-step ahead forecasts. The 80% uncertainty intervals produced by both models are approximately the same width and are marginally narrower than the uncertainty intervals produced by the one-step ahead and 30-step ahead forecasts.

![90-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\90-steps ahead forecasts.png)

Table 3 presents the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 90-step ahead forecasts. The intercept estimate is very small and positive for ARIMA and Prophet, with a marginally smaller intercept estimated for ARIMA. This indicates that both models systematically underestimate the 90-step ahead returns, with ARIMA yielding mean forecasts closer to the observed mean. The slope estimate is small and negative for both models with the slope estimated for ARIMA lying closer to -1. This suggests that the observed 90-steps ahead returns are actually the opposite of the proposed 90-steps ahead forecasts. The F-test for the joint hypothesis of a unity slope and zero intercept fails to be rejected at the 5% significance level for both ARIMA and Prophet, with the p-value from the test being much larger for ARIMA. Hence the forecasts generated by Prophet and ARIMA are unbiased and/or efficient. The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.

![Mincer-Zarnowitz results for 90-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 3.png)

### Comparison over varying forecast horizons
Table 4 and 5 compares the p-values obtained from the Mincer-Zarnowitz tests and the MSE of the forecasts produced by ARIMA and Prophet over the three forecast horizons. In comparing the forecasting accuracy of the ARIMA model over the three forecast horizons, we observe that the p-value gets more significant as the forecast horizon gets longer. This is in contrast to the MSE of the ARIMA forecasts which is smaller at the shortest and longest forecast horizons. In comparing the forecasting accuracy of Prophet, we find that the p-value is more significant for the shortest and longest forecast horizon. This agrees with the MSE of the Prophet forecasts which is smaller at the longest and shortest forecast horizon. Both ARIMA and Prophet generate forecasts with the smallest MSE and the most significant p-value at the longest forecast horizon.
In comparing the forecasting accuracy of ARIMA and Prophet over the three forecast horizons, we notice that the p-value obtained from the Mincer-Zarnowitz results is more significant for ARIMA than Prophet at every forecast horizon. This echoes the results obtained from the MSE which is marginally smaller for ARIMA than Prophet over all forecast horizons.

![Mincer-Zarnowitz results for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 4.png)

![MSE for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 5.png)

## Evaluation on data with missing values and outliers present
By following the Box-Jenkins methodology for selection of an ARIMA model, an ARIMA(3,0,0) was employed to fit the training set which included missing values and outliers. Although the ARIMA(0,0,3) had a slightly smaller AIC, a decision was made to use the ARIMA(3,0,0) instead so that a rolling window forecast may be used. Analysis of the AR parameters indicate that they are statistically significant in the explanation of the movement of BTC/ZAR. The Prophet model fitted to the training set included a linear trend and weekly seasonality. All other parameters were set to Prophet’s default values as described in section \ref{Semi-Automatic Selection}.  The automatic-selection procedure for Prophet and ARIMA selected the same models as those chosen manually.

### 1-step ahead Forecasts
Table 6 shows the MSE of the forecasts generated by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 1-step ahead returns forecasted. We observe that a positive intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically underestimate the 1-step ahead returns. The intercept estimated for ARIMA is smaller than Prophet, indicating that the mean forecasted returns yielded by ARIMA is closer to the mean observed returns. The intercept estimates of 0.011895 and 0.014802 for ARIMA and Prophet respectively are larger than the intercept estimates of 0.009546 and 0.013702, which was based on the “clean” dataset.  This suggests that both models systematically underestimates the returns more when there are missing values and outliers present in the data.

We also observe that the slope estimate is small and negative for both models with the slope estimated for ARIMA lying closer to -1. This suggests that the observed 1-step ahead returns are actually the opposite of the proposed 1-step ahead forecasts. The slope estimates of -0.883228 and -0.676132 for ARIMA and Prophet respectively are less steep than the slope estimates of -2.005624 and -1.709744, which was based on the evaluation of data that had no values missing or outliers present.

![Mincer-Zarnowitz results for 1-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 6.png)

The p-values obtained from the F-test for the joint hypothesis of a unity slope and zero intercept is approximately zero for ARIMA and Prophet, with the p-value for Prophet lying closer to zero than ARIMA. The null hypothesis is thus rejected at the 1% significance level, indicating that the forecasts generated by Prophet and ARIMA are biased and/or inefficient. The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet. The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on the “clean” dataset show similar findings, however led to a smaller MSE of 0.001942 and 0.001958 and a more significant p-value of 0.025449 and 0.007152 for ARIMA and Prophet respectively.

### 30-steps ahead Forecasts
Table 7 shows the MSE of the forecasts produced by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 30-step ahead returns forecasted. We observe that a positive intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically underestimate the 30-step ahead returns. The intercept estimated for ARIMA is larger than Prophet, indicating that the mean forecasted returns yielded by Prophet is closer to the mean observed returns. The intercept estimate of 0.11414 for ARIMA is larger than the estimate of 0.031320 which was based on the evaluation of the “clean” dataset. This suggests that the ARIMA model underestimates the returns more when there are missing values and outliers present in the data.  The intercept estimate of 0.014396 for Prophet is less than the estimate of 0.017169 which was based on the evaluation of data that had no values missing or outliers present. This suggests that Prophet underestimates the returns less when there are missing values and outliers present in the data.

![Mincer-Zarnowitz results for 30-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 7.png)

We also observe that the slope estimate is large and negative for ARIMA and small and negative for Prophet. This suggests that the observed 30-step ahead returns are actually the opposite of the proposed 30-step ahead forecasts. The slope estimate of -12.985335 for ARIMA is more steep than the slope estimate of -12.736757 which was based on the evaluation of data that had no values missing or outliers present.  This suggests that the efficiency of the forecasted returns deteriorates when there are missing values and outliers present in the data. The slope estimate of -0.65161 for Prophet is less steep than the estimate -2.621914, which was based on the evaluation of data that had no values missing or outliers present.  This suggests that the efficiency of the forecasted returns improve when there are missing values and outliers present in the data.

The p-value obtained from the F-test for the joint hypothesis of a unity slope and zero intercept is approximately zero for Prophet. The null hypothesis is thus rejected at the 1% significance level, indicating that the 30-step ahead forecasts generated by Prophet are biased and/or inefficient. The p-value of 0.107532 for ARIMA is significant at the 5% level, suggesting at the 30-step ahead forecasts produced by ARIMA are unbiased and/or efficient. MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a smaller MSE than Prophet. The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on the “clean” dataset, led to a more significant p-value of 0.000377 for Prophet only, and a smaller MSE of 0.001988 and 0.002075 for ARIMA and Prophet respectively.

### 90-steps ahead Forecasts
Table 8 shows the MSE of the forecasts produced by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 90-step ahead returns forecasted. We observe that a negative intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically overestimate the 90-step ahead returns. The intercept estimated for Prophet lies closer to zero, indicating that the mean forecasted returns yielded by Prophet is closer to the mean observed returns. This is in contrast to the positive intercepts that were estimated based on the data that had no values missing or outliers present.

We also observe that the slope estimate is large and positive for ARIMA however small and positive for Prophet. This is in contrast to the small and negative slope estimates which were based on the evaluation of data that had no values missing or outliers present.

![Mincer-Zarnowitz results for 90-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 8.png)

The p-values obtained from the F-test for the joint hypothesis of a unity slope and zero intercept are insignificant for ARIMA and Prophet, with the p-value for Prophet lying close to zero. The null hypothesis is thus rejected at the 5% significance level, indicating that the 90-steps ahead forecasts generated by Prophet and ARIMA are biased and/or inefficient. The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet. The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on data which had no missing values or outliers present contrasts these findings, with a p-value of 0.843066 and 0.408675 for ARIMA and Prophet respectively.

### Comparison over varying forecast horizons
Table 9 and 10 compares the p-values obtained from the Mincer-Zarnowitz tests and the MSE of the forecasts produced by ARIMA and Prophet over the three forecast horizons. In comparing the forecasting accuracy of the ARIMA model over the three forecast horizons, we observe that the p-value is more significant at the longer forecast horizons. This agrees with the MSE of the ARIMA forecasts which is smaller at the longer forecast horizons. In comparing the forecasting accuracy of Prophet, we find that the p-value is more significant for the longest forecast horizon. This agrees with the MSE of the Prophet forecasts which is smaller at the longest forecast horizon.

![Mincer-Zarnowitz results for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 9.png)

![MSE for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 10.png)

In comparing the forecasting accuracy of ARIMA and Prophet over the three forecast horizons, we notice that the p-value obtained from the Mincer-Zarnowitz results is more significant for ARIMA than Prophet at every forecast horizon. This echoes the results obtained from the MSE which is marginally smaller for ARIMA than Prophet over all forecast horizons.

# Discussion and Conclusions \label{Discussion and Conclusions}
The results show a clear difference in the effect of the forecast horizon in the case of the clean data. The results based on the MSE and the p-value obtained from the Mincer-Zarnowitz test surprisingly show that the ARIMA model produces superior forecasts of the returns at the longest forecast horizon. Similarly, Prophet produces superior forecasts of the returns at the longest forecast horizon. We notice that although the ARIMA model outperforms Prophet at every forecast horizon, the difference between the MSE’s of the forecasts generated by both models is insignificant. Furthermore, the Mincer-Zarnowitz p-values for both models are significant at the same forecast horizons, suggesting that the efficiency and biasness of the return forecasts yielded by ARIMA and Prophet are synchronized at each forecast horizon. If a small amount of predictive accuracy is willing to be sacrificed by the analyst in the forecasting of BTC/ZAR, it might be useful to employ Prophet instead of ARIMA. This would enable traders with expert knowledge about Bitcoin, however no statistical knowledge about forecasting models, to incorporate their domain knowledge into the Prophet model through its intuitively adjustable parameters.  Furthermore, as more information about events that may affect the price of Bitcoin becomes available, it may be included in the model and could potentially change the predictive accuracy of Prophet.

The results based on the unclean data show that both ARIMA and Prophet are sensitive to the presence of outliers and missing values in addition to the forecast horizon. We observe that Prophet yields superior forecasts at the longest forecast horizon. Although the MSE suggests that the ARIMA model produces forecasts with minimal error at the longest forecast horizon, analysis of the p-value indicates that the forecasts are efficient and/or unbiased at 30-steps ahead only. It is evident that Prophet produces inferior results in comparison to the ARIMA model over all forecast horizons. This is unexpected as Prophet is said to be robust to the presence of outliers and missing values in the data, without the need for interpolation. We also notice that the forecasts produced by both models deteriorates when there are missing values and outliers present in the data, with the only efficient and/or unbiased forecasts yielded by the ARIMA model at a forecast horizon of 30 days.

The limitation of this study is that it looks specifically at BTC/ZAR. Different results may emerge for fiat currencies or the exchange rate between Bitcoin with other currencies. Hence caution should be exercised when generalising these results to other exchange rates. As a point of further research, more complex techniques such as neural networks could be used as a benchmark for comparison when evaluating the forecasts produced by Prophet.

\newpage

# References

### 2017-09-14 Draft/Write-up/Draft.tex
Final Draft. Sigh

--- 2017-09-14 Draft/Write-up/Draft.Rmd
@@
----
-title: "How does the latest Prophet forecasting model compare to the ARIMA model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz approach"
-Author: "Fiona Ganie"  
-Student: "GNXFIO001"
-bibliography: Tex/ref.bib       
-output:
-  pdf_document:
-    keep_tex: TRUE
-    template: Tex/TexDefault.txt
-    fig_width: 5 
-    fig_height: 4
-abstract: |
-  The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has gained popularity in the exchange rate market due to its ease of implementation and tractability.
-With the evolution of computational power, soft computing techniques have since been used in exchange rate markets.
-The ARIMA model has been used as the standard benchmark model against which these complex methods have been compared.
-Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task.
-Prophet is a sophisticated model that differs to traditional time series models in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process.
-Prophet is also robust to outliers and can handle missing values in the time series without the need for interpolation.
-This paper compares the out-of-sample forecasts of Prophet and ARIMA over varying forecast horizons, using a Mincer-Zarnowitz approach for forecast evaluation.
-A complete dataset of the prices of Bitcoin/Rand as well as a dataset which has missing prices and outliers present is used in the analysis.
-It is found that the ARIMA model produces the most efficient and unbiased forecasts over all forecast horizons, and that the ARIMA model is more robust when faced with missing values and outliers present in the data.
----
+\documentclass[12pt,a4paper]{article}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
+\usepackage{url}
+\usepackage{appendix}
+\usepackage[british]{babel}
+\usepackage{amsmath}
+\usepackage{hyperref}
+\usepackage{graphicx}
+\numberwithin{equation}{section}
+\numberwithin{figure}{section}
+\numberwithin{table}{section}
 
-```
+\usepackage[round]{natbib}
+\bibliographystyle{natbib}
+\def\bibsection{\section*{References}}
 
-# Introduction \label{Introduction}
-The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has attracted attention in financial market forecasting [@khashei2009].
-Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that foreign exchange markets are not efficient and believe that future prices depend on current and past events [@abu1996].
-This has led to the application of the ARIMA model to exchange rate problems, where it has since gained popularity due to its ease of implementation and tractability.
-The ARIMA model is easy to use and yields good forecasts over short forecast horizons, however the linearity of the ARIMA model fails to adequately capture the non-linearity inherent in exchange rate data [@zhang1998].
+% Wrap around which gives all figures included the [H] command, or places it "here". This can be tedious to code in Rmarkdown.
+\usepackage{float}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[H]
+} {
+    \endorigfigure
+}
 
-With the evolution of computational power, non-linear, soft computing techniques have been proposed as a solution.
-The ARIMA model has been used as the standard benchmark model against which these more complex methods have been compared.
-Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task.
-The ARIMA model is tractable and less computationally expensive.
-It has been used as the building blocks for more advanced models and has provided the inspiration for hybrid versions of the model which have been used in exchange rate forecasting.
+\let\origtable\table
+\let\endorigtable\endtable
+\renewenvironment{table}[1][2] {
+    \expandafter\origtable\expandafter[H]
+} {
+    \endorigtable
+}
+
+\def\tightlist{}
+
+\bibpunct[:]{(}{)}{;}{a}{,}{,}
+
+\renewcommand{\baselinestretch}{1.5}
+
+\begin{document}
+
+\begin{titlepage}
+
+\begin{center}
+{\Huge \bf How does the latest Prophet forecasting model compare to the ARIMA
+model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz
+approach}\\
+\today\\
+Fiona Ganie (GNXFIO001)\\
+{\tt GNXFIO001@myuct.ac.za}
+\end{center}
+
+\begin{abstract}
+The Autoregressive Integrated Moving Average (ARIMA) model is one of the
+most widely used time series models that has gained popularity in the
+exchange rate market due to its ease of implementation and tractability.
+With the evolution of computational power, soft computing techniques
+have since been used in exchange rate markets. The ARIMA model has been
+used as the standard benchmark model against which these complex methods
+have been compared. Although they have produced significantly better
+results than the ARIMA model, these models lack interpretability and
+building them is a challenging task. Prophet is a sophisticated model
+that differs to traditional time series models in that it can be
+utilised by non-experts who have little knowledge about the statistical
+intricacies involved in the model, however have domain knowledge about
+the data generating process. Prophet is also robust to outliers and can
+handle missing values in the time series without the need for
+interpolation. This paper compares the out-of-sample forecasts of
+Prophet and ARIMA over varying forecast horizons, using a
+Mincer-Zarnowitz approach for forecast evaluation. A complete dataset of
+the prices of Bitcoin/Rand as well as a dataset which has missing prices
+and outliers present is used in the analysis. It is found that the ARIMA
+model produces the most efficient and unbiased forecasts over all
+forecast horizons, and that the ARIMA model is more robust when faced
+with missing values and outliers present in the data.
+\noindent
+Keywords: Prophet, ARIMA, Mincer-Zarnowitz,   Box-Jenkins, Bitcoin.
+\end{abstract}
+\end{titlepage}
+
+\pagenumbering{arabic}
+
+\section{\texorpdfstring{Introduction
+\label{Introduction}}{Introduction }}\label{introduction}
+
+The Autoregressive Integrated Moving Average (ARIMA) model is one of the
+most widely used time series models that has attracted attention in
+financial market forecasting (Khashei, Bijari, and Ardali 2009).
+Although the Random Walk model has typically been applied to foreign
+exchange markets and has produced superior results, some researchers
+have contended that foreign exchange markets are not efficient and
+believe that future prices depend on current and past events
+(Abu-Mostafa and Atiya 1996). This has led to the application of the
+ARIMA model to exchange rate problems, where it has since gained
+popularity due to its ease of implementation and tractability. The ARIMA
+model is easy to use and yields good forecasts over short forecast
+horizons, however the linearity of the ARIMA model fails to adequately
+capture the non-linearity inherent in exchange rate data (Zhang and Hu
+1998).
+
+With the evolution of computational power, non-linear, soft computing
+techniques have been proposed as a solution. The ARIMA model has been
+used as the standard benchmark model against which these more complex
+methods have been compared. Although they have produced significantly
+better results than the ARIMA model, these models lack interpretability
+and building them is a challenging task. The ARIMA model is tractable
+and less computationally expensive. It has been used as the building
+blocks for more advanced models and has provided the inspiration for
+hybrid versions of the model which have been used in exchange rate
+forecasting.
 
 In December 2016, Facebook open-sourced their forecasting model Prophet.
-Prophet differs to traditional time series model such as the ARIMA, in that it can produce high quality forecasts in a straightforward way.
-Prophet is a sophisticated model that provides informative results, however is configurable and easy to use.
-It can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process.
-This knowledge can be easily incorporated into the model through its intuitively adjustable parameters [@taylor2017].
-Furthermore, Prophet is robust to outliers and can handle missing values in the time series without the need for interpolation [@taylor2017].
-This allows analysts without knowledge on how to pre-process the data, to utilise the model without sacrificing predictive accuracy.
-If Prophet produces forecasts that are as good as the ARIMA model, it can be compared to more complex forecasting methods that are less tractable and flexible.
+Prophet differs to traditional time series model such as the ARIMA, in
+that it can produce high quality forecasts in a straightforward way.
+Prophet is a sophisticated model that provides informative results,
+however is configurable and easy to use. It can be utilised by
+non-experts who have little knowledge about the statistical intricacies
+involved in the model, however have domain knowledge about the data
+generating process. This knowledge can be easily incorporated into the
+model through its intuitively adjustable parameters (Taylor and Letham
+2017). Furthermore, Prophet is robust to outliers and can handle missing
+values in the time series without the need for interpolation (Taylor and
+Letham 2017). This allows analysts without knowledge on how to
+pre-process the data, to utilise the model without sacrificing
+predictive accuracy. If Prophet produces forecasts that are as good as
+the ARIMA model, it can be compared to more complex forecasting methods
+that are less tractable and flexible.
 
-This paper broadly aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of the exchange rate between Bitcoin and the Rand (BTC/ZAR), using a Mincer-Zarnowitz approach of measuring forecast accuracy.
-More specifically, this paper aims to:
+This paper broadly aims to compare the forecasts produced by the
+traditional ARIMA model and Prophet through an evaluation of the
+exchange rate between Bitcoin and the Rand (BTC/ZAR), using a
+Mincer-Zarnowitz approach of measuring forecast accuracy. More
+specifically, this paper aims to:
 
-1. Compare the out-of-sample forecasts of ARIMA and Prophet over varying forecast horizons.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Compare the out-of-sample forecasts of ARIMA and Prophet over varying
+  forecast horizons.
+\item
+  Compare the out-of-sample forecast performance with missing values and
+  outliers present in the data.
+\end{enumerate}
 
-2. Compare the out-of-sample forecast performance with missing values and outliers present in the data.
+Section \ref{Background and Theory} will present the key findings from
+the major works in which comparisons have been made to the ARIMA model
+in exchange rate forecasting, as well as the theory behind the ARIMA and
+Prophet model. Section \ref{Data} will provide a brief explanation and
+motivation for the dataset used in this paper. Section \ref{Methodology}
+will describe the methodology used in selecting the ARIMA and Prophet
+models and the Mincer-Zarnowitz approach of measuring forecast accuracy.
+The results from applying the ARIMA and Prophet model to the dataset is
+presented in Section \ref{Results}. Finally, Section
+\textbackslash{}ref\{Discussion and Conclusions discusses the results in
+a statistical context.
 
-Section \ref{Background and Theory} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting, as well as the theory behind the ARIMA and Prophet model.
-Section \ref{Data} will provide a brief explanation and motivation for the dataset used in this paper.
-Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring forecast accuracy.
-The results from applying the ARIMA and Prophet model to the dataset is presented in Section \ref{Results}.
-Finally, Section \ref{Discussion and Conclusions discusses the results in a statistical context.
+\section{\texorpdfstring{Background and Theory
+\label{Background and Theory}}{Background and Theory }}\label{background-and-theory}
 
-# Background and Theory \label{Background and Theory}
+\subsection{Forecasting Exchange Rates with the ARIMA
+model}\label{forecasting-exchange-rates-with-the-arima-model}
 
-## Forecasting Exchange Rates with the ARIMA model
-ARIMA models became highly popular since its introduction by Box and Jenkins, when it was shown that they could outperform complex econometric models in a variety of situations [@spyros1997].
-The ARIMA model expresses the process {$y_t$} as a function of the weighted average of past values of the process and lagged values of the residuals.
-The weighted average of the past p values of the process represents an autoregressive (AR) process of order p.
-It feeds back past values of the process into the current value, inducing correlation between all lags of the process.
-The weighted average of the q lagged residuals represents a moving average (MA) process of order q.
-The purpose of mixing the MA process with the AR process is to reduce the large number of past values required by AR processes and to control for the autocorrelation which it creates between lagged values of the process.
-The combination of the AR(p) and MA(q) process results in a more parsimonious model, and forms a stationary autoregressive moving average (ARMA(p,q)) process defined as:
+ARIMA models became highly popular since its introduction by Box and
+Jenkins, when it was shown that they could outperform complex
+econometric models in a variety of situations (Hibon and Makridakis
+1997). The ARIMA model expresses the process \{\(y_t\)\} as a function
+of the weighted average of past values of the process and lagged values
+of the residuals. The weighted average of the past p values of the
+process represents an autoregressive (AR) process of order p.~It feeds
+back past values of the process into the current value, inducing
+correlation between all lags of the process. The weighted average of the
+q lagged residuals represents a moving average (MA) process of order q.
+The purpose of mixing the MA process with the AR process is to reduce
+the large number of past values required by AR processes and to control
+for the autocorrelation which it creates between lagged values of the
+process. The combination of the AR(p) and MA(q) process results in a
+more parsimonious model, and forms a stationary autoregressive moving
+average (ARMA(p,q)) process defined as:
 
-$$ 
+\[ 
     y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
-$$
-where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.
+\] where c is a constant and \{\(e_t\)\} is a white noise process with
+zero mean and variance \(\sigma^2\).
 
-The ARIMA(p,d,q) model generalises the ARMA model in that it includes both stationary and non-stationary processes.
-The parameter d is the degree of differencing required to render the process stationary.
-If d is equal to zero the process is stationary and equivalent to an ARMA model, and if d is strictly positive the process requires differencing to make it stationary.
-The ARIMA model can be defined succinctly using the backward shift operator B, which shifts the process back by one unit of time, and is defined as $By_t = y_{t-1}$.
-The ARIMA model has the form:
+The ARIMA(p,d,q) model generalises the ARMA model in that it includes
+both stationary and non-stationary processes. The parameter d is the
+degree of differencing required to render the process stationary. If d
+is equal to zero the process is stationary and equivalent to an ARMA
+model, and if d is strictly positive the process requires differencing
+to make it stationary. The ARIMA model can be defined succinctly using
+the backward shift operator B, which shifts the process back by one unit
+of time, and is defined as \(By_t = y_{t-1}\). The ARIMA model has the
+form:
 
-$$ 
+\[ 
     (1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
-$$
-where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.
+\] where c is a constant and \{\(e_t\)\} is a white noise process with
+zero mean and variance \(\sigma^2\).
 
-ARIMA models have commonly been used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012].
-They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time.
-@kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies, and used the forecast errors as well as the accuracy of the direction of the forecasts to evaluate the performance of three Neural Network models.
-In a similar study, @khashei2012 compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN) to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark by claiming that it is the most important linear model.
-Their comparison was made by investigating the forecasts between the British pound and US dollar over varying forecast horizons.
-Not only has the ARIMA model been useful as a benchmark, but it has also yielded satisfactory results when predicting exchange rates.
-When @nwankwo2014 forecasted the rate between the Nairo and dollar, diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on Akaike’s Information Criterion (AIC).
+ARIMA models have commonly been used in financial forecasting and are
+popular for observing stock prices and exchange rates due to its power
+and statistical properties (C.-S. Lin, Chiu, and Lin 2012). They have
+frequently been used as a benchmark to compare new forecasting
+techniques that have emerged over time. Kamruzzaman and Sarker (2003)
+applied an ARIMA model to forecast the exchange rate between the
+Australian dollar and six other currencies, and used the forecast errors
+as well as the accuracy of the direction of the forecasts to evaluate
+the performance of three Neural Network models. In a similar study,
+Khashei, Bijari, and Ardali (2012) compared the predictive capability of
+their proposed hybrid model consisting of an ARIMA and a Probabilistic
+Neural Network (PNN) to the traditional ARIMA model, and justified the
+use of the ARIMA as a benchmark by claiming that it is the most
+important linear model. Their comparison was made by investigating the
+forecasts between the British pound and US dollar over varying forecast
+horizons. Not only has the ARIMA model been useful as a benchmark, but
+it has also yielded satisfactory results when predicting exchange rates.
+When Nwankwo (2014) forecasted the rate between the Nairo and dollar,
+diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit
+for the data based on Akaike's Information Criterion (AIC).
 
-The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001].
-Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data.
-Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings.
-The most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).
+The ARIMA model is attractive as it is tractable and produces good
+short-term forecasts when more than 100 observations are used (Tseng et
+al. 2001). Although the ARIMA model has the advantage of ease of
+implementation and flexibility, it fails to capture the non-linearity
+and volatility present in exchange rate data. Over time, the ARIMA model
+has evolved to cater for a wider variety of data and to compensate for
+some of its shortcomings. The most popular versions of the ARIMA model
+that has been implemented in exchange rate forecasting is the Seasonal
+ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).
 
-The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing.
-@etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality.
-Their results showed that the SARIMA model adequately described the variation in the exchange rate series.
-The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain.
-The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993].
-This makes the model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993].
-@cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.
+The SARIMA model was introduced to capture the periodic behaviour of
+data and extends the ARIMA class by including a term for seasonal
+differencing. Etuk, Wokoma, and Moffat (2013) modelled the Naira/CFA
+Franc exchange rate which exhibited monthly seasonality using an
+additive SARIMA model, to demonstrate that it can be a useful fit for
+exchange rate data which displays seasonality. Their results showed that
+the SARIMA model adequately described the variation in the exchange rate
+series. The ARFIMA model generalises the ARIMA model in that the degree
+required to make the data stationary can assume any real value, and is
+no longer restricted to the integer domain. The ARFIMA model has the
+ability to capture the dependence between observations that are widely
+spread apart in time (Cheung 1993). This makes the model more
+parsimonious since it can capture long memory in data as well as short
+term dynamics (Cheung 1993). Cheung (1993) fitted the ARFIMA model to
+examine five exchange rates, and found that there was strong evidence of
+long memory in the exchange rate time series.
 
-Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed in an attempt to capture the volatility in financial markets [@anastasakis2009].
-Hsieh (1989) applied a GARCH model to investigate five exchange rates and his results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.
+Generalised autoregressive conditional heteroskedasticity (GARCH) models
+were later developed in an attempt to capture the volatility in
+financial markets (Anastasakis and Mort 2009). Hsieh (1989) applied a
+GARCH model to investigate five exchange rates and his results showed
+that although the GARCH model outperformed the random walk, some
+non-linear information still remained in the residuals.
 
-## Other techniques used to forecast exchange rates
-Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques.
-These more complex techniques are non-linear and can fit complex time series more easily [@castillo2002].
-Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002].
-Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.
+\subsection{Other techniques used to forecast exchange
+rates}\label{other-techniques-used-to-forecast-exchange-rates}
 
-Artificial Neural Networks (ANNs) have had many successful applications in forecasting exchange rates and are more advantageous than other non-linear forecasting methods.
-They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011].
-In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates.
-Superior results were produced when compared to the ARIMA and GARCH model using the root mean square error (RMSE), mean square error (MSE) and mean absolute difference as a measure of performance.
-Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms, and have been successfully used in financial forecasting [@khashei2009].
-Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather than discrete decisions to be provided.
-@santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model.
-He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one step-ahead forecasts.
-By analysing accuracy statistics, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced.
-Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and a Fuzzy ARIMA model.
+Over time, financial forecasting methods have moved away from linear
+models like ARIMA and GARCH, to soft computing techniques. These more
+complex techniques are non-linear and can fit complex time series more
+easily (Castillo and Melin 2002). Unlike the ARIMA model, soft computing
+techniques do not impose structural assumptions on the model apriori
+(Castillo and Melin 2002). Some of the most commonly used artificial
+intelligence methods used to forecast exchange rate data are Neural
+Networks and Fuzzy Logistic Systems.
 
-Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model.
-Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998].
-Like ANNs there is no systematic approach to designing FLSs and they are only understandable when simple.
-Although FLSs has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.
+Artificial Neural Networks (ANNs) have had many successful applications
+in forecasting exchange rates and are more advantageous than other
+non-linear forecasting methods. They are data driven, can adapt to
+non-stationary environments and can approximate any continuous function
+(Khashei and Bijari 2011). In a study done by Fahimifard et al. (2009),
+the ANN was found as an effective way to improve the forecasts of
+exchange rates. Superior results were produced when compared to the
+ARIMA and GARCH model using the root mean square error (RMSE), mean
+square error (MSE) and mean absolute difference as a measure of
+performance. Fuzzy Logistic Systems (FLSs) were initially developed to
+solve problems involving linguistic terms, and have been successfully
+used in financial forecasting (Khashei, Bijari, and Ardali 2009). Fuzzy
+logic tries to imitate human reasoning and the decision-making process
+and allows for finer rather than discrete decisions to be provided.
+Santos, Costa, and Santos Coelho (2007) investigated how well FLSs and
+ANNs perform compared to the traditional ARMA and GARCH model. He
+examined the forecasts of Brazilian exchange rate returns by considering
+different frequencies of the series and comparing their one step-ahead
+forecasts. By analysing accuracy statistics, he found that FLSs and ANNs
+achieved higher returns based on the forecasts they produced. Similar
+results were found by Khashei, Bijari, and Ardali (2009) when he
+analysed the predictive capabilities of FLSs, ANNs, the traditional
+ARIMA model, and a Fuzzy ARIMA model.
 
-## Hybrid ARIMA Models
-Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA model, and of more complex non-linear models.
-By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations.
-Some of the hybrid models which have commonly been used in exchange rate forecasting are the Fuzzy ARIMA and ANN-ARIMA model.
+Although ANNs have been broadly applied in financial forecasting, the
+process of building them is a complex task and there is no consistent
+method of design compared to the traditional Box-Jenkins ARIMA model.
+Unlike ARIMA models, the performance of ANNs is sensitive to many
+modelling factors such as the number of input nodes included and the
+size of the training sample chosen (Zhang and Hu 1998). Like ANNs there
+is no systematic approach to designing FLSs and they are only
+understandable when simple. Although FLSs has the advantage over ARIMA
+models that they can be applied to data with few observations available,
+it gives acceptable rather than accurate results and are more suitable
+for problems which do not require high accuracy.
 
-The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data to yield accurate results [@tseng2001].
-In a world that is constantly changing and with the rapid advancement of technology, access to large amounts of historical data is difficult to obtain.
-On the contrary, a fuzzy regression model requires little historical data however produces wide prediction intervals if extreme values are present in the data.
-The Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to exploit the advantages of both models while simultaneously overcoming their limitations.
-@tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model’s appropriateness and power.
-The Fuzzy ARIMA not only produced forecasts that were superior to the ARIMA and fuzzy time series models but also provided an upper and lower bound which can be used by decision makers to determine the best and worst possible situations.
+\subsection{Hybrid ARIMA Models}\label{hybrid-arima-models}
 
-Although ARIMA models are powerful, they require non-stationary data to be differenced and impose prior assumptions onto the distribution of the data [@ince2006].
-In contrast, machine learning techniques such as ANNs do not impose any assumptions onto the data generating process however, being data-driven, are sensitive to the number of input nodes used.
-The ANN-ARIMA model draws on the strengths of both models and overcomes these individual difficulties.
-@ince2006 created an ANN-ARIMA model by using the ARIMA model to determine the number input nodes required by an ANN for three exchange rates.
-When the ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model outperformed the ARIMA based on the MSE.
+Over time, many researchers began to think of ways in which to harness
+the advantages of tractable linear models such as the ARIMA model, and
+of more complex non-linear models. By combining the ARIMA model with
+other forecasting methods, the advantages of both forecasting methods
+are leveraged while simultaneously improving their limitations. Some of
+the hybrid models which have commonly been used in exchange rate
+forecasting are the Fuzzy ARIMA and ANN-ARIMA model.
 
-Although the traditional ARIMA models produce less superior forecasts than its hybrid forms and other complex non-linear techniques, its forecasts are still satisfactory.
-They are simple models that are easy to implement and have a consistent method of model design and selection.
-ARIMA models are also more robust and efficient than complex structural models in relation to short-run forecasting.
-The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison, justifies it as a good starting point to compare it to Facebook’s forecasting method, Prophet, that was recently released.
+The ARIMA model produces very accurate forecasts over short time
+horizons however it has the limitation of requiring more than 100
+observations of historical data to yield accurate results (Tseng et al.
+2001). In a world that is constantly changing and with the rapid
+advancement of technology, access to large amounts of historical data is
+difficult to obtain. On the contrary, a fuzzy regression model requires
+little historical data however produces wide prediction intervals if
+extreme values are present in the data. The Fuzzy ARIMA model combines
+the ARIMA and fuzzy regression model to exploit the advantages of both
+models while simultaneously overcoming their limitations. Tseng et al.
+(2001) proposed applying a Fuzzy ARIMA model to forecast the exchange
+rate of Taiwan dollars to US dollars to demonstrate the model's
+appropriateness and power. The Fuzzy ARIMA not only produced forecasts
+that were superior to the ARIMA and fuzzy time series models but also
+provided an upper and lower bound which can be used by decision makers
+to determine the best and worst possible situations.
 
-## Forecasting with Prophet 
-The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series.
-Prophet differs to traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process.
-Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared.
-This is Facebook’s definition of forecasting at scale.
-Prophet is similar to a Generalized Additive Model (GAM) - an additive regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017].
-The decomposable model is of the form:
+Although ARIMA models are powerful, they require non-stationary data to
+be differenced and impose prior assumptions onto the distribution of the
+data (Ince and Trafalis 2006). In contrast, machine learning techniques
+such as ANNs do not impose any assumptions onto the data generating
+process however, being data-driven, are sensitive to the number of input
+nodes used. The ANN-ARIMA model draws on the strengths of both models
+and overcomes these individual difficulties. Ince and Trafalis (2006)
+created an ANN-ARIMA model by using the ARIMA model to determine the
+number input nodes required by an ANN for three exchange rates. When the
+ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model
+outperformed the ARIMA based on the MSE.
 
-$$ 
+Although the traditional ARIMA models produce less superior forecasts
+than its hybrid forms and other complex non-linear techniques, its
+forecasts are still satisfactory. They are simple models that are easy
+to implement and have a consistent method of model design and selection.
+ARIMA models are also more robust and efficient than complex structural
+models in relation to short-run forecasting. The fact that they have
+been used as the foundation for more advanced models and have commonly
+been used as a benchmark for comparison, justifies it as a good starting
+point to compare it to Facebook's forecasting method, Prophet, that was
+recently released.
+
+\subsection{Forecasting with Prophet}\label{forecasting-with-prophet}
+
+The techniques that have been considered for exchange rate forecasting
+thus far, require the analyst to have vocational knowledge about time
+series. Prophet differs to traditional time series models in that it is
+flexible and can be customised by a large number of non-experts who have
+little knowledge about time series, however have domain knowledge about
+the data generating process. Prophet allows for a large number of
+forecasts to be produced across a variety of problems and consists of a
+robust evaluation system that allows for a large number of forecasts be
+evaluated and compared. This is Facebook's definition of forecasting at
+scale. Prophet is similar to a Generalized Additive Model (GAM) - an
+additive regression model that consists of non-linear and linear
+regression functions applied to predictor variables (Taylor and Letham
+2017). The decomposable model is of the form:
+
+\[ 
     y(t) = g(t) + s(t) + h(t) + e_t
-$$
-where the components of the model represent the growth, seasonality and holiday respectively, and $e_t$ is white noise.
+\] where the components of the model represent the growth, seasonality
+and holiday respectively, and \(e_t\) is white noise.
 
-Prophet, like the GAM, frames the forecasting problem as a curve fitting exercise and uses backfitting to find the regression functions.
-This allows for the model to be fitted quickly and missing values and large outliers to be handled elegantly.
-The regression model also provides model flexibility and allows the analyst to interactively change model parameters [@taylor2017].
+Prophet, like the GAM, frames the forecasting problem as a curve fitting
+exercise and uses backfitting to find the regression functions. This
+allows for the model to be fitted quickly and missing values and large
+outliers to be handled elegantly. The regression model also provides
+model flexibility and allows the analyst to interactively change model
+parameters (Taylor and Letham 2017).
 
-The growth component of Prophet may be modelled as a linear or non-linear function of time.
-Linear growth is modelled by a piecewise constant function while non-linear growth is modelled similar to population growths which use a logistic growth model [@taylor2017].
-A time-varying upper limit may be specified for logistic growth, at which point the forecasts will saturate.
-This carrying capacity allows the analyst to incorporate their prior knowledge about the maximum obtainable growth level such as the total market or population size into the model.
-Prophet accounts for changes in the trajectory of this trend by automatically detecting and selecting changepoints in the data at which the growth rate is allowed to change.
-These changepoints have a Laplace prior distribution placed on them and its scale parameter may be used to adjust the flexibility of the trend and to choose how aggressively the model should follow historical trend changes.
-Analysts may also adjust the number of potential changepoints included or manually specify their location.
-This allows non-experts with knowledge about events that may affect the growth rate to use the parameter as a knob to either increase or decrease the number of changepoints included [@taylor2017].
-It also allows for the analyst to add changepoints which the automatic selection procedure may have missed or remove changepoints when the model is overfitting historical trends [@taylor2017].
+The growth component of Prophet may be modelled as a linear or
+non-linear function of time. Linear growth is modelled by a piecewise
+constant function while non-linear growth is modelled similar to
+population growths which use a logistic growth model (Taylor and Letham
+2017). A time-varying upper limit may be specified for logistic growth,
+at which point the forecasts will saturate. This carrying capacity
+allows the analyst to incorporate their prior knowledge about the
+maximum obtainable growth level such as the total market or population
+size into the model. Prophet accounts for changes in the trajectory of
+this trend by automatically detecting and selecting changepoints in the
+data at which the growth rate is allowed to change. These changepoints
+have a Laplace prior distribution placed on them and its scale parameter
+may be used to adjust the flexibility of the trend and to choose how
+aggressively the model should follow historical trend changes. Analysts
+may also adjust the number of potential changepoints included or
+manually specify their location. This allows non-experts with knowledge
+about events that may affect the growth rate to use the parameter as a
+knob to either increase or decrease the number of changepoints included
+(Taylor and Letham 2017). It also allows for the analyst to add
+changepoints which the automatic selection procedure may have missed or
+remove changepoints when the model is overfitting historical trends
+(Taylor and Letham 2017).
 
-The decomposable form of the model allows for multiple seasonality components with different periods to be added to the model.
-Seasonality components are modelled by a Fourier series and have a Normal prior distribution placed on its parameters [@taylor2017].
-The spread parameter can be adjusted by analysts to smooth the seasonality and change how much of historical seasonality is projected into the future [@taylor2017].
+The decomposable form of the model allows for multiple seasonality
+components with different periods to be added to the model. Seasonality
+components are modelled by a Fourier series and have a Normal prior
+distribution placed on its parameters (Taylor and Letham 2017). The
+spread parameter can be adjusted by analysts to smooth the seasonality
+and change how much of historical seasonality is projected into the
+future (Taylor and Letham 2017).
 
-The analyst may also provide a list of important events and holidays which have impacted the time series in the past or which they know might impact it in the future.
-The list could include the name, date, and country in which they have taken place or are expected to take place [@taylor2017].
-By specifying the country of occurrence, separate lists can be populated for global events and holidays, and country-specific events and holidays.
-The union of the global and country-specific lists can then be used for forecasting.
-Like seasonality, a Normal prior distribution is placed on the parameters of the holiday component, and the scale parameter can be adjusted by analysts to smooth the holidays [@taylor2017].
+The analyst may also provide a list of important events and holidays
+which have impacted the time series in the past or which they know might
+impact it in the future. The list could include the name, date, and
+country in which they have taken place or are expected to take place
+(Taylor and Letham 2017). By specifying the country of occurrence,
+separate lists can be populated for global events and holidays, and
+country-specific events and holidays. The union of the global and
+country-specific lists can then be used for forecasting. Like
+seasonality, a Normal prior distribution is placed on the parameters of
+the holiday component, and the scale parameter can be adjusted by
+analysts to smooth the holidays (Taylor and Letham 2017).
 
-Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge into the model building process and has produced significantly improved forecasts compared to the ARIMA model.
-@taylor2017 forecasted the number of events on Facebook using Prophet.
-The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend.
-The forecasts produced by Prophet were compared to common forecasting techniques such as exponential smoothing, ARIMA, the seasonal naïve, and the naïve model.
-While exponential smoothing and the seasonal naïve model were quite robust, the ARIMA forecasts were fragile.
-No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.
+Prophet's Bayesian approach to forecasting allows the analyst to
+incorporate their expert knowledge into the model building process and
+has produced significantly improved forecasts compared to the ARIMA
+model. Taylor and Letham (2017) forecasted the number of events on
+Facebook using Prophet. The time series was impacted by holidays, had
+strong multi-period seasonality, and a piecewise trend. The forecasts
+produced by Prophet were compared to common forecasting techniques such
+as exponential smoothing, ARIMA, the seasonal naïve, and the naïve
+model. While exponential smoothing and the seasonal naïve model were
+quite robust, the ARIMA forecasts were fragile. No model besides Prophet
+accounted for the dips around holidays and the upward trend of the time
+series towards later observations.
 
-Prophet’s ability to forecast at scale enables it to model a wide variety of data and may be able to adequately fit exchange rate data.
-Its non-linear components could capture the non-linearity present in exchange rates.
-In contrast, the ARIMA model is a linear function of previous observations and lagged residuals and fails to capture the non-linearity inherent in exchange rate data.
-Prophet performs well on data with strong multiple “human scale” seasonalities and historical trend changes.
-This differs to the ARIMA model which requires the data to be de-trended and the variance stabilised before the model can be fitted.
-Hence Prophet could model the weekly seasonality of closing prices due to low trading activity which occurs around the weekend and high trading activity which occurs mid-week.
+Prophet's ability to forecast at scale enables it to model a wide
+variety of data and may be able to adequately fit exchange rate data.
+Its non-linear components could capture the non-linearity present in
+exchange rates. In contrast, the ARIMA model is a linear function of
+previous observations and lagged residuals and fails to capture the
+non-linearity inherent in exchange rate data. Prophet performs well on
+data with strong multiple ``human scale'' seasonalities and historical
+trend changes. This differs to the ARIMA model which requires the data
+to be de-trended and the variance stabilised before the model can be
+fitted. Hence Prophet could model the weekly seasonality of closing
+prices due to low trading activity which occurs around the weekend and
+high trading activity which occurs mid-week.
 
-Choosing the correct combination of parameters for the ARIMA model is a challenging task due to the array of possible choices.
-Although the auto.arima function in R may be used to automatically select an ARIMA model that best fits the data, completely automatic forecasting methods are too brittle and do not allow for useful assumptions to be incorporated into the model.
-Prophet makes use of a semi-automatic forecasting technique that keeps the analyst-in-the-loop.
-Its default settings are said to generate forecasts that are as accurate as those produced by skilled forecasters.
-If the forecasts produced are unsatisfactory, they can be improved by the analyst by configuring the model through its easily interpretable parameters.
-Furthermore, non-experts who have domain knowledge about factors that affect Bitcoin or if the dates of events which could impact the price of Bitcoin are known, it may be incorporated into the model by the analyst.
-Prophet can produce forecasts over irregular time intervals and allows for missing values in the time series without the need for interpolation [@taylor2017].
-The ARIMA model on the other hand requires large outliers to be removed and handles missing values by interpolation.
-If Prophet produces forecasts that are as good as the ARIMA model when forecasting BTC/ZAR, it can be compared to more complex forecasting methods that are less tractable such as the hybrid models and machine learning techniques seen earlier.
+Choosing the correct combination of parameters for the ARIMA model is a
+challenging task due to the array of possible choices. Although the
+auto.arima function in R may be used to automatically select an ARIMA
+model that best fits the data, completely automatic forecasting methods
+are too brittle and do not allow for useful assumptions to be
+incorporated into the model. Prophet makes use of a semi-automatic
+forecasting technique that keeps the analyst-in-the-loop. Its default
+settings are said to generate forecasts that are as accurate as those
+produced by skilled forecasters. If the forecasts produced are
+unsatisfactory, they can be improved by the analyst by configuring the
+model through its easily interpretable parameters. Furthermore,
+non-experts who have domain knowledge about factors that affect Bitcoin
+or if the dates of events which could impact the price of Bitcoin are
+known, it may be incorporated into the model by the analyst. Prophet can
+produce forecasts over irregular time intervals and allows for missing
+values in the time series without the need for interpolation (Taylor and
+Letham 2017). The ARIMA model on the other hand requires large outliers
+to be removed and handles missing values by interpolation. If Prophet
+produces forecasts that are as good as the ARIMA model when forecasting
+BTC/ZAR, it can be compared to more complex forecasting methods that are
+less tractable such as the hybrid models and machine learning techniques
+seen earlier.
 
-# Data \label{Data}
+\section{\texorpdfstring{Data \label{Data}}{Data }}\label{data}
 
-The data set used in this study consists of the daily closing prices of BTC/ZAR over the period 24 January 2016 to 17 July 2017.
-This comprises of a total of 541 trading days and was the chosen time period for analysis due to constraints in obtaining data over a longer period.
-The data was obtained from Bitcoincharts.
+The data set used in this study consists of the daily closing prices of
+BTC/ZAR over the period 24 January 2016 to 17 July 2017. This comprises
+of a total of 541 trading days and was the chosen time period for
+analysis due to constraints in obtaining data over a longer period. The
+data was obtained from Bitcoincharts.
 
-Bitcoin is of specific interest as it presents an interesting parallel to traditional exchange rate markets.
-The cryptocurrency is built on a decentralised system and as a result its value cannot be directly influenced by a central authority (Fantazzini and Nigmatullin 2016).
-In addition, the Bitcoin market has attracted attention worldwide and is currently the leading cryptocurrency, with awareness and adoption of the currency growing over time (Fantazzini and Nigmatullin 2016).
-Its novelty makes it a highly volatile hence speculative market and provides an opportunity for forecasting.
+Bitcoin is of specific interest as it presents an interesting parallel
+to traditional exchange rate markets. The cryptocurrency is built on a
+decentralised system and as a result its value cannot be directly
+influenced by a central authority (Fantazzini and Nigmatullin 2016). In
+addition, the Bitcoin market has attracted attention worldwide and is
+currently the leading cryptocurrency, with awareness and adoption of the
+currency growing over time (Fantazzini and Nigmatullin 2016). Its
+novelty makes it a highly volatile hence speculative market and provides
+an opportunity for forecasting.
 
-In this study, the analysis of BTC/ZAR forecasts are based on the daily continuous log returns.
-The daily closing prices are transformed into returns by taking the log difference at each time t and is calculated as follows:
+In this study, the analysis of BTC/ZAR forecasts are based on the daily
+continuous log returns. The daily closing prices are transformed into
+returns by taking the log difference at each time t and is calculated as
+follows:
 
-$$ r_t = ln(\frac{p_t}{p_{t-1}}) $$
-where $p_t$ represents the closing price and $r_t$ represents the log return for time t = 1,2,.,T.
+\[ r_t = ln(\frac{p_t}{p_{t-1}}) \] where \(p_t\) represents the closing
+price and \(r_t\) represents the log return for time t = 1,2,.,T.
 
-# Methodology \label{Methodology}
-In this paper, the out-of-sample forecasts of ARIMA and Prophet are compared over varying forecast horizons.
-The data is first split into a training set and test set.
-The training set starts on the 24th January 2016 and ends on the 16th January 2017, while the test set starts on the 17th January 2017 and ends on the 17th July 2017.
-The Box-Jenkins methodology is then used to select the correct ARIMA model while Prophet’s semi-automatic procedure is used for model selection.
-The selected ARIMA and Prophet model are then fitted to the training set and rolling window forecasts are made 1 day, 30 days and 90 days ahead.
-This allows us to examine the forecast horizon effect.
-The Mincer-Zarnowitz test is then used to evaluate the forecasts produced by ARIMA and Prophet against the test set.
-Since there is no consensus on which accuracy statistic best measures the performance of forecasting techniques, the most popular criteria, the Mean Squared Error (MSE) is employed.
-The results obtained from the ARIMA model will be used as a benchmark for comparison.
-This process is then repeated based on data which has outliers and missing values present.
+\section{\texorpdfstring{Methodology
+\label{Methodology}}{Methodology }}\label{methodology}
 
-## The Box-Jenkins Methodology 
-Box and Jenkins proposed a set of guidelines that can be followed when selecting ARIMA models.
-This systematic procedure to designing ARIMA models has made them highly popular [@spyros1997].
-It consists of a four-stage iterative process in which:
+In this paper, the out-of-sample forecasts of ARIMA and Prophet are
+compared over varying forecast horizons. The data is first split into a
+training set and test set. The training set starts on the 24th January
+2016 and ends on the 16th January 2017, while the test set starts on the
+17th January 2017 and ends on the 17th July 2017. The Box-Jenkins
+methodology is then used to select the correct ARIMA model while
+Prophet's semi-automatic procedure is used for model selection. The
+selected ARIMA and Prophet model are then fitted to the training set and
+rolling window forecasts are made 1 day, 30 days and 90 days ahead. This
+allows us to examine the forecast horizon effect. The Mincer-Zarnowitz
+test is then used to evaluate the forecasts produced by ARIMA and
+Prophet against the test set. Since there is no consensus on which
+accuracy statistic best measures the performance of forecasting
+techniques, the most popular criteria, the Mean Squared Error (MSE) is
+employed. The results obtained from the ARIMA model will be used as a
+benchmark for comparison. This process is then repeated based on data
+which has outliers and missing values present.
 
-1. The process is either transformed or differenced to de-trend and stabilise the variance of the data.
+\subsection{The Box-Jenkins
+Methodology}\label{the-box-jenkins-methodology}
 
-2. The autocorrelation and partial autocorrelation plots are used to determine the order of p and q. 
-3. The parameters of the model are then estimated.
+Box and Jenkins proposed a set of guidelines that can be followed when
+selecting ARIMA models. This systematic procedure to designing ARIMA
+models has made them highly popular (Hibon and Makridakis 1997). It
+consists of a four-stage iterative process in which:
 
-4. A diagnostic check is performed to ensure that the residuals are a white noise process.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  The process is either transformed or differenced to de-trend and
+  stabilise the variance of the data.
+\item
+  The autocorrelation and partial autocorrelation plots are used to
+  determine the order of p and q.
+\item
+  The parameters of the model are then estimated.
+\item
+  A diagnostic check is performed to ensure that the residuals are a
+  white noise process.
+\end{enumerate}
 
-If the residuals are not white noise, steps 2-4 are repeated until a satisfactory model is identified.
-On the contrary, if the diagnostic check shows that the residuals are random, the developed model will be the final model used for forecasting.
+If the residuals are not white noise, steps 2-4 are repeated until a
+satisfactory model is identified. On the contrary, if the diagnostic
+check shows that the residuals are random, the developed model will be
+the final model used for forecasting.
 
-## Semi-Automatic Selection \label{Semi-Automatic Selection}
-When a large number of forecasts are produced, manually identifying problematic forecasts becomes a time consuming and difficult task.
-Prophet provides a semi-automated forecast evaluation system that selects the best model which fits the data.
-When there are large forecast errors, the forecasts are flagged so that the analyst can explore the cause of the errors, identify and remove potential outliers and either adjust the model or choose a more appropriate model [@taylor2017].
-This keeps the analyst-in-the-loop.
-Prophet has the following default settings:
+\subsection{\texorpdfstring{Semi-Automatic Selection
+\label{Semi-Automatic Selection}}{Semi-Automatic Selection }}\label{semi-automatic-selection}
 
-* The trend is set to be linear
-* The width of the uncertainty intervals is set to 80% 
-* Weekly and yearly seasonality are automatically detected and included in the model if present 
-* The smoothing parameter for holidays and seasonality is set at 10 
-* The smoothing parameter for the trend is set to 0.05
-* The number of potential changepoints is set to 25
+When a large number of forecasts are produced, manually identifying
+problematic forecasts becomes a time consuming and difficult task.
+Prophet provides a semi-automated forecast evaluation system that
+selects the best model which fits the data. When there are large
+forecast errors, the forecasts are flagged so that the analyst can
+explore the cause of the errors, identify and remove potential outliers
+and either adjust the model or choose a more appropriate model (Taylor
+and Letham 2017). This keeps the analyst-in-the-loop. Prophet has the
+following default settings:
 
-This paper makes use of Prophet’s default settings to fit the data, however weekly seasonality is included in the model.
-This allows us to account for our prior knowledge about the closing prices of Bitcoin which tend to be lower around the weekend however higher mid-week, due to fluctuations in trading activity.
-Furthermore, a linear trend is appropriate since Bitcoin does not have an upper limit on its closing prices.
+\begin{itemize}
+\tightlist
+\item
+  The trend is set to be linear
+\item
+  The width of the uncertainty intervals is set to 80\%
+\item
+  Weekly and yearly seasonality are automatically detected and included
+  in the model if present
+\item
+  The smoothing parameter for holidays and seasonality is set at 10
+\item
+  The smoothing parameter for the trend is set to 0.05
+\item
+  The number of potential changepoints is set to 25
+\end{itemize}
 
-## Rolling Window Forecasts
-Rolling window forecasts are useful in evaluating the robustness of a forecasting method.
-In a rolling window forecast, the forecasts are made h-steps at a time and the actual observation rather than the predicted value is used for the next prediction in the forecast horizon.
-In this way, a poor forecast will not have negative consequences on future forecasts to be made, since the observed values will be used to correct itself for the remaining forecasts [@zhang1998].
-In this paper, we first fit the ARIMA and Prophet model to the training set and forecast 1 day ahead.
-We then refit the model to the training set, increasing its size by one observation, and the next 1 day ahead forecast is made.
-This process is repeated until all forecasts have been made into the test set.
-The 1 day ahead forecasts are then compared to the 1 day ahead observed values in the test set and the error statistics are calculated.
-This procedure is then repeated for the 30 day and 90 day ahead forecasts.
+This paper makes use of Prophet's default settings to fit the data,
+however weekly seasonality is included in the model. This allows us to
+account for our prior knowledge about the closing prices of Bitcoin
+which tend to be lower around the weekend however higher mid-week, due
+to fluctuations in trading activity. Furthermore, a linear trend is
+appropriate since Bitcoin does not have an upper limit on its closing
+prices.
 
-## Mincer-Zarnowitz Approach to Forecast Evaluation 
-The Mincer-Zarnowitz approach to evaluating forecast accuracy is commonly used and can be useful when comparing the forecasts produced by Prophet and the ARIMA model.
-@mincer1969 proposed an absolute measure to evaluate forecast accuracy which considers the distance between the actual and predicted values.
-To analyse the absolute errors produced by the forecasts, the observed values $r_t$ are regressed against the predicted values $r_t$, i.e.
+\subsection{Rolling Window Forecasts}\label{rolling-window-forecasts}
 
-$$r_t = \alpha + \beta\cdot\hat{r_t} + e_t$$
+Rolling window forecasts are useful in evaluating the robustness of a
+forecasting method. In a rolling window forecast, the forecasts are made
+h-steps at a time and the actual observation rather than the predicted
+value is used for the next prediction in the forecast horizon. In this
+way, a poor forecast will not have negative consequences on future
+forecasts to be made, since the observed values will be used to correct
+itself for the remaining forecasts (Zhang and Hu 1998). In this paper,
+we first fit the ARIMA and Prophet model to the training set and
+forecast 1 day ahead. We then refit the model to the training set,
+increasing its size by one observation, and the next 1 day ahead
+forecast is made. This process is repeated until all forecasts have been
+made into the test set. The 1 day ahead forecasts are then compared to
+the 1 day ahead observed values in the test set and the error statistics
+are calculated. This procedure is then repeated for the 30 day and 90
+day ahead forecasts.
 
-where $\alpha$ represents the mean distance between the observed and predicted values while $\beta$ represents the correlation between the forecasted errors and the predicted values.
-When alpha = 0 it implies that the forecasts are unbiased and do not systematically overestimate or underestimate the data.
-When b=1 it implies that the forecasts are efficient and uncorrelated with the forecasted errors.
-A joint hypothesis test of $H_0: \alpha = 0 \cup \beta = 1$ is performed to check the efficiency and bias of the forecasts and the model which produces the best results generates superior forecasts.
+\subsection{Mincer-Zarnowitz Approach to Forecast
+Evaluation}\label{mincer-zarnowitz-approach-to-forecast-evaluation}
 
-## Evaluation of forecasts based on data with missing values and outliers
-The original dataset is modified in order to examine how robust Prophet and ARIMA are when fitted to data with missing values and outliers.
-20 values are randomly removed from the training set and 5 outliers are randomly inserted in place of existing values.
-The models are then fitted to this “dirty” dataset and evaluating of forecasts are carried out as before.
+The Mincer-Zarnowitz approach to evaluating forecast accuracy is
+commonly used and can be useful when comparing the forecasts produced by
+Prophet and the ARIMA model. Mincer and Zarnowitz (1969) proposed an
+absolute measure to evaluate forecast accuracy which considers the
+distance between the actual and predicted values. To analyse the
+absolute errors produced by the forecasts, the observed values \(r_t\)
+are regressed against the predicted values \(r_t\), i.e.
 
-# Results \label{Results}
+\[r_t = \alpha + \beta\cdot\hat{r_t} + e_t\]
 
-## Evaluation over varying forecast horizons
-Although the Box-Jenkins methodology resulted in an ARIMA(0,0,1) model being the best fit to the training set, an ARIMA(1,0,0) is employed as the benchmark for comparison so that a rolling window forecast may be conducted.
-The AIC of the ARIMA(1,0,0) is marginally higher than that of the ARIMA(0,0,1) and analysis of the p-value indicates that the AR parameter is statistically significant in the explanation of the movement of BTC/ZAR.
-The Prophet model fitted to the training set included a linear trend and weekly seasonality, with all other parameters set to the default values as described in section \ref{Semi-Automatic Selection}.
-The automatic-selection procedure for ARIMA and Prophet selected the same model as those chosen manually.
+where \(\alpha\) represents the mean distance between the observed and
+predicted values while \(\beta\) represents the correlation between the
+forecasted errors and the predicted values. When alpha = 0 it implies
+that the forecasts are unbiased and do not systematically overestimate
+or underestimate the data. When b=1 it implies that the forecasts are
+efficient and uncorrelated with the forecasted errors. A joint
+hypothesis test of \(H_0: \alpha = 0 \cup \beta = 1\) is performed to
+check the efficiency and bias of the forecasts and the model which
+produces the best results generates superior forecasts.
 
-### 1-step ahead Forecasts
-Figure 1 depicts the resulting forecasts from fitting both models.
-The forecasts produced by the ARIMA model and Prophet lie within a band centred around zero.
-Both models fail to capture the large movement of returns away from zero.
-The 80% uncertainty intervals produced by both models are approximately the same width and begin to widen as forecasts are made further into the test set.
+\subsection{Evaluation of forecasts based on data with missing values
+and
+outliers}\label{evaluation-of-forecasts-based-on-data-with-missing-values-and-outliers}
 
-![1-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\1-step ahead forecasts.png)
+The original dataset is modified in order to examine how robust Prophet
+and ARIMA are when fitted to data with missing values and outliers. 20
+values are randomly removed from the training set and 5 outliers are
+randomly inserted in place of existing values. The models are then
+fitted to this ``dirty'' dataset and evaluating of forecasts are carried
+out as before.
 
-Table 1 displays the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the one-step ahead forecasts.
-The intercept estimate is small and positive for ARIMA and Prophet, with the intercept for ARIMA being closer to zero than Prophet.
-This indicates that both models systematically underestimate the 1-step ahead returns, with ARIMA yielding mean forecasts closer to the observed mean.
-The slope estimate is negative for both models indicating that the observed returns are actually the opposite of the suggested forecast.
-The F-test for the joint hypothesis of a unity slope and zero intercept is rejected at the 5% significance level for both models, with the p-value for ARIMA being much larger than Prophet.
-Hence the forecasts generated by both ARIMA and Prophet are biased and/or inefficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
+\section{\texorpdfstring{Results
+\label{Results}}{Results }}\label{results}
 
-![Mincer-Zarnowitz results for 1-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 1.png)
+\subsection{Evaluation over varying forecast
+horizons}\label{evaluation-over-varying-forecast-horizons}
 
-### 30-steps ahead Forecasts
-Figure 2 illustrates the resulting 30-step ahead forecasts produced from fitting ARIMA and Prophet.
-As in the 1-step ahead case, both models fail to capture the large movement of returns away from zero with its forecasts lying within a band centred around zero.
-The 80% uncertainty intervals produced by both models are approximately the same width and marginally wider than the uncertainty intervals produced by the 1-step ahead forecasts.
+Although the Box-Jenkins methodology resulted in an ARIMA(0,0,1) model
+being the best fit to the training set, an ARIMA(1,0,0) is employed as
+the benchmark for comparison so that a rolling window forecast may be
+conducted. The AIC of the ARIMA(1,0,0) is marginally higher than that of
+the ARIMA(0,0,1) and analysis of the p-value indicates that the AR
+parameter is statistically significant in the explanation of the
+movement of BTC/ZAR. The Prophet model fitted to the training set
+included a linear trend and weekly seasonality, with all other
+parameters set to the default values as described in section
+\ref{Semi-Automatic Selection}. The automatic-selection procedure for
+ARIMA and Prophet selected the same model as those chosen manually.
 
-![30-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\30-step ahead forecasts.png)
+\subsubsection{1-step ahead Forecasts}\label{step-ahead-forecasts}
 
-Table 2 displays the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 30-step ahead forecasts.
-The intercept estimate is small and positive for ARIMA and Prophet, with a smaller intercept estimated for Prophet.
-This indicates that both models systematically underestimate the 30-step ahead returns, with Prophet yielding mean forecasts closer to the observed mean.
-The slope estimate is large and negative for both models with the slope estimated to be as low as -12.74 for ARIMA.
-This suggests that the observed returns are actually the opposite of the proposed forecast.
-The F-test for the joint hypothesis of a unity slope and zero intercept is rejected at the 5% significance level for Prophet only.
-The p-value of 0.0614 from the joint hypothesis test is slightly significant for ARIMA.
-Hence the forecasts generated by Prophet are biased and/or inefficient while the forecasts generated by ARIMA are unbiased and/or efficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
+Figure 1 depicts the resulting forecasts from fitting both models. The
+forecasts produced by the ARIMA model and Prophet lie within a band
+centred around zero. Both models fail to capture the large movement of
+returns away from zero. The 80\% uncertainty intervals produced by both
+models are approximately the same width and begin to widen as forecasts
+are made further into the test set.
 
-![Mincer-Zarnowitz results for 30-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 2.png)
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/1-step ahead forecasts.png}
+\caption{1-step ahead forecasts from the ARIMA(1,0,0) and Prophet model}
+\end{figure}
 
-### 90-steps ahead Forecasts
-Figure 3 illustrates the 90-step ahead forecasts produced by ARIMA and Prophet.
-Both models fail to capture large movement of returns away from zero as in the case of the 1-step ahead and 30-step ahead forecasts.
-The 80% uncertainty intervals produced by both models are approximately the same width and are marginally narrower than the uncertainty intervals produced by the one-step ahead and 30-step ahead forecasts.
+Table 1 displays the MSE and the Mincer-Zarnowitz results obtained from
+regressing the test set returns against the one-step ahead forecasts.
+The intercept estimate is small and positive for ARIMA and Prophet, with
+the intercept for ARIMA being closer to zero than Prophet. This
+indicates that both models systematically underestimate the 1-step ahead
+returns, with ARIMA yielding mean forecasts closer to the observed mean.
+The slope estimate is negative for both models indicating that the
+observed returns are actually the opposite of the suggested forecast.
+The F-test for the joint hypothesis of a unity slope and zero intercept
+is rejected at the 5\% significance level for both models, with the
+p-value for ARIMA being much larger than Prophet. Hence the forecasts
+generated by both ARIMA and Prophet are biased and/or inefficient. The
+MSE produced by both models agree with the Mincer-Zarnowitz results,
+with ARIMA producing a marginally smaller MSE than Prophet.
 
-![90-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\90-steps ahead forecasts.png)
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 1.png}
+\caption{Mincer-Zarnowitz results for 1-step ahead forecasts}
+\end{figure}
 
-Table 3 presents the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 90-step ahead forecasts.
-The intercept estimate is very small and positive for ARIMA and Prophet, with a marginally smaller intercept estimated for ARIMA.
-This indicates that both models systematically underestimate the 90-step ahead returns, with ARIMA yielding mean forecasts closer to the observed mean.
-The slope estimate is small and negative for both models with the slope estimated for ARIMA lying closer to -1.
-This suggests that the observed 90-steps ahead returns are actually the opposite of the proposed 90-steps ahead forecasts.
-The F-test for the joint hypothesis of a unity slope and zero intercept fails to be rejected at the 5% significance level for both ARIMA and Prophet, with the p-value from the test being much larger for ARIMA.
-Hence the forecasts generated by Prophet and ARIMA are unbiased and/or efficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
+\subsubsection{30-steps ahead Forecasts}\label{steps-ahead-forecasts}
 
-![Mincer-Zarnowitz results for 90-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 3.png)
+Figure 2 illustrates the resulting 30-step ahead forecasts produced from
+fitting ARIMA and Prophet. As in the 1-step ahead case, both models fail
+to capture the large movement of returns away from zero with its
+forecasts lying within a band centred around zero. The 80\% uncertainty
+intervals produced by both models are approximately the same width and
+marginally wider than the uncertainty intervals produced by the 1-step
+ahead forecasts.
 
-### Comparison over varying forecast horizons
-Table 4 and 5 compares the p-values obtained from the Mincer-Zarnowitz tests and the MSE of the forecasts produced by ARIMA and Prophet over the three forecast horizons.
-In comparing the forecasting accuracy of the ARIMA model over the three forecast horizons, we observe that the p-value gets more significant as the forecast horizon gets longer.
-This is in contrast to the MSE of the ARIMA forecasts which is smaller at the shortest and longest forecast horizons.
-In comparing the forecasting accuracy of Prophet, we find that the p-value is more significant for the shortest and longest forecast horizon.
-This agrees with the MSE of the Prophet forecasts which is smaller at the longest and shortest forecast horizon.
-Both ARIMA and Prophet generate forecasts with the smallest MSE and the most significant p-value at the longest forecast horizon.
-In comparing the forecasting accuracy of ARIMA and Prophet over the three forecast horizons, we notice that the p-value obtained from the Mincer-Zarnowitz results is more significant for ARIMA than Prophet at every forecast horizon.
-This echoes the results obtained from the MSE which is marginally smaller for ARIMA than Prophet over all forecast horizons.
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/30-step ahead forecasts.png}
+\caption{30-step ahead forecasts from the ARIMA(1,0,0) and Prophet
+model}
+\end{figure}
 
-![Mincer-Zarnowitz results for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 4.png)
+Table 2 displays the MSE and the Mincer-Zarnowitz results obtained from
+regressing the test set returns against the 30-step ahead forecasts. The
+intercept estimate is small and positive for ARIMA and Prophet, with a
+smaller intercept estimated for Prophet. This indicates that both models
+systematically underestimate the 30-step ahead returns, with Prophet
+yielding mean forecasts closer to the observed mean. The slope estimate
+is large and negative for both models with the slope estimated to be as
+low as -12.74 for ARIMA. This suggests that the observed returns are
+actually the opposite of the proposed forecast. The F-test for the joint
+hypothesis of a unity slope and zero intercept is rejected at the 5\%
+significance level for Prophet only. The p-value of 0.0614 from the
+joint hypothesis test is slightly significant for ARIMA. Hence the
+forecasts generated by Prophet are biased and/or inefficient while the
+forecasts generated by ARIMA are unbiased and/or efficient. The MSE
+produced by both models agree with the Mincer-Zarnowitz results, with
+ARIMA producing a marginally smaller MSE than Prophet.
 
-![MSE for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 5.png)
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 2.png}
+\caption{Mincer-Zarnowitz results for 30-step ahead forecasts}
+\end{figure}
 
-## Evaluation on data with missing values and outliers present
-By following the Box-Jenkins methodology for selection of an ARIMA model, an ARIMA(3,0,0) was employed to fit the training set which included missing values and outliers.
-Although the ARIMA(0,0,3) had a slightly smaller AIC, a decision was made to use the ARIMA(3,0,0) instead so that a rolling window forecast may be used.
-Analysis of the AR parameters indicate that they are statistically significant in the explanation of the movement of BTC/ZAR.
-The Prophet model fitted to the training set included a linear trend and weekly seasonality.
-All other parameters were set to Prophet’s default values as described in section \ref{Semi-Automatic Selection}.
-The automatic-selection procedure for Prophet and ARIMA selected the same models as those chosen manually.
+\subsubsection{90-steps ahead Forecasts}\label{steps-ahead-forecasts-1}
 
-### 1-step ahead Forecasts
-Table 6 shows the MSE of the forecasts generated by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 1-step ahead returns forecasted.
-We observe that a positive intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically underestimate the 1-step ahead returns.
-The intercept estimated for ARIMA is smaller than Prophet, indicating that the mean forecasted returns yielded by ARIMA is closer to the mean observed returns.
-The intercept estimates of 0.011895 and 0.014802 for ARIMA and Prophet respectively are larger than the intercept estimates of 0.009546 and 0.013702, which was based on the “clean” dataset.
-This suggests that both models systematically underestimates the returns more when there are missing values and outliers present in the data.
+Figure 3 illustrates the 90-step ahead forecasts produced by ARIMA and
+Prophet. Both models fail to capture large movement of returns away from
+zero as in the case of the 1-step ahead and 30-step ahead forecasts. The
+80\% uncertainty intervals produced by both models are approximately the
+same width and are marginally narrower than the uncertainty intervals
+produced by the one-step ahead and 30-step ahead forecasts.
 
-We also observe that the slope estimate is small and negative for both models with the slope estimated for ARIMA lying closer to -1.
-This suggests that the observed 1-step ahead returns are actually the opposite of the proposed 1-step ahead forecasts.
-The slope estimates of -0.883228 and -0.676132 for ARIMA and Prophet respectively are less steep than the slope estimates of -2.005624 and -1.709744, which was based on the evaluation of data that had no values missing or outliers present.
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/90-steps ahead forecasts.png}
+\caption{90-step ahead forecasts from the ARIMA(1,0,0) and Prophet
+model}
+\end{figure}
 
-![Mincer-Zarnowitz results for 1-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 6.png)
+Table 3 presents the MSE and the Mincer-Zarnowitz results obtained from
+regressing the test set returns against the 90-step ahead forecasts. The
+intercept estimate is very small and positive for ARIMA and Prophet,
+with a marginally smaller intercept estimated for ARIMA. This indicates
+that both models systematically underestimate the 90-step ahead returns,
+with ARIMA yielding mean forecasts closer to the observed mean. The
+slope estimate is small and negative for both models with the slope
+estimated for ARIMA lying closer to -1. This suggests that the observed
+90-steps ahead returns are actually the opposite of the proposed
+90-steps ahead forecasts. The F-test for the joint hypothesis of a unity
+slope and zero intercept fails to be rejected at the 5\% significance
+level for both ARIMA and Prophet, with the p-value from the test being
+much larger for ARIMA. Hence the forecasts generated by Prophet and
+ARIMA are unbiased and/or efficient. The MSE produced by both models
+agree with the Mincer-Zarnowitz results, with ARIMA producing a
+marginally smaller MSE than Prophet.
 
-The p-values obtained from the F-test for the joint hypothesis of a unity slope and zero intercept is approximately zero for ARIMA and Prophet, with the p-value for Prophet lying closer to zero than ARIMA.
-The null hypothesis is thus rejected at the 1% significance level, indicating that the forecasts generated by Prophet and ARIMA are biased and/or inefficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
-The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on the “clean” dataset show similar findings, however led to a smaller MSE of 0.001942 and 0.001958 and a more significant p-value of 0.025449 and 0.007152 for ARIMA and Prophet respectively.
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 3.png}
+\caption{Mincer-Zarnowitz results for 90-step ahead forecasts}
+\end{figure}
 
-### 30-steps ahead Forecasts
-Table 7 shows the MSE of the forecasts produced by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 30-step ahead returns forecasted.
-We observe that a positive intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically underestimate the 30-step ahead returns.
-The intercept estimated for ARIMA is larger than Prophet, indicating that the mean forecasted returns yielded by Prophet is closer to the mean observed returns.
-The intercept estimate of 0.11414 for ARIMA is larger than the estimate of 0.031320 which was based on the evaluation of the “clean” dataset.
-This suggests that the ARIMA model underestimates the returns more when there are missing values and outliers present in the data.
-The intercept estimate of 0.014396 for Prophet is less than the estimate of 0.017169 which was based on the evaluation of data that had no values missing or outliers present.
-This suggests that Prophet underestimates the returns less when there are missing values and outliers present in the data.
+\subsubsection{Comparison over varying forecast
+horizons}\label{comparison-over-varying-forecast-horizons}
 
-![Mincer-Zarnowitz results for 30-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 7.png)
+Table 4 and 5 compares the p-values obtained from the Mincer-Zarnowitz
+tests and the MSE of the forecasts produced by ARIMA and Prophet over
+the three forecast horizons. In comparing the forecasting accuracy of
+the ARIMA model over the three forecast horizons, we observe that the
+p-value gets more significant as the forecast horizon gets longer. This
+is in contrast to the MSE of the ARIMA forecasts which is smaller at the
+shortest and longest forecast horizons. In comparing the forecasting
+accuracy of Prophet, we find that the p-value is more significant for
+the shortest and longest forecast horizon. This agrees with the MSE of
+the Prophet forecasts which is smaller at the longest and shortest
+forecast horizon. Both ARIMA and Prophet generate forecasts with the
+smallest MSE and the most significant p-value at the longest forecast
+horizon. In comparing the forecasting accuracy of ARIMA and Prophet over
+the three forecast horizons, we notice that the p-value obtained from
+the Mincer-Zarnowitz results is more significant for ARIMA than Prophet
+at every forecast horizon. This echoes the results obtained from the MSE
+which is marginally smaller for ARIMA than Prophet over all forecast
+horizons.
 
-We also observe that the slope estimate is large and negative for ARIMA and small and negative for Prophet.
-This suggests that the observed 30-step ahead returns are actually the opposite of the proposed 30-step ahead forecasts.
-The slope estimate of -12.985335 for ARIMA is more steep than the slope estimate of -12.736757 which was based on the evaluation of data that had no values missing or outliers present.
-This suggests that the efficiency of the forecasted returns deteriorates when there are missing values and outliers present in the data.
-The slope estimate of -0.65161 for Prophet is less steep than the estimate -2.621914, which was based on the evaluation of data that had no values missing or outliers present.
-This suggests that the efficiency of the forecasted returns improve when there are missing values and outliers present in the data.
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 4.png}
+\caption{Mincer-Zarnowitz results for all forecast horizons}
+\end{figure}
 
-The p-value obtained from the F-test for the joint hypothesis of a unity slope and zero intercept is approximately zero for Prophet.
-The null hypothesis is thus rejected at the 1% significance level, indicating that the 30-step ahead forecasts generated by Prophet are biased and/or inefficient.
-The p-value of 0.107532 for ARIMA is significant at the 5% level, suggesting at the 30-step ahead forecasts produced by ARIMA are unbiased and/or efficient.
-MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a smaller MSE than Prophet.
-The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on the “clean” dataset, led to a more significant p-value of 0.000377 for Prophet only, and a smaller MSE of 0.001988 and 0.002075 for ARIMA and Prophet respectively.
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 5.png}
+\caption{MSE for all forecast horizons}
+\end{figure}
 
-### 90-steps ahead Forecasts
-Table 8 shows the MSE of the forecasts produced by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 90-step ahead returns forecasted.
-We observe that a negative intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically overestimate the 90-step ahead returns.
-The intercept estimated for Prophet lies closer to zero, indicating that the mean forecasted returns yielded by Prophet is closer to the mean observed returns.
-This is in contrast to the positive intercepts that were estimated based on the data that had no values missing or outliers present.
+\subsection{Evaluation on data with missing values and outliers
+present}\label{evaluation-on-data-with-missing-values-and-outliers-present}
 
-We also observe that the slope estimate is large and positive for ARIMA however small and positive for Prophet.
-This is in contrast to the small and negative slope estimates which were based on the evaluation of data that had no values missing or outliers present.
+By following the Box-Jenkins methodology for selection of an ARIMA
+model, an ARIMA(3,0,0) was employed to fit the training set which
+included missing values and outliers. Although the ARIMA(0,0,3) had a
+slightly smaller AIC, a decision was made to use the ARIMA(3,0,0)
+instead so that a rolling window forecast may be used. Analysis of the
+AR parameters indicate that they are statistically significant in the
+explanation of the movement of BTC/ZAR. The Prophet model fitted to the
+training set included a linear trend and weekly seasonality. All other
+parameters were set to Prophet's default values as described in section
+\ref{Semi-Automatic Selection}. The automatic-selection procedure for
+Prophet and ARIMA selected the same models as those chosen manually.
 
-![Mincer-Zarnowitz results for 90-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 8.png)
+\subsubsection{1-step ahead Forecasts}\label{step-ahead-forecasts-1}
 
-The p-values obtained from the F-test for the joint hypothesis of a unity slope and zero intercept are insignificant for ARIMA and Prophet, with the p-value for Prophet lying close to zero.
-The null hypothesis is thus rejected at the 5% significance level, indicating that the 90-steps ahead forecasts generated by Prophet and ARIMA are biased and/or inefficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
-The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on data which had no missing values or outliers present contrasts these findings, with a p-value of 0.843066 and 0.408675 for ARIMA and Prophet respectively.
+Table 6 shows the MSE of the forecasts generated by ARIMA and Prophet
+and the Mincer-Zarnowitz results obtained from regressing the test set
+returns against the 1-step ahead returns forecasted. We observe that a
+positive intercept is estimated in the case of ARIMA and Prophet,
+indicating that both models systematically underestimate the 1-step
+ahead returns. The intercept estimated for ARIMA is smaller than
+Prophet, indicating that the mean forecasted returns yielded by ARIMA is
+closer to the mean observed returns. The intercept estimates of 0.011895
+and 0.014802 for ARIMA and Prophet respectively are larger than the
+intercept estimates of 0.009546 and 0.013702, which was based on the
+``clean'' dataset. This suggests that both models systematically
+underestimates the returns more when there are missing values and
+outliers present in the data.
 
-### Comparison over varying forecast horizons
-Table 9 and 10 compares the p-values obtained from the Mincer-Zarnowitz tests and the MSE of the forecasts produced by ARIMA and Prophet over the three forecast horizons.
-In comparing the forecasting accuracy of the ARIMA model over the three forecast horizons, we observe that the p-value is more significant at the longer forecast horizons.
-This agrees with the MSE of the ARIMA forecasts which is smaller at the longer forecast horizons.
-In comparing the forecasting accuracy of Prophet, we find that the p-value is more significant for the longest forecast horizon.
-This agrees with the MSE of the Prophet forecasts which is smaller at the longest forecast horizon.
+We also observe that the slope estimate is small and negative for both
+models with the slope estimated for ARIMA lying closer to -1. This
+suggests that the observed 1-step ahead returns are actually the
+opposite of the proposed 1-step ahead forecasts. The slope estimates of
+-0.883228 and -0.676132 for ARIMA and Prophet respectively are less
+steep than the slope estimates of -2.005624 and -1.709744, which was
+based on the evaluation of data that had no values missing or outliers
+present.
 
-![Mincer-Zarnowitz results for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 9.png)
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 6.png}
+\caption{Mincer-Zarnowitz results for 1-step ahead forecasts}
+\end{figure}
 
-![MSE for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 10.png)
+The p-values obtained from the F-test for the joint hypothesis of a
+unity slope and zero intercept is approximately zero for ARIMA and
+Prophet, with the p-value for Prophet lying closer to zero than ARIMA.
+The null hypothesis is thus rejected at the 1\% significance level,
+indicating that the forecasts generated by Prophet and ARIMA are biased
+and/or inefficient. The MSE produced by both models agree with the
+Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE
+than Prophet. The results obtained from evaluating the forecasting
+accuracy of ARIMA and Prophet based on the ``clean'' dataset show
+similar findings, however led to a smaller MSE of 0.001942 and 0.001958
+and a more significant p-value of 0.025449 and 0.007152 for ARIMA and
+Prophet respectively.
 
-In comparing the forecasting accuracy of ARIMA and Prophet over the three forecast horizons, we notice that the p-value obtained from the Mincer-Zarnowitz results is more significant for ARIMA than Prophet at every forecast horizon.
-This echoes the results obtained from the MSE which is marginally smaller for ARIMA than Prophet over all forecast horizons.
+\subsubsection{30-steps ahead Forecasts}\label{steps-ahead-forecasts-2}
 
-# Discussion and Conclusions \label{Discussion and Conclusions}
-The results show a clear difference in the effect of the forecast horizon in the case of the clean data.
-The results based on the MSE and the p-value obtained from the Mincer-Zarnowitz test surprisingly show that the ARIMA model produces superior forecasts of the returns at the longest forecast horizon.
-Similarly, Prophet produces superior forecasts of the returns at the longest forecast horizon.
-We notice that although the ARIMA model outperforms Prophet at every forecast horizon, the difference between the MSE’s of the forecasts generated by both models is insignificant.
-Furthermore, the Mincer-Zarnowitz p-values for both models are significant at the same forecast horizons, suggesting that the efficiency and biasness of the return forecasts yielded by ARIMA and Prophet are synchronized at each forecast horizon.
-If a small amount of predictive accuracy is willing to be sacrificed by the analyst in the forecasting of BTC/ZAR, it might be useful to employ Prophet instead of ARIMA.
-This would enable traders with expert knowledge about Bitcoin, however no statistical knowledge about forecasting models, to incorporate their domain knowledge into the Prophet model through its intuitively adjustable parameters.
-Furthermore, as more information about events that may affect the price of Bitcoin becomes available, it may be included in the model and could potentially change the predictive accuracy of Prophet.
+Table 7 shows the MSE of the forecasts produced by ARIMA and Prophet and
+the Mincer-Zarnowitz results obtained from regressing the test set
+returns against the 30-step ahead returns forecasted. We observe that a
+positive intercept is estimated in the case of ARIMA and Prophet,
+indicating that both models systematically underestimate the 30-step
+ahead returns. The intercept estimated for ARIMA is larger than Prophet,
+indicating that the mean forecasted returns yielded by Prophet is closer
+to the mean observed returns. The intercept estimate of 0.11414 for
+ARIMA is larger than the estimate of 0.031320 which was based on the
+evaluation of the ``clean'' dataset. This suggests that the ARIMA model
+underestimates the returns more when there are missing values and
+outliers present in the data. The intercept estimate of 0.014396 for
+Prophet is less than the estimate of 0.017169 which was based on the
+evaluation of data that had no values missing or outliers present. This
+suggests that Prophet underestimates the returns less when there are
+missing values and outliers present in the data.
 
-The results based on the unclean data show that both ARIMA and Prophet are sensitive to the presence of outliers and missing values in addition to the forecast horizon.
-We observe that Prophet yields superior forecasts at the longest forecast horizon.
-Although the MSE suggests that the ARIMA model produces forecasts with minimal error at the longest forecast horizon, analysis of the p-value indicates that the forecasts are efficient and/or unbiased at 30-steps ahead only.
-It is evident that Prophet produces inferior results in comparison to the ARIMA model over all forecast horizons.
-This is unexpected as Prophet is said to be robust to the presence of outliers and missing values in the data, without the need for interpolation.
-We also notice that the forecasts produced by both models deteriorates when there are missing values and outliers present in the data, with the only efficient and/or unbiased forecasts yielded by the ARIMA model at a forecast horizon of 30 days.
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 7.png}
+\caption{Mincer-Zarnowitz results for 30-step ahead forecasts}
+\end{figure}
+
+We also observe that the slope estimate is large and negative for ARIMA
+and small and negative for Prophet. This suggests that the observed
+30-step ahead returns are actually the opposite of the proposed 30-step
+ahead forecasts. The slope estimate of -12.985335 for ARIMA is more
+steep than the slope estimate of -12.736757 which was based on the
+evaluation of data that had no values missing or outliers present. This
+suggests that the efficiency of the forecasted returns deteriorates when
+there are missing values and outliers present in the data. The slope
+estimate of -0.65161 for Prophet is less steep than the estimate
+-2.621914, which was based on the evaluation of data that had no values
+missing or outliers present. This suggests that the efficiency of the
+forecasted returns improve when there are missing values and outliers
+present in the data.
+
+The p-value obtained from the F-test for the joint hypothesis of a unity
+slope and zero intercept is approximately zero for Prophet. The null
+hypothesis is thus rejected at the 1\% significance level, indicating
+that the 30-step ahead forecasts generated by Prophet are biased and/or
+inefficient. The p-value of 0.107532 for ARIMA is significant at the 5\%
+level, suggesting at the 30-step ahead forecasts produced by ARIMA are
+unbiased and/or efficient. MSE produced by both models agree with the
+Mincer-Zarnowitz results, with ARIMA producing a smaller MSE than
+Prophet. The results obtained from evaluating the forecasting accuracy
+of ARIMA and Prophet based on the ``clean'' dataset, led to a more
+significant p-value of 0.000377 for Prophet only, and a smaller MSE of
+0.001988 and 0.002075 for ARIMA and Prophet respectively.
+
+\subsubsection{90-steps ahead Forecasts}\label{steps-ahead-forecasts-3}
+
+Table 8 shows the MSE of the forecasts produced by ARIMA and Prophet and
+the Mincer-Zarnowitz results obtained from regressing the test set
+returns against the 90-step ahead returns forecasted. We observe that a
+negative intercept is estimated in the case of ARIMA and Prophet,
+indicating that both models systematically overestimate the 90-step
+ahead returns. The intercept estimated for Prophet lies closer to zero,
+indicating that the mean forecasted returns yielded by Prophet is closer
+to the mean observed returns. This is in contrast to the positive
+intercepts that were estimated based on the data that had no values
+missing or outliers present.
+
+We also observe that the slope estimate is large and positive for ARIMA
+however small and positive for Prophet. This is in contrast to the small
+and negative slope estimates which were based on the evaluation of data
+that had no values missing or outliers present.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 8.png}
+\caption{Mincer-Zarnowitz results for 90-step ahead forecasts}
+\end{figure}
+
+The p-values obtained from the F-test for the joint hypothesis of a
+unity slope and zero intercept are insignificant for ARIMA and Prophet,
+with the p-value for Prophet lying close to zero. The null hypothesis is
+thus rejected at the 5\% significance level, indicating that the
+90-steps ahead forecasts generated by Prophet and ARIMA are biased
+and/or inefficient. The MSE produced by both models agree with the
+Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE
+than Prophet. The results obtained from evaluating the forecasting
+accuracy of ARIMA and Prophet based on data which had no missing values
+or outliers present contrasts these findings, with a p-value of 0.843066
+and 0.408675 for ARIMA and Prophet respectively.
+
+\subsubsection{Comparison over varying forecast
+horizons}\label{comparison-over-varying-forecast-horizons-1}
+
+Table 9 and 10 compares the p-values obtained from the Mincer-Zarnowitz
+tests and the MSE of the forecasts produced by ARIMA and Prophet over
+the three forecast horizons. In comparing the forecasting accuracy of
+the ARIMA model over the three forecast horizons, we observe that the
+p-value is more significant at the longer forecast horizons. This agrees
+with the MSE of the ARIMA forecasts which is smaller at the longer
+forecast horizons. In comparing the forecasting accuracy of Prophet, we
+find that the p-value is more significant for the longest forecast
+horizon. This agrees with the MSE of the Prophet forecasts which is
+smaller at the longest forecast horizon.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 9.png}
+\caption{Mincer-Zarnowitz results for all forecast horizons}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 10.png}
+\caption{MSE for all forecast horizons}
+\end{figure}
+
+In comparing the forecasting accuracy of ARIMA and Prophet over the
+three forecast horizons, we notice that the p-value obtained from the
+Mincer-Zarnowitz results is more significant for ARIMA than Prophet at
+every forecast horizon. This echoes the results obtained from the MSE
+which is marginally smaller for ARIMA than Prophet over all forecast
+horizons.
+
+\section{\texorpdfstring{Discussion and Conclusions
+\label{Discussion and Conclusions}}{Discussion and Conclusions }}\label{discussion-and-conclusions}
+
+The results show a clear difference in the effect of the forecast
+horizon in the case of the clean data. The results based on the MSE and
+the p-value obtained from the Mincer-Zarnowitz test surprisingly show
+that the ARIMA model produces superior forecasts of the returns at the
+longest forecast horizon. Similarly, Prophet produces superior forecasts
+of the returns at the longest forecast horizon. We notice that although
+the ARIMA model outperforms Prophet at every forecast horizon, the
+difference between the MSE's of the forecasts generated by both models
+is insignificant. Furthermore, the Mincer-Zarnowitz p-values for both
+models are significant at the same forecast horizons, suggesting that
+the efficiency and biasness of the return forecasts yielded by ARIMA and
+Prophet are synchronized at each forecast horizon. If a small amount of
+predictive accuracy is willing to be sacrificed by the analyst in the
+forecasting of BTC/ZAR, it might be useful to employ Prophet instead of
+ARIMA. This would enable traders with expert knowledge about Bitcoin,
+however no statistical knowledge about forecasting models, to
+incorporate their domain knowledge into the Prophet model through its
+intuitively adjustable parameters. Furthermore, as more information
+about events that may affect the price of Bitcoin becomes available, it
+may be included in the model and could potentially change the predictive
+accuracy of Prophet.
+
+The results based on the unclean data show that both ARIMA and Prophet
+are sensitive to the presence of outliers and missing values in addition
+to the forecast horizon. We observe that Prophet yields superior
+forecasts at the longest forecast horizon. Although the MSE suggests
+that the ARIMA model produces forecasts with minimal error at the
+longest forecast horizon, analysis of the p-value indicates that the
+forecasts are efficient and/or unbiased at 30-steps ahead only. It is
+evident that Prophet produces inferior results in comparison to the
+ARIMA model over all forecast horizons. This is unexpected as Prophet is
+said to be robust to the presence of outliers and missing values in the
+data, without the need for interpolation. We also notice that the
+forecasts produced by both models deteriorates when there are missing
+values and outliers present in the data, with the only efficient and/or
+unbiased forecasts yielded by the ARIMA model at a forecast horizon of
+30 days.
 
 The limitation of this study is that it looks specifically at BTC/ZAR.
-Different results may emerge for fiat currencies or the exchange rate between Bitcoin with other currencies.
-Hence caution should be exercised when generalising these results to other exchange rates.
-As a point of further research, more complex techniques such as neural networks could be used as a benchmark for comparison when evaluating the forecasts produced by Prophet.
+Different results may emerge for fiat currencies or the exchange rate
+between Bitcoin with other currencies. Hence caution should be exercised
+when generalising these results to other exchange rates. As a point of
+further research, more complex techniques such as neural networks could
+be used as a benchmark for comparison when evaluating the forecasts
+produced by Prophet.
 
@@
 
-# References
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-abu1996}{}
+Abu-Mostafa, Yaser S, and Amir F Atiya. 1996. ``Introduction to
+Financial Forecasting.'' \emph{Applied Intelligence} 6 (3). Springer:
+205--13.
+
+\hypertarget{ref-anastasakis2009}{}
+Anastasakis, Leonidas, and Neil Mort. 2009. ``Exchange Rate Forecasting
+Using a Combined Parametric and Nonparametric Self-Organising Modelling
+Approach.'' \emph{Expert Systems with Applications} 36 (10). Elsevier:
+12001--11.
+
+\hypertarget{ref-castillo2002}{}
+Castillo, Oscar, and Patricia Melin. 2002. ``Hybrid Intelligent Systems
+for Time Series Prediction Using Neural Networks, Fuzzy Logic, and
+Fractal Theory.'' \emph{IEEE Transactions on Neural Networks} 13 (6).
+IEEE: 1395--1408.
+
+\hypertarget{ref-cheung1993}{}
+Cheung, Yin-Wong. 1993. ``Long Memory in Foreign-Exchange Rates.''
+\emph{Journal of Business \& Economic Statistics} 11 (1). Taylor \&
+Francis: 93--101.
+
+\hypertarget{ref-etuk2013}{}
+Etuk, Ette Harrison, Dagogo SA Wokoma, and Imoh Udo Moffat. 2013.
+``Additive Sarima Modelling of Monthly Nigerian Naira-Cfa Franc Exchange
+Rates.'' \emph{European Journal of Statistics and Probability} 1 (1):
+1--12.
+
+\hypertarget{ref-fahimifard2009}{}
+Fahimifard, SM, Masuod Homayounifar, M Sabouhi, and AR Moghaddamnia.
+2009. ``Comparison of Anfis, Ann, Garch and Arima Techniques to Exchange
+Rate Forecasting.'' \emph{Journal of Applied Sciences} 9 (20): 3641--51.
+
+\hypertarget{ref-spyros1997}{}
+Hibon, Michael, and Spyros Makridakis. 1997. ``ARMA Models and the
+Box--Jenkins Methodology.'' John Wiley \& Sons, Ltd.
+
+\hypertarget{ref-ince2006}{}
+Ince, Huseyin, and Theodore B Trafalis. 2006. ``A Hybrid Model for
+Exchange Rate Prediction.'' \emph{Decision Support Systems} 42 (2).
+Elsevier: 1054--62.
+
+\hypertarget{ref-kamruzzaman2003}{}
+Kamruzzaman, Joarder, and Ruhul A Sarker. 2003. ``Forecasting of
+Currency Exchange Rates Using Ann: A Case Study.'' In \emph{Neural
+Networks and Signal Processing, 2003. Proceedings of the 2003
+International Conference on}, 1:793--97. IEEE.
+
+\hypertarget{ref-khashei2011}{}
+Khashei, Mehdi, and Hehdi Bijari. 2011. ``Exchange Rate Forecasting
+Better with Hybrid Artificial Neural Networks Models.'' \emph{Journal of
+Mathematical and Computational Science} 1 (1). Science \& Knowledge
+Publishing Corporation Limited (SCIK): 103.
+
+\hypertarget{ref-khashei2009}{}
+Khashei, Mehdi, Mehdi Bijari, and Gholam Ali Raissi Ardali. 2009.
+``Improvement of Auto-Regressive Integrated Moving Average Models Using
+Fuzzy Logic and Artificial Neural Networks (Anns).''
+\emph{Neurocomputing} 72 (4). Elsevier: 956--67.
+
+\hypertarget{ref-khashei2012}{}
+---------. 2012. ``Hybridization of Autoregressive Integrated Moving
+Average (Arima) with Probabilistic Neural Networks (Pnns).''
+\emph{Computers \& Industrial Engineering} 63 (1). Elsevier: 37--45.
+
+\hypertarget{ref-lin2012}{}
+Lin, Chiun-Sin, Sheng-Hsiung Chiu, and Tzu-Yu Lin. 2012. ``Empirical
+Mode Decomposition--based Least Squares Support Vector Regression for
+Foreign Exchange Rate Forecasting.'' \emph{Economic Modelling} 29 (6).
+Elsevier: 2583--90.
+
+\hypertarget{ref-mincer1969}{}
+Mincer, Jacob A, and Victor Zarnowitz. 1969. ``The Evaluation of
+Economic Forecasts.'' In \emph{Economic Forecasts and Expectations:
+Analysis of Forecasting Behavior and Performance}, 3--46. NBER.
+
+\hypertarget{ref-nwankwo2014}{}
+Nwankwo, Steve C. 2014. ``Autoregressive Integrated Moving Average
+(Arima) Model for Exchange Rate (Naira to Dollar).'' \emph{Academic
+Journal of Interdisciplinary Studies} 3 (4): 429.
+
+\hypertarget{ref-santos2007}{}
+Santos, André Alves Portela, Newton Carneiro Affonso da Costa, and
+Leandro dos Santos Coelho. 2007. ``Computational Intelligence Approaches
+and Linear Models in Case Studies of Forecasting Exchange Rates.''
+\emph{Expert Systems with Applications} 33 (4). Elsevier: 816--23.
+
+\hypertarget{ref-taylor2017}{}
+Taylor, Sean J, and Benjamin Letham. 2017. ``Forecasting at Scale.''
+
+\hypertarget{ref-tseng2001}{}
+Tseng, Fang-Mei, Gwo-Hshiung Tzeng, Hsiao-Cheng Yu, and Benjamin JC
+Yuan. 2001. ``Fuzzy Arima Model for Forecasting the Foreign Exchange
+Market.'' \emph{Fuzzy Sets and Systems} 118 (1). Elsevier: 9--19.
+
+\hypertarget{ref-zhang1998}{}
+Zhang, Gioqinang, and Michael Y Hu. 1998. ``Neural Network Forecasting
+of the British Pound/Us Dollar Exchange Rate.'' \emph{Omega} 26 (4).
+Elsevier: 495--506.
+
+\newpage
+\renewcommand{\baselinestretch}{1}
+\nocite{*}
+\bibliography{}
+
+\end{document}

### 2017-09-14 Draft/Code/missing values and outliers.R

@@ -1,114 +1,16 @@
+#----------HOW ROBUST IS PROPHET WHEN THERE ARE MISSING VALUES AND OUTLIERS COMPARED TO ARIMA?-----------------
 
-# ------------------------------- READING IN DATASET --------------------------------------
+# Prophet is said to be robust when there are missing values whereas arima requires Kalman smoothing and interpolation
+# Let's test this out
 
-# Loading required packages
-library(rmsfuns)
-packagestoload <- c("stats", "xts", "dplyr", "Dplyr2Xts", "PerformanceAnalytics", "ggplot2", "readr", "gridExtra", "lmtest")
-load_pkg(packagelist = packagestoload) 
+missing_eval <- function(Data){
+  
+# Insert missing values
+ExchangeRates <- ExchangeRates %>% select(TradeDate, DailyLogReturns)
+ExchangeRates[c(4, 10, 17, 20, 35, 48, 50, 100, 110, 177, 140, 157, 200, 230, 245, 270, 286, 300, 323, 330),2] <- NA
 
-# Set relative path
-setwd("~/R/UCT_hons_proj/Draft/Data")
-ExchangeRates <- read.csv("bitxZAR.csv.gz", header = FALSE, sep = ",", dec = ".")
-
-# Looking at first few rows of dataset
-ExchangeRates %>% 
-  data.frame %>% 
-  head
-
-# Looking at structure of dataset
-dim(ExchangeRates)
-str(ExchangeRates)
-
-#--------------------------------- CLEANING THE DATASET --------------------------------------
-
-# Make unixtime stamp a POSIX object 
-ExchangeRates$V1 <- as.POSIXct(ExchangeRates$V1, tz = "GMT", origin = "1970-01-01")
-
-# Convert dataframe to xts
-ExchangeRates <- xts(ExchangeRates[,-1], order.by = ExchangeRates[,1])
-
-# Agregate prices by day
-ExchangeRates <- to.daily(ExchangeRates, period = "days", drop.time = TRUE, OHCL = TRUE, k = 1)
-
-# Convert xts back to dataframe
-ExchangeRates <- data.frame(date = index(ExchangeRates), coredata(ExchangeRates))
-
-# Set column names of dataframe
-ExchangeRates <- ExchangeRates %>% 
-  select(TradeDate = date, 
-         OpenPrice = ExchangeRates.Open, 
-         HighPrice = ExchangeRates.High, 
-         LowPrice = ExchangeRates.Low, 
-         ClosePrice = ExchangeRates.Close)
-
[…]

### 2017-10-27 Draft/Write-up/Draft.Rmd
Fixed about 80% of the mess. Still fixing the rest

@@
 ---
-title: "How does the latest Prophet forecasting model compare to the ARIMA model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz approach"
+title: "A comparison between the latest Prophet forecasting model and the ARIMA model: An evaluation of BTC/ZAR predictability"
 Author: "Fiona Ganie"  
@@
     template: Tex/TexDefault.txt
-    fig_width: 5 
-    fig_height: 4
+    fig_width: 4 
+    fig_height: 3
 abstract: |
@@
 Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task.
-Prophet is a sophisticated model that differs to traditional time series models in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process.
+Prophet is a sophisticated model that differs from traditional time series models in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process.
 Prophet is also robust to outliers and can handle missing values in the time series without the need for interpolation.
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 4, fig.height = 3, fig.pos="H", fig.pos = 'H')
 
+options(knitr.table.format = "latex")
 ```
@@
 # Introduction \label{Introduction}
-The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has attracted attention in financial market forecasting [@khashei2009].
+The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that have attracted attention in financial market forecasting [@khashei2009].
 Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that foreign exchange markets are not efficient and believe that future prices depend on current and past events [@abu1996].
@@
 The ARIMA model is tractable and less computationally expensive.
-It has been used as the building blocks for more advanced models and has provided the inspiration for hybrid versions of the model which have been used in exchange rate forecasting.
+It has been used as the building block for more advanced models and has provided the inspiration for hybrid versions of the model which have been used in exchange rate forecasting.
 
 In December 2016, Facebook open-sourced their forecasting model Prophet.
-Prophet differs to traditional time series model such as the ARIMA, in that it can produce high quality forecasts in a straightforward way.
+Prophet differs from traditional time series model such as the ARIMA in that it can produce high quality forecasts in a straightforward way.
 Prophet is a sophisticated model that provides informative results, however is configurable and easy to use.
 It can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process.
-This knowledge can be easily incorporated into the model through its intuitively adjustable parameters [@taylor2017].
+This knowledge can easily be incorporated into the model through its intuitively adjustable parameters [@taylor2017].
 Furthermore, Prophet is robust to outliers and can handle missing values in the time series without the need for interpolation [@taylor2017].
@@
 
-This paper broadly aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of the exchange rate between Bitcoin and the Rand (BTC/ZAR), using a Mincer-Zarnowitz approach of measuring forecast accuracy.
-More specifically, this paper aims to:
+This paper broadly aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of the exchange rate between Bitcoin and the Rand (BTC/ZAR), using the Model Confidence Set and a Mincer-Zarnowitz approach of measuring forecast accuracy.
+More specifically, this paper aims at:
 
-1. Compare the out-of-sample forecasts of ARIMA and Prophet over varying forecast horizons.
+1. Comparing the out-of-sample forecasts of ARIMA and Prophet over varying forecast horizons.
 
-2. Compare the out-of-sample forecast performance with missing values and outliers present in the data.
+2. Comparing the out-of-sample forecast performance with missing values and outliers present in the data.
 
-Section \ref{Background and Theory} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting, as well as the theory behind the ARIMA and Prophet model.
+Section \ref{Background} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting.
 Section \ref{Data} will provide a brief explanation and motivation for the dataset used in this paper.
-Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring forecast accuracy.
+Section \ref{Forecasting Methods} explains and compares the formulation of the ARIMA and Prophet model.
+Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models, the Model Confidence Set and the Mincer-Zarnowitz approach of measuring forecast accuracy.
 The results from applying the ARIMA and Prophet model to the dataset is presented in Section \ref{Results}.
-Finally, Section \ref{Discussion and Conclusions discusses the results in a statistical context.
+Finally, section \ref{Discussion and Conclusions} discusses the results and its financial implications in a statistical context.
 
-# Background and Theory \label{Background and Theory}
+# Background \label{Background}
 
 ## Forecasting Exchange Rates with the ARIMA model
-ARIMA models became highly popular since its introduction by Box and Jenkins, when it was shown that they could outperform complex econometric models in a variety of situations [@spyros1997].
+ARIMA models have commonly been used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012].
+The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001].
+They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time and have yielded satisfactory results when predicting exchange rates.
+@nwankwo2014 forecasted the exchange rate between the Naira and dollar, using Akaike's Information Criterion (AIC) as a measure of performance.
+Diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data.
+
+Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data.
+Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings.
+Some of the most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA (ARFIMA) model.
+
+The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing.
+@etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality.
+Their results showed that the SARIMA model adequately described the variation in the exchange rate series.
+
+The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain.
+The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993].
+This makes the model parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993].
+@cheung1993 fitted the ARFIMA model to examine five exchange rates and found that there was strong evidence of long memory in the exchange rate time series.
+
+## Other techniques used to forecast exchange rates
+Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed due to the failure of ARIMA models to capture the volatility in financial markets  [@anastasakis2009].
+@fahimifard2009 applied a GARCH model to the Rial/USD and Rial/EUR to compare the forecasts yielded by the GARCH and ARIMA model over varying forecast horizons.
+Their results illustrated that the GARCH model outperformed the ARIMA model when using the Root Mean Square Error (RMSE), Mean Square Error (MSE) and Mean Absolute Deviation (MAD) as a performance criteria.
+
+Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques.
+These complex techniques are non-linear and can fit complex time series more easily [@castillo2002].
+Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002].
+Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.
+
+Artificial Neural Networks (ANNs) mimic the structure of the brain and have had many successful applications in forecasting exchange rates.
+They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011].
+In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates.
+Superior results were produced when compared to the ARIMA and GARCH model using the RMSE, MSE and MAD as a measure of performance.
+
+Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms and have successfully been used in financial forecasting [@khashei2009].
+Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather than discrete decisions to be provided.
+@santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model.
+He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one-step ahead forecasts.
+By analysing accuracy statistics, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced.
+Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model and a Fuzzy ARIMA model.
+
+Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model.
+The performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998].
+Like ANNs, there is no systematic approach for designing FLSs and they are only understandable when simple.
+Although FLSs has the advantage over ARIMA models in that they can be applied to data with few observations available, it yields acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.
+
+The traditional ARIMA model produces less superior forecasts than its hybrid forms and other complex non-linear techniques, however its forecasts are still satisfactory.
+It is a simple model that is easy to implement and has a consistent method of model design and selection.
+ARIMA models are more robust and efficient than complex structural models in relation to short-run forecasting.
+The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison, justifies it as a good starting point to compare it to Facebook's forecasting method, Prophet, that was recently released.
+
+## Forecasting with Prophet
+The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series.
+Prophet differs from traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process.
+Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared.
+This is Facebook’s definition of forecasting at scale.
+
+Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge into the model building process and has produced significantly improved forecasts compared to the ARIMA model.
+@taylor2017 forecasted the number of events on Facebook using Prophet.
+The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend.
+The forecasts produced by Prophet were compared to common forecasting techniques such as exponential smoothing, ARIMA, the seasonal naive, and the naive model.
+While exponential smoothing and the seasonal naive model were quite robust, the ARIMA forecasts were fragile.
+No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.
+If Prophet produces forecasts that are as good as the ARIMA model when forecasting BTC/ZAR, it can be compared to more complex forecasting methods that are less tractable such as the hybrid models and machine learning techniques seen earlier.
+
+# Data \label{Data}
+
+The dataset used in this study consists of the daily closing prices of BTC/ZAR over the period 24 January 2016 to 17 July 2017.
+This comprises of a total of 541 trading days and was the chosen time period for analysis due to constraints in obtaining data over a longer period.
+The data was obtained from Bitcoincharts.
+
+Bitcoin is of specific interest as it presents an interesting parallel to traditional exchange rate markets.
+The cryptocurrency is built on a decentralised system and as a result its value cannot be directly influenced by a central authority [@fantazzini2016].
+In addition, the Bitcoin market has attracted attention worldwide and is currently the leading cryptocurrency, with awareness and adoption of the currency growing over time [@fantazzini2016].
+Its novelty makes it a highly volatile hence speculative market and provides an opportunity for forecasting.
+
+In this study, the analysis of BTC/ZAR forecasts are based on the daily continuous log returns.
+The daily closing prices are transformed into returns by taking the log difference at each time t and is calculated as follows:
+
+$$ r_t = ln(\frac{p_t}{p_{t-1}}) $$
+where $p_t$ represents the closing price and $r_t$ represents the log return for time t = 1,2,.,T.
+
+# Forecasting Methods \label{Forecasting Methods}
+## The ARIMA model
 The ARIMA model expresses the process {$y_t$} as a function of the weighted average of past values of the process and lagged values of the residuals.
@@
 The ARIMA model can be defined succinctly using the backward shift operator B, which shifts the process back by one unit of time, and is defined as $By_t = y_{t-1}$.
-The ARIMA model has the form:
+The ARIMA model has the form [@hyndman2014]:
 
@@
 
-ARIMA models have commonly been used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012].
-They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time.
-@kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies, and used the forecast errors as well as the accuracy of the direction of the forecasts to evaluate the performance of three Neural Network models.
-In a similar study, @khashei2012 compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN) to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark by claiming that it is the most important linear model.
-Their comparison was made by investigating the forecasts between the British pound and US dollar over varying forecast horizons.
-Not only has the ARIMA model been useful as a benchmark, but it has also yielded satisfactory results when predicting exchange rates.
-When @nwankwo2014 forecasted the rate between the Nairo and dollar, diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on Akaike’s Information Criterion (AIC).
-
-The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001].
-Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data.
-Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings.
-The most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).
-
-The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing.
-@etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality.
-Their results showed that the SARIMA model adequately described the variation in the exchange rate series.
-The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain.
-The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993].
-This makes the model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993].
-@cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.
-
-Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed in an attempt to capture the volatility in financial markets [@anastasakis2009].
-Hsieh (1989) applied a GARCH model to investigate five exchange rates and his results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.
-
-## Other techniques used to forecast exchange rates
-Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques.
-These more complex techniques are non-linear and can fit complex time series more easily [@castillo2002].
-Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002].
-Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.
-
-Artificial Neural Networks (ANNs) have had many successful applications in forecasting exchange rates and are more advantageous than other non-linear forecasting methods.
-They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011].
-In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates.
-Superior results were produced when compared to the ARIMA and GARCH model using the root mean square error (RMSE), mean square error (MSE) and mean absolute difference as a measure of performance.
-Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms, and have been successfully used in financial forecasting [@khashei2009].
-Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather than discrete decisions to be provided.
-@santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model.
-He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one step-ahead forecasts.
-By analysing accuracy statistics, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced.
-Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and a Fuzzy ARIMA model.
-
-Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model.
-Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998].
-Like ANNs there is no systematic approach to designing FLSs and they are only understandable when simple.
-Although FLSs has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.
-
-## Hybrid ARIMA Models
-Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA model, and of more complex non-linear models.
-By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations.
-Some of the hybrid models which have commonly been used in exchange rate forecasting are the Fuzzy ARIMA and ANN-ARIMA model.
-
-The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data to yield accurate results [@tseng2001].
-In a world that is constantly changing and with the rapid advancement of technology, access to large amounts of historical data is difficult to obtain.
-On the contrary, a fuzzy regression model requires little historical data however produces wide prediction intervals if extreme values are present in the data.
-The Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to exploit the advantages of both models while simultaneously overcoming their limitations.
-@tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model’s appropriateness and power.
-The Fuzzy ARIMA not only produced forecasts that were superior to the ARIMA and fuzzy time series models but also provided an upper and lower bound which can be used by decision makers to determine the best and worst possible situations.
-
-Although ARIMA models are powerful, they require non-stationary data to be differenced and impose prior assumptions onto the distribution of the data [@ince2006].
-In contrast, machine learning techniques such as ANNs do not impose any assumptions onto the data generating process however, being data-driven, are sensitive to the number of input nodes used.
-The ANN-ARIMA model draws on the strengths of both models and overcomes these individual difficulties.
-@ince2006 created an ANN-ARIMA model by using the ARIMA model to determine the number input nodes required by an ANN for three exchange rates.
-When the ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model outperformed the ARIMA based on the MSE.
-
-Although the traditional ARIMA models produce less superior forecasts than its hybrid forms and other complex non-linear techniques, its forecasts are still satisfactory.
-They are simple models that are easy to implement and have a consistent method of model design and selection.
-ARIMA models are also more robust and efficient than complex structural models in relation to short-run forecasting.
-The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison, justifies it as a good starting point to compare it to Facebook’s forecasting method, Prophet, that was recently released.
-
-## Forecasting with Prophet 
-The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series.
-Prophet differs to traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process.
-Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared.
-This is Facebook’s definition of forecasting at scale.
+## The Prophet model 
 Prophet is similar to a Generalized Additive Model (GAM) - an additive regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017].
@@
 The regression model also provides model flexibility and allows the analyst to interactively change model parameters [@taylor2017].
-
 The growth component of Prophet may be modelled as a linear or non-linear function of time.
@@
 
-Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge into the model building process and has produced significantly improved forecasts compared to the ARIMA model.
-@taylor2017 forecasted the number of events on Facebook using Prophet.
-The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend.
-The forecasts produced by Prophet were compared to common forecasting techniques such as exponential smoothing, ARIMA, the seasonal naïve, and the naïve model.
-While exponential smoothing and the seasonal naïve model were quite robust, the ARIMA forecasts were fragile.
-No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.
-
 Prophet’s ability to forecast at scale enables it to model a wide variety of data and may be able to adequately fit exchange rate data.
@@
 
-# Data \label{Data}
-
-The data set used in this study consists of the daily closing prices of BTC/ZAR over the period 24 January 2016 to 17 July 2017.
-This comprises of a total of 541 trading days and was the chosen time period for analysis due to constraints in obtaining data over a longer period.
-The data was obtained from Bitcoincharts.
-
-Bitcoin is of specific interest as it presents an interesting parallel to traditional exchange rate markets.
-The cryptocurrency is built on a decentralised system and as a result its value cannot be directly influenced by a central authority (Fantazzini and Nigmatullin 2016).
-In addition, the Bitcoin market has attracted attention worldwide and is currently the leading cryptocurrency, with awareness and adoption of the currency growing over time (Fantazzini and Nigmatullin 2016).
-Its novelty makes it a highly volatile hence speculative market and provides an opportunity for forecasting.
-
-In this study, the analysis of BTC/ZAR forecasts are based on the daily continuous log returns.
-The daily closing prices are transformed into returns by taking the log difference at each time t and is calculated as follows:
-
-$$ r_t = ln(\frac{p_t}{p_{t-1}}) $$
-where $p_t$ represents the closing price and $r_t$ represents the log return for time t = 1,2,.,T.
-
 # Methodology \label{Methodology}
@@
 This allows us to examine the forecast horizon effect.
-The Mincer-Zarnowitz test is then used to evaluate the forecasts produced by ARIMA and Prophet against the test set.
-Since there is no consensus on which accuracy statistic best measures the performance of forecasting techniques, the most popular criteria, the Mean Squared Error (MSE) is employed.
+The Model Confidence Set and Mincer-Zarnowitz test is then used to evaluate the forecasts produced by ARIMA and Prophet against the test set.
+Since there is no consensus on which accuracy statistic best measures the performance of forecasting techniques, the most popular criteria, the Root Mean Squared Error (RMSE) is employed.
 The results obtained from the ARIMA model will be used as a benchmark for comparison.
@@
 ## The Box-Jenkins Methodology 
-Box and Jenkins proposed a set of guidelines that can be followed when selecting ARIMA models.
-This systematic procedure to designing ARIMA models has made them highly popular [@spyros1997].
-It consists of a four-stage iterative process in which:
+@box1970 proposed a set of guidelines that can be followed when selecting ARIMA models. It consists of a four-stage iterative process in which:
 
 1. The process is either transformed or differenced to de-trend and stabilise the variance of the data. 
-
 2. The autocorrelation and partial autocorrelation plots are used to determine the order of p and q. 
 3. The parameters of the model are then estimated.
-
 4. A diagnostic check is performed to ensure that the residuals are a white noise process.
@@
 If the residuals are not white noise, steps 2-4 are repeated until a satisfactory model is identified.
-On the contrary, if the diagnostic check shows that the residuals are random, the developed model will be the final model used for forecasting.
+On the contrary, if the diagnostic check reveals that the residuals are random, the developed model will be the final model used for forecasting.
 
@@
 
-* The trend is set to be linear
-* The width of the uncertainty intervals is set to 80% 
+* The trend is set to be linear.
+* The width of the uncertainty intervals is set to 80%. 
 * Weekly and yearly seasonality are automatically detected and included in the model if present 
-* The smoothing parameter for holidays and seasonality is set at 10 
-* The smoothing parameter for the trend is set to 0.05
-* The number of potential changepoints is set to 25
+* The smoothing parameter for holidays and seasonality is set at 10 while the smoothing parameter for the trend is set to 0.05.
+* The number of potential changepoints is set to 25.
 
 This paper makes use of Prophet’s default settings to fit the data, however weekly seasonality is included in the model.
-This allows us to account for our prior knowledge about the closing prices of Bitcoin which tend to be lower around the weekend however higher mid-week, due to fluctuations in trading activity.
+This allows us to account for our prior knowledge about the closing prices of Bitcoin which tend to be lower around the weekend however higher mid-week due to fluctuations in trading activity.
 Furthermore, a linear trend is appropriate since Bitcoin does not have an upper limit on its closing prices.
@@
 In a rolling window forecast, the forecasts are made h-steps at a time and the actual observation rather than the predicted value is used for the next prediction in the forecast horizon.
-In this way, a poor forecast will not have negative consequences on future forecasts to be made, since the observed values will be used to correct itself for the remaining forecasts [@zhang1998].
-In this paper, we first fit the ARIMA and Prophet model to the training set and forecast 1 day ahead.
-We then refit the model to the training set, increasing its size by one observation, and the next 1 day ahead forecast is made.
-This process is repeated until all forecasts have been made into the test set.
-The 1 day ahead forecasts are then compared to the 1 day ahead observed values in the test set and the error statistics are calculated.
-This procedure is then repeated for the 30 day and 90 day ahead forecasts.
+In this way, a poor forecast will not have negative consequences on future forecasts to be made since the observed values will be used to correct itself for the remaining forecasts [@zhang1998].
+In this paper, the ARIMA and Prophet model is first fitted to the training set and the returns are forecasted h-days ahead for h = 1, 30, 90.
+The size of the training set is then increased by one observation and the models are refitted.
+The next h-day ahead return is forecasted and this process is repeated until all forecasts have been made into the test set.
+The h-day ahead forecasts are then compared to the h-day ahead observed values in the test set and the forecast errors are calculated to determine the RMSE.
+
+## Model Confidence Set
+The Model Confidence Set (MCS) procedure developed by @hansen2011 is a sequential procedure which starts with an initial set of models $M_0$ of size m.
+The worse model is then eliminated at each step until the null hypothesis of equal predictive ability (EPA) fails to be rejected for all the models belonging to the set at a given confidence level $1 - \alpha$.
+This smaller set of models $M^*$ is the superior set of models (SSM) with the best case occurring when the SSM consists of a single best model.
+
+The null hypothesis for EPA for a set of models $M$ is given by: 
+$$H_0: \mu_i,_j = 0 \;\;\; \forall \;\; i,j = 1,...,m$$
+
+where $\mu_i,_j = E[d_i,_j]$ represents the expected loss differential, $d_i,_j,_t = l_i,_t - l_j,_t$ represents the loss differential between models $i$ and $j$ at time t, and $l_i,_t$ and $l_j,_t$ are the loss functions associated with model $i$ and $j$ at time t respectively.
+This paper makes use of the squared forecast error as a loss function, however any arbitrary loss function that satisfies the weak stationarity conditions described in @hansen2011 may be used.
+The null hypothesis is tested by constructing the following test statistics:
+
+$$t_i,_j = \frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}$$ \vspace{1pt}
+$$ T_R = \max_{i,j\in M} \;\; |t_i,_j|$$
+where $d_i,_j = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }$ measures the sample average loss differential between model $i$ and $j$ and  $\hat{var[\bar{d_i,_j}]}$ is estimated through bootstrapping as described in @hansen2011.
+
+This paper uses a block–bootstrap procedure of 5000 resamples as in @hansen2011.
+The block length $p$ is chosen by fitting an AR(p) model to the $d_i,_j$ terms and determining the maximum number of significant AR parameters obtained.
+While $t_i,_j$ has a t-distribution, $T_R$ has a non-standard distribution under the null hypothesis hence is estimated by implementing a bootstrapping procedure similar to that used to estimate $\hat{var[\bar{d_i,_j}]}$.
+
+The worst model is then eliminated according to an elimination rule which is coherent with the calculated t-statistic and is defined as:
+
+$$e_R = \arg max_{i}\bigg(\sup_{j\in M}\frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}\bigg)$$
 
@@
 The Mincer-Zarnowitz approach to evaluating forecast accuracy is commonly used and can be useful when comparing the forecasts produced by Prophet and the ARIMA model.
-@mincer1969 proposed an absolute measure to evaluate forecast accuracy which considers the distance between the actual and predicted values.
-To analyse the absolute errors produced by the forecasts, the observed values $r_t$ are regressed against the predicted values $r_t$, i.e.
-
+@mincer1969 proposed an absolute measure to evaluate forecast accuracy that considers the distance between the actual and predicted values.
+To analyse the absolute errors produced by the forecasts, the observed values $r_t$ are regressed against the predicted values $\hat{r_t}$, i.e.
 $$r_t = \alpha + \beta\cdot\hat{r_t} + e_t$$
-
-where $\alpha$ represents the mean distance between the observed and predicted values while $\beta$ represents the correlation between the forecasted errors and the predicted values.
-When alpha = 0 it implies that the forecasts are unbiased and do not systematically overestimate or underestimate the data.
-When b=1 it implies that the forecasts are efficient and uncorrelated with the forecasted errors.
+where $\alpha$ represents the mean distance between the observed and predicted values while $\beta$ represents the correlation between the forecasted errors and predicted values.
+When $\alpha = 0$ it implies that the forecasts are unbiased and do not systematically overestimate or underestimate the data.
+When $\beta = 1$ it implies that the forecasts are efficient and uncorrelated with the forecasted errors.
 A joint hypothesis test of $H_0: \alpha = 0 \cup \beta = 1$ is performed to check the efficiency and bias of the forecasts and the model which produces the best results generates superior forecasts.
@@
 ## Evaluation of forecasts based on data with missing values and outliers
-The original dataset is modified in order to examine how robust Prophet and ARIMA are when fitted to data with missing values and outliers.
+The original dataset is modified in order to examine how robust the ARIMA and Prophet model are when faced with missing values and outliers.
 20 values are randomly removed from the training set and 5 outliers are randomly inserted in place of existing values.
-The models are then fitted to this “dirty” dataset and evaluating of forecasts are carried out as before.
+Both models are then fitted to this “dirty” dataset and the forecasts are evaluated in the same way as the "clean" dataset.
 
@@
 
-## Evaluation over varying forecast horizons
-Although the Box-Jenkins methodology resulted in an ARIMA(0,0,1) model being the best fit to the training set, an ARIMA(1,0,0) is employed as the benchmark for comparison so that a rolling window forecast may be conducted.
-The AIC of the ARIMA(1,0,0) is marginally higher than that of the ARIMA(0,0,1) and analysis of the p-value indicates that the AR parameter is statistically significant in the explanation of the movement of BTC/ZAR.
-The Prophet model fitted to the training set included a linear trend and weekly seasonality, with all other parameters set to the default values as described in section \ref{Semi-Automatic Selection}.
-The automatic-selection procedure for ARIMA and Prophet selected the same model as those chosen manually.
+## Evaluation of clean dataset over varying forecast horizons
+By following the Box-Jenkins methodology, an ARIMA(0,0,1) model was found to be the best fit for the clean training set.
+Making use of a mean model would result in a returns forecast of zero in a rolling window context.
+This is due to the white noise behaviour of log returns.
+Hence a decision was made to employ the ARIMA(1,0,0) model instead.
+The ARIMA(1,0,0) model has a marginally higher AIC and an analysis of the p-value indicated that the AR parameter is statistically significant in the explanation of the movement of BTC/ZAR.
+The Prophet model fitted to the clean training set included a linear trend and weekly seasonality with all other parameters set to the default values as described in section \ref{Semi-Automatic Selection}.
+The models chosen manually are consistent with the models selected by the automatic-selection procedure for ARIMA and Prophet.
 
-### 1-step ahead Forecasts
-Figure 1 depicts the resulting forecasts from fitting both models.
-The forecasts produced by the ARIMA model and Prophet lie within a band centred around zero.
-Both models fail to capture the large movement of returns away from zero.
-The 80% uncertainty intervals produced by both models are approximately the same width and begin to widen as forecasts are made further into the test set.
+Table \ref{tab1} displays the MZ results obtained by regressing the test set returns against the forecasted returns, the RMSE, and the models that are included in the MCS.
+It is evident that the ARIMA model ranks ahead of the Prophet model over all forecast horizons.
+The MCS p-values indicates that both models lie within the 90% confidence interval at the 1 day and 3 month forecast horizon, while the Prophet model is eliminated from the MCS at the 1 month forecast horizon.
+This suggests that the ARIMA and Prophet model have equal predictive accuracy at the 1 day and 3 month forecast horizon, and that the ARIMA model produces superior forecasts at the 1 month forecast horizon.
+Furthermore, the MCS p-values indicates that while ARIMA model lies within the 90% confidence interval with certainty across all forecast horizons, the Prophet model lies within the 90% confidence interval 52.44% of the time at the 1 day forecast horizon and 45.78% of the time at the 3 month forecast horizon.
 
-![1-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\1-step ahead forecasts.png)
+```{r results = 'asis'}
 
-Table 1 displays the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the one-step ahead forecasts.
-The intercept estimate is small and positive for ARIMA and Prophet, with the intercept for ARIMA being closer to zero than Prophet.
-This indicates that both models systematically underestimate the 1-step ahead returns, with ARIMA yielding mean forecasts closer to the observed mean.
-The slope estimate is negative for both models indicating that the observed returns are actually the opposite of the suggested forecast.
-The F-test for the joint hypothesis of a unity slope and zero intercept is rejected at the 5% significance level for both models, with the p-value for ARIMA being much larger than Prophet.
-Hence the forecasts generated by both ARIMA and Prophet are biased and/or inefficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
+library(dplyr)
 
-![Mincer-Zarnowitz results for 1-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 1.png)
+data <- read.csv2("~/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Clean Results.csv", header = TRUE, sep = ";", dec = ".")
 
-### 30-steps ahead Forecasts
-Figure 2 illustrates the resulting 30-step ahead forecasts produced from fitting ARIMA and Prophet.
-As in the 1-step ahead case, both models fail to capture the large movement of returns away from zero with its forecasts lying within a band centred around zero.
-The 80% uncertainty intervals produced by both models are approximately the same width and marginally wider than the uncertainty intervals produced by the 1-step ahead forecasts.
+colnames(data) <- c("Forecast", "Model", "Model", "MCS", "$\\alpha$", "$\\beta$", "MZ-test", "RMSE")
 
-![30-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\30-step ahead forecasts.png)
+library(xtable)
 
-Table 2 displays the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 30-step ahead forecasts.
-The intercept estimate is small and positive for ARIMA and Prophet, with a smaller intercept estimated for Prophet.
-This indicates that both models systematically underestimate the 30-step ahead returns, with Prophet yielding mean forecasts closer to the observed mean.
-The slope estimate is large and negative for both models with the slope estimated to be as low as -12.74 for ARIMA.
-This suggests that the observed returns are actually the opposite of the proposed forecast.
-The F-test for the joint hypothesis of a unity slope and zero intercept is rejected at the 5% significance level for Prophet only.
-The p-value of 0.0614 from the joint hypothesis test is slightly significant for ARIMA.
-Hence the forecasts generated by Prophet are biased and/or inefficient while the forecasts generated by ARIMA are unbiased and/or efficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
+table <- xtable(data, caption = "MCS, MZ-test and RMSE results based on clean data \\label{tab1}", auto = T)
 
-![Mincer-Zarnowitz results for 30-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 2.png)
+align(table) <- rep("c", 9)
 
-### 90-steps ahead Forecasts
-Figure 3 illustrates the 90-step ahead forecasts produced by ARIMA and Prophet.
-Both models fail to capture large movement of returns away from zero as in the case of the 1-step ahead and 30-step ahead forecasts.
-The 80% uncertainty intervals produced by both models are approximately the same width and are marginally narrower than the uncertainty intervals produced by the one-step ahead and 30-step ahead forecasts.
+print.xtable(table, 
+             floating = TRUE,
+             table.placement = 'H', 
+             comment = FALSE,
+             include.rownames = getOption("xtable.include.rownames", F),
+             caption.placement = 'top',
+             digits = 5,
+             scalebox = 0.9,
+             sanitize.text.function = function(x){x},
+             hline.after = c(-1,1,3,5,7)
+             )
+```
 
-![90-step ahead forecasts from the ARIMA(1,0,0) and Prophet model](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\90-steps ahead forecasts.png)
+The intercept estimate represented by $\alpha$ is small and positive for ARIMA and Prophet across all forecast horizons, with the smallest estimates obtained at the longest forecast horizon.
+This suggests that both models systematically underestimates the returns, but to a smaller degree at long forecast horizons.
+The intercept estimate for ARIMA is closer to zero compared to Prophet at the 1 day and 3 month forecast horizon, indicating that the mean forecasts yielded by ARIMA is closer to the observed mean than Prophet.
+The slope estimate represented by $\beta$ is negative for both models across all forecast horizons, indicating that the observed returns are actually the opposite of the suggested forecast and that both models fail to capture the explosive behaviour of BTC/ZAR.
 
-Table 3 presents the MSE and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 90-step ahead forecasts.
-The intercept estimate is very small and positive for ARIMA and Prophet, with a marginally smaller intercept estimated for ARIMA.
-This indicates that both models systematically underestimate the 90-step ahead returns, with ARIMA yielding mean forecasts closer to the observed mean.
-The slope estimate is small and negative for both models with the slope estimated for ARIMA lying closer to -1.
-This suggests that the observed 90-steps ahead returns are actually the opposite of the proposed 90-steps ahead forecasts.
-The F-test for the joint hypothesis of a unity slope and zero intercept fails to be rejected at the 5% significance level for both ARIMA and Prophet, with the p-value from the test being much larger for ARIMA.
-Hence the forecasts generated by Prophet and ARIMA are unbiased and/or efficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
+The MZ p-values at the 1 day and 1 month forecast horizon show that the joint hypothesis of a unity slope and zero intercept is rejected at the 10% significance level for both models.
+This suggests that the forecasts generated by ARIMA and Prophet are biased and/or inefficient.
+Both models only produce unbiased and/or efficient forecasts at the longest forecast horizon, with the ARIMA model yielding a considerably larger p-value compared to Prophet.
+The RMSE produced by the ARIMA model is marginally smaller than Prophet across all forecasts horizons, with the smallest RMSE calculated at the longest forecast horizon.
+These results agree with the results obtained from the MCS and the MZ-test.
 
-![Mincer-Zarnowitz results for 90-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 3.png)
+## Evaluation of data with missing values and outliers present
+By following the Box-Jenkins methodology, an ARIMA(0,0,3) model was found to be the best fit for the dirty training set.
+A decision was made to employ the ARIMA(3,0,0) model instead.
+The motivation behind this choice is the same as that provided in the case of the clean data.
+The ARIMA(3,0,0) model has a marginally higher AIC and an analysis of the p-value indicated that the AR parameters are statistically significant in the explanation of the movement of BTC/ZAR.
+The Prophet model fitted to the clean training set included a linear trend and weekly seasonality with all other parameters set to the default values as described in section \ref{Semi-Automatic Selection}.
+The models chosen manually are consistent with the models selected by the automatic-selection procedure for ARIMA and Prophet.
 
-### Comparison over varying forecast horizons
-Table 4 and 5 compares the p-values obtained from the Mincer-Zarnowitz tests and the MSE of the forecasts produced by ARIMA and Prophet over the three forecast horizons.
-In comparing the forecasting accuracy of the ARIMA model over the three forecast horizons, we observe that the p-value gets more significant as the forecast horizon gets longer.
-This is in contrast to the MSE of the ARIMA forecasts which is smaller at the shortest and longest forecast horizons.
-In comparing the forecasting accuracy of Prophet, we find that the p-value is more significant for the shortest and longest forecast horizon.
-This agrees with the MSE of the Prophet forecasts which is smaller at the longest and shortest forecast horizon.
-Both ARIMA and Prophet generate forecasts with the smallest MSE and the most significant p-value at the longest forecast horizon.
-In comparing the forecasting accuracy of ARIMA and Prophet over the three forecast horizons, we notice that the p-value obtained from the Mincer-Zarnowitz results is more significant for ARIMA than Prophet at every forecast horizon.
-This echoes the results obtained from the MSE which is marginally smaller for ARIMA than Prophet over all forecast horizons.
+Table \ref{tab2} displays the MZ results obtained by regressing the test set returns against the forecasted returns, the RMSE and the models that are included in the MCS.
+We observe that the presence of outliers and missing values in the data has no effect on the rank of the ARIMA and Prophet model.
+Contrary to the results from the clean data, the Prophet model lies within the 90% confidence interval at the 1 day forecast horizon only.
+This suggests that the ARIMA and Prophet model only have equal predictive accuracy at the shortest forecast horizon.
+The MCS p-values indicates that while the ARIMA model always lies within the 90% confidence interval with certainty, the Prophet model only lies within the confidence interval at the 1 day forecast horizon 41.3% of the time.
+This differs to the results from the clean data in which the Prophet model lies within the 90% confidence interval 52.44% of the time at the 1 day forecast horizon.
 
-![Mincer-Zarnowitz results for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 4.png)
+```{r table2, results = 'asis'}
 
-![MSE for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 5.png)
+library(dplyr)
 
-## Evaluation on data with missing values and outliers present
-By following the Box-Jenkins methodology for selection of an ARIMA model, an ARIMA(3,0,0) was employed to fit the training set which included missing values and outliers.
-Although the ARIMA(0,0,3) had a slightly smaller AIC, a decision was made to use the ARIMA(3,0,0) instead so that a rolling window forecast may be used.
-Analysis of the AR parameters indicate that they are statistically significant in the explanation of the movement of BTC/ZAR.
-The Prophet model fitted to the training set included a linear trend and weekly seasonality.
-All other parameters were set to Prophet’s default values as described in section \ref{Semi-Automatic Selection}.
-The automatic-selection procedure for Prophet and ARIMA selected the same models as those chosen manually.
+data2 <- read.csv2("~/R/UCT_hons_proj/Draft/Figures/Evaluation with missing values and outliers/Dirty Results.csv", header = TRUE, sep = ";", dec = ".")
 
-### 1-step ahead Forecasts
-Table 6 shows the MSE of the forecasts generated by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 1-step ahead returns forecasted.
-We observe that a positive intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically underestimate the 1-step ahead returns.
-The intercept estimated for ARIMA is smaller than Prophet, indicating that the mean forecasted returns yielded by ARIMA is closer to the mean observed returns.
-The intercept estimates of 0.011895 and 0.014802 for ARIMA and Prophet respectively are larger than the intercept estimates of 0.009546 and 0.013702, which was based on the “clean” dataset.
-This suggests that both models systematically underestimates the returns more when there are missing values and outliers present in the data.
+colnames(data2) <- c("Forecast", "Model", "Model", "MCS", "$\\alpha$", "$\\beta$", "MZ-test", "RMSE")
 
-We also observe that the slope estimate is small and negative for both models with the slope estimated for ARIMA lying closer to -1.
-This suggests that the observed 1-step ahead returns are actually the opposite of the proposed 1-step ahead forecasts.
-The slope estimates of -0.883228 and -0.676132 for ARIMA and Prophet respectively are less steep than the slope estimates of -2.005624 and -1.709744, which was based on the evaluation of data that had no values missing or outliers present.
+library(xtable)
 
-![Mincer-Zarnowitz results for 1-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 6.png)
+table2 <- xtable(data2, caption = "MCS, MZ-test and RMSE results based on dirty data \\label{tab2}", auto = T)
 
-The p-values obtained from the F-test for the joint hypothesis of a unity slope and zero intercept is approximately zero for ARIMA and Prophet, with the p-value for Prophet lying closer to zero than ARIMA.
-The null hypothesis is thus rejected at the 1% significance level, indicating that the forecasts generated by Prophet and ARIMA are biased and/or inefficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
-The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on the “clean” dataset show similar findings, however led to a smaller MSE of 0.001942 and 0.001958 and a more significant p-value of 0.025449 and 0.007152 for ARIMA and Prophet respectively.
+align(table2) <- rep("c", 9)
 
-### 30-steps ahead Forecasts
-Table 7 shows the MSE of the forecasts produced by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 30-step ahead returns forecasted.
-We observe that a positive intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically underestimate the 30-step ahead returns.
-The intercept estimated for ARIMA is larger than Prophet, indicating that the mean forecasted returns yielded by Prophet is closer to the mean observed returns.
-The intercept estimate of 0.11414 for ARIMA is larger than the estimate of 0.031320 which was based on the evaluation of the “clean” dataset.
-This suggests that the ARIMA model underestimates the returns more when there are missing values and outliers present in the data.
-The intercept estimate of 0.014396 for Prophet is less than the estimate of 0.017169 which was based on the evaluation of data that had no values missing or outliers present.
-This suggests that Prophet underestimates the returns less when there are missing values and outliers present in the data.
+print.xtable(table2, 
+             floating = TRUE,
+             table.placement = 'H', 
+             comment = FALSE,
+             include.rownames = getOption("xtable.include.rownames", F),
+             caption.placement = 'top',
+             digits = 5,
+             scalebox = 0.9,
+             sanitize.text.function = function(x){x},
+             hline.after = c(-1,1,3,5,7)
+             )
+```
 
-![Mincer-Zarnowitz results for 30-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 7.png)
+We notice that a positive intercept and negative slope is still estimated at the 1 day and 1 month forecast horizon when the data is dirty.
+The intercept estimates are larger than the case of the clean data, suggesting that the biasness of the forecasts yielded by both models increases when there are outliers and missing values present in the data.
+Contrary to the results from the clean data, Prophet’s intercept estimate lies closer to zero at the 1 month and 3 month forecast horizon.
+This indicates that the mean forecasted returns yielded by Prophet is closer to the mean observed returns when outliers and missing values are present in the data.
+The negative intercept and positive slope estimated for the 3 month horizon contrasts with the positive intercept and negative slope estimated in the case of the clean data.
+This shows that although the forecasts produced by both models are systematically overestimated at long forecast horizons when the data is dirty, they are no longer the opposite of the observed returns.
 
-We also observe that the slope estimate is large and negative for ARIMA and small and negative for Prophet.
-This suggests that the observed 30-step ahead returns are actually the opposite of the proposed 30-step ahead forecasts.
-The slope estimate of -12.985335 for ARIMA is more steep than the slope estimate of -12.736757 which was based on the evaluation of data that had no values missing or outliers present.
-This suggests that the efficiency of the forecasted returns deteriorates when there are missing values and outliers present in the data.
-The slope estimate of -0.65161 for Prophet is less steep than the estimate -2.621914, which was based on the evaluation of data that had no values missing or outliers present.
-This suggests that the efficiency of the forecasted returns improve when there are missing values and outliers present in the data.
-
-The p-value obtained from the F-test for the joint hypothesis of a unity slope and zero intercept is approximately zero for Prophet.
-The null hypothesis is thus rejected at the 1% significance level, indicating that the 30-step ahead forecasts generated by Prophet are biased and/or inefficient.
-The p-value of 0.107532 for ARIMA is significant at the 5% level, suggesting at the 30-step ahead forecasts produced by ARIMA are unbiased and/or efficient.
-MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a smaller MSE than Prophet.
-The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on the “clean” dataset, led to a more significant p-value of 0.000377 for Prophet only, and a smaller MSE of 0.001988 and 0.002075 for ARIMA and Prophet respectively.
-
-### 90-steps ahead Forecasts
-Table 8 shows the MSE of the forecasts produced by ARIMA and Prophet and the Mincer-Zarnowitz results obtained from regressing the test set returns against the 90-step ahead returns forecasted.
-We observe that a negative intercept is estimated in the case of ARIMA and Prophet, indicating that both models systematically overestimate the 90-step ahead returns.
-The intercept estimated for Prophet lies closer to zero, indicating that the mean forecasted returns yielded by Prophet is closer to the mean observed returns.
-This is in contrast to the positive intercepts that were estimated based on the data that had no values missing or outliers present.
-
-We also observe that the slope estimate is large and positive for ARIMA however small and positive for Prophet.
-This is in contrast to the small and negative slope estimates which were based on the evaluation of data that had no values missing or outliers present.
-
-![Mincer-Zarnowitz results for 90-step ahead forecasts](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 8.png)
-
-The p-values obtained from the F-test for the joint hypothesis of a unity slope and zero intercept are insignificant for ARIMA and Prophet, with the p-value for Prophet lying close to zero.
-The null hypothesis is thus rejected at the 5% significance level, indicating that the 90-steps ahead forecasts generated by Prophet and ARIMA are biased and/or inefficient.
-The MSE produced by both models agree with the Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE than Prophet.
-The results obtained from evaluating the forecasting accuracy of ARIMA and Prophet based on data which had no missing values or outliers present contrasts these findings, with a p-value of 0.843066 and 0.408675 for ARIMA and Prophet respectively.
-
-### Comparison over varying forecast horizons
-Table 9 and 10 compares the p-values obtained from the Mincer-Zarnowitz tests and the MSE of the forecasts produced by ARIMA and Prophet over the three forecast horizons.
-In comparing the forecasting accuracy of the ARIMA model over the three forecast horizons, we observe that the p-value is more significant at the longer forecast horizons.
-This agrees with the MSE of the ARIMA forecasts which is smaller at the longer forecast horizons.
-In comparing the forecasting accuracy of Prophet, we find that the p-value is more significant for the longest forecast horizon.
-This agrees with the MSE of the Prophet forecasts which is smaller at the longest forecast horizon.
-
-![Mincer-Zarnowitz results for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 9.png)
-
-![MSE for all forecast horizons](C:\Users\Fiona\Documents\R\UCT_hons_proj\Draft\Figures\Evaluation of Prophet and ARIMA forecasts\Table 10.png)
-
-In comparing the forecasting accuracy of ARIMA and Prophet over the three forecast horizons, we notice that the p-value obtained from the Mincer-Zarnowitz results is more significant for ARIMA than Prophet at every forecast horizon.
-This echoes the results obtained from the MSE which is marginally smaller for ARIMA than Prophet over all forecast horizons.
+The p-values obtained from the F-test for the joint hypothesis of a unity slope and zero intercept is approximately zero for ARIMA and Prophet across all forecast horizons, with the p-value for Prophet lying closer to zero than ARIMA.
+The null hypothesis is thus rejected at the 10% significance level, indicating that the forecasts generated by Prophet and ARIMA are always biased and/or inefficient.
+This contrasts with the clean data which generated unbiased and/or efficient forecasts at the longest forecast horizon.
+The RMSE produced by both models agrees with the MZ-test and MCS results, with ARIMA producing a marginally smaller RMSE than Prophet across all forecast horizons, and the RMSE being larger than the case of the clean data.
 
 # Discussion and Conclusions \label{Discussion and Conclusions}
-The results show a clear difference in the effect of the forecast horizon in the case of the clean data.
-The results based on the MSE and the p-value obtained from the Mincer-Zarnowitz test surprisingly show that the ARIMA model produces superior forecasts of the returns at the longest forecast horizon.
-Similarly, Prophet produces superior forecasts of the returns at the longest forecast horizon.
-We notice that although the ARIMA model outperforms Prophet at every forecast horizon, the difference between the MSE’s of the forecasts generated by both models is insignificant.
+The results show a clear difference in the predictive accuracy of the ARIMA and Prophet model in addition to the forecast horizon effect.
+The ARIMA model ranks ahead of Prophet when forecasting the returns of BTC/ZAR across all forecast horizons.
+These results hold true irrespective of whether there are missing values and outliers present in the data.
+Nevertheless, it is reassuring to see that both models have equal predictive accuracy when forecasting at short and long forecast horizons when the data is clean and at short forecast horizons when the data is dirty.
 Furthermore, the Mincer-Zarnowitz p-values for both models are significant at the same forecast horizons, suggesting that the efficiency and biasness of the return forecasts yielded by ARIMA and Prophet are synchronized at each forecast horizon.
-If a small amount of predictive accuracy is willing to be sacrificed by the analyst in the forecasting of BTC/ZAR, it might be useful to employ Prophet instead of ARIMA.
-This would enable traders with expert knowledge about Bitcoin, however no statistical knowledge about forecasting models, to incorporate their domain knowledge into the Prophet model through its intuitively adjustable parameters.
-Furthermore, as more information about events that may affect the price of Bitcoin becomes available, it may be included in the model and could potentially change the predictive accuracy of Prophet.
+Hence, if the analyst is engaged in short-term or long-term trading, they could choose to employ the Prophet model instead of the ARIMA model.
+This would allow them to produce return forecasts that are as good as the ARIMA model, without having any expert knowledge about model building and selection.
+It would also provide them with the flexibility to customise the model to suit their needs and incorporate their domain knowledge about BTC/ZAR through the model’s intuitively adjustable parameters.
+If the analyst is interested in generating unbiased and/or efficient forecasts with the least amount of forecast error, they could utilise either model, however should only forecast BTC/ZAR returns over long forecast horizons with data that do not include any outliers or missing values.
 
-The results based on the unclean data show that both ARIMA and Prophet are sensitive to the presence of outliers and missing values in addition to the forecast horizon.
-We observe that Prophet yields superior forecasts at the longest forecast horizon.
-Although the MSE suggests that the ARIMA model produces forecasts with minimal error at the longest forecast horizon, analysis of the p-value indicates that the forecasts are efficient and/or unbiased at 30-steps ahead only.
-It is evident that Prophet produces inferior results in comparison to the ARIMA model over all forecast horizons.
-This is unexpected as Prophet is said to be robust to the presence of outliers and missing values in the data, without the need for interpolation.
-We also notice that the forecasts produced by both models deteriorates when there are missing values and outliers present in the data, with the only efficient and/or unbiased forecasts yielded by the ARIMA model at a forecast horizon of 30 days.
+The results from the dirty data show that the ARIMA and Prophet model are sensitive to the presence of outliers and missing values, with both models producing unbiased and/or inefficient forecasts and higher RMSEs across all forecast horizons.
+This is disappointing as Prophet is said to be robust when faced with outliers and can handle missing values without the need for interpolation unlike the ARIMA model.
+Since both models have equal predictive accuracy at the shortest forecast horizon, an analyst engaged in short-term trading who has no knowledge on how to pre-process BTC/ZAR data before forecasting, could apply the Prophet model as it would provide them with flexibility and ease of implementation without sacrificing any predictive accuracy.
 
-The limitation of this study is that it looks specifically at BTC/ZAR.
-Different results may emerge for fiat currencies or the exchange rate between Bitcoin with other currencies.
-Hence caution should be exercised when generalising these results to other exchange rates.
-As a point of further research, more complex techniques such as neural networks could be used as a benchmark for comparison when evaluating the forecasts produced by Prophet.
+Although the Prophet model was eliminated from the Model Confidence Set at other forecast horizons, these results might change as more data and information about events which might impact the price of BTC/ZAR becomes available and is incorporated into the Prophet model.
+Furthermore, the RMSE shows that the difference between the forecast errors of ARIMA and Prophet is marginal, even in the case where both models don’t have equal predictive accuracy.
+Hence, if the analyst is willing to sacrifice a small amount of predictive accuracy in return for the tractability and ease of implementation which Prophet provides, then they might favour Prophet over the ARIMA model.
+
+The attractive features of Prophet were not fully utilised in this paper due to the stationarity of log returns.
+Different results may emerge when forecasting the direction of returns or closing prices.
+Furthermore, the reader should exercise caution when generalising these results to other exchange rate data or to exchange rates which involve fiat currencies as results may differ.
+As a point of further research, more complex techniques such as Neural Networks could be used as a benchmark for comparison when evaluating the forecasts produced by Prophet.
 

### 2017-10-27 Draft/Write-up/Draft.tex
Fixed about 80% of the mess. Still fixing the rest

@@
 \begin{center}
-{\Huge \bf How does the latest Prophet forecasting model compare to the ARIMA
-model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz
-approach}\\
+{\Huge \bf A comparison between the latest Prophet forecasting model and the ARIMA
+model: An evaluation of BTC/ZAR predictability}\\
 \today\\
@@
 building them is a challenging task. Prophet is a sophisticated model
-that differs to traditional time series models in that it can be
+that differs from traditional time series models in that it can be
 utilised by non-experts who have little knowledge about the statistical
@@
 with missing values and outliers present in the data.
-\noindent
-Keywords: Prophet, ARIMA, Mincer-Zarnowitz,   Box-Jenkins, Bitcoin.
+\noindent\\
+Keywords: Prophet, ARIMA, Mincer-Zarnowitz,   Box-Jenkins, Bitcoin, Model Confidence Set
 \end{abstract}
@@
 The Autoregressive Integrated Moving Average (ARIMA) model is one of the
-most widely used time series models that has attracted attention in
+most widely used time series models that have attracted attention in
 financial market forecasting (Khashei, Bijari, and Ardali 2009).
@@
 and less computationally expensive. It has been used as the building
-blocks for more advanced models and has provided the inspiration for
+block for more advanced models and has provided the inspiration for
 hybrid versions of the model which have been used in exchange rate
@@
 In December 2016, Facebook open-sourced their forecasting model Prophet.
-Prophet differs to traditional time series model such as the ARIMA, in
+Prophet differs from traditional time series model such as the ARIMA in
 that it can produce high quality forecasts in a straightforward way.
@@
 involved in the model, however have domain knowledge about the data
-generating process. This knowledge can be easily incorporated into the
+generating process. This knowledge can easily be incorporated into the
 model through its intuitively adjustable parameters (Taylor and Letham
@@
 traditional ARIMA model and Prophet through an evaluation of the
-exchange rate between Bitcoin and the Rand (BTC/ZAR), using a
-Mincer-Zarnowitz approach of measuring forecast accuracy. More
-specifically, this paper aims to:
+exchange rate between Bitcoin and the Rand (BTC/ZAR), using the Model
+Confidence Set and a Mincer-Zarnowitz approach of measuring forecast
+accuracy. More specifically, this paper aims at:
 
@@
 \item
-  Compare the out-of-sample forecasts of ARIMA and Prophet over varying
-  forecast horizons.
-\item
-  Compare the out-of-sample forecast performance with missing values and
-  outliers present in the data.
+  Comparing the out-of-sample forecasts of ARIMA and Prophet over
+  varying forecast horizons.
+\item
+  Comparing the out-of-sample forecast performance with missing values
+  and outliers present in the data.
 \end{enumerate}
 
-Section \ref{Background and Theory} will present the key findings from
-the major works in which comparisons have been made to the ARIMA model
-in exchange rate forecasting, as well as the theory behind the ARIMA and
-Prophet model. Section \ref{Data} will provide a brief explanation and
-motivation for the dataset used in this paper. Section \ref{Methodology}
-will describe the methodology used in selecting the ARIMA and Prophet
-models and the Mincer-Zarnowitz approach of measuring forecast accuracy.
-The results from applying the ARIMA and Prophet model to the dataset is
-presented in Section \ref{Results}. Finally, Section
-\textbackslash{}ref\{Discussion and Conclusions discusses the results in
-a statistical context.
-
-\section{\texorpdfstring{Background and Theory
-\label{Background and Theory}}{Background and Theory }}\label{background-and-theory}
+Section \ref{Background} will present the key findings from the major
+works in which comparisons have been made to the ARIMA model in exchange
+rate forecasting. Section \ref{Data} will provide a brief explanation
+and motivation for the dataset used in this paper. Section
+\ref{Forecasting Methods} explains and compares the formulation of the
+ARIMA and Prophet model. Section \ref{Methodology} will describe the
+methodology used in selecting the ARIMA and Prophet models, the Model
+Confidence Set and the Mincer-Zarnowitz approach of measuring forecast
+accuracy. The results from applying the ARIMA and Prophet model to the
+dataset is presented in Section \ref{Results}. Finally, section
+\ref{Discussion and Conclusions} discusses the results and its financial
+implications in a statistical context.
+
+\section{\texorpdfstring{Background
+\label{Background}}{Background }}\label{background}
 
@@
 
-ARIMA models became highly popular since its introduction by Box and
-Jenkins, when it was shown that they could outperform complex
-econometric models in a variety of situations (Hibon and Makridakis
-1997). The ARIMA model expresses the process \{\(y_t\)\} as a function
-of the weighted average of past values of the process and lagged values
-of the residuals. The weighted average of the past p values of the
-process represents an autoregressive (AR) process of order p.~It feeds
-back past values of the process into the current value, inducing
-correlation between all lags of the process. The weighted average of the
-q lagged residuals represents a moving average (MA) process of order q.
-The purpose of mixing the MA process with the AR process is to reduce
-the large number of past values required by AR processes and to control
-for the autocorrelation which it creates between lagged values of the
+ARIMA models have commonly been used in financial forecasting and are
+popular for observing stock prices and exchange rates due to its power
+and statistical properties (C.-S. Lin, Chiu, and Lin 2012). The ARIMA
+model is attractive as it is tractable and produces good short-term
+forecasts when more than 100 observations are used (Tseng et al. 2001).
+They have frequently been used as a benchmark to compare new forecasting
+techniques that have emerged over time and have yielded satisfactory
+results when predicting exchange rates. Nwankwo (2014) forecasted the
+exchange rate between the Naira and dollar, using Akaike's Information
+Criterion (AIC) as a measure of performance. Diagnostic testing revealed
+that the ARIMA(1,0,0) model was the best fit for the data.
+
+Although the ARIMA model has the advantage of ease of implementation and
+flexibility, it fails to capture the non-linearity and volatility
+present in exchange rate data. Over time, the ARIMA model has evolved to
+cater for a wider variety of data and to compensate for some of its
+shortcomings. Some of the most popular versions of the ARIMA model that
+has been implemented in exchange rate forecasting is the Seasonal ARIMA
+(SARIMA) and Fractional ARIMA (ARFIMA) model.
+
+The SARIMA model was introduced to capture the periodic behaviour of
+data and extends the ARIMA class by including a term for seasonal
+differencing. Etuk, Wokoma, and Moffat (2013) modelled the Naira/CFA
+Franc exchange rate which exhibited monthly seasonality using an
+additive SARIMA model, to demonstrate that it can be a useful fit for
+exchange rate data which displays seasonality. Their results showed that
+the SARIMA model adequately described the variation in the exchange rate
+series.
+
+The ARFIMA model generalises the ARIMA model in that the degree required
+to make the data stationary can assume any real value, and is no longer
+restricted to the integer domain. The ARFIMA model has the ability to
+capture the dependence between observations that are widely spread apart
+in time (Cheung 1993). This makes the model parsimonious since it can
+capture long memory in data as well as short term dynamics (Cheung
+1993). Cheung (1993) fitted the ARFIMA model to examine five exchange
+rates and found that there was strong evidence of long memory in the
+exchange rate time series.
+
+\subsection{Other techniques used to forecast exchange
+rates}\label{other-techniques-used-to-forecast-exchange-rates}
+
+Generalised autoregressive conditional heteroskedasticity (GARCH) models
+were later developed due to the failure of ARIMA models to capture the
+volatility in financial markets (Anastasakis and Mort 2009). Fahimifard
+et al. (2009) applied a GARCH model to the Rial/USD and Rial/EUR to
+compare the forecasts yielded by the GARCH and ARIMA model over varying
+forecast horizons. Their results illustrated that the GARCH model
+outperformed the ARIMA model when using the Root Mean Square Error
+(RMSE), Mean Square Error (MSE) and Mean Absolute Deviation (MAD) as a
+performance criteria.
+
+Over time, financial forecasting methods have moved away from linear
+models like ARIMA and GARCH, to soft computing techniques. These complex
+techniques are non-linear and can fit complex time series more easily
+(Castillo and Melin 2002). Unlike the ARIMA model, soft computing
+techniques do not impose structural assumptions on the model apriori
+(Castillo and Melin 2002). Some of the most commonly used artificial
+intelligence methods used to forecast exchange rate data are Neural
+Networks and Fuzzy Logistic Systems.
+
+Artificial Neural Networks (ANNs) mimic the structure of the brain and
+have had many successful applications in forecasting exchange rates.
+They are data driven, can adapt to non-stationary environments and can
+approximate any continuous function (Khashei and Bijari 2011). In a
+study done by Fahimifard et al. (2009), the ANN was found as an
+effective way to improve the forecasts of exchange rates. Superior
+results were produced when compared to the ARIMA and GARCH model using
+the RMSE, MSE and MAD as a measure of performance.
+
+Fuzzy Logistic Systems (FLSs) were initially developed to solve problems
+involving linguistic terms and have successfully been used in financial
+forecasting (Khashei, Bijari, and Ardali 2009). Fuzzy logic tries to
+imitate human reasoning and the decision-making process and allows for
+finer rather than discrete decisions to be provided. Santos, Costa, and
+Santos Coelho (2007) investigated how well FLSs and ANNs perform
+compared to the traditional ARMA and GARCH model. He examined the
+forecasts of Brazilian exchange rate returns by considering different
+frequencies of the series and comparing their one-step ahead forecasts.
+By analysing accuracy statistics, he found that FLSs and ANNs achieved
+higher returns based on the forecasts they produced. Similar results
+were found by Khashei, Bijari, and Ardali (2009) when he analysed the
+predictive capabilities of FLSs, ANNs, the traditional ARIMA model and a
+Fuzzy ARIMA model.
+
+Although ANNs have been broadly applied in financial forecasting, the
+process of building them is a complex task and there is no consistent
+method of design compared to the traditional Box-Jenkins ARIMA model.
+The performance of ANNs is sensitive to many modelling factors such as
+the number of input nodes included and the size of the training sample
+chosen (Zhang and Hu 1998). Like ANNs, there is no systematic approach
+for designing FLSs and they are only understandable when simple.
+Although FLSs has the advantage over ARIMA models in that they can be
+applied to data with few observations available, it yields acceptable
+rather than accurate results and are more suitable for problems which do
+not require high accuracy.
+
+The traditional ARIMA model produces less superior forecasts than its
+hybrid forms and other complex non-linear techniques, however its
+forecasts are still satisfactory. It is a simple model that is easy to
+implement and has a consistent method of model design and selection.
+ARIMA models are more robust and efficient than complex structural
+models in relation to short-run forecasting. The fact that they have
+been used as the foundation for more advanced models and have commonly
+been used as a benchmark for comparison, justifies it as a good starting
+point to compare it to Facebook's forecasting method, Prophet, that was
+recently released.
+
+\subsection{Forecasting with Prophet}\label{forecasting-with-prophet}
+
+The techniques that have been considered for exchange rate forecasting
+thus far, require the analyst to have vocational knowledge about time
+series. Prophet differs from traditional time series models in that it
+is flexible and can be customised by a large number of non-experts who
+have little knowledge about time series, however have domain knowledge
+about the data generating process. Prophet allows for a large number of
+forecasts to be produced across a variety of problems and consists of a
+robust evaluation system that allows for a large number of forecasts be
+evaluated and compared. This is Facebook's definition of forecasting at
+scale.
+
+Prophet's Bayesian approach to forecasting allows the analyst to
+incorporate their expert knowledge into the model building process and
+has produced significantly improved forecasts compared to the ARIMA
+model. Taylor and Letham (2017) forecasted the number of events on
+Facebook using Prophet. The time series was impacted by holidays, had
+strong multi-period seasonality, and a piecewise trend. The forecasts
+produced by Prophet were compared to common forecasting techniques such
+as exponential smoothing, ARIMA, the seasonal naive, and the naive
+model. While exponential smoothing and the seasonal naive model were
+quite robust, the ARIMA forecasts were fragile. No model besides Prophet
+accounted for the dips around holidays and the upward trend of the time
+series towards later observations. If Prophet produces forecasts that
+are as good as the ARIMA model when forecasting BTC/ZAR, it can be
+compared to more complex forecasting methods that are less tractable
+such as the hybrid models and machine learning techniques seen earlier.
+
+\section{\texorpdfstring{Data \label{Data}}{Data }}\label{data}
+
+The dataset used in this study consists of the daily closing prices of
+BTC/ZAR over the period 24 January 2016 to 17 July 2017. This comprises
+of a total of 541 trading days and was the chosen time period for
+analysis due to constraints in obtaining data over a longer period. The
+data was obtained from Bitcoincharts.
+
+Bitcoin is of specific interest as it presents an interesting parallel
+to traditional exchange rate markets. The cryptocurrency is built on a
+decentralised system and as a result its value cannot be directly
+influenced by a central authority (Fantazzini et al. 2016). In addition,
+the Bitcoin market has attracted attention worldwide and is currently
+the leading cryptocurrency, with awareness and adoption of the currency
+growing over time (Fantazzini et al. 2016). Its novelty makes it a
+highly volatile hence speculative market and provides an opportunity for
+forecasting.
+
+In this study, the analysis of BTC/ZAR forecasts are based on the daily
+continuous log returns. The daily closing prices are transformed into
+returns by taking the log difference at each time t and is calculated as
+follows:
+
+\[ r_t = ln(\frac{p_t}{p_{t-1}}) \] where \(p_t\) represents the closing
+price and \(r_t\) represents the log return for time t = 1,2,.,T.
+
+\section{\texorpdfstring{Forecasting Methods
+\label{Forecasting Methods}}{Forecasting Methods }}\label{forecasting-methods}
+
+\subsection{The ARIMA model}\label{the-arima-model}
+
+The ARIMA model expresses the process \{\(y_t\)\} as a function of the
+weighted average of past values of the process and lagged values of the
+residuals. The weighted average of the past p values of the process
+represents an autoregressive (AR) process of order p.~It feeds back past
+values of the process into the current value, inducing correlation
+between all lags of the process. The weighted average of the q lagged
+residuals represents a moving average (MA) process of order q. The
+purpose of mixing the MA process with the AR process is to reduce the
+large number of past values required by AR processes and to control for
+the autocorrelation which it creates between lagged values of the
 process. The combination of the AR(p) and MA(q) process results in a
@@
 of time, and is defined as \(By_t = y_{t-1}\). The ARIMA model has the
-form:
+form (Hyndman and Athanasopoulos 2014):
 
@@
 
-ARIMA models have commonly been used in financial forecasting and are
-popular for observing stock prices and exchange rates due to its power
-and statistical properties (C.-S. Lin, Chiu, and Lin 2012). They have
-frequently been used as a benchmark to compare new forecasting
-techniques that have emerged over time. Kamruzzaman and Sarker (2003)
-applied an ARIMA model to forecast the exchange rate between the
-Australian dollar and six other currencies, and used the forecast errors
-as well as the accuracy of the direction of the forecasts to evaluate
-the performance of three Neural Network models. In a similar study,
-Khashei, Bijari, and Ardali (2012) compared the predictive capability of
-their proposed hybrid model consisting of an ARIMA and a Probabilistic
-Neural Network (PNN) to the traditional ARIMA model, and justified the
-use of the ARIMA as a benchmark by claiming that it is the most
-important linear model. Their comparison was made by investigating the
-forecasts between the British pound and US dollar over varying forecast
-horizons. Not only has the ARIMA model been useful as a benchmark, but
-it has also yielded satisfactory results when predicting exchange rates.
-When Nwankwo (2014) forecasted the rate between the Nairo and dollar,
-diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit
-for the data based on Akaike's Information Criterion (AIC).
-
-The ARIMA model is attractive as it is tractable and produces good
-short-term forecasts when more than 100 observations are used (Tseng et
-al. 2001). Although the ARIMA model has the advantage of ease of
-implementation and flexibility, it fails to capture the non-linearity
-and volatility present in exchange rate data. Over time, the ARIMA model
-has evolved to cater for a wider variety of data and to compensate for
-some of its shortcomings. The most popular versions of the ARIMA model
-that has been implemented in exchange rate forecasting is the Seasonal
-ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).
-
-The SARIMA model was introduced to capture the periodic behaviour of
-data and extends the ARIMA class by including a term for seasonal
-differencing. Etuk, Wokoma, and Moffat (2013) modelled the Naira/CFA
-Franc exchange rate which exhibited monthly seasonality using an
-additive SARIMA model, to demonstrate that it can be a useful fit for
-exchange rate data which displays seasonality. Their results showed that
-the SARIMA model adequately described the variation in the exchange rate
-series. The ARFIMA model generalises the ARIMA model in that the degree
-required to make the data stationary can assume any real value, and is
-no longer restricted to the integer domain. The ARFIMA model has the
-ability to capture the dependence between observations that are widely
-spread apart in time (Cheung 1993). This makes the model more
-parsimonious since it can capture long memory in data as well as short
-term dynamics (Cheung 1993). Cheung (1993) fitted the ARFIMA model to
-examine five exchange rates, and found that there was strong evidence of
-long memory in the exchange rate time series.
-
-Generalised autoregressive conditional heteroskedasticity (GARCH) models
-were later developed in an attempt to capture the volatility in
-financial markets (Anastasakis and Mort 2009). Hsieh (1989) applied a
-GARCH model to investigate five exchange rates and his results showed
-that although the GARCH model outperformed the random walk, some
-non-linear information still remained in the residuals.
-
-\subsection{Other techniques used to forecast exchange
-rates}\label{other-techniques-used-to-forecast-exchange-rates}
-
-Over time, financial forecasting methods have moved away from linear
-models like ARIMA and GARCH, to soft computing techniques. These more
-complex techniques are non-linear and can fit complex time series more
-easily (Castillo and Melin 2002). Unlike the ARIMA model, soft computing
-techniques do not impose structural assumptions on the model apriori
-(Castillo and Melin 2002). Some of the most commonly used artificial
-intelligence methods used to forecast exchange rate data are Neural
-Networks and Fuzzy Logistic Systems.
-
-Artificial Neural Networks (ANNs) have had many successful applications
-in forecasting exchange rates and are more advantageous than other
-non-linear forecasting methods. They are data driven, can adapt to
-non-stationary environments and can approximate any continuous function
-(Khashei and Bijari 2011). In a study done by Fahimifard et al. (2009),
-the ANN was found as an effective way to improve the forecasts of
-exchange rates. Superior results were produced when compared to the
-ARIMA and GARCH model using the root mean square error (RMSE), mean
-square error (MSE) and mean absolute difference as a measure of
-performance. Fuzzy Logistic Systems (FLSs) were initially developed to
-solve problems involving linguistic terms, and have been successfully
-used in financial forecasting (Khashei, Bijari, and Ardali 2009). Fuzzy
-logic tries to imitate human reasoning and the decision-making process
-and allows for finer rather than discrete decisions to be provided.
-Santos, Costa, and Santos Coelho (2007) investigated how well FLSs and
-ANNs perform compared to the traditional ARMA and GARCH model. He
-examined the forecasts of Brazilian exchange rate returns by considering
-different frequencies of the series and comparing their one step-ahead
-forecasts. By analysing accuracy statistics, he found that FLSs and ANNs
-achieved higher returns based on the forecasts they produced. Similar
-results were found by Khashei, Bijari, and Ardali (2009) when he
-analysed the predictive capabilities of FLSs, ANNs, the traditional
-ARIMA model, and a Fuzzy ARIMA model.
-
-Although ANNs have been broadly applied in financial forecasting, the
-process of building them is a complex task and there is no consistent
-method of design compared to the traditional Box-Jenkins ARIMA model.
-Unlike ARIMA models, the performance of ANNs is sensitive to many
-modelling factors such as the number of input nodes included and the
-size of the training sample chosen (Zhang and Hu 1998). Like ANNs there
-is no systematic approach to designing FLSs and they are only
-understandable when simple. Although FLSs has the advantage over ARIMA
-models that they can be applied to data with few observations available,
-it gives acceptable rather than accurate results and are more suitable
-for problems which do not require high accuracy.
-
-\subsection{Hybrid ARIMA Models}\label{hybrid-arima-models}
-
-Over time, many researchers began to think of ways in which to harness
-the advantages of tractable linear models such as the ARIMA model, and
-of more complex non-linear models. By combining the ARIMA model with
-other forecasting methods, the advantages of both forecasting methods
-are leveraged while simultaneously improving their limitations. Some of
-the hybrid models which have commonly been used in exchange rate
-forecasting are the Fuzzy ARIMA and ANN-ARIMA model.
-
-The ARIMA model produces very accurate forecasts over short time
-horizons however it has the limitation of requiring more than 100
-observations of historical data to yield accurate results (Tseng et al.
-2001). In a world that is constantly changing and with the rapid
-advancement of technology, access to large amounts of historical data is
-difficult to obtain. On the contrary, a fuzzy regression model requires
-little historical data however produces wide prediction intervals if
-extreme values are present in the data. The Fuzzy ARIMA model combines
-the ARIMA and fuzzy regression model to exploit the advantages of both
-models while simultaneously overcoming their limitations. Tseng et al.
-(2001) proposed applying a Fuzzy ARIMA model to forecast the exchange
-rate of Taiwan dollars to US dollars to demonstrate the model's
-appropriateness and power. The Fuzzy ARIMA not only produced forecasts
-that were superior to the ARIMA and fuzzy time series models but also
-provided an upper and lower bound which can be used by decision makers
-to determine the best and worst possible situations.
-
-Although ARIMA models are powerful, they require non-stationary data to
-be differenced and impose prior assumptions onto the distribution of the
-data (Ince and Trafalis 2006). In contrast, machine learning techniques
-such as ANNs do not impose any assumptions onto the data generating
-process however, being data-driven, are sensitive to the number of input
-nodes used. The ANN-ARIMA model draws on the strengths of both models
-and overcomes these individual difficulties. Ince and Trafalis (2006)
-created an ANN-ARIMA model by using the ARIMA model to determine the
-number input nodes required by an ANN for three exchange rates. When the
-ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model
-outperformed the ARIMA based on the MSE.
-
-Although the traditional ARIMA models produce less superior forecasts
-than its hybrid forms and other complex non-linear techniques, its
-forecasts are still satisfactory. They are simple models that are easy
-to implement and have a consistent method of model design and selection.
-ARIMA models are also more robust and efficient than complex structural
-models in relation to short-run forecasting. The fact that they have
-been used as the foundation for more advanced models and have commonly
-been used as a benchmark for comparison, justifies it as a good starting
-point to compare it to Facebook's forecasting method, Prophet, that was
-recently released.
-
-\subsection{Forecasting with Prophet}\label{forecasting-with-prophet}
-
-The techniques that have been considered for exchange rate forecasting
-thus far, require the analyst to have vocational knowledge about time
-series. Prophet differs to traditional time series models in that it is
-flexible and can be customised by a large number of non-experts who have
-little knowledge about time series, however have domain knowledge about
-the data generating process. Prophet allows for a large number of
-forecasts to be produced across a variety of problems and consists of a
-robust evaluation system that allows for a large number of forecasts be
-evaluated and compared. This is Facebook's definition of forecasting at
-scale. Prophet is similar to a Generalized Additive Model (GAM) - an
-additive regression model that consists of non-linear and linear
-regression functions applied to predictor variables (Taylor and Letham
-2017). The decomposable model is of the form:
+\subsection{The Prophet model}\label{the-prophet-model}
+
+Prophet is similar to a Generalized Additive Model (GAM) - an additive
+regression model that consists of non-linear and linear regression
+functions applied to predictor variables (Taylor and Letham 2017). The
+decomposable model is of the form:
 
@@
 model flexibility and allows the analyst to interactively change model
-parameters (Taylor and Letham 2017).
-
-The growth component of Prophet may be modelled as a linear or
-non-linear function of time. Linear growth is modelled by a piecewise
-constant function while non-linear growth is modelled similar to
-population growths which use a logistic growth model (Taylor and Letham
-2017). A time-varying upper limit may be specified for logistic growth,
-at which point the forecasts will saturate. This carrying capacity
-allows the analyst to incorporate their prior knowledge about the
-maximum obtainable growth level such as the total market or population
-size into the model. Prophet accounts for changes in the trajectory of
-this trend by automatically detecting and selecting changepoints in the
-data at which the growth rate is allowed to change. These changepoints
-have a Laplace prior distribution placed on them and its scale parameter
-may be used to adjust the flexibility of the trend and to choose how
-aggressively the model should follow historical trend changes. Analysts
-may also adjust the number of potential changepoints included or
-manually specify their location. This allows non-experts with knowledge
-about events that may affect the growth rate to use the parameter as a
-knob to either increase or decrease the number of changepoints included
-(Taylor and Letham 2017). It also allows for the analyst to add
-changepoints which the automatic selection procedure may have missed or
-remove changepoints when the model is overfitting historical trends
-(Taylor and Letham 2017).
+parameters (Taylor and Letham 2017). The growth component of Prophet may
+be modelled as a linear or non-linear function of time. Linear growth is
+modelled by a piecewise constant function while non-linear growth is
+modelled similar to population growths which use a logistic growth model
+(Taylor and Letham 2017). A time-varying upper limit may be specified
+for logistic growth, at which point the forecasts will saturate. This
+carrying capacity allows the analyst to incorporate their prior
+knowledge about the maximum obtainable growth level such as the total
+market or population size into the model. Prophet accounts for changes
+in the trajectory of this trend by automatically detecting and selecting
+changepoints in the data at which the growth rate is allowed to change.
+These changepoints have a Laplace prior distribution placed on them and
+its scale parameter may be used to adjust the flexibility of the trend
+and to choose how aggressively the model should follow historical trend
+changes. Analysts may also adjust the number of potential changepoints
+included or manually specify their location. This allows non-experts
+with knowledge about events that may affect the growth rate to use the
+parameter as a knob to either increase or decrease the number of
+changepoints included (Taylor and Letham 2017). It also allows for the
+analyst to add changepoints which the automatic selection procedure may
+have missed or remove changepoints when the model is overfitting
+historical trends (Taylor and Letham 2017).
 
@@
 analysts to smooth the holidays (Taylor and Letham 2017).
-
-Prophet's Bayesian approach to forecasting allows the analyst to
-incorporate their expert knowledge into the model building process and
-has produced significantly improved forecasts compared to the ARIMA
-model. Taylor and Letham (2017) forecasted the number of events on
-Facebook using Prophet. The time series was impacted by holidays, had
-strong multi-period seasonality, and a piecewise trend. The forecasts
-produced by Prophet were compared to common forecasting techniques such
-as exponential smoothing, ARIMA, the seasonal naïve, and the naïve
-model. While exponential smoothing and the seasonal naïve model were
-quite robust, the ARIMA forecasts were fragile. No model besides Prophet
-accounted for the dips around holidays and the upward trend of the time
-series towards later observations.
 
@@
 
-\section{\texorpdfstring{Data \label{Data}}{Data }}\label{data}
-
-The data set used in this study consists of the daily closing prices of
-BTC/ZAR over the period 24 January 2016 to 17 July 2017. This comprises
-of a total of 541 trading days and was the chosen time period for
-analysis due to constraints in obtaining data over a longer period. The
-data was obtained from Bitcoincharts.
-
-Bitcoin is of specific interest as it presents an interesting parallel
-to traditional exchange rate markets. The cryptocurrency is built on a
-decentralised system and as a result its value cannot be directly
-influenced by a central authority (Fantazzini and Nigmatullin 2016). In
-addition, the Bitcoin market has attracted attention worldwide and is
-currently the leading cryptocurrency, with awareness and adoption of the
-currency growing over time (Fantazzini and Nigmatullin 2016). Its
-novelty makes it a highly volatile hence speculative market and provides
-an opportunity for forecasting.
-
-In this study, the analysis of BTC/ZAR forecasts are based on the daily
-continuous log returns. The daily closing prices are transformed into
-returns by taking the log difference at each time t and is calculated as
-follows:
-
-\[ r_t = ln(\frac{p_t}{p_{t-1}}) \] where \(p_t\) represents the closing
-price and \(r_t\) represents the log return for time t = 1,2,.,T.
-
 \section{\texorpdfstring{Methodology
@@
 rolling window forecasts are made 1 day, 30 days and 90 days ahead. This
-allows us to examine the forecast horizon effect. The Mincer-Zarnowitz
-test is then used to evaluate the forecasts produced by ARIMA and
-Prophet against the test set. Since there is no consensus on which
-accuracy statistic best measures the performance of forecasting
-techniques, the most popular criteria, the Mean Squared Error (MSE) is
-employed. The results obtained from the ARIMA model will be used as a
-benchmark for comparison. This process is then repeated based on data
-which has outliers and missing values present.
+allows us to examine the forecast horizon effect. The Model Confidence
+Set and Mincer-Zarnowitz test is then used to evaluate the forecasts
+produced by ARIMA and Prophet against the test set. Since there is no
+consensus on which accuracy statistic best measures the performance of
+forecasting techniques, the most popular criteria, the Root Mean Squared
+Error (RMSE) is employed. The results obtained from the ARIMA model will
+be used as a benchmark for comparison. This process is then repeated
+based on data which has outliers and missing values present.
 
@@
 
-Box and Jenkins proposed a set of guidelines that can be followed when
-selecting ARIMA models. This systematic procedure to designing ARIMA
-models has made them highly popular (Hibon and Makridakis 1997). It
-consists of a four-stage iterative process in which:
+Box, Jenkins, and Reinsel (1970) proposed a set of guidelines that can
+be followed when selecting ARIMA models. It consists of a four-stage
+iterative process in which:
 
@@
 \def\labelenumi{\arabic{enumi}.}
+\tightlist
 \item
@@
 satisfactory model is identified. On the contrary, if the diagnostic
-check shows that the residuals are random, the developed model will be
+check reveals that the residuals are random, the developed model will be
 the final model used for forecasting.
@@
 \item
-  The trend is set to be linear
-\item
-  The width of the uncertainty intervals is set to 80\%
+  The trend is set to be linear.
+\item
+  The width of the uncertainty intervals is set to 80\%.
 \item
@@
   The smoothing parameter for holidays and seasonality is set at 10
-\item
-  The smoothing parameter for the trend is set to 0.05
-\item
-  The number of potential changepoints is set to 25
+  while the smoothing parameter for the trend is set to 0.05.
+\item
+  The number of potential changepoints is set to 25.
 \end{itemize}
@@
 account for our prior knowledge about the closing prices of Bitcoin
-which tend to be lower around the weekend however higher mid-week, due
-to fluctuations in trading activity. Furthermore, a linear trend is
+which tend to be lower around the weekend however higher mid-week due to
+fluctuations in trading activity. Furthermore, a linear trend is
 appropriate since Bitcoin does not have an upper limit on its closing
@@
 way, a poor forecast will not have negative consequences on future
-forecasts to be made, since the observed values will be used to correct
+forecasts to be made since the observed values will be used to correct
 itself for the remaining forecasts (Zhang and Hu 1998). In this paper,
-we first fit the ARIMA and Prophet model to the training set and
-forecast 1 day ahead. We then refit the model to the training set,
-increasing its size by one observation, and the next 1 day ahead
-forecast is made. This process is repeated until all forecasts have been
-made into the test set. The 1 day ahead forecasts are then compared to
-the 1 day ahead observed values in the test set and the error statistics
-are calculated. This procedure is then repeated for the 30 day and 90
-day ahead forecasts.
+the ARIMA and Prophet model is first fitted to the training set and the
+returns are forecasted h-days ahead for h = 1, 30, 90. The size of the
+training set is then increased by one observation and the models are
+refitted. The next h-day ahead return is forecasted and this process is
+repeated until all forecasts have been made into the test set. The h-day
+ahead forecasts are then compared to the h-day ahead observed values in
+the test set and the forecast errors are calculated to determine the
+RMSE.
+
+\subsection{Model Confidence Set}\label{model-confidence-set}
+
+The Model Confidence Set (MCS) procedure developed by Hansen, Lunde, and
+Nason (2011) is a sequential procedure which starts with an initial set
+of models \(M_0\) of size m. The worse model is then eliminated at each
+step until the null hypothesis of equal predictive ability (EPA) fails
+to be rejected for all the models belonging to the set at a given
+confidence level \(1 - \alpha\). This smaller set of models \(M^*\) is
+the superior set of models (SSM) with the best case occurring when the
+SSM consists of a single best model.
+
+The null hypothesis for EPA for a set of models \(M\) is given by:
+\[H_0: \mu_i,_j = 0 \;\;\; \forall \;\; i,j = 1,...,m\]
+
+where \(\mu_i,_j = E[d_i,_j]\) represents the expected loss
+differential, \(d_i,_j,_t = l_i,_t - l_j,_t\) represents the loss
+differential between models \(i\) and \(j\) at time t, and \(l_i,_t\)
+and \(l_j,_t\) are the loss functions associated with model \(i\) and
+\(j\) at time t respectively. This paper makes use of the squared
+forecast error as a loss function, however any arbitrary loss function
+that satisfies the weak stationarity conditions described in Hansen,
+Lunde, and Nason (2011) may be used.\\
+The null hypothesis is tested by constructing the following test
+statistics:
+
+\[t_i,_j = \frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}\]
+\vspace{1pt} \[ T_R = \max_{i,j\in M} \;\; |t_i,_j|\] where
+\(d_i,_j = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }\) measures the sample
+average loss differential between model \(i\) and \(j\) and
+\(\hat{var[\bar{d_i,_j}]}\) is estimated through bootstrapping as
+described in Hansen, Lunde, and Nason (2011).
+
+This paper uses a block--bootstrap procedure of 5000 resamples as in
+Hansen, Lunde, and Nason (2011). The block length \(p\) is chosen by
+fitting an AR(p) model to the \(d_i,_j\) terms and determining the
+maximum number of significant AR parameters obtained. While \(t_i,_j\)
+has a t-distribution, \(T_R\) has a non-standard distribution under the
+null hypothesis hence is estimated by implementing a bootstrapping
+procedure similar to that used to estimate \(\hat{var[\bar{d_i,_j}]}\).
+
+The worst model is then eliminated according to an elimination rule
+which is coherent with the calculated t-statistic and is defined as:
+
+\[e_R = \arg max_{i}\bigg(\sup_{j\in M}\frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}\bigg)\]
 
@@
 Prophet and the ARIMA model. Mincer and Zarnowitz (1969) proposed an
-absolute measure to evaluate forecast accuracy which considers the
+absolute measure to evaluate forecast accuracy that considers the
 distance between the actual and predicted values. To analyse the
 absolute errors produced by the forecasts, the observed values \(r_t\)
-are regressed against the predicted values \(r_t\), i.e.
-
-\[r_t = \alpha + \beta\cdot\hat{r_t} + e_t\]
-
-where \(\alpha\) represents the mean distance between the observed and
-predicted values while \(\beta\) represents the correlation between the
-forecasted errors and the predicted values. When alpha = 0 it implies
-that the forecasts are unbiased and do not systematically overestimate
-or underestimate the data. When b=1 it implies that the forecasts are
-efficient and uncorrelated with the forecasted errors. A joint
-hypothesis test of \(H_0: \alpha = 0 \cup \beta = 1\) is performed to
-check the efficiency and bias of the forecasts and the model which
-produces the best results generates superior forecasts.
+are regressed against the predicted values \(\hat{r_t}\), i.e.
+\[r_t = \alpha + \beta\cdot\hat{r_t} + e_t\] where \(\alpha\) represents
+the mean distance between the observed and predicted values while
+\(\beta\) represents the correlation between the forecasted errors and
+predicted values. When \(\alpha = 0\) it implies that the forecasts are
+unbiased and do not systematically overestimate or underestimate the
+data. When \(\beta = 1\) it implies that the forecasts are efficient and
+uncorrelated with the forecasted errors. A joint hypothesis test of
+\(H_0: \alpha = 0 \cup \beta = 1\) is performed to check the efficiency
+and bias of the forecasts and the model which produces the best results
+generates superior forecasts.
 
@@
 
-The original dataset is modified in order to examine how robust Prophet
-and ARIMA are when fitted to data with missing values and outliers. 20
-values are randomly removed from the training set and 5 outliers are
-randomly inserted in place of existing values. The models are then
-fitted to this ``dirty'' dataset and evaluating of forecasts are carried
-out as before.
+The original dataset is modified in order to examine how robust the
+ARIMA and Prophet model are when faced with missing values and outliers.
+20 values are randomly removed from the training set and 5 outliers are
+randomly inserted in place of existing values. Both models are then
+fitted to this ``dirty'' dataset and the forecasts are evaluated in the
+same way as the ``clean'' dataset.
 
@@
 
-\subsection{Evaluation over varying forecast
-horizons}\label{evaluation-over-varying-forecast-horizons}
-
-Although the Box-Jenkins methodology resulted in an ARIMA(0,0,1) model
-being the best fit to the training set, an ARIMA(1,0,0) is employed as
-the benchmark for comparison so that a rolling window forecast may be
-conducted. The AIC of the ARIMA(1,0,0) is marginally higher than that of
-the ARIMA(0,0,1) and analysis of the p-value indicates that the AR
-parameter is statistically significant in the explanation of the
-movement of BTC/ZAR. The Prophet model fitted to the training set
-included a linear trend and weekly seasonality, with all other
-parameters set to the default values as described in section
-\ref{Semi-Automatic Selection}. The automatic-selection procedure for
-ARIMA and Prophet selected the same model as those chosen manually.
-
-\subsubsection{1-step ahead Forecasts}\label{step-ahead-forecasts}
-
-Figure 1 depicts the resulting forecasts from fitting both models. The
-forecasts produced by the ARIMA model and Prophet lie within a band
-centred around zero. Both models fail to capture the large movement of
-returns away from zero. The 80\% uncertainty intervals produced by both
-models are approximately the same width and begin to widen as forecasts
-are made further into the test set.
-
-\begin{figure}[htbp]
+\subsection{Evaluation of clean dataset over varying forecast
+horizons}\label{evaluation-of-clean-dataset-over-varying-forecast-horizons}
+
+By following the Box-Jenkins methodology, an ARIMA(0,0,1) model was
+found to be the best fit for the clean training set. Making use of a
+mean model would result in a returns forecast of zero in a rolling
+window context. This is due to the white noise behaviour of log returns.
+Hence a decision was made to employ the ARIMA(1,0,0) model instead. The
+ARIMA(1,0,0) model has a marginally higher AIC and an analysis of the
+p-value indicated that the AR parameter is statistically significant in
+the explanation of the movement of BTC/ZAR. The Prophet model fitted to
+the clean training set included a linear trend and weekly seasonality
+with all other parameters set to the default values as described in
+section \ref{Semi-Automatic Selection}. The models chosen manually are
+consistent with the models selected by the automatic-selection procedure
+for ARIMA and Prophet.
+
+Table \ref{tab1} displays the MZ results obtained by regressing the test
+set returns against the forecasted returns, the RMSE, and the models
+that are included in the MCS. It is evident that the ARIMA model ranks
+ahead of the Prophet model over all forecast horizons. The MCS p-values
+indicates that both models lie within the 90\% confidence interval at
+the 1 day and 3 month forecast horizon, while the Prophet model is
+eliminated from the MCS at the 1 month forecast horizon. This suggests
+that the ARIMA and Prophet model have equal predictive accuracy at the 1
+day and 3 month forecast horizon, and that the ARIMA model produces
+superior forecasts at the 1 month forecast horizon. Furthermore, the MCS
+p-values indicates that while ARIMA model lies within the 90\%
+confidence interval with certainty across all forecast horizons, the
+Prophet model lies within the 90\% confidence interval 52.44\% of the
+time at the 1 day forecast horizon and 45.78\% of the time at the 3
+month forecast horizon.
+
+\begin{table}[H]
 \centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/1-step ahead forecasts.png}
-\caption{1-step ahead forecasts from the ARIMA(1,0,0) and Prophet model}
-\end{figure}
-
-Table 1 displays the MSE and the Mincer-Zarnowitz results obtained from
-regressing the test set returns against the one-step ahead forecasts.
-The intercept estimate is small and positive for ARIMA and Prophet, with
-the intercept for ARIMA being closer to zero than Prophet. This
-indicates that both models systematically underestimate the 1-step ahead
-returns, with ARIMA yielding mean forecasts closer to the observed mean.
-The slope estimate is negative for both models indicating that the
-observed returns are actually the opposite of the suggested forecast.
-The F-test for the joint hypothesis of a unity slope and zero intercept
-is rejected at the 5\% significance level for both models, with the
-p-value for ARIMA being much larger than Prophet. Hence the forecasts
-generated by both ARIMA and Prophet are biased and/or inefficient. The
-MSE produced by both models agree with the Mincer-Zarnowitz results,
-with ARIMA producing a marginally smaller MSE than Prophet.
-
-\begin{figure}[htbp]
+\caption{MCS, MZ-test and RMSE results based on clean data \label{tab1}} 
+\scalebox{0.9}{
+\begin{tabular}{cccccccc}
+  \hline
+Forecast & Model & Model & MCS & $\alpha$ & $\beta$ & MZ-test & RMSE \\ 
+ Horizon &  & Rank & p-values &  &  & p-values &  \\ 
+   \hline
+1 & ARIMA & 1 & 1 & 0.009546 & -2.005624 & 0.02545 & 0.04406 \\ 
+   & Prophet & 2 & 0.5244 & 0.013702 & -1.709744 & 0.007152 & 0.04425 \\ 
+   \hline
+30 & ARIMA & 1 & 1 & 0.031320 & -12.736757 & 0.06136 & 0.04459 \\ 
+   & Prophet & 2 & 0.023 * & 0.017169 & -2.621914 & 0.0003773 & 0.04556 \\ 
+   \hline
+90 & ARIMA & 1 & 1 & 0.006012 & -0.938687 & 0.8430658 & 0.04231 \\ 
+   & Prophet & 2 & 0.4578 & 0.006953 & -0.615435 & 0.4086752 & 0.04259 \\ 
+   \hline
+\end{tabular}
+}
+\end{table}
+
+The intercept estimate represented by \(\alpha\) is small and positive
+for ARIMA and Prophet across all forecast horizons, with the smallest
+estimates obtained at the longest forecast horizon. This suggests that
+both models systematically underestimates the returns, but to a smaller
+degree at long forecast horizons. The intercept estimate for ARIMA is
+closer to zero compared to Prophet at the 1 day and 3 month forecast
+horizon, indicating that the mean forecasts yielded by ARIMA is closer
+to the observed mean than Prophet. The slope estimate represented by
+\(\beta\) is negative for both models across all forecast horizons,
+indicating that the observed returns are actually the opposite of the
+suggested forecast and that both models fail to capture the explosive
+behaviour of BTC/ZAR.
+
+The MZ p-values at the 1 day and 1 month forecast horizon show that the
+joint hypothesis of a unity slope and zero intercept is rejected at the
+10\% significance level for both models. This suggests that the
+forecasts generated by ARIMA and Prophet are biased and/or inefficient.
+Both models only produce unbiased and/or efficient forecasts at the
+longest forecast horizon, with the ARIMA model yielding a considerably
+larger p-value compared to Prophet. The RMSE produced by the ARIMA model
+is marginally smaller than Prophet across all forecasts horizons, with
+the smallest RMSE calculated at the longest forecast horizon. These
+results agree with the results obtained from the MCS and the MZ-test.
+
+\subsection{Evaluation of data with missing values and outliers
+present}\label{evaluation-of-data-with-missing-values-and-outliers-present}
+
+By following the Box-Jenkins methodology, an ARIMA(0,0,3) model was
+found to be the best fit for the dirty training set. A decision was made
+to employ the ARIMA(3,0,0) model instead. The motivation behind this
+choice is the same as that provided in the case of the clean data. The
+ARIMA(3,0,0) model has a marginally higher AIC and an analysis of the
+p-value indicated that the AR parameters are statistically significant
+in the explanation of the movement of BTC/ZAR. The Prophet model fitted
+to the clean training set included a linear trend and weekly seasonality
+with all other parameters set to the default values as described in
+section \ref{Semi-Automatic Selection}. The models chosen manually are
+consistent with the models selected by the automatic-selection procedure
+for ARIMA and Prophet.
+
+Table \ref{tab2} displays the MZ results obtained by regressing the test
+set returns against the forecasted returns, the RMSE and the models that
+are included in the MCS. We observe that the presence of outliers and
+missing values in the data has no effect on the rank of the ARIMA and
+Prophet model. Contrary to the results from the clean data, the Prophet
+model lies within the 90\% confidence interval at the 1 day forecast
+horizon only. This suggests that the ARIMA and Prophet model only have
+equal predictive accuracy at the shortest forecast horizon. The MCS
+p-values indicates that while the ARIMA model always lies within the
+90\% confidence interval with certainty, the Prophet model only lies
+within the confidence interval at the 1 day forecast horizon 41.3\% of
+the time. This differs to the results from the clean data in which the
+Prophet model lies within the 90\% confidence interval 52.44\% of the
+time at the 1 day forecast horizon.
+
+\begin{table}[H]
 \centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 1.png}
-\caption{Mincer-Zarnowitz results for 1-step ahead forecasts}
-\end{figure}
-
-\subsubsection{30-steps ahead Forecasts}\label{steps-ahead-forecasts}
-
-Figure 2 illustrates the resulting 30-step ahead forecasts produced from
-fitting ARIMA and Prophet. As in the 1-step ahead case, both models fail
-to capture the large movement of returns away from zero with its
-forecasts lying within a band centred around zero. The 80\% uncertainty
-intervals produced by both models are approximately the same width and
-marginally wider than the uncertainty intervals produced by the 1-step
-ahead forecasts.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/30-step ahead forecasts.png}
-\caption{30-step ahead forecasts from the ARIMA(1,0,0) and Prophet
-model}
-\end{figure}
-
-Table 2 displays the MSE and the Mincer-Zarnowitz results obtained from
-regressing the test set returns against the 30-step ahead forecasts. The
-intercept estimate is small and positive for ARIMA and Prophet, with a
-smaller intercept estimated for Prophet. This indicates that both models
-systematically underestimate the 30-step ahead returns, with Prophet
-yielding mean forecasts closer to the observed mean. The slope estimate
-is large and negative for both models with the slope estimated to be as
-low as -12.74 for ARIMA. This suggests that the observed returns are
-actually the opposite of the proposed forecast. The F-test for the joint
-hypothesis of a unity slope and zero intercept is rejected at the 5\%
-significance level for Prophet only. The p-value of 0.0614 from the
-joint hypothesis test is slightly significant for ARIMA. Hence the
-forecasts generated by Prophet are biased and/or inefficient while the
-forecasts generated by ARIMA are unbiased and/or efficient. The MSE
-produced by both models agree with the Mincer-Zarnowitz results, with
-ARIMA producing a marginally smaller MSE than Prophet.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 2.png}
-\caption{Mincer-Zarnowitz results for 30-step ahead forecasts}
-\end{figure}
-
-\subsubsection{90-steps ahead Forecasts}\label{steps-ahead-forecasts-1}
-
-Figure 3 illustrates the 90-step ahead forecasts produced by ARIMA and
-Prophet. Both models fail to capture large movement of returns away from
-zero as in the case of the 1-step ahead and 30-step ahead forecasts. The
-80\% uncertainty intervals produced by both models are approximately the
-same width and are marginally narrower than the uncertainty intervals
-produced by the one-step ahead and 30-step ahead forecasts.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/90-steps ahead forecasts.png}
-\caption{90-step ahead forecasts from the ARIMA(1,0,0) and Prophet
-model}
-\end{figure}
-
-Table 3 presents the MSE and the Mincer-Zarnowitz results obtained from
-regressing the test set returns against the 90-step ahead forecasts. The
-intercept estimate is very small and positive for ARIMA and Prophet,
-with a marginally smaller intercept estimated for ARIMA. This indicates
-that both models systematically underestimate the 90-step ahead returns,
-with ARIMA yielding mean forecasts closer to the observed mean. The
-slope estimate is small and negative for both models with the slope
-estimated for ARIMA lying closer to -1. This suggests that the observed
-90-steps ahead returns are actually the opposite of the proposed
-90-steps ahead forecasts. The F-test for the joint hypothesis of a unity
-slope and zero intercept fails to be rejected at the 5\% significance
-level for both ARIMA and Prophet, with the p-value from the test being
-much larger for ARIMA. Hence the forecasts generated by Prophet and
-ARIMA are unbiased and/or efficient. The MSE produced by both models
-agree with the Mincer-Zarnowitz results, with ARIMA producing a
-marginally smaller MSE than Prophet.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 3.png}
-\caption{Mincer-Zarnowitz results for 90-step ahead forecasts}
-\end{figure}
-
-\subsubsection{Comparison over varying forecast
-horizons}\label{comparison-over-varying-forecast-horizons}
-
-Table 4 and 5 compares the p-values obtained from the Mincer-Zarnowitz
-tests and the MSE of the forecasts produced by ARIMA and Prophet over
-the three forecast horizons. In comparing the forecasting accuracy of
-the ARIMA model over the three forecast horizons, we observe that the
-p-value gets more significant as the forecast horizon gets longer. This
-is in contrast to the MSE of the ARIMA forecasts which is smaller at the
-shortest and longest forecast horizons. In comparing the forecasting
-accuracy of Prophet, we find that the p-value is more significant for
-the shortest and longest forecast horizon. This agrees with the MSE of
-the Prophet forecasts which is smaller at the longest and shortest
-forecast horizon. Both ARIMA and Prophet generate forecasts with the
-smallest MSE and the most significant p-value at the longest forecast
-horizon. In comparing the forecasting accuracy of ARIMA and Prophet over
-the three forecast horizons, we notice that the p-value obtained from
-the Mincer-Zarnowitz results is more significant for ARIMA than Prophet
-at every forecast horizon. This echoes the results obtained from the MSE
-which is marginally smaller for ARIMA than Prophet over all forecast
-horizons.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 4.png}
-\caption{Mincer-Zarnowitz results for all forecast horizons}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 5.png}
-\caption{MSE for all forecast horizons}
-\end{figure}
-
-\subsection{Evaluation on data with missing values and outliers
-present}\label{evaluation-on-data-with-missing-values-and-outliers-present}
-
-By following the Box-Jenkins methodology for selection of an ARIMA
-model, an ARIMA(3,0,0) was employed to fit the training set which
-included missing values and outliers. Although the ARIMA(0,0,3) had a
-slightly smaller AIC, a decision was made to use the ARIMA(3,0,0)
-instead so that a rolling window forecast may be used. Analysis of the
-AR parameters indicate that they are statistically significant in the
-explanation of the movement of BTC/ZAR. The Prophet model fitted to the
-training set included a linear trend and weekly seasonality. All other
-parameters were set to Prophet's default values as described in section
-\ref{Semi-Automatic Selection}. The automatic-selection procedure for
-Prophet and ARIMA selected the same models as those chosen manually.
-
-\subsubsection{1-step ahead Forecasts}\label{step-ahead-forecasts-1}
-
-Table 6 shows the MSE of the forecasts generated by ARIMA and Prophet
-and the Mincer-Zarnowitz results obtained from regressing the test set
-returns against the 1-step ahead returns forecasted. We observe that a
-positive intercept is estimated in the case of ARIMA and Prophet,
-indicating that both models systematically underestimate the 1-step
-ahead returns. The intercept estimated for ARIMA is smaller than
-Prophet, indicating that the mean forecasted returns yielded by ARIMA is
-closer to the mean observed returns. The intercept estimates of 0.011895
-and 0.014802 for ARIMA and Prophet respectively are larger than the
-intercept estimates of 0.009546 and 0.013702, which was based on the
-``clean'' dataset. This suggests that both models systematically
-underestimates the returns more when there are missing values and
-outliers present in the data.
-
-We also observe that the slope estimate is small and negative for both
-models with the slope estimated for ARIMA lying closer to -1. This
-suggests that the observed 1-step ahead returns are actually the
-opposite of the proposed 1-step ahead forecasts. The slope estimates of
--0.883228 and -0.676132 for ARIMA and Prophet respectively are less
-steep than the slope estimates of -2.005624 and -1.709744, which was
-based on the evaluation of data that had no values missing or outliers
-present.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 6.png}
-\caption{Mincer-Zarnowitz results for 1-step ahead forecasts}
-\end{figure}
+\caption{MCS, MZ-test and RMSE results based on dirty data \label{tab2}} 
+\scalebox{0.9}{
+\begin{tabular}{cccccccc}
+  \hline
+Forecast & Model & Model & MCS & $\alpha$ & $\beta$ & MZ-test & RMSE \\ 
+ Horizon &  & Rank & p-values &  &  & p-values &  \\ 
+   \hline
+1 & ARIMA & 1 & 1 & 0.011895 & -0.8832 & 0.00003182 & 0.045530 \\ 
+   & Prophet & 2 & 0.413 & 0.014802 & -0.6761 & 0.00000532 & 0.046210 \\ 
+   \hline
+30 & ARIMA & 1 & 1 & 0.114100 & -12.9853 & 0.1075 & 0.044758 \\ 
+   & Prophet & 2 & 0.0016* & 0.014400 & -0.6516 & 0.00000148 & 0.048180 \\ 
+   \hline
+90 & ARIMA & 1 & 1 & -0.226700 & 27.0635 & 0.01294 & 0.042302 \\ 
+   & Prophet & 2 & 0.0336* & -0.009291 & 0.6438 & 0.000764 & 0.045061 \\ 
+   \hline
+\end{tabular}
+}
+\end{table}
+
+We notice that a positive intercept and negative slope is still
+estimated at the 1 day and 1 month forecast horizon when the data is
+dirty. The intercept estimates are larger than the case of the clean
+data, suggesting that the biasness of the forecasts yielded by both
+models increases when there are outliers and missing values present in
+the data. Contrary to the results from the clean data, Prophet's
+intercept estimate lies closer to zero at the 1 month and 3 month
+forecast horizon. This indicates that the mean forecasted returns
+yielded by Prophet is closer to the mean observed returns when outliers
+and missing values are present in the data. The negative intercept and
+positive slope estimated for the 3 month horizon contrasts with the
+positive intercept and negative slope estimated in the case of the clean
+data. This shows that although the forecasts produced by both models are
+systematically overestimated at long forecast horizons when the data is
+dirty, they are no longer the opposite of the observed returns.
 
@@
 unity slope and zero intercept is approximately zero for ARIMA and
-Prophet, with the p-value for Prophet lying closer to zero than ARIMA.
-The null hypothesis is thus rejected at the 1\% significance level,
-indicating that the forecasts generated by Prophet and ARIMA are biased
-and/or inefficient. The MSE produced by both models agree with the
-Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE
-than Prophet. The results obtained from evaluating the forecasting
-accuracy of ARIMA and Prophet based on the ``clean'' dataset show
-similar findings, however led to a smaller MSE of 0.001942 and 0.001958
-and a more significant p-value of 0.025449 and 0.007152 for ARIMA and
-Prophet respectively.
-
-\subsubsection{30-steps ahead Forecasts}\label{steps-ahead-forecasts-2}
-
-Table 7 shows the MSE of the forecasts produced by ARIMA and Prophet and
-the Mincer-Zarnowitz results obtained from regressing the test set
-returns against the 30-step ahead returns forecasted. We observe that a
-positive intercept is estimated in the case of ARIMA and Prophet,
-indicating that both models systematically underestimate the 30-step
-ahead returns. The intercept estimated for ARIMA is larger than Prophet,
-indicating that the mean forecasted returns yielded by Prophet is closer
-to the mean observed returns. The intercept estimate of 0.11414 for
-ARIMA is larger than the estimate of 0.031320 which was based on the
-evaluation of the ``clean'' dataset. This suggests that the ARIMA model
-underestimates the returns more when there are missing values and
-outliers present in the data. The intercept estimate of 0.014396 for
-Prophet is less than the estimate of 0.017169 which was based on the
-evaluation of data that had no values missing or outliers present. This
-suggests that Prophet underestimates the returns less when there are
-missing values and outliers present in the data.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 7.png}
-\caption{Mincer-Zarnowitz results for 30-step ahead forecasts}
-\end{figure}
-
-We also observe that the slope estimate is large and negative for ARIMA
-and small and negative for Prophet. This suggests that the observed
-30-step ahead returns are actually the opposite of the proposed 30-step
-ahead forecasts. The slope estimate of -12.985335 for ARIMA is more
-steep than the slope estimate of -12.736757 which was based on the
-evaluation of data that had no values missing or outliers present. This
-suggests that the efficiency of the forecasted returns deteriorates when
-there are missing values and outliers present in the data. The slope
-estimate of -0.65161 for Prophet is less steep than the estimate
--2.621914, which was based on the evaluation of data that had no values
-missing or outliers present. This suggests that the efficiency of the
-forecasted returns improve when there are missing values and outliers
-present in the data.
-
-The p-value obtained from the F-test for the joint hypothesis of a unity
-slope and zero intercept is approximately zero for Prophet. The null
-hypothesis is thus rejected at the 1\% significance level, indicating
-that the 30-step ahead forecasts generated by Prophet are biased and/or
-inefficient. The p-value of 0.107532 for ARIMA is significant at the 5\%
-level, suggesting at the 30-step ahead forecasts produced by ARIMA are
-unbiased and/or efficient. MSE produced by both models agree with the
-Mincer-Zarnowitz results, with ARIMA producing a smaller MSE than
-Prophet. The results obtained from evaluating the forecasting accuracy
-of ARIMA and Prophet based on the ``clean'' dataset, led to a more
-significant p-value of 0.000377 for Prophet only, and a smaller MSE of
-0.001988 and 0.002075 for ARIMA and Prophet respectively.
-
-\subsubsection{90-steps ahead Forecasts}\label{steps-ahead-forecasts-3}
-
-Table 8 shows the MSE of the forecasts produced by ARIMA and Prophet and
-the Mincer-Zarnowitz results obtained from regressing the test set
-returns against the 90-step ahead returns forecasted. We observe that a
-negative intercept is estimated in the case of ARIMA and Prophet,
-indicating that both models systematically overestimate the 90-step
-ahead returns. The intercept estimated for Prophet lies closer to zero,
-indicating that the mean forecasted returns yielded by Prophet is closer
-to the mean observed returns. This is in contrast to the positive
-intercepts that were estimated based on the data that had no values
-missing or outliers present.
-
-We also observe that the slope estimate is large and positive for ARIMA
-however small and positive for Prophet. This is in contrast to the small
-and negative slope estimates which were based on the evaluation of data
-that had no values missing or outliers present.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 8.png}
-\caption{Mincer-Zarnowitz results for 90-step ahead forecasts}
-\end{figure}
-
-The p-values obtained from the F-test for the joint hypothesis of a
-unity slope and zero intercept are insignificant for ARIMA and Prophet,
-with the p-value for Prophet lying close to zero. The null hypothesis is
-thus rejected at the 5\% significance level, indicating that the
-90-steps ahead forecasts generated by Prophet and ARIMA are biased
-and/or inefficient. The MSE produced by both models agree with the
-Mincer-Zarnowitz results, with ARIMA producing a marginally smaller MSE
-than Prophet. The results obtained from evaluating the forecasting
-accuracy of ARIMA and Prophet based on data which had no missing values
-or outliers present contrasts these findings, with a p-value of 0.843066
-and 0.408675 for ARIMA and Prophet respectively.
-
-\subsubsection{Comparison over varying forecast
-horizons}\label{comparison-over-varying-forecast-horizons-1}
-
-Table 9 and 10 compares the p-values obtained from the Mincer-Zarnowitz
-tests and the MSE of the forecasts produced by ARIMA and Prophet over
-the three forecast horizons. In comparing the forecasting accuracy of
-the ARIMA model over the three forecast horizons, we observe that the
-p-value is more significant at the longer forecast horizons. This agrees
-with the MSE of the ARIMA forecasts which is smaller at the longer
-forecast horizons. In comparing the forecasting accuracy of Prophet, we
-find that the p-value is more significant for the longest forecast
-horizon. This agrees with the MSE of the Prophet forecasts which is
-smaller at the longest forecast horizon.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 9.png}
-\caption{Mincer-Zarnowitz results for all forecast horizons}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{C:/Users/Fiona/Documents/R/UCT_hons_proj/Draft/Figures/Evaluation of Prophet and ARIMA forecasts/Table 10.png}
-\caption{MSE for all forecast horizons}
-\end{figure}
-
-In comparing the forecasting accuracy of ARIMA and Prophet over the
-three forecast horizons, we notice that the p-value obtained from the
-Mincer-Zarnowitz results is more significant for ARIMA than Prophet at
-every forecast horizon. This echoes the results obtained from the MSE
-which is marginally smaller for ARIMA than Prophet over all forecast
-horizons.
+Prophet across all forecast horizons, with the p-value for Prophet lying
+closer to zero than ARIMA. The null hypothesis is thus rejected at the
+10\% significance level, indicating that the forecasts generated by
+Prophet and ARIMA are always biased and/or inefficient. This contrasts
+with the clean data which generated unbiased and/or efficient forecasts
+at the longest forecast horizon. The RMSE produced by both models agrees
+with the MZ-test and MCS results, with ARIMA producing a marginally
+smaller RMSE than Prophet across all forecast horizons, and the RMSE
+being larger than the case of the clean data.
 
@@
 
-The results show a clear difference in the effect of the forecast
-horizon in the case of the clean data. The results based on the MSE and
-the p-value obtained from the Mincer-Zarnowitz test surprisingly show
-that the ARIMA model produces superior forecasts of the returns at the
-longest forecast horizon. Similarly, Prophet produces superior forecasts
-of the returns at the longest forecast horizon. We notice that although
-the ARIMA model outperforms Prophet at every forecast horizon, the
-difference between the MSE's of the forecasts generated by both models
-is insignificant. Furthermore, the Mincer-Zarnowitz p-values for both
+The results show a clear difference in the predictive accuracy of the
+ARIMA and Prophet model in addition to the forecast horizon effect. The
+ARIMA model ranks ahead of Prophet when forecasting the returns of
+BTC/ZAR across all forecast horizons. These results hold true
+irrespective of whether there are missing values and outliers present in
+the data. Nevertheless, it is reassuring to see that both models have
+equal predictive accuracy when forecasting at short and long forecast
+horizons when the data is clean and at short forecast horizons when the
+data is dirty. Furthermore, the Mincer-Zarnowitz p-values for both
 models are significant at the same forecast horizons, suggesting that
 the efficiency and biasness of the return forecasts yielded by ARIMA and
-Prophet are synchronized at each forecast horizon. If a small amount of
-predictive accuracy is willing to be sacrificed by the analyst in the
-forecasting of BTC/ZAR, it might be useful to employ Prophet instead of
-ARIMA. This would enable traders with expert knowledge about Bitcoin,
-however no statistical knowledge about forecasting models, to
-incorporate their domain knowledge into the Prophet model through its
-intuitively adjustable parameters. Furthermore, as more information
-about events that may affect the price of Bitcoin becomes available, it
-may be included in the model and could potentially change the predictive
-accuracy of Prophet.
-
-The results based on the unclean data show that both ARIMA and Prophet
-are sensitive to the presence of outliers and missing values in addition
-to the forecast horizon. We observe that Prophet yields superior
-forecasts at the longest forecast horizon. Although the MSE suggests
-that the ARIMA model produces forecasts with minimal error at the
-longest forecast horizon, analysis of the p-value indicates that the
-forecasts are efficient and/or unbiased at 30-steps ahead only. It is
-evident that Prophet produces inferior results in comparison to the
-ARIMA model over all forecast horizons. This is unexpected as Prophet is
-said to be robust to the presence of outliers and missing values in the
-data, without the need for interpolation. We also notice that the
-forecasts produced by both models deteriorates when there are missing
-values and outliers present in the data, with the only efficient and/or
-unbiased forecasts yielded by the ARIMA model at a forecast horizon of
-30 days.
-
-The limitation of this study is that it looks specifically at BTC/ZAR.
-Different results may emerge for fiat currencies or the exchange rate
-between Bitcoin with other currencies. Hence caution should be exercised
-when generalising these results to other exchange rates. As a point of
-further research, more complex techniques such as neural networks could
-be used as a benchmark for comparison when evaluating the forecasts
-produced by Prophet.
+Prophet are synchronized at each forecast horizon. Hence, if the analyst
+is engaged in short-term or long-term trading, they could choose to
+employ the Prophet model instead of the ARIMA model. This would allow
+them to produce return forecasts that are as good as the ARIMA model,
+without having any expert knowledge about model building and selection.
+It would also provide them with the flexibility to customise the model
+to suit their needs and incorporate their domain knowledge about BTC/ZAR
+through the model's intuitively adjustable parameters. If the analyst is
+interested in generating unbiased and/or efficient forecasts with the
+least amount of forecast error, they could utilise either model, however
+should only forecast BTC/ZAR returns over long forecast horizons with
+data that do not include any outliers or missing values.
+
+The results from the dirty data show that the ARIMA and Prophet model
+are sensitive to the presence of outliers and missing values, with both
+models producing unbiased and/or inefficient forecasts and higher RMSEs
+across all forecast horizons. This is disappointing as Prophet is said
+to be robust when faced with outliers and can handle missing values
+without the need for interpolation unlike the ARIMA model. Since both
+models have equal predictive accuracy at the shortest forecast horizon,
+an analyst engaged in short-term trading who has no knowledge on how to
+pre-process BTC/ZAR data before forecasting, could apply the Prophet
+model as it would provide them with flexibility and ease of
+implementation without sacrificing any predictive accuracy.
+
+Although the Prophet model was eliminated from the Model Confidence Set
+at other forecast horizons, these results might change as more data and
+information about events which might impact the price of BTC/ZAR becomes
+available and is incorporated into the Prophet model. Furthermore, the
+RMSE shows that the difference between the forecast errors of ARIMA and
+Prophet is marginal, even in the case where both models don't have equal
+predictive accuracy. Hence, if the analyst is willing to sacrifice a
+small amount of predictive accuracy in return for the tractability and
+ease of implementation which Prophet provides, then they might favour
+Prophet over the ARIMA model.
+
+The attractive features of Prophet were not fully utilised in this paper
+due to the stationarity of log returns. Different results may emerge
+when forecasting the direction of returns or closing prices.
+Furthermore, the reader should exercise caution when generalising these
+results to other exchange rate data or to exchange rates which involve
+fiat currencies as results may differ. As a point of further research,
+more complex techniques such as Neural Networks could be used as a
+benchmark for comparison when evaluating the forecasts produced by
+Prophet.
 
@@
 
+\hypertarget{ref-box1970}{}
+Box, George EP, Gwilym M Jenkins, and G Reinsel. 1970. ``Forecasting and
+Control.'' \emph{Time Series Analysis} 3: 75.
+
 \hypertarget{ref-castillo2002}{}
@@
 
-\hypertarget{ref-spyros1997}{}
-Hibon, Michael, and Spyros Makridakis. 1997. ``ARMA Models and the
-Box--Jenkins Methodology.'' John Wiley \& Sons, Ltd.
-
-\hypertarget{ref-ince2006}{}
-Ince, Huseyin, and Theodore B Trafalis. 2006. ``A Hybrid Model for
-Exchange Rate Prediction.'' \emph{Decision Support Systems} 42 (2).
-Elsevier: 1054--62.
-
-\hypertarget{ref-kamruzzaman2003}{}
-Kamruzzaman, Joarder, and Ruhul A Sarker. 2003. ``Forecasting of
-Currency Exchange Rates Using Ann: A Case Study.'' In \emph{Neural
-Networks and Signal Processing, 2003. Proceedings of the 2003
-International Conference on}, 1:793--97. IEEE.
+\hypertarget{ref-fantazzini2016}{}
+Fantazzini, Dean, Erik Nigmatullin, Vera Sukhanovskaya, and Sergey
+Ivliev. 2016. ``Everything You Always Wanted to Know About Bitcoin
+Modelling but Were Afraid to Ask.''
+
+\hypertarget{ref-hansen2011}{}
+Hansen, Peter R, Asger Lunde, and James M Nason. 2011. ``The Model
+Confidence Set.'' \emph{Econometrica} 79 (2). Wiley Online Library:
+453--97.
+
+\hypertarget{ref-hyndman2014}{}
+Hyndman, Rob J, and George Athanasopoulos. 2014. \emph{Forecasting:
+Principles and Practice}. OTexts.
 
@@
 
-\hypertarget{ref-khashei2012}{}
----------. 2012. ``Hybridization of Autoregressive Integrated Moving
-Average (Arima) with Probabilistic Neural Networks (Pnns).''
-\emph{Computers \& Industrial Engineering} 63 (1). Elsevier: 37--45.
-
 \hypertarget{ref-lin2012}{}

### 2017-10-27 Draft/Code/draftv2.R

@@ -335,7 +335,7 @@ plot_one_step_arima <- ggplot(ExchangeRates, aes(x = TradeDate, y = DailyLogRetu
   xlab("Trade Date") +
   ylab("Daily Log Returns") +
   theme_minimal() + 
-  labs(title  = paste("One-step ahead forecasts from ARIMA(1,0,0)"))
+  labs(title  = paste("1-step ahead forecasts from ARIMA(1,0,0)"))
 
 # Plot of 1 step-ahead Prophet forecasts
 plot_one_step_proph <- ggplot(ExchangeRates, aes(x = TradeDate, y = DailyLogReturns)) +
@@ -345,11 +345,31 @@ plot_one_step_proph <- ggplot(ExchangeRates, aes(x = TradeDate, y = DailyLogRetu
   xlab("Trade Date") +
   ylab("Daily Log Returns") +
   theme_minimal() + 
-  labs(title  = paste("One-step ahead Prophet forecasts"))
+  labs(title  = paste("1-step ahead Prophet forecasts"))
 
 grid.arrange(plot_one_step_arima,
              plot_one_step_proph,
              nrow = 2)
+
+# One-step Model Confidence Set
+library(MCS)
+
+# One-step squared error losses
+one_step_arima_losses <- LossLevel(realized = One_Step_Results %>% select(actual),
+                                   evaluated = One_Step_Results %>% select(arima_forecast), 
+                                   which = "SE")
+
+one_step_proph_losses <- LossLevel(realized = One_Step_Results %>% select(actual),
+                                   evaluated = One_Step_Results %>% select(proph_forecast), 
+                                   which = "SE")
+
+one_step_losses <- data.frame(one_step_arima_losses, one_step_proph_losses)
+
+one_step_losses <- one_step_losses %>% 
+  select(ARIMA = arima_forecast, Prophet = proph_forecast)
+
+# One-step MCS
+one_step_MCS <- MCSprocedure(Loss = one_step_losses, alpha = 0.10, B = 5000, statistic = "TR")
 
 # Forecasting 30-steps ahead for 6 month out-of-sample 
 h = 30
@@ -438,6 +458,25 @@ grid.arrange(plot_Thirty_step_arima,
              plot_Thirty_step_proph,
              nrow = 2)
 
+# Thirty-step Model Confidence Set
+
+# Thirty-step squared error losses
+thirty_step_arima_losses <- LossLevel(realized = Thirty_Step_Results %>% select(actual),
[…]

### 2017-10-27 Draft/Code/missing values and outliers.R

@@ -4,7 +4,7 @@
 # Let's test this out
 
 missing_eval <- function(Data){
-  
+
 # Insert missing values
 ExchangeRates <- ExchangeRates %>% select(TradeDate, DailyLogReturns)
 ExchangeRates[c(4, 10, 17, 20, 35, 48, 50, 100, 110, 177, 140, 157, 200, 230, 245, 270, 286, 300, 323, 330),2] <- NA
@@ -204,7 +204,7 @@ plot_one_step_arima <- ggplot(ExchangeRates, aes(x = TradeDate, y = DailyLogRetu
   xlab("Trade Date") +
   ylab("Daily Log Returns") +
   theme_minimal() + 
-  labs(title  = paste("One-step ahead forecasts from ARIMA(3,0,0)"))
+  labs(title  = paste("1-step ahead forecasts from ARIMA(3,0,0)"))
 
 # Plot of 1 step-ahead Prophet forecasts
 plot_one_step_proph <- ggplot(ExchangeRates, aes(x = TradeDate, y = DailyLogReturns)) +
@@ -214,11 +214,31 @@ plot_one_step_proph <- ggplot(ExchangeRates, aes(x = TradeDate, y = DailyLogRetu
   xlab("Trade Date") +
   ylab("Daily Log Returns") +
   theme_minimal() + 
-  labs(title  = paste("One-step ahead Prophet forecasts"))
+  labs(title  = paste("1-step ahead Prophet forecasts"))
 
 grid.arrange(plot_one_step_arima,
              plot_one_step_proph,
              nrow = 2)
+
+# One-step Model Confidence Set
+library(MCS)
+
+# One-step squared error losses
+one_step_arima_losses <- LossLevel(realized = One_Step_Results %>% select(actual),
+                                   evaluated = One_Step_Results %>% select(arima_forecast), 
+                                   which = "SE")
+
+one_step_proph_losses <- LossLevel(realized = One_Step_Results %>% select(actual),
+                                   evaluated = One_Step_Results %>% select(proph_forecast), 
+                                   which = "SE")
+
+one_step_losses <- data.frame(one_step_arima_losses, one_step_proph_losses)
+
+one_step_losses <- one_step_losses %>% 
+  select(ARIMA = arima_forecast, Prophet = proph_forecast)
+
+# One-step MCS
[…]

### 2017-10-27 Draft/Write-up/Draft.Rmd
Gaah! Typos fixed

@@
 The Model Confidence Set (MCS) procedure developed by @hansen2011 is a sequential procedure which starts with an initial set of models $M_0$ of size m.
-The worse model is then eliminated at each step until the null hypothesis of equal predictive ability (EPA) fails to be rejected for all the models belonging to the set at a given confidence level $1 - \alpha$.
+The worst model is then eliminated at each step until the null hypothesis of equal predictive ability (EPA) fails to be rejected for all the models belonging to the set at a given confidence level $1 - \alpha$.
 This smaller set of models $M^*$ is the superior set of models (SSM) with the best case occurring when the SSM consists of a single best model.
@@
 
-where $\mu_i,_j = E[d_i,_j]$ represents the expected loss differential, $d_i,_j,_t = l_i,_t - l_j,_t$ represents the loss differential between models $i$ and $j$ at time t, and $l_i,_t$ and $l_j,_t$ are the loss functions associated with model $i$ and $j$ at time t respectively.
+where $\mu_i,_j = E[d_i,_j,_t]$ represents the expected loss differential, $d_i,_j,_t = l_i,_t - l_j,_t$ represents the loss differential between models $i$ and $j$ at time t, and $l_i,_t$ and $l_j,_t$ are the loss functions associated with model $i$ and $j$ at time t respectively.
 This paper makes use of the squared forecast error as a loss function, however any arbitrary loss function that satisfies the weak stationarity conditions described in @hansen2011 may be used.
@@
 $$ T_R = \max_{i,j\in M} \;\; |t_i,_j|$$
-where $d_i,_j = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }$ measures the sample average loss differential between model $i$ and $j$ and  $\hat{var[\bar{d_i,_j}]}$ is estimated through bootstrapping as described in @hansen2011.
+where $\bar{d_i,_j} = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }$ measures the sample average loss differential between model $i$ and $j$ and  $\hat{var[\bar{d_i,_j}]}$ is estimated through bootstrapping as described in @hansen2011.
 
 This paper uses a block–bootstrap procedure of 5000 resamples as in @hansen2011.
-The block length $p$ is chosen by fitting an AR(p) model to the $d_i,_j$ terms and determining the maximum number of significant AR parameters obtained.
+The block length $p$ is chosen by fitting an AR(p) model to the $d_i,_j,_t$ terms and determining the maximum number of significant AR parameters obtained.
 While $t_i,_j$ has a t-distribution, $T_R$ has a non-standard distribution under the null hypothesis hence is estimated by implementing a bootstrapping procedure similar to that used to estimate $\hat{var[\bar{d_i,_j}]}$.

### 2017-10-27 Draft/Write-up/Draft.tex
Gaah! Typos fixed

@@
 Nason (2011) is a sequential procedure which starts with an initial set
-of models \(M_0\) of size m. The worse model is then eliminated at each
+of models \(M_0\) of size m. The worst model is then eliminated at each
 step until the null hypothesis of equal predictive ability (EPA) fails
@@
 
-where \(\mu_i,_j = E[d_i,_j]\) represents the expected loss
+where \(\mu_i,_j = E[d_i,_j,_t]\) represents the expected loss
 differential, \(d_i,_j,_t = l_i,_t - l_j,_t\) represents the loss
@@
 \vspace{1pt} \[ T_R = \max_{i,j\in M} \;\; |t_i,_j|\] where
-\(d_i,_j = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }\) measures the sample
+\(\bar{d_i,_j} = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }\) measures the sample
 average loss differential between model \(i\) and \(j\) and
@@
 Hansen, Lunde, and Nason (2011). The block length \(p\) is chosen by
-fitting an AR(p) model to the \(d_i,_j\) terms and determining the
+fitting an AR(p) model to the \(d_i,_j,_t\) terms and determining the
 maximum number of significant AR parameters obtained. While \(t_i,_j\)

### 2017-11-06 Draft/Write-up/Draft.Rmd
Final Paper

@@
 Prophet is also robust to outliers and can handle missing values in the time series without the need for interpolation.
-This paper compares the out-of-sample forecasts of Prophet and ARIMA over varying forecast horizons, using a Mincer-Zarnowitz approach for forecast evaluation.
+This paper compares the out-of-sample forecasts of Prophet and ARIMA over varying forecast horizons, using the Model Confidence Set and Mincer-Zarnowitz test to evaluate forecast accuracy.
 A complete dataset of the prices of Bitcoin/Rand as well as a dataset which has missing prices and outliers present is used in the analysis.
-It is found that the ARIMA model produces the most efficient and unbiased forecasts over all forecast horizons, and that the ARIMA model is more robust when faced with missing values and outliers present in the data.
+It is found that the ARIMA model ranks ahead of Prophet over all forecast horizons and is more robust when faced with missing values and outliers.
+In addition, both models having equal predictive accuracy when forecasting at short and long forecast horizons when the data is clean and at short forecast horizons when the data is dirty, with unbiased and/or efficient forecasts only generated over the longest forecast horizon in the case of the clean data.
 ---
@@
 # Introduction \label{Introduction}
-The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that have attracted attention in financial market forecasting [@khashei2009].
+The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has attracted attention in financial market forecasting [@khashei2009].
 Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that foreign exchange markets are not efficient and believe that future prices depend on current and past events [@abu1996].
@@
 In December 2016, Facebook open-sourced their forecasting model Prophet.
-Prophet differs from traditional time series model such as the ARIMA in that it can produce high quality forecasts in a straightforward way.
+Prophet differs from traditional time series models such as the ARIMA in that it can produce high quality forecasts in a straightforward way.
 Prophet is a sophisticated model that provides informative results, however is configurable and easy to use.
@@
 @fahimifard2009 applied a GARCH model to the Rial/USD and Rial/EUR to compare the forecasts yielded by the GARCH and ARIMA model over varying forecast horizons.
-Their results illustrated that the GARCH model outperformed the ARIMA model when using the Root Mean Square Error (RMSE), Mean Square Error (MSE) and Mean Absolute Deviation (MAD) as a performance criteria.
+Their results illustrated that the GARCH model outperformed the ARIMA model when using the Root Mean Square Error (RMSE), Mean Square Error (MSE) and Mean Absolute Deviation (MAD) as performance criteria.
 
@@
 ## Model Confidence Set
-The Model Confidence Set (MCS) procedure developed by @hansen2011 is a sequential procedure which starts with an initial set of models $M_0$ of size m.
+The Model Confidence Set (MCS) for a set of models is analogous to a confidence interval for a model parameter.
+The MCS procedure developed by @hansen2011 is a sequential procedure that starts with an initial set of models $M_0$ of size m.
 The worst model is then eliminated at each step until the null hypothesis of equal predictive ability (EPA) fails to be rejected for all the models belonging to the set at a given confidence level $1 - \alpha$.
@@
 $$ T_R = \max_{i,j\in M} \;\; |t_i,_j|$$
-where $\bar{d_i,_j} = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }$ measures the sample average loss differential between model $i$ and $j$ and  $\hat{var[\bar{d_i,_j}]}$ is estimated through bootstrapping as described in @hansen2011.
+where $\bar{d_i,_j} = m^{-1}\sum_{t=1}^{m}{d_i,_j,_t }$ measures the sample average loss differential between model $i$ and $j$ and  $\hat{var[\bar{d_i,_j}]}$ is estimated through the bootstrapping procedure described in @hansen2011.
 
-This paper uses a block–bootstrap procedure of 5000 resamples as in @hansen2011.
+This paper uses a block-bootstrap procedure of 5000 resamples.
 The block length $p$ is chosen by fitting an AR(p) model to the $d_i,_j,_t$ terms and determining the maximum number of significant AR parameters obtained.
-While $t_i,_j$ has a t-distribution, $T_R$ has a non-standard distribution under the null hypothesis hence is estimated by implementing a bootstrapping procedure similar to that used to estimate $\hat{var[\bar{d_i,_j}]}$.
+While $t_i,_j$ has a t-distribution, $T_R$ has a non-standard asymptotic distribution under the null hypothesis hence is estimated by implementing a bootstrapping procedure similar to that used to estimate ${var[\bar{d_i,_j}]}$.
 
-The worst model is then eliminated according to an elimination rule which is coherent with the calculated t-statistic and is defined as:
+The worst model is then eliminated according to an elimination rule that is coherent with the calculated t-statistic and is defined as:
 
 $$e_R = \arg max_{i}\bigg(\sup_{j\in M}\frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}\bigg)$$
-
 ## Mincer-Zarnowitz Approach to Forecast Evaluation 

### 2017-11-06 Draft/Write-up/Draft.tex
Final Paper

@@
 interpolation. This paper compares the out-of-sample forecasts of
-Prophet and ARIMA over varying forecast horizons, using a
-Mincer-Zarnowitz approach for forecast evaluation. A complete dataset of
-the prices of Bitcoin/Rand as well as a dataset which has missing prices
-and outliers present is used in the analysis. It is found that the ARIMA
-model produces the most efficient and unbiased forecasts over all
-forecast horizons, and that the ARIMA model is more robust when faced
-with missing values and outliers present in the data.
+Prophet and ARIMA over varying forecast horizons, using the Model
+Confidence Set and Mincer-Zarnowitz test to evaluate forecast accuracy.
+A complete dataset of the prices of Bitcoin/Rand as well as a dataset
+which has missing prices and outliers present is used in the analysis.
+It is found that the ARIMA model ranks ahead of Prophet over all
+forecast horizons and is more robust when faced with missing values and
+outliers. In addition, both models having equal predictive accuracy when
+forecasting at short and long forecast horizons when the data is clean
+and at short forecast horizons when the data is dirty, with unbiased
+and/or efficient forecasts only generated over the longest forecast
+horizon in the case of the clean data.
 \noindent\\
@@
 The Autoregressive Integrated Moving Average (ARIMA) model is one of the
-most widely used time series models that have attracted attention in
+most widely used time series models that has attracted attention in
 financial market forecasting (Khashei, Bijari, and Ardali 2009).
@@
 In December 2016, Facebook open-sourced their forecasting model Prophet.
-Prophet differs from traditional time series model such as the ARIMA in
+Prophet differs from traditional time series models such as the ARIMA in
 that it can produce high quality forecasts in a straightforward way.
@@
 outperformed the ARIMA model when using the Root Mean Square Error
-(RMSE), Mean Square Error (MSE) and Mean Absolute Deviation (MAD) as a
+(RMSE), Mean Square Error (MSE) and Mean Absolute Deviation (MAD) as
 performance criteria.
@@
 robust evaluation system that allows for a large number of forecasts be
-evaluated and compared. This is Facebook's definition of forecasting at
-scale.
+evaluated and compared. This is Facebookâs definition of forecasting
+at scale.
 
-Prophet's Bayesian approach to forecasting allows the analyst to
+Prophetâs Bayesian approach to forecasting allows the analyst to
 incorporate their expert knowledge into the model building process and
@@
 
-Prophet's ability to forecast at scale enables it to model a wide
+Prophetâs ability to forecast at scale enables it to model a wide
 variety of data and may be able to adequately fit exchange rate data.
@@
 non-linearity inherent in exchange rate data. Prophet performs well on
-data with strong multiple ``human scale'' seasonalities and historical
+data with strong multiple âhuman scaleâ seasonalities and historical
 trend changes. This differs to the ARIMA model which requires the data
@@
 methodology is then used to select the correct ARIMA model while
-Prophet's semi-automatic procedure is used for model selection. The
+Prophetâs semi-automatic procedure is used for model selection. The
 selected ARIMA and Prophet model are then fitted to the training set and
@@
 
-This paper makes use of Prophet's default settings to fit the data,
+This paper makes use of Prophetâs default settings to fit the data,
 however weekly seasonality is included in the model. This allows us to
@@
 
-The Model Confidence Set (MCS) procedure developed by Hansen, Lunde, and
-Nason (2011) is a sequential procedure which starts with an initial set
-of models \(M_0\) of size m. The worst model is then eliminated at each
-step until the null hypothesis of equal predictive ability (EPA) fails
-to be rejected for all the models belonging to the set at a given
-confidence level \(1 - \alpha\). This smaller set of models \(M^*\) is
-the superior set of models (SSM) with the best case occurring when the
-SSM consists of a single best model.
+The Model Confidence Set (MCS) for a set of models is analogous to a
+confidence interval for a model parameter. The MCS procedure developed
+by Hansen, Lunde, and Nason (2011) is a sequential procedure that starts
+with an initial set of models \(M_0\) of size m. The worst model is then
+eliminated at each step until the null hypothesis of equal predictive
+ability (EPA) fails to be rejected for all the models belonging to the
+set at a given confidence level \(1 - \alpha\). This smaller set of
+models \(M^*\) is the superior set of models (SSM) with the best case
+occurring when the SSM consists of a single best model.
 
@@
 average loss differential between model \(i\) and \(j\) and
-\(\hat{var[\bar{d_i,_j}]}\) is estimated through bootstrapping as
-described in Hansen, Lunde, and Nason (2011).
+\(\hat{var[\bar{d_i,_j}]}\) is estimated through the bootstrapping
+procedure described in Hansen, Lunde, and Nason (2011).
 
-This paper uses a block--bootstrap procedure of 5000 resamples as in
-Hansen, Lunde, and Nason (2011). The block length \(p\) is chosen by
-fitting an AR(p) model to the \(d_i,_j,_t\) terms and determining the
-maximum number of significant AR parameters obtained. While \(t_i,_j\)
-has a t-distribution, \(T_R\) has a non-standard distribution under the
-null hypothesis hence is estimated by implementing a bootstrapping
-procedure similar to that used to estimate \(\hat{var[\bar{d_i,_j}]}\).
+This paper uses a block-bootstrap procedure of 5000 resamples. The block
+length \(p\) is chosen by fitting an AR(p) model to the \(d_i,_j,_t\)
+terms and determining the maximum number of significant AR parameters
+obtained. While \(t_i,_j\) has a t-distribution, \(T_R\) has a
+non-standard asymptotic distribution under the null hypothesis hence is
+estimated by implementing a bootstrapping procedure similar to that used
+to estimate \({var[\bar{d_i,_j}]}\).
 
-The worst model is then eliminated according to an elimination rule
-which is coherent with the calculated t-statistic and is defined as:
+The worst model is then eliminated according to an elimination rule that
+is coherent with the calculated t-statistic and is defined as:
 
 \[e_R = \arg max_{i}\bigg(\sup_{j\in M}\frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}\bigg)\]
-
-\subsection{Mincer-Zarnowitz Approach to Forecast
-Evaluation}\label{mincer-zarnowitz-approach-to-forecast-evaluation}
-
-The Mincer-Zarnowitz approach to evaluating forecast accuracy is
-commonly used and can be useful when comparing the forecasts produced by
-Prophet and the ARIMA model. Mincer and Zarnowitz (1969) proposed an
-absolute measure to evaluate forecast accuracy that considers the
-distance between the actual and predicted values. To analyse the
-absolute errors produced by the forecasts, the observed values \(r_t\)
-are regressed against the predicted values \(\hat{r_t}\), i.e.
+\#\# Mincer-Zarnowitz Approach to Forecast Evaluation The
+Mincer-Zarnowitz approach to evaluating forecast accuracy is commonly
+used and can be useful when comparing the forecasts produced by Prophet
+and the ARIMA model. Mincer and Zarnowitz (1969) proposed an absolute
+measure to evaluate forecast accuracy that considers the distance
+between the actual and predicted values. To analyse the absolute errors
+produced by the forecasts, the observed values \(r_t\) are regressed
+against the predicted values \(\hat{r_t}\), i.e.
 \[r_t = \alpha + \beta\cdot\hat{r_t} + e_t\] where \(\alpha\) represents
@@
 randomly inserted in place of existing values. Both models are then
-fitted to this ``dirty'' dataset and the forecasts are evaluated in the
-same way as the ``clean'' dataset.
+fitted to this âdirtyâ dataset and the forecasts are evaluated in
+the same way as the ``clean'' dataset.
 
@@
 models increases when there are outliers and missing values present in
-the data. Contrary to the results from the clean data, Prophet's
+the data. Contrary to the results from the clean data, Prophetâs
 intercept estimate lies closer to zero at the 1 month and 3 month
@@
 to suit their needs and incorporate their domain knowledge about BTC/ZAR
-through the model's intuitively adjustable parameters. If the analyst is
-interested in generating unbiased and/or efficient forecasts with the
+through the modelâs intuitively adjustable parameters. If the analyst
+is interested in generating unbiased and/or efficient forecasts with the
 least amount of forecast error, they could utilise either model, however
@@
 RMSE shows that the difference between the forecast errors of ARIMA and
-Prophet is marginal, even in the case where both models don't have equal
-predictive accuracy. Hence, if the analyst is willing to sacrifice a
-small amount of predictive accuracy in return for the tractability and
+Prophet is marginal, even in the case where both models donât have
+equal predictive accuracy. Hence, if the analyst is willing to sacrifice
+a small amount of predictive accuracy in return for the tractability and
 ease of implementation which Prophet provides, then they might favour

### 2017-11-06 Draft/Code/draftv2.R

@@ -999,115 +999,6 @@ kable(table_MSE)
 source("~/R/UCT_hons_proj/Draft/Code/missing values and outliers.R")
 missing_eval(ExchangeRates)
 
-# --------- Is THERE AN OPPORTUNITY IN FORECASTING BITCOIN? A LOOK AT THE EMH ---------
-
-# 1-step ahead trading profit with prophet
-one_step_proph_profit <- sum(with(One_Step_Results, (proph_forecast>0)*actual + (proph_forecast<0)*-actual))
-one_step_proph_profit
-# profit = 0.1894012
-
-# 1-step ahead trading profit with arima
-one_step_arima_profit <- sum(with(One_Step_Results, (arima_forecast>0)*actual + (arima_forecast<0)*-actual))
-one_step_arima_profit
-# profit = 0.4903809
-
-# Profit from random trade signals
-num_simulations <- 10000
-
-random_return <- numeric(length = num_simulations)
-
-for(i in seq(num_simulations)) {
-  random_vector <- rnorm(nrow(One_Step_Results))
-  
-  random_return[i] <- sum(with(One_Step_Results, (random_vector>0)*actual +(random_vector<0)*-actual))
-}
-
-# Comparison of 1-step ahead profit
-comp_1_step_profit <- data.frame(random_return, one_step_proph_profit, one_step_arima_profit)
-
-plot_1_step_profit <- ggplot(comp_1_step_profit, aes(random_return)) +
-  geom_histogram() +
-  geom_vline(aes(xintercept = one_step_proph_profit, colour = "prophet profit"), size = 1) +
-  geom_vline(aes(xintercept = one_step_arima_profit, colour = "arima profit"), size = 1) +
-  scale_color_manual("", breaks = c("prophet profit", "arima profit"), values = c("prophet profit" = "blue", "arima profit" = "purple")) +
-  labs(title = paste("Comparison of Profit from Random Trade Signals")) +
-  xlab("returns from random trade signals") +
-  ylab("no. of simulations")
-
-plot_1_step_profit
-
-# 30-step ahead trading profit with prophet
-thirty_step_proph_profit <- sum(with(Thirty_Step_Results, (proph_forecast>0)*actual + (proph_forecast<0)*-actual))
-thirty_step_proph_profit
-# profit = 0.2005354
-
-# 30-step ahead trading profit with arima
[…]

### 2017-11-06 Draft/Write-up/Draft.Rmd
Definitely Final Paper

@@
 
-Prophet, like the GAM, frames the forecasting problem as a curve fitting exercise and uses backfitting to find the regression functions.
+Prophet, like the GAM, frames the forecasting problem as a curve-fitting exercise and uses backfitting to find the regression functions.
 This allows for the model to be fitted quickly and missing values and large outliers to be handled elegantly.
@@
 
-Prophet’s ability to forecast at scale enables it to model a wide variety of data and may be able to adequately fit exchange rate data.
+Prophet's ability to forecast at scale enables it to model a wide variety of data and may be able to adequately fit exchange rate data.
 Its non-linear components could capture the non-linearity present in exchange rates.
 In contrast, the ARIMA model is a linear function of previous observations and lagged residuals and fails to capture the non-linearity inherent in exchange rate data.
-Prophet performs well on data with strong multiple “human scale” seasonalities and historical trend changes.
+Prophet performs well on data with strong multiple "human scale" seasonalities and historical trend changes.
 This differs to the ARIMA model which requires the data to be de-trended and the variance stabilised before the model can be fitted.
@@
 The training set starts on the 24th January 2016 and ends on the 16th January 2017, while the test set starts on the 17th January 2017 and ends on the 17th July 2017.
-The Box-Jenkins methodology is then used to select the correct ARIMA model while Prophet’s semi-automatic procedure is used for model selection.
+The Box-Jenkins methodology is then used to select the correct ARIMA model while Prophet's semi-automatic procedure is used for model selection.
 The selected ARIMA and Prophet model are then fitted to the training set and rolling window forecasts are made 1 day, 30 days and 90 days ahead.
@@
 
-This paper makes use of Prophet’s default settings to fit the data, however weekly seasonality is included in the model.
+This paper makes use of Prophet's default settings to fit the data, however weekly seasonality is included in the model.
 This allows us to account for our prior knowledge about the closing prices of Bitcoin which tend to be lower around the weekend however higher mid-week due to fluctuations in trading activity.
@@
 In this way, a poor forecast will not have negative consequences on future forecasts to be made since the observed values will be used to correct itself for the remaining forecasts [@zhang1998].
-In this paper, the ARIMA and Prophet model is first fitted to the training set and the returns are forecasted h-days ahead for h = 1, 30, 90.
+In this paper, the ARIMA and Prophet models are first fitted to the training set and the returns are forecasted h-days ahead for h = 1, 30, 90.
 The size of the training set is then increased by one observation and the models are refitted.
@@
 $$e_R = \arg max_{i}\bigg(\sup_{j\in M}\frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}\bigg)$$
+
 ## Mincer-Zarnowitz Approach to Forecast Evaluation 

### 2017-11-06 Draft/Write-up/Draft.tex
Definitely Final Paper

@@
 robust evaluation system that allows for a large number of forecasts be
-evaluated and compared. This is Facebookâs definition of forecasting
-at scale.
+evaluated and compared. This is Facebook's definition of forecasting at
+scale.
 
-Prophetâs Bayesian approach to forecasting allows the analyst to
+Prophet's Bayesian approach to forecasting allows the analyst to
 incorporate their expert knowledge into the model building process and
@@
 
-Prophet, like the GAM, frames the forecasting problem as a curve fitting
+Prophet, like the GAM, frames the forecasting problem as a curve-fitting
 exercise and uses backfitting to find the regression functions. This
@@
 
-Prophetâs ability to forecast at scale enables it to model a wide
+Prophet's ability to forecast at scale enables it to model a wide
 variety of data and may be able to adequately fit exchange rate data.
@@
 non-linearity inherent in exchange rate data. Prophet performs well on
-data with strong multiple âhuman scaleâ seasonalities and historical
+data with strong multiple ``human scale'' seasonalities and historical
 trend changes. This differs to the ARIMA model which requires the data
@@
 methodology is then used to select the correct ARIMA model while
-Prophetâs semi-automatic procedure is used for model selection. The
+Prophet's semi-automatic procedure is used for model selection. The
 selected ARIMA and Prophet model are then fitted to the training set and
@@
 
-This paper makes use of Prophetâs default settings to fit the data,
+This paper makes use of Prophet's default settings to fit the data,
 however weekly seasonality is included in the model. This allows us to
@@
 itself for the remaining forecasts (Zhang and Hu 1998). In this paper,
-the ARIMA and Prophet model is first fitted to the training set and the
-returns are forecasted h-days ahead for h = 1, 30, 90. The size of the
-training set is then increased by one observation and the models are
+the ARIMA and Prophet models are first fitted to the training set and
+the returns are forecasted h-days ahead for h = 1, 30, 90. The size of
+the training set is then increased by one observation and the models are
 refitted. The next h-day ahead return is forecasted and this process is
@@
 \[e_R = \arg max_{i}\bigg(\sup_{j\in M}\frac{\bar{d_i,_j}}{\sqrt{\hat{var[\bar{d_i,_j}]}}}\bigg)\]
-\#\# Mincer-Zarnowitz Approach to Forecast Evaluation The
-Mincer-Zarnowitz approach to evaluating forecast accuracy is commonly
-used and can be useful when comparing the forecasts produced by Prophet
-and the ARIMA model. Mincer and Zarnowitz (1969) proposed an absolute
-measure to evaluate forecast accuracy that considers the distance
-between the actual and predicted values. To analyse the absolute errors
-produced by the forecasts, the observed values \(r_t\) are regressed
-against the predicted values \(\hat{r_t}\), i.e.
+
+\subsection{Mincer-Zarnowitz Approach to Forecast
+Evaluation}\label{mincer-zarnowitz-approach-to-forecast-evaluation}
+
+The Mincer-Zarnowitz approach to evaluating forecast accuracy is
+commonly used and can be useful when comparing the forecasts produced by
+Prophet and the ARIMA model. Mincer and Zarnowitz (1969) proposed an
+absolute measure to evaluate forecast accuracy that considers the
+distance between the actual and predicted values. To analyse the
+absolute errors produced by the forecasts, the observed values \(r_t\)
+are regressed against the predicted values \(\hat{r_t}\), i.e.
 \[r_t = \alpha + \beta\cdot\hat{r_t} + e_t\] where \(\alpha\) represents
@@
 randomly inserted in place of existing values. Both models are then
-fitted to this âdirtyâ dataset and the forecasts are evaluated in
-the same way as the ``clean'' dataset.
+fitted to this ``dirty'' dataset and the forecasts are evaluated in the
+same way as the ``clean'' dataset.
 
@@
 models increases when there are outliers and missing values present in
-the data. Contrary to the results from the clean data, Prophetâs
+the data. Contrary to the results from the clean data, Prophet's
 intercept estimate lies closer to zero at the 1 month and 3 month
@@
 to suit their needs and incorporate their domain knowledge about BTC/ZAR
-through the modelâs intuitively adjustable parameters. If the analyst
-is interested in generating unbiased and/or efficient forecasts with the
+through the model's intuitively adjustable parameters. If the analyst is
+interested in generating unbiased and/or efficient forecasts with the
 least amount of forecast error, they could utilise either model, however
@@
 RMSE shows that the difference between the forecast errors of ARIMA and
-Prophet is marginal, even in the case where both models donât have
-equal predictive accuracy. Hence, if the analyst is willing to sacrifice
-a small amount of predictive accuracy in return for the tractability and
+Prophet is marginal, even in the case where both models don't have equal
+predictive accuracy. Hence, if the analyst is willing to sacrifice a
+small amount of predictive accuracy in return for the tractability and
 ease of implementation which Prophet provides, then they might favour
